---
title: |
  CSAP Code Notebook 1.1
subtitle: |
  Data Preparation
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Building the Analysis Frame

This notebook turns the published **`eu25games2019`** dataset into the long-format
analysis frame this project estimates on: one row per respondent × game × round.

All wrangling that used to live here — merging Hahm et al.'s analysis frame with the
raw per-country Dynata exports, recovering the thermometer slots from the
country-specific `Q196_*` columns, and hand-coding ~1,750 party spellings against Party
Facts — has been extracted into a standalone, citable data repository:

> Muno, Tristan, and Thomas König. *eu25games2019: A cleaned three-wave EU elections
> survey (2019) across 25 countries*. Zenodo.
> <https://doi.org/10.5281/zenodo.21294634>

Source repository: <https://github.com/LS-Konig/eu25games2019> (v1.0.0).
The survey design and data collection underlying it are due to
[Hahm, Hilpert & König (2024)](https://doi.org/10.1017/S0003055423000266).

Consequently, notebooks `1.2_parties.qmd` (Party Facts crosswalk) and
`1.3_codebook.qmd` (variable labels + codebook) no longer exist: Party Facts IDs ship
with the data as `ext_*_pf_name` / `ext_*_pf_id`, and the codebook is generated upstream
at `eu25games2019/code/08_codebook.html`. That codebook — not this notebook — is the
authoritative variable reference.

## The input

`data/01_raw/eu25games2019.rds`: 103,685 × 847, **wide**, one row per respondent × wave,
xz-compressed to ~29 MB (>1 GB in memory once decompressed — budget for it).

| Prefix | Category |
|--------|----------|
| `q_` | Coded survey responses; `q_*_raw` keeps the original response string |
| `cj_` | Conjoint experiment attributes and game outcomes (waves 2–3) |
| `prime_` | Priming-experiment paradata |
| `meta_` | Respondent / wave metadata |
| `ext_` | Externally merged Party Facts IDs and names |
| `der_` | Derived variables (upstream: the three attention-check flags only) |

## What this notebook does

1. Load the published file (optionally re-fetching it from GitHub).
2. Check the shipped file still has the shape this code was written against.
3. Fold the French questionnaire fork (`*_fra`) back into the base columns.
4. Apply respondent-level cleaning.
5. Carry wave-1 covariates forward into missing wave-2/3 cells.
6. Reduce to one game-playing wave per respondent.
7. Reshape the six game slots from wide to long.

Output: `data/02_processed/eu25games2019_long.rds`, consumed by `2.1_key_variables.qmd`.

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)

# Install and load required packages
p_required <- c(
  "tidyverse", # dplyr, ggplot, tibbles etc
  "here", # relative and sys agnostic file paths
  "knitr", # kable()
  "sessioninfo" # documentation
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load Data

`download_fresh` controls where the data comes from. Set it to `TRUE` to pull the
current `main` of the source repository and overwrite the local copy; otherwise the
committed copy in `data/01_raw/` is read. The committed copy is what the manuscript
renders against, so leave this `FALSE` unless you deliberately want to move to a newer
upstream release — and re-run the contract check below if you do.

```{r}
#| label: load-data

download_fresh <- FALSE

if (download_fresh) {
  raw_url <- paste0(
    "https://raw.githubusercontent.com/LS-Konig/eu25games2019/",
    "main/data/03_final/eu25games2019.rds?cache_bust=1"
  )

  tmp <- tempfile(fileext = ".rds")
  download.file(raw_url, tmp, mode = "wb")

  # decompresses on read, takes a few seconds
  eu25games2019 <- read_rds(tmp)

  write_rds(
    eu25games2019,
    file = here("data", "01_raw", "eu25games2019.rds"),
    compress = "xz"
  )

  rm(raw_url, tmp)
} else {
  eu25games2019 <- read_rds(
    file = here("data", "01_raw", "eu25games2019.rds")
  )
}

dim(eu25games2019)
```

## Contract check

Everything downstream is written against a specific shape of the published file. An
upstream release that renames or drops a column would otherwise fail somewhere deep in
the analysis notebooks with a confusing error. These assertions fail fast instead.

The `_nationality_drawn` / `_nationality_shown` split matters in particular: the drawn
nationality is a flat background lottery, whereas the *shown* nationality is what the
respondent actually saw, because the treatment branch overrides the draw in every
condition except `4_somenat`. Earlier releases conflated the two under `cj_<slot>_nat`.

```{r}
#| label: verify-input-contract

slots <- c(
  "cj_dict1",
  "cj_dict2",
  "cj_dict3",
  "cj_trust1",
  "cj_trust2",
  "cj_trust3"
)

# note: the conjoint party attribute ships only as `_party_raw` (the displayed
# spelling) plus its Party Facts siblings -- there is no coded `_party` column
new_cols <- as.vector(outer(
  slots,
  c(
    "_nationality_drawn",
    "_nationality_shown",
    "_treatment",
    "_party_raw",
    "_pl1",
    "_pl2"
  ),
  paste0
))

old_cols <- c(
  paste0(slots, "_nat"),
  paste0(slots, "_nat_raw"),
  paste0("der_", slots, "_nationality")
)

pf_cols <- as.vector(outer(
  paste0("ext_", slots, "_party"),
  c("_pf_name", "_pf_id"),
  paste0
))

stopifnot(
  # dimensions roughly stable
  abs(nrow(eu25games2019) - 103685) < 50,
  abs(ncol(eu25games2019) - 847) < 5,
  # game columns present under their current names
  all(new_cols %in% names(eu25games2019)),
  # no superseded conflated names survive
  !any(old_cols %in% names(eu25games2019)),
  # Party Facts siblings for the conjoint party attribute
  all(pf_cols %in% names(eu25games2019)),
  # thermometer slots and their Party Facts siblings
  all(paste0("q_thermo_", 1:6) %in% names(eu25games2019)),
  all(paste0("ext_q_thermo_", 1:6, "_pf_name") %in% names(eu25games2019)),
  # attention checks
  all(paste0("der_att_check_", 1:3) %in% names(eu25games2019)),
  # no dictator/trust games in wave 1: all game-slot columns NA there
  # (the separate cj_task* conjoint instrument does run in wave 1)
  all(vapply(
    eu25games2019[
      eu25games2019$meta_wave == 1,
      grepl("^cj_(dict|trust)[123]_", names(eu25games2019))
    ],
    \(x) all(is.na(x)),
    logical(1)
  ))
)

rm(new_cols, old_cols, pf_cols)
```

# Code the two label-string scales

Two seven-point items — willingness to take risks and willingness to give something up
today for the future — were left in the published data as the *untranslated label
string* rather than as a code, e.g. `"(1) Completely unwilling to take risks"`,
`"(1) Völlig unwillens, Risiken einzugehen"`. Every language embeds the response code as
a leading digit run, which is the same rule the upstream pipeline applies to its other
ordinal scales, so we extract it here. Values outside 1–7 are scale-entry errors and
become `NA`.

This has to happen before the French reconciliation below, because for these two items
the French fork *is* already coded as an integer, so base and fork would otherwise be of
incompatible types.

```{r}
#| label: code-label-string-scales

label_string_scales <- c("q_risk_taking", "q_future_discount")

eu25games2019 <- eu25games2019 |>
  mutate(across(
    all_of(label_string_scales),
    ~ {
      code <- as.integer(str_extract(.x, "\\d+"))
      if_else(code >= 1 & code <= 7, code, NA_integer_)
    }
  ))

eu25games2019 |>
  select(all_of(label_string_scales)) |>
  pivot_longer(everything(), names_to = "variable", values_to = "code") |>
  count(variable, code) |>
  pivot_wider(names_from = variable, values_from = n) |>
  kable(caption = "Coded values of the two label-string scales.")

stopifnot(
  all(vapply(
    eu25games2019[label_string_scales],
    \(x) is.integer(x) && all(is.na(x) | (x >= 1 & x <= 7)),
    logical(1)
  ))
)
```

# Reconcile the French questionnaire fork

France was fielded through a separate Qualtrics block, so a subset of French responses
lands in parallel `*_fra` columns rather than in the base column. This is not cosmetic:
in wave 2 the base vote-recall column holds **8** French responses while the fork holds
**600**. Left unreconciled, France would effectively vanish from the vote-anchored group
— exactly the group this project is about.

We `coalesce()` each fork into its base column (base value wins where both exist) and
then drop the fork. `q_vote_choice_ep2014_fra` is the French label for the 2014 EP
recall, so it folds into `q_vote_choice_past2014`.

```{r}
#| label: coalesce-french-fork

# base <- fork, for the Party Facts sibling pairs
pf_fork_map <- tribble(
  ~base,                        ~fork,
  "ext_q_party_id",             "ext_q_party_id_fra",
  "ext_q_party_id_other",       "ext_q_party_id_other_fra",
  "ext_q_vote_choice_past",     "ext_q_vote_choice_past_fra",
  "ext_q_vote_choice_past_other", "ext_q_vote_choice_past_other_fra",
  "ext_q_vote_choice_hypo",     "ext_q_vote_choice_hypo_fra",
  "ext_q_vote_choice_hypo_other", "ext_q_vote_choice_hypo_other_fra",
  "ext_q_vote_choice_past2014", "ext_q_vote_choice_ep2014_fra",
  "ext_q_vote_choice_past2014_other", "ext_q_vote_choice_ep2014_other_fra"
)

coalesce_fork <- function(df, base, fork) {
  for (sfx in c("_pf_name", "_pf_id")) {
    b <- paste0(base, sfx)
    f <- paste0(fork, sfx)
    if (all(c(b, f) %in% names(df))) {
      df[[b]] <- coalesce(df[[b]], df[[f]])
      df[[f]] <- NULL
    }
  }
  df
}

eu25games2019 <- reduce2(
  pf_fork_map$base,
  pf_fork_map$fork,
  coalesce_fork,
  .init = eu25games2019
)

# plain attitudinal / demographic items: q_<name>_fra -> q_<name>
fra_items <- names(eu25games2019) |>
  str_subset("^q_.*_fra$") |>
  setdiff(str_subset(names(eu25games2019), "_raw$"))

type_clashes <- character()

for (f in fra_items) {
  b <- str_remove(f, "_fra$")
  if (!(b %in% names(eu25games2019))) next
  # a base/fork pair of differing type means the upstream recode treated the two
  # questionnaire versions differently; fold it in by hand rather than guess
  if (!identical(class(eu25games2019[[b]]), class(eu25games2019[[f]]))) {
    type_clashes <- c(type_clashes, b)
    next
  }
  eu25games2019[[b]] <- coalesce(eu25games2019[[b]], eu25games2019[[f]])
}

stopifnot(length(type_clashes) == 0)

# drop every remaining fork column, forks and their raw siblings alike
eu25games2019 <- eu25games2019 |>
  select(-matches("_fra(_raw)?$"), -matches("_fra_pf_(name|id)$"))
```

France must now be non-empty on the party-attachment item and on at least one vote item
in every wave in which those items were asked.

```{r}
#| label: verify-french-fork

fr_check <- eu25games2019 |>
  filter(meta_country == "France") |>
  summarise(
    n = n(),
    party_id = sum(!is.na(ext_q_party_id_pf_name)),
    vote_past = sum(!is.na(ext_q_vote_choice_past_pf_name)),
    vote_hypo = sum(!is.na(ext_q_vote_choice_hypo_pf_name)),
    vote_intended = sum(!is.na(ext_q_vote_choice_intended_pf_name)),
    .by = meta_wave
  ) |>
  arrange(meta_wave)

fr_check |>
  kable(caption = "France: non-missing responses per wave after folding in the fork.")

stopifnot(
  # attachment asked in all three waves
  all(fr_check$party_id > 0),
  # vote recall recovered in the game waves
  all(fr_check$vote_past[fr_check$meta_wave %in% c(2, 3)] > 100)
)
```

# Respondent cleaning

The old pipeline inherited an already-cleaned frame from Hahm et al. The published data
is the full uncleaned survey, so the cleaning decisions are now ours and are made
explicitly here. The rules follow the companion project `MBEU25`
(`code/02_data_prep.qmd`), so the two projects analyse comparable samples.

`apply_attention_check` is broken out as a switch because it is by far the most
consequential filter — see the sample-accounting table at the end of this section.

```{r}
#| label: cleaning-switches

apply_attention_check <- TRUE
```

Wave-1 respondents who never returned carry no games, but their wave-1 rows are the
source for the covariate backfill further down, so they are kept for now and dropped
after backfilling.

```{r}
#| label: filter-wave1-only

n_start <- nrow(eu25games2019)

pids_in_games <- eu25games2019 |>
  filter(meta_wave %in% c(2, 3)) |>
  distinct(meta_pid) |>
  pull(meta_pid)

eu25games2019 <- eu25games2019 |>
  filter(meta_wave %in% c(2, 3) | meta_pid %in% pids_in_games)

n_after_wave1only <- nrow(eu25games2019)
n_after_wave1only
```

The attention check asked respondents to ignore a question and type a fixed phrase into
an "Other" box. `der_att_check_3` is `TRUE` only when both the categorical response and
the free-text box were answered as instructed. It was fielded in all three waves.

```{r}
#| label: filter-attention-check

eu25games2019 |>
  count(meta_wave, der_att_check_3) |>
  pivot_wider(names_from = der_att_check_3, values_from = n) |>
  kable(caption = "Attention-check outcomes by wave, before filtering.")

if (apply_attention_check) {
  eu25games2019 <- eu25games2019 |>
    filter(der_att_check_3)
}

n_after_attention <- nrow(eu25games2019)
n_after_attention
```

Incomplete responses are dropped next. A small number of panelists then still appear
twice within a wave *with a completed game* — response timings suggest an invitation
link that was clicked twice rather than genuine repeat participation, and the duplicate
records are near-identical. They cannot be adjudicated, so both copies go.

```{r}
#| label: filter-completeness-and-duplicates

to_remove <- eu25games2019 |>
  filter(!is.na(cj_dict1_pl2), meta_finished) |>
  count(meta_wave, meta_pid) |>
  filter(n > 1) |>
  pull(meta_pid)

length(to_remove)

eu25games2019 <- eu25games2019 |>
  # only complete responses
  filter(meta_finished) |>
  # remove the within-wave duplicate panelists identified above
  filter(!(meta_pid %in% to_remove)) |>
  # keep only respondents who appear in a game wave, plus their wave-1 rows
  filter(meta_pid %in% meta_pid[meta_wave %in% c(2, 3)]) |>
  # three known duplicate wave-1 IDs
  filter(
    !(meta_wave == 1 &
      meta_pid %in% c("1078674746", "1277956255", "1516614375"))
  ) |>
  # and any wave-1 row with no ID at all
  filter(!(meta_wave == 1 & is.na(meta_pid)))

n_after_completeness <- nrow(eu25games2019)
n_after_completeness
```

# Carry wave-1 covariates forward

The games start in wave 2, so every item asked in wave 1 is a valid *pre-treatment*
measure for the same panelist. Where a wave-2/3 cell is missing and the panelist's
wave-1 value exists, we carry it forward, matched on `(meta_pid, meta_country)`.

Only `NA` cells are touched — existing values are never overwritten — and the audit
table below reports exactly how many cells each variable recovered. The recovered items
are almost entirely attitudinal (efficacy, populism and identity scales); carrying them
forward assumes temporal stability over the few months separating the waves, which is
acceptable because they enter only as respondent-level covariates, never as outcomes.

```{r}
#| label: backfill-wave1

# pre-game respondent items that exist in wave 1 and are used downstream
vars_pre_game <- c(
  "q_gender",
  "q_age",
  "q_identity_country",
  "q_identity_eu",
  "q_identity_europe",
  "q_religion",
  "q_class",
  "q_lr_self",
  "q_euint_self",
  "q_nativ_econ",
  "q_nativ_cult",
  "q_nat_int",
  "q_eu_int",
  "q_satis_demo_nat",
  "q_nat_efficacy_understand",
  "q_nat_efficacy_care",
  "q_nat_efficacy_nosay",
  "q_eu_efficacy_understand",
  "q_eu_efficacy_care",
  "q_eu_efficacy_nosay",
  "q_party_harm",
  "q_pop_will",
  "q_pop_people",
  "q_pop_citizen",
  "q_pop_reps",
  "q_pop_goodevil",
  "q_people_incompetent",
  "q_dem_compromise",
  "q_dem_listen",
  "q_tech_experts",
  "q_tech_leaders",
  "q_risk_taking",
  "q_future_discount",
  "q_rural_urban",
  "q_edu_age_stop"
)

stopifnot(all(vars_pre_game %in% names(eu25games2019)))

# each panelist's wave-1 values
w1_values <- eu25games2019 |>
  filter(meta_wave == 1) |>
  select(meta_pid, meta_country, all_of(vars_pre_game))

na_before <- eu25games2019 |>
  filter(meta_wave %in% c(2, 3)) |>
  summarise(across(all_of(vars_pre_game), ~ sum(is.na(.x)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "na_before")

# carry one variable forward from its wave-1 partner
backfill_w1 <- function(df, v) {
  w1 <- paste0(v, "_w1")
  df |>
    mutate(
      "filled_{v}" := meta_wave %in%
        c(2, 3) &
        is.na(.data[[v]]) &
        !is.na(.data[[w1]])
    ) |>
    mutate("{v}" := coalesce(.data[[v]], .data[[w1]]))
}

eu25games2019 <- eu25games2019 |>
  left_join(
    w1_values,
    by = c("meta_pid", "meta_country"),
    suffix = c("", "_w1")
  ) |>
  reduce(vars_pre_game, backfill_w1, .init = _)

na_after <- eu25games2019 |>
  filter(meta_wave %in% c(2, 3)) |>
  summarise(across(all_of(vars_pre_game), ~ sum(is.na(.x)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "na_after")

n_filled <- eu25games2019 |>
  filter(meta_wave %in% c(2, 3)) |>
  summarise(across(starts_with("filled_"), ~ sum(.x))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n_filled") |>
  mutate(variable = str_remove(variable, "^filled_"))

fill_audit <- na_before |>
  left_join(na_after, by = "variable") |>
  left_join(n_filled, by = "variable") |>
  filter(n_filled > 0) |>
  arrange(desc(n_filled))

fill_audit |>
  kable(
    caption = "Wave-1 backfill: wave-2/3 missing cells before and after, and cells recovered."
  )

cat("Total cells filled from wave 1:", sum(fill_audit$n_filled), "\n")

stopifnot(all(fill_audit$na_after == fill_audit$na_before - fill_audit$n_filled))

# drop helper columns
eu25games2019 <- eu25games2019 |>
  select(-ends_with("_w1"), -starts_with("filled_"))
```

# One game-playing wave per respondent

Wave-1 rows have now served their purpose and carry no games, so they are dropped.

What remains needs one more decision that `MBEU25` does not have to make. Roughly 9,500
panelists played the games in **both** wave 2 and wave 3. `MBEU25` keeps both
observations, because its models carry respondent-level random effects. This project is
different: its descriptive tables and its partisan-type framework are *respondent*-level
(`distinct(meta_pid)` throughout notebook 3.1), so a panelist present twice would be
counted twice, with an arbitrary wave silently winning each tabulation.

Hahm et al. resolved this the same way we do — their analysis frame contains each
respondent exactly once. We keep each panelist's **earliest** game wave, so wave 2 is
preferred wherever available and the retained measurement is the one closest to the
2019 European election.

```{r}
#| label: one-wave-per-respondent

eu25games2019 <- eu25games2019 |>
  # wave 1 carries no games
  filter(meta_wave %in% c(2, 3)) |>
  # and neither do wave-2/3 respondents who abandoned before the games
  filter(!is.na(cj_dict1_pl2)) |>
  # keep the earliest game wave per panelist
  slice_min(meta_wave, n = 1, by = meta_pid)

stopifnot(
  # one row per respondent
  nrow(eu25games2019) == n_distinct(eu25games2019$meta_pid),
  # both game waves survive
  all(c(2, 3) %in% eu25games2019$meta_wave)
)

eu25games2019 |>
  count(meta_wave) |>
  kable(caption = "Retained respondents by game wave.")
```

## Sample accounting

```{r}
#| label: tbl-sample-accounting

tibble(
  step = c(
    "published file (respondent x wave)",
    "wave 2/3, plus wave-1 rows of returning panelists",
    if (apply_attention_check) "passed the attention check" else NULL,
    "complete responses, within-wave duplicates removed",
    "played the games, earliest game wave only"
  ),
  n_rows = c(
    n_start,
    n_after_wave1only,
    if (apply_attention_check) n_after_attention else NULL,
    n_after_completeness,
    nrow(eu25games2019)
  )
) |>
  mutate(dropped = lag(n_rows) - n_rows) |>
  kable(caption = "Rows remaining after each cleaning step.")

cat(
  "Respondents in the analysis frame:",
  n_distinct(eu25games2019$meta_pid),
  "\n"
)
```

# Reshape the games to long format

The published data is wide: each of the six game slots (`cj_dict1..3`, `cj_trust1..3`)
carries a full set of attribute and outcome columns. Estimation needs one row per game
played.

Two details are worth stating, because both are easy to get wrong:

- The Party Facts siblings of the conjoint party attribute are named
  `ext_cj_<slot>_party_pf_*` and therefore do **not** match a `^cj_(dict|trust)\d+_`
  pattern. They are renamed into the slot namespace before the pivot and renamed back
  afterwards, so that `ext_cj_party_pf_name` — the column notebooks 2.1 and 3.2 expect —
  survives the reshape.
- The columns held fixed during the pivot must be excluded with
  `starts_with("cj_treatment")`, never `starts_with("cj_tr")`: the shorter prefix also
  matches `cj_trust*` and silently drops the entire Trust game.

```{r}
#| label: game-outcomes-to-long

n_respondents <- n_distinct(eu25games2019$meta_pid)

eu25_long <- eu25games2019 |>
  # move the Party Facts siblings into the slot namespace so the pivot sees them
  rename_with(
    ~ str_replace(.x, "^ext_(cj_(?:dict|trust)\\d+)_party_pf_", "\\1_partypf"),
    matches("^ext_cj_(dict|trust)\\d+_party_pf_(name|id)$")
  ) |>
  pivot_longer(
    cols = matches("^cj_(dict|trust)\\d+_"),
    names_to = c("game_type", "round", ".value"),
    names_pattern = "^(cj_dict|cj_trust)(\\d+)_(.+)$"
  ) |>
  # restore the cj_ prefix on the pivoted stems
  rename_with(
    ~ paste0("cj_", .x),
    -c(
      starts_with("meta_"),
      starts_with("q_"),
      starts_with("ext_"),
      starts_with("der_"),
      starts_with("prime_"),
      starts_with("cj_attr"),
      starts_with("cj_att_pos"),
      starts_with("cj_rand"),
      starts_with("cj_task"),
      starts_with("cj_treatment"),
      contains("_instr_")
    )
  ) |>
  rename(
    ext_cj_party_pf_name = cj_partypfname,
    ext_cj_party_pf_id = cj_partypfid
  ) |>
  mutate(
    # game identifiers in the form the analysis notebooks use
    meta_game = if_else(cj_game_type == "cj_dict", "dict", "trust"),
    meta_round = as.integer(cj_round)
  ) |>
  select(-cj_game_type, -cj_round)

stopifnot(
  # six slots per respondent, nothing dropped or duplicated
  nrow(eu25_long) == 6 * n_respondents,
  # evenly split across the two games
  all(table(eu25_long$meta_game) == 3 * n_respondents),
  # the outcome and the reshaped attributes came through
  all(
    c(
      "cj_pl1",
      "cj_pl2",
      "cj_treatment",
      "cj_nationality_shown",
      "cj_party_raw",
      "cj_eupos",
      "cj_rel",
      "cj_age",
      "cj_sex",
      "cj_class",
      "ext_cj_party_pf_name"
    ) %in%
      names(eu25_long)
  ),
  # allocations are on the 0-10 token scale
  min(eu25_long$cj_pl2, na.rm = TRUE) >= 0,
  max(eu25_long$cj_pl2, na.rm = TRUE) <= 10
)

dim(eu25_long)
```

## Equivalence with the previous pipeline

Three variables that used to be inherited from Hahm et al.'s derived analysis frame are
now reconstructed from the published raw survey. The correspondence is exact rather than
approximate, and is documented upstream:

| previously (Hahm et al.) | now (published) | note |
|---|---|---|
| `cj_token` (`pl2gets`) | `cj_pl2` | tokens the co-player receives, 0–10 |
| `cj_trmnt` (`trmnt`) | `cj_treatment` | same four conditions, `-` → `_` in the level names |
| `der_conational` (`nationality`) | `cj_nationality_shown` | `co-national` → `own_country`, `out-national-EU` → `eu`, `out-national-non-EU` → `non_eu` |
| `der_nopid` (`nopid`) | `ext_q_party_id_pf_name == "none"` | |

`eu25games2019/code/07_recode.qmd` states that `cj_<slot>_nationality_shown` reproduces
the effective `nationality` of Hilpert's analysis frame
(`prepare_dat_consld.R:2374–2391`), so this is a documented reimplementation of the
original coding, not an independent guess at it.

```{r}
#| label: tbl-treatment-check

eu25_long |>
  count(cj_treatment, cj_nationality_shown) |>
  kable(
    caption = "Shown nationality by treatment condition. The conational conditions force `own_country`, `3_eunat_eustance` forces `eu`, and only `4_somenat` passes the background draw through."
  )

stopifnot(
  all(
    eu25_long$cj_nationality_shown[
      grepl("conat", eu25_long$cj_treatment)
    ] ==
      "own_country"
  ),
  all(
    eu25_long$cj_nationality_shown[
      eu25_long$cj_treatment == "3_eunat_eustance"
    ] ==
      "eu"
  )
)
```

```{r}
#| label: tbl-token-distribution

eu25_long |>
  summarise(
    n = sum(!is.na(cj_pl2)),
    mean = round(mean(cj_pl2, na.rm = TRUE), 3),
    sd = round(sd(cj_pl2, na.rm = TRUE), 3),
    median = median(cj_pl2, na.rm = TRUE),
    .by = meta_game
  ) |>
  kable(caption = "Token allocations by game.")
```

# Save

```{r}
#| label: save-data

write_rds(
  eu25_long,
  file = here(
    "data",
    "02_processed",
    "eu25games2019_long.rds"
  ),
  compress = "xz"
)
```

# Session Info

```{r}
#| label: session-info

session_info()
```

# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds."
))
```
