---
title: |
  CSAP Code Notebook 1.1
subtitle: |
  ESS cumulative datafile: CSV to partitioned Parquet
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

This notebook converts the raw ESS extract `data/01_raw/ess/Datafile-subset.csv`
(1.5 GB, 2,836 columns, 430,581 rows, rounds 1-11) into a ZSTD-compressed Parquet
dataset partitioned by `cntry` and `essround`.

**It is a pure format conversion. Nothing is cleaned, recoded, filtered or dropped.**
Every ESS missing code (7/8/9, 66/77/88/99, 6666/7777/8888/9999 and the wider variants)
survives untouched, because the correct code depends on each variable's field width and
that decision belongs in a later layer, not in the raw one.

The point of the exercise is column pruning. Any given analysis touches a few dozen of
the 2,836 variables; Parquet lets a reader fetch exactly those, from exactly the
countries it needs, without parsing 1.5 GB of text.

What this notebook does:

1. Checks whether the Parquet dataset is already current, and skips everything if so.
2. Freezes the CSV schema once, using full-file type inference.
3. Audits that inference for the one failure mode that would silently corrupt data.
4. Streams the CSV into partitioned Parquet without ever holding it in R memory.
5. Publishes the result atomically and records what it did.
6. Validates row counts, column counts, schema, on-disk size and read performance.

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)

# A fresh R 4.6 user library has no CRAN mirror configured, and this notebook may be
# the first thing ever run against it. Without this, install.packages() errors out in
# a non-interactive session.
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Install and load required packages
#
# dplyr rather than the whole tidyverse: the only tidy verbs used here are filter(),
# select() and collect() in the arrow demo read, which does not justify pulling the
# metapackage into a virgin library.
p_required <- c(
  "here",
  "DBI",
  "duckdb",
  "arrow",
  "dplyr",
  "knitr",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Paths and Parameters

All paths are anchored on the project root with `here::here()`, so the notebook renders
from anywhere and never calls `setwd()`.

`conversion_mode` selects the write strategy. `"staged"` materialises the CSV into an
on-disk DuckDB table and then writes one round at a time; `"direct"` is the simpler
single-shot `COPY` straight off the CSV. Why `"staged"` is the default is explained
under *Conversion*.

```{r}
#| label: paths

csv_path      <- here::here("data", "01_raw", "ess", "Datafile-subset.csv")
ess_dir       <- dirname(csv_path)
pq_root       <- file.path(ess_dir, "parquet")
pq_tmp        <- file.path(ess_dir, ".parquet-tmp")
manifest_path <- file.path(ess_dir, "parquet-manifest.rds")

# Behaviour switches
conversion_mode     <- "staged"  # "staged" (default) or "direct"
audit_leading_zeros <- TRUE      # one extra CSV scan; see Type-Inference Audit
duckdb_threads      <- 8L
duckdb_memory       <- "32GB"

# Scratch space for the staging database and any spill. Outside the repo.
scratch_dir <- file.path(tempdir(), "ess_convert")
dir.create(scratch_dir, recursive = TRUE, showWarnings = FALSE)
stage_db  <- file.path(scratch_dir, "ess_stage.duckdb")
spill_dir <- file.path(scratch_dir, "duckdb_spill")

# DuckDB wants forward slashes inside SQL string literals on Windows.
sqlp <- function(x) gsub("\\\\", "/", x)

knitr::kable(
  data.frame(
    parameter = c(
      "csv_path", "pq_root", "manifest_path", "conversion_mode",
      "audit_leading_zeros", "duckdb_threads", "duckdb_memory", "scratch_dir"
    ),
    value = c(
      csv_path, pq_root, manifest_path, conversion_mode,
      audit_leading_zeros, duckdb_threads, duckdb_memory, scratch_dir
    )
  ),
  row.names = FALSE
)
```

# Is the Parquet Dataset Already Current?

The conversion is expensive and the input never changes on its own, so a re-render
should cost seconds, not minutes.

Three guards have to agree before the conversion is skipped. The manifest records the
CSV's size and mtime at conversion time, so a swapped or truncated input is caught even
if its mtime is preserved. The file count must still match. And the *oldest* Parquet
part must postdate the CSV -- `min()`, not `max()`, because "newer than the CSV" read
loosely would happily pass a dataset in which one early partition is stale.

```{r}
#| label: idempotency-check

csv_mtime <- file.mtime(csv_path)
csv_size  <- file.size(csv_path)

pq_files <- if (dir.exists(pq_root)) {
  list.files(pq_root, pattern = "[.]parquet$", recursive = TRUE, full.names = TRUE)
} else {
  character(0)
}

up_to_date <- FALSE
if (length(pq_files) > 0 && file.exists(manifest_path)) {
  manifest <- readRDS(manifest_path)
  up_to_date <-
    isTRUE(manifest$csv_size == csv_size) &&
    isTRUE(manifest$csv_mtime == csv_mtime) &&
    isTRUE(manifest$n_files == length(pq_files)) &&
    min(file.mtime(pq_files)) > csv_mtime
}

if (up_to_date) {
  message("Parquet dataset is current (", length(pq_files),
          " files). Skipping conversion.")
} else {
  message("Parquet dataset is missing or stale. Converting.")
}
```

# Freezing the CSV Schema

## Why full-file inference is not optional here

DuckDB's CSV sniffer normally reads a sample. On this file a sample gets two things
wrong, and both would be silent:

- `idno` reaches 510,130,400,606, which overflows a 32-bit integer. A sample that saw
  only small IDs would pick the narrow type.
- Fifteen columns hold quoted ISO country and language codes -- `cntbrtha`, `ctzshipa`,
  `fbrncnt`, `lnghomb`, `lnghom2`, `mbrncnt` and their variants -- with values like
  `"000"` and `"01"` sitting alongside `"ABK"` and `"AE"`. **DuckDB strips quotes before
  it decides a type**, so a sample that happened to see only the numeric-looking values
  would type the column as an integer and turn `"000"` into `0`. Reading the whole file
  finds the alphabetic values and keeps the column as text.

`sample_size = -1` reads the entire file for inference. It costs one full scan, once.

## Sniffing once and reusing the result

`sniff_csv()` returns a `Prompt` column holding a complete, ready-to-run `read_csv()`
call with the full `columns={...}` struct already written out. Using DuckDB's own
generated clause removes any chance of a hand-built schema disagreeing with what the
sniffer actually found, and every later query reuses that one frozen clause -- so no
subsequent read can re-infer anything differently.

```{r}
#| label: connect-duckdb
#| eval: !expr '!up_to_date'

con <- DBI::dbConnect(duckdb::duckdb())

invisible(DBI::dbExecute(con, "SET preserve_insertion_order = false"))
invisible(DBI::dbExecute(con, paste0("SET threads = ", duckdb_threads)))
invisible(DBI::dbExecute(con, paste0("SET memory_limit = '", duckdb_memory, "'")))
invisible(DBI::dbExecute(con, paste0("SET temp_directory = '", sqlp(spill_dir), "'")))
invisible(DBI::dbExecute(con, "SET max_temp_directory_size = '60GB'"))
invisible(DBI::dbExecute(con, "SET partitioned_write_max_open_files = 256"))
invisible(DBI::dbExecute(con, "SET partitioned_write_flush_threshold = 65536"))

knitr::kable(
  DBI::dbGetQuery(con, "
    SELECT name, value FROM duckdb_settings()
    WHERE name IN ('threads', 'memory_limit', 'temp_directory',
                   'preserve_insertion_order',
                   'partitioned_write_max_open_files',
                   'partitioned_write_flush_threshold')
    ORDER BY name"),
  row.names = FALSE
)
```

```{r}
#| label: sniff-schema
#| eval: !expr '!up_to_date'

t_sniff <- system.time({
  DBI::dbExecute(con, paste0(
    "CREATE OR REPLACE TEMP TABLE ess_sniff AS ",
    "SELECT * FROM sniff_csv('", sqlp(csv_path), "', sample_size = -1)"
  ))
})

# The generated FROM clause is 150-250 KB of SQL. Build queries with paste0(), never
# sprintf() -- a stray '%' in that text would be read as a format specifier.
from_clause <- DBI::dbGetQuery(con, "SELECT Prompt FROM ess_sniff")$Prompt
from_clause <- sub(";[[:space:]]*$", "", from_clause)

schema_csv <- DBI::dbGetQuery(con, "
  SELECT row_number() OVER () AS pos,
         s.col.name AS column_name,
         s.col.type AS column_type
  FROM (SELECT unnest(Columns) AS col FROM ess_sniff) s")

message("Sniff took ", round(t_sniff[["elapsed"]], 1), " s; ",
        nrow(schema_csv), " columns inferred.")
message("Generated FROM clause is ", format(nchar(from_clause), big.mark = ","),
        " characters. First 300:")
message(substr(from_clause, 1, 300))
```

```{r}
#| label: schema-type-summary
#| eval: !expr '!up_to_date'

knitr::kable(
  as.data.frame(sort(table(schema_csv$column_type), decreasing = TRUE)),
  col.names = c("inferred type", "columns"),
  row.names = FALSE
)
```

## Type-Inference Audit

Full-file inference handles the quoted-code columns, but it cannot help where a column
is *genuinely* all-numeric and still carries a leading zero. An ISCO occupation code
like `0110` would be read as `110` -- a different occupation -- with no error raised.

The audit below reads the file once with `all_varchar = true` and asks, for every column
the sniffer typed as an integer, whether any value has a leading zero. It is pure string
comparison with no casting, so it is cheap. Any column it names is rewritten to
`VARCHAR` in the frozen clause before the write.

```{r}
#| label: leading-zero-audit
#| eval: !expr '!up_to_date && audit_leading_zeros'

int_cols <- schema_csv$column_name[
  schema_csv$column_type %in% c("BIGINT", "INTEGER", "HUGEINT", "SMALLINT", "TINYINT")
]

offenders <- character(0)
if (length(int_cols) > 0) {
  aggs <- paste0(
    'max(CASE WHEN length("', int_cols, '") > 1 AND ("', int_cols,
    '" LIKE \'0%\' OR "', int_cols, '" LIKE \'-0%\') THEN 1 ELSE 0 END) AS "',
    int_cols, '"'
  )
  t_audit <- system.time({
    lz <- DBI::dbGetQuery(con, paste0(
      "SELECT ", paste(aggs, collapse = ", "),
      " FROM read_csv('", sqlp(csv_path), "', all_varchar = true)"
    ))
  })
  vals <- unlist(lz)
  offenders <- names(lz)[!is.na(vals) & vals == 1L]
  message("Audited ", length(int_cols), " integer-typed columns in ",
          round(t_audit[["elapsed"]], 1), " s.")
}

if (length(offenders) > 0) {
  message("Leading zeros found -- forcing these columns to VARCHAR:")
  print(offenders)
} else {
  message("No integer-typed column carries a leading zero. Nothing to patch.")
}
```

`edition` needs an override the audit cannot find. It holds version labels -- `"2.3"`,
`"4.6"`, `"6.7"` -- which parse cleanly as numbers, so the sniffer types them `DOUBLE`.
They are labels, not quantities, and pinning them to `VARCHAR` preserves the literal.
That is preservation, not recoding.

`proddate` and fifteen interview date/time columns are typed `DATE` and `TIMESTAMP`, and
that is left alone. The parse is unambiguous and correct -- `23.11.2023` becomes
`2023-11-23`, `02.07.2026` becomes `2026-07-02`, so the sniffer has read the day-first
format the right way round -- and a date column is the honest type for a date. It is
recorded here as a decision rather than left to be discovered later.

One thing deliberately *not* set: `nullstr = ''`. DuckDB already maps an unquoted empty
field to `NULL`, which is what the CSV means. Setting `nullstr` would additionally
collapse a quoted `""` into `NULL`, conflating "empty string" with "missing" -- a
semantic change, and out of scope for a format conversion.

```{r}
#| label: patch-schema
#| eval: !expr '!up_to_date'

force_varchar <- unique(c(offenders, "edition"))

patch_type <- function(clause, col, new_type = "VARCHAR") {
  pattern <- paste0("('", col, "'[[:space:]]*:[[:space:]]*)'[^']*'")
  sub(pattern, paste0("\\1'", new_type, "'"), clause)
}

for (cl in force_varchar) {
  from_clause <- patch_type(from_clause, cl)
  schema_csv$column_type[schema_csv$column_name == cl] <- "VARCHAR"
}

message("Forced to VARCHAR: ", paste(force_varchar, collapse = ", "))
```

# Conversion

## Why the obvious single COPY is not used

The natural statement is one `COPY ... TO ... (PARTITION_BY (cntry, essround))` reading
straight from the CSV. On this file that is a real risk of an out-of-memory kill, for a
reason specific to its shape.

DuckDB's partitioned writer keeps a 2,048-row buffer per partition per thread. At 2,836
columns that is roughly 46 MB per buffer, and this file has **222** non-empty
`cntry` x `essround` cells. Eight threads times 222 partitions is on the order of 80 GB
of allocator memory, which cannot spill to disk. Two defaults compound it:
`partitioned_write_flush_threshold` is 524,288 rows -- more than the entire file -- so
no thread ever flushes early, and `partitioned_write_max_open_files` is 100, below 222.

The fix is not to shrink the buffer but to cut how many partitions are open at once.
Staging the CSV into an on-disk DuckDB table and then writing one round at a time means
each `COPY` touches only the 18-23 countries present in that round. It also makes schema
consistency structural rather than hopeful: all eleven writes read the same table, so
there is no second inference pass that could disagree with the first.

The cost is two full CSV scans either way, because the frozen clause means the write
pass sniffs nothing.

```{r}
#| label: stage-table
#| eval: !expr '!up_to_date && conversion_mode == "staged"'

if (file.exists(stage_db)) unlink(stage_db, force = TRUE)
invisible(DBI::dbExecute(con, paste0("ATTACH '", sqlp(stage_db), "' AS stage")))

t_stage <- system.time({
  DBI::dbExecute(con, paste0(
    "CREATE OR REPLACE TABLE stage.ess_raw AS SELECT * ", from_clause
  ))
})

n_staged <- DBI::dbGetQuery(con, "SELECT count(*) AS n FROM stage.ess_raw")$n
message("Staged ", format(n_staged, big.mark = ","), " rows in ",
        round(t_stage[["elapsed"]], 1), " s (",
        round(file.size(stage_db) / 1024^3, 2), " GB on disk).")
```

```{r}
#| label: partitioned-copy
#| eval: !expr '!up_to_date'

if (dir.exists(pq_tmp)) unlink(pq_tmp, recursive = TRUE, force = TRUE)

copy_opts <- paste(
  "FORMAT PARQUET",
  "PARTITION_BY (cntry, essround)",
  "COMPRESSION zstd",
  "APPEND true",
  sep = ", "
)

t_write <- system.time({
  if (conversion_mode == "staged") {
    rounds <- DBI::dbGetQuery(
      con, "SELECT DISTINCT essround FROM stage.ess_raw ORDER BY essround"
    )$essround
    written <- vapply(rounds, function(k) {
      n <- DBI::dbExecute(con, paste0(
        "COPY (SELECT * FROM stage.ess_raw WHERE essround = ", k, ") ",
        "TO '", sqlp(pq_tmp), "' (", copy_opts, ")"
      ))
      message("  round ", k, ": ", format(n, big.mark = ","), " rows")
      as.numeric(n)
    }, numeric(1))
    n_rows_written <- sum(written)
  } else {
    n_rows_written <- as.numeric(DBI::dbExecute(con, paste0(
      "COPY (SELECT * ", from_clause, ") ",
      "TO '", sqlp(pq_tmp), "' (", copy_opts, ")"
    )))
  }
})

message("Wrote ", format(n_rows_written, big.mark = ","), " rows in ",
        round(t_write[["elapsed"]], 1), " s.")
```

# Publishing the Dataset

The write goes to `.parquet-tmp/` and is renamed to `parquet/` only once it has
succeeded. That is what makes the idempotency guard trustworthy: a `parquet/` directory
can only exist if a run finished, so there is no half-written state to detect. The
leading dot keeps `arrow` from treating an in-progress directory as part of the dataset.

DuckDB has to release its file handles before the rename, or Windows refuses it.

```{r}
#| label: rename-and-manifest
#| eval: !expr '!up_to_date'

duckdb_version <- as.character(utils::packageVersion("duckdb"))
DBI::dbDisconnect(con, shutdown = TRUE)

if (dir.exists(pq_root)) unlink(pq_root, recursive = TRUE, force = TRUE)
ok <- file.rename(pq_tmp, pq_root)
if (!ok) {
  Sys.sleep(2)
  ok <- file.rename(pq_tmp, pq_root)
}
stopifnot("directory rename failed - .parquet-tmp left in place" = ok)

pq_files <- list.files(pq_root, pattern = "[.]parquet$",
                       recursive = TRUE, full.names = TRUE)

manifest <- list(
  csv_path       = "data/01_raw/ess/Datafile-subset.csv",
  csv_size       = csv_size,
  csv_mtime      = csv_mtime,
  n_rows_csv     = n_rows_written,
  n_cols         = nrow(schema_csv),
  n_files        = length(pq_files),
  parquet_bytes  = sum(file.size(pq_files)),
  mode           = conversion_mode,
  forced_varchar = force_varchar,
  elapsed_sniff  = unname(t_sniff[["elapsed"]]),
  elapsed_write  = unname(t_write[["elapsed"]]),
  duckdb_version = duckdb_version,
  r_version      = R.version.string,
  converted_at   = Sys.time(),
  schema         = schema_csv
)
saveRDS(manifest, manifest_path)

# Scratch space is outside the repo, but there is no reason to leave gigabytes behind.
unlink(scratch_dir, recursive = TRUE, force = TRUE)

message("Published ", length(pq_files), " Parquet files to ", pq_root)
```

# Validation

Everything below runs on every render, warm or cold, reading only the Parquet dataset
and the manifest. The CSV is never touched again.

```{r}
#| label: validation-connect

con <- DBI::dbConnect(duckdb::duckdb())
manifest <- readRDS(manifest_path)
pq_files <- list.files(pq_root, pattern = "[.]parquet$",
                       recursive = TRUE, full.names = TRUE)
pq_glob  <- sqlp(file.path(pq_root, "**", "*.parquet"))

hive_args <- paste0(
  "hive_partitioning = true, ",
  "hive_types = {'cntry': 'VARCHAR', 'essround': 'BIGINT'}"
)
```

## Row Counts

The Parquet count comes from the file footers, not the data pages, so it is a metadata
read. It must equal the row count the conversion reported.

```{r}
#| label: validate-rows

n_parquet <- DBI::dbGetQuery(con, paste0(
  "SELECT count(*) AS n FROM read_parquet('", pq_glob, "', ", hive_args, ")"
))$n

knitr::kable(
  data.frame(
    source = c("CSV (rows written at conversion)", "Parquet (footer count)"),
    rows   = format(c(manifest$n_rows_csv, n_parquet), big.mark = ",")
  ),
  row.names = FALSE
)

stopifnot("row count mismatch between CSV and Parquet" =
            n_parquet == manifest$n_rows_csv)
message("Row counts agree: ", format(n_parquet, big.mark = ","))
```

## Columns and Schema

Two column counts are worth printing, because they differ for a good reason.
`WRITE_PARTITION_COLUMNS` defaults to false, so `cntry` and `essround` are *not* stored
inside the files -- they are encoded in the directory names and restored on read. A
single file therefore reports 2,834 columns while the dataset reports 2,836.

The restored partition keys are appended at the end of the schema, whereas in the CSV
`essround` is column 2 and `cntry` is column 6. So the check is set equality of names,
not position.

```{r}
#| label: validate-cols

schema_pq <- DBI::dbGetQuery(con, paste0(
  "DESCRIBE SELECT * FROM read_parquet('", pq_glob, "', ", hive_args, ")"
))
schema_one <- DBI::dbGetQuery(con, paste0(
  "DESCRIBE SELECT * FROM read_parquet('", sqlp(pq_files[1]), "', hive_partitioning = false)"
))

knitr::kable(
  data.frame(
    measure = c("dataset, hive keys restored", "single file, as stored",
                "CSV, at conversion"),
    columns = c(nrow(schema_pq), nrow(schema_one), manifest$n_cols)
  ),
  row.names = FALSE
)

stopifnot(
  "column names differ between CSV and Parquet" =
    setequal(manifest$schema$column_name, schema_pq$column_name)
)
message("All ", nrow(schema_pq), " column names round-trip intact.")
```

```{r}
#| label: schema-summary

knitr::kable(
  as.data.frame(sort(table(schema_pq$column_type), decreasing = TRUE)),
  col.names = c("stored type", "columns"),
  row.names = FALSE
)
```

The complete schema follows. It is long, so it sits in a scrolling panel. The
authoritative variable reference remains the ESS codebook that ships beside the CSV,
`data/01_raw/ess/Datafile-subset codebook.html`.

::: {style="max-height: 500px; overflow-y: auto; font-size: 0.8em;"}

```{r}
#| label: schema-table

knitr::kable(
  data.frame(
    pos    = seq_len(nrow(schema_pq)),
    column = schema_pq$column_name,
    type   = schema_pq$column_type
  ),
  row.names = FALSE
)
```

:::

## On-Disk Size

```{r}
#| label: validate-size

csv_bytes <- manifest$csv_size
pq_bytes  <- sum(file.size(pq_files))
mean_mb   <- pq_bytes / length(pq_files) / 1024^2

knitr::kable(
  data.frame(
    measure = c("CSV", "Parquet (total)", "compression ratio",
                "Parquet files", "mean file size"),
    value = c(
      paste0(round(csv_bytes / 1024^3, 2), " GB"),
      paste0(round(pq_bytes / 1024^3, 2), " GB"),
      paste0(round(csv_bytes / pq_bytes, 1), "x"),
      format(length(pq_files), big.mark = ","),
      paste0(round(mean_mb, 2), " MB")
    )
  ),
  row.names = FALSE
)

if (mean_mb < 5) {
  message(
    "Note: mean partition is ", round(mean_mb, 2), " MB. With ", nrow(schema_pq),
    " columns, a large share of each file is Parquet footer metadata rather than ",
    "data. This layout is chosen for partition pruning, not scan throughput. If ",
    "whole-dataset scans ever dominate, partitioning on cntry alone (~40 files) ",
    "would cut the metadata overhead substantially."
  )
}
```

## Timed Demo Read

The point of the conversion, measured. Six columns out of 2,836, one country out of the
partition tree: `cntry == "DE"` touches 11 files of 222, and only the requested column
chunks are read from each.

`clsprty` and `prtdgcl` are the two-item partisan-attachment battery this project is
built on, `lrscale` is the ideology anchor, and `cntry` / `essround` / `idno` are the
key.

```{r}
#| label: demo-duckdb

demo_cols <- c("cntry", "essround", "idno", "clsprty", "prtdgcl", "lrscale")

t_duck <- system.time({
  demo_duck <- DBI::dbGetQuery(con, paste0(
    "SELECT ", paste(demo_cols, collapse = ", "),
    " FROM read_parquet('", pq_glob, "', ", hive_args, ")",
    " WHERE cntry = 'DE'"
  ))
})

message("DuckDB: ", nrow(demo_duck), " rows x ", ncol(demo_duck),
        " cols in ", round(t_duck[["elapsed"]], 3), " s")

knitr::kable(as.data.frame(table(demo_duck$essround)),
             col.names = c("essround", "rows"), row.names = FALSE)
```

```{r}
#| label: demo-arrow

if (requireNamespace("arrow", quietly = TRUE)) {
  ds <- arrow::open_dataset(pq_root, format = "parquet")
  t_arrow <- system.time({
    demo_arrow <- ds |>
      dplyr::filter(cntry == "DE") |>
      dplyr::select(dplyr::all_of(demo_cols)) |>
      dplyr::collect()
  })
  message("arrow:  ", nrow(demo_arrow), " rows x ", ncol(demo_arrow),
          " cols in ", round(t_arrow[["elapsed"]], 3), " s")

  # Free cross-engine consistency check.
  a <- as.data.frame(demo_arrow)
  a <- a[order(a$essround, a$idno), demo_cols]
  d <- demo_duck[order(demo_duck$essround, demo_duck$idno), demo_cols]
  message("arrow and duckdb agree: ",
          isTRUE(all.equal(a, d, check.attributes = FALSE)))
} else {
  message("arrow is not installed; skipping the arrow route. ",
          "The duckdb route above covers the same ground.")
}
```

## Missing Codes and Wide Values Survived

Spot checks on what a careless conversion would have broken: the quoted language code
`"000"` must still be a string, the widest `idno` must not have been truncated into a
narrower integer type, the `clsprty` missing codes must still be sitting in the data
rather than having been turned into `NA`, and `edition` must still read as a label.

```{r}
#| label: validate-fidelity

fidelity <- DBI::dbGetQuery(con, paste0(
  "SELECT
     max(idno)                                    AS max_idno,
     count(*) FILTER (WHERE lnghom2 = '000')      AS lnghom2_000,
     count(*) FILTER (WHERE clsprty IN (7, 8, 9)) AS clsprty_missing_codes,
     count(DISTINCT edition)                      AS n_editions,
     count(DISTINCT name)                         AS n_source_files
   FROM read_parquet('", pq_glob, "', ", hive_args, ")"
))

knitr::kable(
  data.frame(
    check = c("max(idno), expected 510130400606",
              "rows with lnghom2 == '000' kept as text",
              "rows where clsprty still carries code 7/8/9",
              "distinct edition labels (VARCHAR, not DOUBLE)",
              "distinct source files, expected 12"),
    value = format(as.numeric(fidelity[1, ]), big.mark = ",", scientific = FALSE)
  ),
  row.names = FALSE
)
```

The twelfth source file is the reason `essround` alone does not identify a sample. ESS
round 10 was released as two of them -- `ESS10e03_3` and the self-completion
`ESS10SCe03_2` -- and both carry `essround == 10`, so they land in the same partitions
and are separable only by `name`. Nothing is filtered here; it is simply a fact any
later layer has to know.

```{r}
#| label: validate-round10

knitr::kable(
  DBI::dbGetQuery(con, paste0(
    "SELECT name, essround, count(*) AS rows, count(DISTINCT cntry) AS countries
     FROM read_parquet('", pq_glob, "', ", hive_args, ")
     GROUP BY name, essround ORDER BY essround, name"
  )),
  format.args = list(big.mark = ","),
  row.names = FALSE
)
```

# Conversion Record

```{r}
#| label: print-manifest

knitr::kable(
  data.frame(
    field = c("converted at", "mode", "source CSV", "CSV size", "rows", "columns",
              "partitions", "Parquet size", "forced to VARCHAR",
              "sniff time", "write time", "duckdb", "R"),
    value = c(
      format(manifest$converted_at, "%Y-%m-%d %H:%M:%S"),
      manifest$mode,
      manifest$csv_path,
      paste0(round(manifest$csv_size / 1024^3, 2), " GB"),
      format(manifest$n_rows_csv, big.mark = ","),
      format(manifest$n_cols, big.mark = ","),
      format(manifest$n_files, big.mark = ","),
      paste0(round(manifest$parquet_bytes / 1024^3, 2), " GB"),
      paste(manifest$forced_varchar, collapse = ", "),
      paste0(round(manifest$elapsed_sniff, 1), " s"),
      paste0(round(manifest$elapsed_write, 1), " s"),
      manifest$duckdb_version,
      manifest$r_version
    )
  ),
  row.names = FALSE
)
```

```{r}
#| label: validation-disconnect

DBI::dbDisconnect(con, shutdown = TRUE)
```

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
