---
title: |
  CSAP Code Notebook 1.2
subtitle: |
  ESS party items coded to Party Facts
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

This notebook gives every ESS party item in the 25 democracies a pair of Party Facts variables, the way the eu25games release does (`eu25games2019/code/07_recode.qmd`, section *Party Facts crosswalk*):

- `ext_<item>_pf_name`: the Party Facts core name, `"<ESS label> [og]"` for a party that is not in Party Facts, or a sentinel for a non-party answer (`none`, `dont-know`, `refused`, `other`, `independent`, `did-not-vote`, `blank-invalid`, ...);
- `ext_<item>_pf_id`: the Party Facts id, `NA` for `[og]` and sentinels.

The items are party attachment (`prtcl*`, stub `prtcl`) and national vote.
The main vote (stub `prtv`) is the list vote: Germany's second vote (`prtv*de2`), Lithuania's multi-member list vote (`prtv*lt1`), and `prtvt*` elsewhere, as in notebook 2.4.
The other German and Lithuanian votes get their own stubs: `prtv_de1` (first vote), `prtv_lt2` and `prtv_lt3` (single-member constituency, first and second round).

Two outputs, both in `data/02_processed/`:

- `ess_party_pf.csv`: the crosswalk, one row per country × round × variable × code that occurs in the data, with the codebook label and the `source` of the coding (`link`, `name`, `hand`, `missing`);
- `ess_party_pf.parquet`: one row per respondent (`cntry`, `essround`, `name`, `idno`) with the `ext_*` variables.

# Setup

```{r}
#| label: setup

start_time <- Sys.time()
options(width = 80)

p_required <- c(
  "tidyverse",
  "here",
  "rvest",
  "arrow",
  "stringi",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

source(here("code", "00_helper", "pid_tree.R")) # countries_25
source(here("code", "00_helper", "pf_code.R"))
```

# Codebook Labels

Every `prtcl*` and `prtv*` variable is one block in the ESS codebook, with a table of codes and value labels (parser as in notebook 2.4).
All codes are kept here, the missing codes included, because some codes above 66 are real answers ("Other", "Blank vote").

```{r}
#| label: codebook-labels

codebook <- read_html(
  here("data", "01_raw", "ess", "Datafile-subset codebook.html")
)
blocks <- html_elements(
  codebook,
  xpath = "//h3[starts-with(@id,'prtcl') or starts-with(@id,'prtv')]/.."
)

labels <- blocks |>
  map(\(b) {
    tibble(
      variable = html_attr(html_element(b, "h3"), "id"),
      code = as.integer(html_text2(html_elements(b, "td.nowrap"))),
      label = html_text2(html_elements(b, "td:not(.nowrap)"))
    )
  }) |>
  list_rbind()

labels |>
  summarise(variables = n_distinct(variable), labels = n())
```

# Read the Party Items

One Parquet read per country takes every party variable of that country.
Each item is then stacked long; a respondent has at most one non-empty variable per item, because ESS opens a new variable whenever a country's party list is renumbered.

```{r}
#| label: read

ess <- open_dataset(
  here("data", "01_raw", "ess", "parquet"),
  partitioning = hive_partition(cntry = utf8(), essround = int64())
)
ess_cols <- names(ess)

# Variable pattern per item stub
item_patterns <- function(cc) {
  lc <- str_to_lower(cc)
  main_vote <- case_when(
    cc == "DE" ~ "^prtv[a-z]?de2$",
    cc == "LT" ~ "^prtv[a-z]?lt1$",
    .default = paste0("^prtvt[a-z]?", lc, "$")
  )
  c(
    prtcl = paste0("^prtcl[a-z]?", lc, "$"),
    prtv = main_vote,
    prtv_de1 = if (cc == "DE") "^prtv[a-z]?de1$",
    prtv_lt2 = if (cc == "LT") "^prtv[a-z]?lt2$",
    prtv_lt3 = if (cc == "LT") "^prtv[a-z]?lt3$"
  )
}

read_country <- function(cc) {
  pats <- item_patterns(cc)
  vars <- map(pats, \(p) str_subset(ess_cols, p))

  raw <- ess |>
    filter(cntry == cc) |>
    select(cntry, essround, name, idno, all_of(unlist(vars, use.names = FALSE))) |>
    collect() |>
    mutate(across(-c(cntry, essround, name, idno), as.integer))

  imap(vars, \(v, stub) {
    raw |>
      select(cntry, essround, name, idno, all_of(v)) |>
      pivot_longer(
        all_of(v),
        names_to = "variable",
        values_to = "code",
        values_drop_na = TRUE
      ) |>
      mutate(stub = stub)
  }) |>
    list_rbind() |>
    bind_rows(distinct(raw, cntry, essround, name, idno))
}

# Long: one row per respondent × item (plus one empty row per respondent,
# so that respondents without any party answer are kept)
items <- map(countries_25, read_country) |>
  list_rbind()

# One answer per respondent and item
stopifnot(
  items |>
    filter(!is.na(stub)) |>
    count(cntry, essround, name, idno, stub) |>
    pull(n) |>
    max() == 1
)

# The cells to code: every country × round × variable × code in the data
cells <- items |>
  filter(!is.na(stub)) |>
  count(cntry, essround, variable, code, name = "n") |>
  left_join(labels, by = c("variable", "code")) |>
  mutate(
    type = case_when(
      str_starts(variable, "prtcl") ~ "c",
      str_detect(variable, "(de|lt)[0-9]$") ~ paste0("v-", str_sub(variable, -3)),
      .default = "v"
    ),
    key = norm_key(label)
  )

# Every code in the data has a codebook label
stopifnot(!anyNA(cells$label))

cells |>
  summarise(
    respondents = n_distinct(items$cntry, items$essround, items$name, items$idno),
    variables = n_distinct(variable),
    cells = n(),
    labels = n_distinct(cntry, key)
  )
```

# Party Facts Links

Party Facts links the ESS attachment and vote lists as datasets `essprtc` and `essprtv`.
The link id is `CC-<round>-<code>-<type>`, with type `c` (attachment), `v` (vote), or `v-de1`, `v-de2`, `v-lt1`, ... for the split votes.
It records the round in which a party **first** appears on a country's list, and its code in that round.
When ESS renumbers a list later, no new link is written, so a code in a later round can belong to a different party (Austria's code 5 is LIF in rounds 1–2 but the Greens in round 3).
A link is therefore trusted only at its exact position, and the party it names is carried to other rounds through the label.

```{r}
#| label: links

core <- pf_core()

links <- bind_rows(pf_links("essprtc", core), pf_links("essprtv", core)) |>
  extract(
    dataset_party_id,
    c("cntry", "start", "code", "type"),
    "^([A-Z]{2})-([0-9]+)-([0-9]+)-(.+)$",
    convert = TRUE
  ) |>
  # no links to parties that ended before 1950 (see pf_historic())
  filter(!pf_id %in% pf_historic(core))

# Links at their exact position
placed <- cells |>
  inner_join(links, by = c("cntry", "essround" = "start", "code", "type"))

tibble(
  links = nrow(filter(links, cntry %in% countries_25)),
  at_position = nrow(distinct(placed, cntry, essround, code, type)),
  technical = sum(!is.na(placed[["technical"]]))
)
```

A label that is linked to more than one party, or only to a technical Party Facts id ("1 percent", "other", ...), is left to the hand map below.
Every other linked label is checked against all names Party Facts holds for the linked party, in any dataset: after lower-casing and stripping accents and punctuation, one name must contain the other as whole words.
Labels in Greek script cannot pass that test against transliterated names and are printed for checking by eye.

```{r}
#| label: link-check

# All names Party Facts has for each id
pf_names <- read_csv(
  file.path(pf_dir, "partyfacts-external-parties.csv"),
  col_types = cols(.default = col_character())
) |>
  transmute(pf_id = as.integer(partyfacts_id), name_short, name, name_english) |>
  bind_rows(
    core |>
      transmute(pf_id, name_short = pf_short, name = pf_name, name_english = pf_english)
  ) |>
  pivot_longer(-pf_id, values_to = "nk", values_drop_na = TRUE) |>
  mutate(nk = normalise_party(nk)) |>
  filter(!is.na(pf_id), nchar(nk) >= 2) |>
  distinct(pf_id, nk)

# One party per label, from real (not technical) links
label_links <- placed |>
  distinct(cntry, key, label, pf_name, pf_id, technical) |>
  filter(n() == 1, is.na(technical), .by = c(cntry, key))

link_check <- label_links |>
  mutate(nl = normalise_party(label)) |>
  left_join(pf_names, by = "pf_id", relationship = "many-to-many") |>
  summarise(
    ok = any(contains_words(nl, nk) | contains_words(nk, nl), na.rm = TRUE),
    .by = c(cntry, key, label, pf_name, pf_id)
  )

# Failures: only Greek-script labels
link_check |>
  filter(!ok) |>
  select(cntry, label, pf_id, pf_name)
stopifnot(all(stri_detect_regex(filter(link_check, !ok)$label, "\\p{Greek}")))

map_link <- label_links |>
  transmute(cntry, key, pf_name, pf_id, source = "link")
```

Labels without a usable link are matched by exact normalised name against the country's Party Facts parties: their names in Party Facts core and in the ESS links.

```{r}
#| label: name-match

iso2 <- set_names(names(pf_iso3), pf_iso3)
country_names <- bind_rows(
  links |>
    filter(is.na(technical)) |>
    select(cntry, pf_name, pf_id, link_short, link_name, link_english),
  core |>
    filter(is.na(technical), country %in% pf_iso3) |>
    transmute(
      cntry = iso2[country],
      pf_name,
      pf_id,
      core_name = pf_name,
      pf_short,
      pf_english
    )
) |>
  pivot_longer(
    -c(cntry, pf_name, pf_id),
    values_to = "nk",
    values_drop_na = TRUE
  ) |>
  mutate(nk = normalise_party(nk)) |>
  filter(nchar(nk) >= 2) |>
  distinct(cntry, nk, pf_name, pf_id) |>
  # "PP" is also the short name of Spain's Partido Progresista (1834-74)
  filter(!pf_id %in% pf_historic(core))

map_name <- cells |>
  distinct(cntry, key, label) |>
  anti_join(map_link, by = c("cntry", "key")) |>
  # labels linked only to several parties or to a technical id go to the hand map
  anti_join(placed, by = c("cntry", "key")) |>
  mutate(nl = normalise_party(label)) |>
  inner_join(country_names, by = c("cntry", "nl" = "nk")) |>
  distinct(cntry, key, pf_name, pf_id) |>
  filter(n() == 1, .by = c(cntry, key)) |>
  mutate(source = "name")

nrow(map_name)
```

# Hand Map

Everything else is coded by hand: labels linked to more than one party, labels linked only to a technical id, and labels with no link and no exact name match.
Keys are country and label, as in the eu25games map, so one row covers every round and variable in which a label occurs.
Missing codes (66 not applicable, 77 refusal, 88 don't know, 99 no answer) are handled by code, not here.

```{r}
#| label: hand-map

map_hand <- tribble(
  ~cntry, ~label, ~pf_name, ~pf_id,
  "AT", "JETZT", "Liste Peter Pilz", 6137L, # JETZT is the 2018 rename of Liste Pilz
  "AT", "Other", "other", NA,
  "AT", "Piratenpartei Österreich", "Piratenpartei Österreich [og]", NA,
  "BE", "Blank", "blank-invalid", NA,
  "BE", "CDH", "Parti Social Chrétien — Centre démocrate humaniste", 633L,
  "BE", "Groen!", "Agalev — Groen", 528L,
  "BE", "Invalid", "blank-invalid", NA,
  "BE", "Open VLD", "Partij voor Vrijheid en Vooruitgang — Open Vlaamse Liberalen en Democraten", 49L,
  "BE", "Open VLD + Vivant", "Open VLD + Vivant [og]", NA, # 2003 VLD-Vivant joint list; no PF alliance entry
  "BE", "Other", "other", NA,
  "BE", "SP.A. + Vlaams - Progressieven (Spirit)", "Socialistische Partij Anders / Sociaal-Liberale Partij / Spirit", 1586L, # PF has the SP.A-Spirit cartel entry
  "BE", "Sp.a (Vooruit)", "Socialistische Partij Anders — Vooruit", 1680L,
  "BE", "Vlaams - Progressieven (Spirit)", "Sociaal-Liberale Partij", 7015L, # Spirit renamed VlaamsProgressieven 2008
  "BE", "cdH (Les Engagés)", "Parti Social Chrétien — Centre démocrate humaniste", 633L, # cdH renamed Les Engagés 2022, same PF id
  "BG", "Alternativa za bulgarsko vuzrazhdane (ABV)", "Alternativa za balgarsko vazrazhdane", 3187L,
  "BG", "Bulgarski Vazhod", "Balgarski vazhod", 9060L,
  "BG", "Bulgarski natsionalen suyuz - Nova demokratsiya", "Balgarski Natsionalen Saiuz — ND", 8260L,
  "BG", "DPS - Dvijenie za prava i svobodi", "Dvizhenie za prava i svobodi", 982L,
  "BG", "Demokrati za Silna Balgariia", "Demokrati za Silna Balgarija", 1195L,
  "BG", "Demokrati za silna Bulgaria (DSB)", "Demokrati za Silna Balgarija", 1195L,
  "BG", "Druga partia/koalitsia", "other", NA,
  "BG", "Druga politicheska sila", "other", NA,
  "BG", "Dvijenie 'Gergiovden'", "Dvizhenie Gergyovden", 1820L,
  "BG", "Dvijenie za Prava i Svobodi", "Dvizhenie za prava i svobodi", 982L,
  "BG", "Grazhdani za evropeysko razvitie na Balgariya – Sayuz na demokratichnite sili (GERB - SDS)", "Grazhdani za evropeysko razvitie na Balgariya – Sayuz na demokratichnite sili (GERB - SDS) [og]", NA, # GERB-SDS coalition; no PF alliance entry
  "BG", "Grazhdansko obedinenie za realna demokratsia (S.Binev)", "Grazhdansko obedinenie za realna demokratsia (S.Binev) [og]", NA,
  "BG", "I do not support anyone", "blank-invalid", NA, # ballot option 'I do not support anyone'
  "BG", "Invalid", "blank-invalid", NA,
  "BG", "KP Sayuz na desnite sili", "KP Sayuz na desnite sili [og]", NA, # cannot identify with a PF entry
  "BG", "Kandidati, izdignati ot Initsiativni komiteti", "independent", NA, # candidates nominated by initiative committees = independents
  "BG", "Koalitsia za Balgaria (BSP; PBS; Zemedelski saiuz) Al. Stamb", "Koalicija za Balgarija — BSP za Bŭlgariya", 1665L,
  "BG", "Koalitsiya ABV – Dvizhenie 21", "Koalitsiya ABV – Dvizhenie 21 [og]", NA, # ABV-Movement 21 coalition (2017); no PF alliance entry
  "BG", "Koalitsiya Dvizhenie za Bulgariya", "Koalitsiya Dvizhenie za Bulgariya [og]", NA,
  "BG", "Koalitziia BNS", "Balgarski Naroden Sajuz", 263L,
  "BG", "Koalitziia na Rozata", "Koalitsiya na Rozata", 2054L,
  "BG", "Levitsata!", "Levitsata! [og]", NA,
  "BG", "NFSB-Volya", "NFSB-Volya [og]", NA, # NFSB-Volya coalition; no PF alliance entry
  "BG", "Narodna partia „Svoboda i dostoynstvo“", "Narodna partia „Svoboda i dostoynstvo“ [og]", NA, # People's Party Freedom and Dignity is not PF 4095 (Centre-Freedom and Dignity coalition)
  "BG", "Natsionalno dvizhenie za stabilnost i vazhod (NDSV)", "Nacionalno dviženie (Simeon Vtori) za stabilnost i văzhod", 374L,
  "BG", "Nezavisim kandidat", "independent", NA,
  "BG", "Nova republika - DSB - Suyuz za Plovdiv - BDO", "Nova republika - DSB - Suyuz za Plovdiv - BDO [og]", NA, # 2017 coalition; no PF alliance entry
  "BG", "Novo vreme", "Hoboto Bpeme", 2022L, # PF transliterates Novo vreme as 'Hoboto Bpeme'
  "BG", "Other", "other", NA,
  "BG", "PP Balgarska demokratichna obshtnost", "PP Balgarska demokratichna obshtnost [og]", NA,
  "BG", "PP Natsionalno dvizhenie Edinstvo", "PP Natsionalno dvizhenie Edinstvo [og]", NA,
  "BG", "PP Partia za horata ot naroda", "PP Partia za horata ot naroda [og]", NA,
  "BG", "PP Red, zakonnost i spravedlivost (RZS)", "Red, zakonnost i spravedlivost", 272L,
  "BG", "Partia Roma", "Partia Roma [og]", NA,
  "BG", "Prodalzhavame Promyanata – Demokratichna Balgariya (PP-DB)", "Prodalzhavame Promyanata – Demokratichna Balgariya (PP-DB) [og]", NA, # PP-DB coalition; PF has PP and DB only separately
  "BG", "Sayuz na demokratichnite sili (SDS)", "Sayuz na Demokratichni Sili", 482L,
  "BG", "Sayuz na svobodnite demokrati (SSD)", "Sajuz na svobodnite demokrati", 8033L,
  "BG", "VMRO", "VMRO — Bulgarsko Natsionalno Dvizhenie", 2056L,
  "CZ", "Naděje (Hope)", "Naděje (Hope) [og]", NA,
  "CZ", "Other", "other", NA,
  "CZ", "SV-SOS P.of Countryside-Assoc.Civic Forces", "SV-SOS P.of Countryside-Assoc.Civic Forces [og]", NA,
  "DE", "National Democratic Party (NPD)", "Nationaldemokratische Partei Deutschlands", 1723L,
  "DE", "Other", "other", NA,
  "DE", "The Left Party.PDS (Die Linkspartei.PDS)", "Partei des Demokratischen Sozialismus", 86L, # Linkspartei.PDS (2005-07) is the PDS id; Die Linke starts 2007
  "DK", "Andet - other", "other", NA,
  "DK", "Dansk Folkeparti - Danish People's Party", "Dansk Folkeparti", 1022L,
  "DK", "Dansk Folkeparti - Danish peoples party", "Dansk Folkeparti", 1022L,
  "DK", "Det Konservative Folkeparti - Conservative", "Konservative Folkeparti", 536L,
  "DK", "Det Konservative Folkeparti - Conservative People's Party", "Konservative Folkeparti", 536L,
  "DK", "Det Radikale Venstre - Danish Social-Liberal Party", "Det Radikale Venstre", 1507L,
  "DK", "Det Radikale Venstre - The Radical Liberal Party", "Det Radikale Venstre", 1507L,
  "DK", "Enhedslisten - Unity List - The Red-Green Alliance", "Enhedslisten / De Rød-Grønne", 1527L,
  "DK", "Enhedslisten, De Rød-Grønne - The Red-Green Alliance", "Enhedslisten / De Rød-Grønne", 1527L,
  "DK", "Kristendemokraterne", "Kristeligt Folkeparti / Kristendemokraterne", 53L,
  "DK", "Kristendemokraterne - Christian democrats", "Kristeligt Folkeparti / Kristendemokraterne", 53L,
  "DK", "Kristendemokraterne - Christian democtrats", "Kristeligt Folkeparti / Kristendemokraterne", 53L,
  "DK", "Liberal Alliance - Liberal Alliance", "[Ny] Liberal Alliance", 212L,
  "DK", "Other", "other", NA,
  "DK", "SF Socialistisk Folkeparti - Socialist People's Party", "Socialistisk Folkeparti", 329L,
  "DK", "SF- Socialistisk Folkeparti - the Socialist People's Party", "Socialistisk Folkeparti", 329L,
  "DK", "Socialdemokraterne - the Danish social democrats", "Socialdemokraterne", 379L,
  "DK", "Socialdemokraterne - the Danish social democtrats", "Socialdemokraterne", 379L,
  "DK", "Socialdemokratiet - The Social democrats", "Socialdemokraterne", 379L,
  "DK", "Venstre, Danmarks Liberale Parti - The Liberal Party", "Venstre", 1204L,
  "DK", "Venstre, Danmarks Liberale Parti - Venstre", "Venstre", 1204L,
  "EE", "Eesti Iseseisvuspartei", "Eesti Iseseisvuspartei", 8922L, # not 1299 (ERSP 1992-95)
  "EE", "Eesti Vabaduspartei – Põllumeeste Kogu", "Eesti Vabaduspartei – Põllumeeste Kogu [og]", NA, # Estonian Free Party-Farmers' Assembly (2020s), not the 1990s PF 831
  "EE", "Eestimaa Ühendatud Vasakpartei", "Eestimaa Ühendatud Vasakpartei", 9124L,
  "EE", "Isamaa Erakond", "Erakond Isamaa ja Res Publica Liit", 685L, # Isamaa (2018 rename of IRL)
  "EE", "Mitte kellegi poolt", "blank-invalid", NA, # 'for nobody'; Estonia has no none-option, treated as blank vote
  "EE", "Other", "other", NA,
  "EE", "Rahva Ühtsuse Erakond", "Rahva Ühtsuse Erakond [og]", NA,
  "EE", "Tulevikuerakond", "Tulevikuerakond [og]", NA,
  "EE", "Üksikkandidaadid", "independent", NA,
  "EE", "Üksikkandidaadid või muud", "other", NA, # mixed 'independent candidates or other'
  "ES", "AMAIUR-Bildu", "Amaiur", 1671L,
  "ES", "Alianza Catalana", "Alianza Catalana [og]", NA,
  "ES", "Blank vote", "blank-invalid", NA,
  "ES", "CC (Coalición Canaria)", "Coalición Canaria", 81L,
  "ES", "CC-PCC (Coalición Canaria)", "Coalición Canaria", 81L, # CC-PNC alliance; PF codes CC alliances as 81
  "ES", "CC-PNC", "Coalición Canaria", 81L, # CC-PNC alliance; PF codes CC alliances as 81
  "ES", "Ciudadanos", "Ciudadanos — Partido de la Ciudadanía", 3217L,
  "ES", "Coalición Canaria - Nueva Canarias", "Coalición Canaria", 81L, # PF (manifesto link) codes CC-NC as 81
  "ES", "Compromís", "Coalició Compromís", 5623L,
  "ES", "Drago Canarias", "Drago Canarias [og]", NA,
  "ES", "EAJ-PNV", "Partido Nacionalista Vasco — Euzko Alderdi Jeltzalea", 1637L,
  "ES", "Esquerra Republicana de Catalunya (ERC)", "Esquerra Republicana de Catalunya", 848L,
  "ES", "Esquerra Republicana de Catalunya - ERC", "Esquerra Republicana de Catalunya", 848L,
  "ES", "Invalid vote", "blank-invalid", NA,
  "ES", "Izquierda Unida (IU) - (ICV en Cataluña)", "Izquierda Unida", 247L, # IU; ICV is its Catalan partner
  "ES", "Izquierda Unida (IU)-(ICV en Cataluña)", "Izquierda Unida", 247L, # IU; ICV is its Catalan partner
  "ES", "Junts per Catalunya/PDeCAT", "Junts per Catalunya — Partit Demòcrata Europeu Català", 9097L, # 2019 JxCat-PDeCAT list entry
  "ES", "Junts x Cat", "Junts per Catalunya", 7421L,
  "ES", "Més Compromis", "Coalició Compromís", 5623L, # Més Compromís = Compromís list
  "ES", "Navarra Suma", "Navarra Suma [og]", NA,
  "ES", "Other", "other", NA,
  "ES", "Otro", "other", NA,
  "ES", "Otros", "other", NA,
  "ES", "PNC (Nueva Canaria)", "PNC (Nueva Canaria) [og]", NA, # label conflates PNC (PF 8261) and Nueva Canarias; ambiguous
  "ES", "PRC", "PRC [og]", NA,
  "ES", "Partido Nacionalista Basco (PNV)", "Partido Nacionalista Vasco — Euzko Alderdi Jeltzalea", 1637L,
  "ES", "Partido Popular", "Alianza-Partido Popular / Partido Popular", 441L,
  "ES", "Partido Popular - PP (con UPN en Navarra)", "Alianza-Partido Popular / Partido Popular", 441L,
  "ES", "SUMAR", "SUMAR [og]", NA,
  "ES", "Spoiled vote", "blank-invalid", NA,
  "ES", "Teruel Existe", "Teruel Existe [og]", NA,
  "ES", "Votó en blanco", "blank-invalid", NA,
  "ES", "Votó nulo", "blank-invalid", NA,
  "FI", "Animal Justice Party", "Animal Justice Party [og]", NA,
  "FI", "Change 2011", "Change 2011 [og]", NA,
  "FI", "Crystal Party", "Crystal Party [og]", NA,
  "FI", "Feminist Party", "Feminist Party [og]", NA,
  "FI", "Finnish People's Blue-whites", "Finnish People's Blue-whites [og]", NA,
  "FI", "For the Poor", "For the Poor [og]", NA,
  "FI", "Freedom Alliance", "Freedom Alliance [og]", NA,
  "FI", "Freedom Party", "Freedom Party [og]", NA,
  "FI", "För Åland", "För Åland [og]", NA, # För Åland list; PF 1062 Åländsk Samling not clearly the same list
  "FI", "Independence Party", "Independence Party [og]", NA,
  "FI", "Liberal Party - Freedom to Choose", "Liberal Party - Freedom to Choose [og]", NA, # post-2011 Liberal Party, not PF 249 (LKP)
  "FI", "Natural law party", "Natural law party [og]", NA,
  "FI", "Open Party", "Open Party [og]", NA,
  "FI", "Other", "other", NA,
  "FI", "Power Belongs to the People", "Power Belongs to the People [og]", NA,
  "FI", "Senior Citizens' Party", "Senior Citizens' Party [og]", NA,
  "FI", "The Communist Workers' Party", "The Communist Workers' Party [og]", NA, # KTP, not PF 7030 (SKP)
  "FI", "The Liberal Party - Freedom to Choose", "The Liberal Party - Freedom to Choose [og]", NA, # post-2011 Liberal Party, not PF 249 (LKP)
  "FI", "Workers' Party", "Workers' Party [og]", NA,
  "FR", "Autre", "other", NA,
  "FR", "Autre (PRECISEZ)", "other", NA,
  "FR", "Blanc", "blank-invalid", NA,
  "FR", "Ensemble ! (LREM, MODEM et Horizons)", "Ensemble ! (LREM, MODEM et Horizons) [og]", NA, # Ensemble alliance; no PF alliance entry
  "FR", "Le Rassemblement National (RN, ex Front National)", "Front National / Rassemblement national", 433L,
  "FR", "Le Rassemblement national (ex Front National)", "Front National / Rassemblement national", 433L,
  "FR", "Les Républicains et l’UDI", "Les Républicains et l’UDI [og]", NA, # LR-UDI joint label; no PF alliance entry
  "FR", "MRC (Mouvement Rèpublicain et Citoyen)", "Mouvement des citoyens / Pôle républicain", 3156L, # MRC = renamed Mouvement des citoyens
  "FR", "MRC (Mouvement républicain et citoyen)", "Mouvement des citoyens / Pôle républicain", 3156L, # MRC = renamed Mouvement des citoyens
  "FR", "MoDem (Mouvement Démocrate)", "Mouvement démocrate", 496L,
  "FR", "NPA (Nouveau Parti Anti-capitaliste)", "Nouveau Parti Anticapitaliste", 6684L,
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS)", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", NA, # NUPES alliance; no PF alliance entry
  "FR", "Nul", "blank-invalid", NA,
  "FR", "PCF (Parti Communiste Francais)", "Parti communiste française", 1251L,
  "FR", "PCF (Parti Communiste Français)", "Parti communiste française", 1251L,
  "FR", "PRG (Parti radical de gauche)", "Parti radical de gauche", 1783L,
  "FR", "RPF (Rassemblement du Peuple Français)", "RPF (Rassemblement du Peuple Français) [og]", NA, # likely Pasqua's Rassemblement pour la France (not 1947 RPF 8042); PF external links conflict
  "FR", "Renaissance (ex La République en Marche !)", "La République En Marche! / Renaissance", 5857L,
  "FR", "UMP (L’Union)", "Union pour un mouvement populaire", 1595L,
  "FR", "UMP (Union pour la majorité présidentielle)", "Union pour un mouvement populaire", 1595L,
  "FR", "UMP (Union pour un Mouvement Populaire)", "Union pour un mouvement populaire", 1595L,
  "FR", "Un candidat divers gauche", "divers gauche", 126L, # PF pseudo-party 'divers gauche' (non-technical)
  "FR", "l’Union pour la France (Debout la France, Les Patriotes et Génération Frexit)", "l’Union pour la France (Debout la France, Les Patriotes et Génération Frexit) [og]", NA, # DLF-led alliance; no PF alliance entry
  "GB", "Independent", "independent", NA,
  "GB", "Independent candidate", "independent", NA,
  "GB", "Independent(s) (nir)", "independent", NA,
  "GB", "Other", "other", NA,
  "GB", "Other (nir)", "other", NA,
  "GB", "Traditional Unionist (nir)", "Traditional Unionist (nir) [og]", NA, # TUV only a 1-percent technical id in Party Facts
  "GR", "Blank", "blank-invalid", NA,
  "GR", "Communist Party - KKE", "Kommounistikó Kómma Elládas", 48L,
  "GR", "Independent Citizens' Movement (SPITHA)", "Independent Citizens' Movement (SPITHA) [og]", NA, # SPITHA (Theodorakis) only 1-percent technical in PF
  "GR", "Invalid", "blank-invalid", NA,
  "GR", "New Democracy (ND)", "Néa Đimokratía", 794L,
  "GR", "New Democracy - ND", "Néa Đimokratía", 794L,
  "GR", "No one", "none", NA, # ambiguous: attachment "no party" vs ballot "none"; treated as no party
  "GR", "Null", "blank-invalid", NA,
  "GR", "Other", "other", NA,
  "GR", "Other parties", "other", NA,
  "GR", "Panhellenic Socialist Movement – PASOK", "Panellinio Sosialistikó Kínima", 1468L,
  "GR", "Popular Orthodox Rally - LAOS", "Laïkós Orthódoxos Synagermós", 965L,
  "GR", "Smokers' Groups for Art and Visual Formation (KOTES)", "Smokers' Groups for Art and Visual Formation (KOTES) [og]", NA, # KOTES not in PF
  "GR", "The Coalition of the Radical Left (SYRIZA)", "Synaspismós Rizospastikís Aristerás", 301L,
  "GR", "The Communist Party of Greece (KKE)", "Kommounistikó Kómma Elládas", 48L,
  "GR", "The Ecologist Greens (OP)", "Oikologoi Prasinoi", 378L,
  "GR", "The Panhellenic Socialist Movement (PASOK)", "Panellinio Sosialistikó Kínima", 1468L,
  "GR", "The Popular Orthodox Rally (LAOS)", "Laïkós Orthódoxos Synagermós", 965L,
  "GR", "Ανταρσία", "Antikapitalistiki Aristeri Synergasia gia tin Anatropi", 95L,
  "GR", "ΚΙΝΑ.ΑΛ", "Kinima Allagis", 7909L,
  "GR", "ΚΚΕ", "Kommounistikó Kómma Elládas", 48L,
  "GR", "ΛΑΟΣ", "Laïkós Orthódoxos Synagermós", 965L,
  "GR", "ΝΔ", "Néa Đimokratía", 794L,
  "GR", "Οικολόγοι Πράσινοι", "Oikologoi Prasinoi", 378L,
  "GR", "Χρυσή Αυγή", "Laïkós Sýndesmos / Chrysí Avgí", 1660L,
  "HR", "Centar", "Pametno", 7336L, # Centar is the renamed Pametno (PF wikipedia link)
  "HR", "DP Miroslava Škore, Hrvatski suverenisti, Blok za Hrvatsku, HKS, Hrast, SU, Zelena Lista", "DP Miroslava Škore, Hrvatski suverenisti, Blok za Hrvatsku, HKS, Hrast, SU, Zelena Lista [og]", NA, # DP-led coalition; no PF alliance entry
  "HR", "DP, HS, Blok za Hrvatsku, HKS, Hrast", "DP, HS, Blok za Hrvatsku, HKS, Hrast [og]", NA, # DP-led coalition; no PF alliance entry
  "HR", "Fokus", "Fokus", 9480L,
  "HR", "HDZ, HSLS", "HDZ, HSLS [og]", NA, # HDZ-HSLS coalition
  "HR", "HNS", "Hrvatska narodna stranka — Liberalni Demokrati", 799L,
  "HR", "Hrvatska demokršćanska stranka (HDS)", "Hrvatska kršćanska demokratska stranka", 5458L, # PF lists HDS as name_other of HKDS
  "HR", "Hrvatska socijalno liberalna stranka (HSLS)", "Hrvatska socijalno liberalna stranka", 214L,
  "HR", "Koalicija HDSSB-HKS", "Koalicija HDSSB-HKS [og]", NA, # coalition
  "HR", "Koalicija IDS, PGS, LRI", "Koalicija IDS, PGS, LRI [og]", NA, # coalition
  "HR", "Koalicija Pametno i Za grad", "Koalicija Pametno i Za grad [og]", NA, # coalition
  "HR", "Koalicija Živi zid, Promijenimo Hrvatsku, Abeceda, HDSS", "Koalicija Živi zid, Promijenimo Hrvatsku, Abeceda, HDSS [og]", NA, # coalition
  "HR", "Lista Mislav Kolakušić", "Lista Mislav Kolakušić [og]", NA, # Kolakušić list; PF codes it as independent
  "HR", "Možemo! - politička platforma, Zageb je naš, NL, RF, ORAH, Za Grad", "Možemo! - politička platforma, Zageb je naš, NL, RF, ORAH, Za Grad [og]", NA, # Možemo-led coalition; not the party alone
  "HR", "Možemo, Zagreb je naš, NLJ, RF, Orah", "Možemo, Zagreb je naš, NLJ, RF, Orah [og]", NA, # Možemo-led coalition; not the party alone
  "HR", "Narodna stranka - Reformisti - Hrvatska seljačka stranka braće Radić, Stranka hrvatskih umirovljenika", "Narodna stranka - Reformisti - Hrvatska seljačka stranka braće Radić, Stranka hrvatskih umirovljenika [og]", NA, # coalition
  "HR", "Nova ljevica", "Nova ljevica [og]", NA, # Nova ljevica not in PF (1-percent)
  "HR", "Other", "other", NA,
  "HR", "Radnička fronta", "Radnička fronta [og]", NA,
  "HR", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS [og]", NA, # coalition
  "HR", "SDP, HSS, HSU", "SDP, HSS, HSU [og]", NA, # coalition
  "HR", "START - Stranka antikorupcije, razvoja i transparentnosti", "START - Stranka antikorupcije, razvoja i transparentnosti [og]", NA,
  "HR", "SiP, Pametno, Fokus", "SiP, Pametno, Fokus [og]", NA, # coalition
  "HU", "6 FIDESZ-KDNP (Fidesz Magyar Polgári Szövetség-Kereszténydemokrata Néppárt)", "Fidesz — KDNP pártszövetség / Magyar Szolidaritás Szövetsége", 6366L,
  "HU", "CM (Civil Mozgalom)", "CM (Civil Mozgalom) [og]", NA,
  "HU", "Centrum Part", "Centrumpárt", 1779L,
  "HU", "Centrum Párt (Centrum Összefogás Magyarországért)", "Centrumpárt", 1779L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD)", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", NA, # 2022 opposition alliance; PF 8265 is the 2014 alliance
  "HU", "FKNP (FÜGGETLEN KISGAZDA-, NEMZETI EGYSÉG PÁRT)", "FKNP (FÜGGETLEN KISGAZDA-, NEMZETI EGYSÉG PÁRT) [og]", NA,
  "HU", "Fidesz - KDNP (Fidesz – Magyar Polgári Szövetség Keresztényd)", "Fidesz — KDNP pártszövetség / Magyar Szolidaritás Szövetsége", 6366L,
  "HU", "Fidesz - KDNP (Fidesz – Magyar Polgári Szövetség Kereszténydemokrata Néppárt)", "Fidesz — KDNP pártszövetség / Magyar Szolidaritás Szövetsége", 6366L,
  "HU", "HTJP-Hungarian Truth", "HTJP-Hungarian Truth [og]", NA, # possibly MIÉP (Hungarian Truth and Life) but abbreviation differs
  "HU", "KDNP (Kereszténydemokrata Néppárt)", "Kereszténydemokrata Néppárt", 1412L,
  "HU", "Keresztenydemokrata Part - Kereszteny Centrum Osszefogas", "Keresztenydemokrata Part - Kereszteny Centrum Osszefogas [og]", NA, # unclear KDNP splinter/Centrum list, PF 1-percent
  "HU", "LMP (Lehet Más A Politika)", "Lehet Más a Politika — LMP – Magyarország Zöld Pártja", 1650L,
  "HU", "MDF (Magyar Demokrata Fórum)", "Magyar Demokrata Fórum", 1697L,
  "HU", "MIÉP - Jobbik a Harmadik Út", "MIÉP — Jobbik a Harmadik Út pártszövetség", 8264L,
  "HU", "MRDP (MAGYARORSZÁGI ROMÁK DEMOKRATIKUS PÁRTJA)", "MRDP (MAGYARORSZÁGI ROMÁK DEMOKRATIKUS PÁRTJA) [og]", NA,
  "HU", "MSZDP (Magyar Szociáldemokrata Párt)", "Magyarországi Szociáldemokrata Párt", 1726L,
  "HU", "MSZP (Magyar Szocialista Párt)", "Magyar Szocialista Párt", 1408L,
  "HU", "MUNKÁSPÁRT", "Magyar Kommunista Munkáspárt", 859L,
  "HU", "Magyar Videk es Polgari Part", "Magyar Videk es Polgari Part [og]", NA,
  "HU", "Magyar Vidék és Polgári Párt", "Magyar Vidék és Polgári Párt [og]", NA,
  "HU", "Munkaspart", "Magyar Kommunista Munkáspárt", 859L,
  "HU", "Munkáspárt (Magyar Kommunista Munkáspárt)", "Magyar Kommunista Munkáspárt", 859L,
  "HU", "None of them", "blank-invalid", NA, # ballot option "none of them"
  "HU", "NÉP (Normális Élet Pártja)", "NÉP (Normális Élet Pártja) [og]", NA,
  "HU", "Other", "other", NA,
  "HU", "SZDSZ (Szabad Demokraták Szövetsége - A Magyar Liberális Párt)", "Szabad Demokraták Szövetsége", 910L,
  "IE", "Independent", "independent", NA,
  "IE", "Other", "other", NA,
  "IE", "Socialist Party", "Páirtí Sóisialach", 1448L,
  "IE", "Solidarity/People Before Profit (Solidarity/PBP)", "Solidarity — People Before Profit", 8241L,
  "IT", "Alleanza Verdi e Sinistra", "Alleanza Verdi e Sinistra [og]", NA, # AVS alliance, no PF entry
  "IT", "Altro", "other", NA,
  "IT", "Altro (SPECIFICARE)", "other", NA,
  "IT", "Altro (scrivere)", "other", NA,
  "IT", "Blank vote", "blank-invalid", NA,
  "IT", "Forza Italia", "Forza Italia [og]", NA, # ESS entry spans 2002-2022: FI (1626) vs FI 2013 (8058)
  "IT", "Fratelli d'Italia - An", "Fratelli d'Italia", 2280L, # FdI-AN is 2013-14 name of FdI
  "IT", "Fratelli d'Italia con Giorgia Meloni", "Fratelli d'Italia", 2280L,
  "IT", "Ha votato solo al maggioritario", "did-not-vote", NA,
  "IT", "Il popolo della famiglia", "Il popolo della famiglia [og]", NA,
  "IT", "La destra", "La Destra", 8188L,
  "IT", "Lega", "Lega Nord", 1221L,
  "IT", "Liberi e Uguali (LEU)", "Liberi e Uguali", 8246L,
  "IT", "Lista di Pietro", "Lista di Pietro — Italia del Valori", 768L,
  "IT", "Noi con l'Italia - UDC", "Noi con l'Italia - UDC [og]", NA, # joint list 2018
  "IT", "Partido Democratico", "Partito Democratico", 802L,
  "IT", "Partito Comunista", "Partito Comunista [og]", NA, # Rizzo's PC (2014), not PCI
  "IT", "Partito Socialista", "Partito Socialista [og]", NA, # PSI (2007) not in PF; 742 ends 1994
  "IT", "Popolo delle Libertà (PDL)", "Il Popolo della Libertà", 365L,
  "IT", "Terzo Polo (Azione-Italia Viva)", "Terzo Polo (Azione-Italia Viva) [og]", NA, # Azione-IV alliance
  "IT", "Unione Popolare", "Unione Popolare [og]", NA,
  "IT", "Vox Italia", "Vox Italia [og]", NA,
  "LT", "Anti-corruption coalition of K. Krivickas and N. Puteikis (Lithuanian Centre Party, Lithuanian Pensioners' Party) (APKK)", "Anti-corruption coalition of K. Krivickas and N. Puteikis (Lithuanian Centre Party, Lithuanian Pensioners' Party) (APKK) [og]", NA,
  "LT", "Centre Party - Nationalists (CPT)", "Lietuvos Centro Partija", 5455L, # PF: Centro partija - Tautininkai = LCP
  "LT", "Christian Union (KS)", "Christian Union (KS) [og]", NA, # Krikščionių sąjunga (2016-) not in PF
  "LT", "Coalition of S. Buškevičius and the Nationalists 'Against corruption and poverty' (Party 'Young Lithuania', Nationalis", "Coalition of S. Buškevičius and the Nationalists 'Against corruption and poverty' (Party 'Young Lithuania', Nationalis [og]", NA, # 2012 electoral coalition
  "LT", "Democratic Labour and Unity Party (DDVP)", "Democratic Labour and Unity Party (DDVP) [og]", NA,
  "LT", "Did not vote for a multi-member nationwide candidate list", "did-not-vote", NA,
  "LT", "Did not vote for a single-member constituency candidate list", "did-not-vote", NA,
  "LT", "Did not vote for a single-member constituency candidate list, 2nd round", "did-not-vote", NA,
  "LT", "Don't know if voted for a multi-member nationwide candidate list", "dont-know", NA,
  "LT", "Don't know if voted for a single-member constituency candidate list", "dont-know", NA,
  "LT", "Don't know if voted for a single-member constituency candidate list, 2nd round", "dont-know", NA,
  "LT", "Electoral Action of Poles in Lithuania-Christian Families Alliance (LLRA-KSS)", "Lietuvos lenkų rinkimų akcija", 556L, # LLRA-KŠS list coded as LLRA
  "LT", "Independent/Self-Nominated", "independent", NA,
  "LT", "Labourist Party (LP)", "Labourist Party (LP) [og]", NA, # unclear which party; PF 1-percent
  "LT", "Lithuanian Peasant and Greens Union (LVZS)", "Lietuvos valstiečių ir žaliųjų sąjunga / Lietuvos valstiečių liaudininkų sąjunga", 1490L,
  "LT", "Lithuanian People's Party (LZP)", "Lietuvos žalioji partija", 1407L, # ESS mislabel of Green Party (LZP); PF links it to 1407
  "LT", "National Unity Union (TVS)", "National Unity Union (TVS) [og]", NA, # PF 1-percent
  "LT", "Nationalist Union (TS)", "Lietuvių tautininkų ir respublikonų sąjunga", 9107L, # 377 is the interwar party (1924–40), a Party Facts link error
  "LT", "No 2nd round in single-member constituency", "did-not-vote", NA,
  "LT", "Other", "other", NA,
  "LT", "Party 'Lithuania - For everyone' (LV)", "Party 'Lithuania - For everyone' (LV) [og]", NA,
  "LT", "Political Party 'Lithuanian Green Movement' (LZS)", "Lietuvos žalioji partija", 1407L, # Green party founded 2011; PF links round 6 entry to 1407
  "LT", "Refused to tell if voted for a multi-member nationwide candidate list", "refused", NA,
  "LT", "Refused to tell if voted for a single-member constituency candidate list", "refused", NA,
  "LT", "Refused to tell if voted for a single-member constituency candidate list, 2nd round", "refused", NA,
  "LT", "Republican Party (RP)", "Republican Party (RP) [og]", NA, # PF 1-percent
  "LT", "Union of Intergenerational Solidarity - Cohesion for Lithuania (KSSL)", "Union of Intergenerational Solidarity - Cohesion for Lithuania (KSSL) [og]", NA,
  "LV", "Apvienotais saraksts — Latvijas Zaļā partija, Latvijas Reģionu Apvienība, Liepājas partija", "Apvienotais saraksts — Latvijas Zaļā partija, Latvijas Reģionu Apvienība, Liepājas partija [og]", NA, # United List 2022, no PF entry
  "LV", "Apvienība Latvijai", "Apvienība Latvijai [og]", NA,
  "LV", "Konservatīvie", "Jaunā konservatīvā partija", 6355L, # JKP renamed Konservatīvie 2022
  "LV", "Kristīgi Progresīvā Partija", "Kristīgi Progresīvā Partija [og]", NA,
  "LV", "LSDSP/KDS/GKL", "LSDSP/KDS/GKL [og]", NA, # 2018 joint list
  "LV", "Latviešu Nacionālisti", "Latviešu Nacionālisti [og]", NA,
  "LV", "Latvijas centriskā partija", "Latvijas centriskā partija [og]", NA,
  "LV", "Other", "other", NA,
  "LV", "Par alternatīvu", "Par alternatīvu [og]", NA,
  "LV", "Partija 'Vienoti Latvijai'", "Vienoti Latvijai", 3195L,
  "LV", "Politiskā partija 'Katram un katrai'", "Katram un katrai", 9076L,
  "LV", "Politiskā partija 'Republika'", "Republika", 9079L,
  "LV", "Politiskā partija 'Stabilitātei!'", "Stabilitātei!", 9080L,
  "LV", "Politiskā partija 'Tautas varas spēks'", "Spēks vienotībā / Alternative / Tautas varas spēks", 3214L,
  "NL", "Article 1", "Article 1 [og]", NA, # Artikel 1 (later Bij1), not in PF
  "NL", "Artikel1/Bij1", "Artikel1/Bij1 [og]", NA, # not in PF
  "NL", "Blanc", "blank-invalid", NA,
  "NL", "Group Geert Wilders", "Partij voor de Vrijheid", 298L, # Groep Wilders = PVV precursor (2006); PF ccs links it to 298
  "NL", "Other", "other", NA,
  "NL", "Pieter Omtzigt", "independent", NA, # ESS10 (2021-22): Omtzigt sat as independent MP before NSC; PF essprtc links to independent
  "NL", "Pirate Party", "Pirate Party [og]", NA, # Piratenpartij not in PF
  "PL", "Alternatywa Ruch Społeczny", "Alternatywa Ruch Społeczny [og]", NA,
  "PL", "Demokratyczna Partia Lewicy", "Demokratyczna Partia Lewicy [og]", NA,
  "PL", "Dom Ojczysty", "Dom Ojczysty [og]", NA,
  "PL", "Koalicja Bezpartyjni i Samorządowcy", "Koalicja Bezpartyjni i Samorządowcy [og]", NA,
  "PL", "Konfederacja Wolność i Niepodległość", "Konfederacja Wolność i Niepodległość", 9026L,
  "PL", "Nowa Lewica", "Nowa Lewica", 9349L,
  "PL", "Nowa Prawica", "Kongres Nowej Prawicy", 4713L, # Korwin-Mikke's Nowa Prawica = KNP
  "PL", "Ogólnopolska Koalicja Obywatelska", "Ogólnopolska Koalicja Obywatelska [og]", NA, # 2005 minor list, not KO (9122)
  "PL", "Other", "other", NA,
  "PL", "Partia Ludowo-Demokratyczna", "Partia Ludowo-Demokratyczna [og]", NA,
  "PL", "Polska Partia Pracy – Sierpień '80", "Polska Partia Pracy", 3699L,
  "PL", "Polska Unia Gospodarcza", "Polska Unia Gospodarcza [og]", NA,
  "PL", "Polska Wspólnota Narodowa", "Polska Wspólnota Narodowa [og]", NA,
  "PL", "Polskie Stronnictwo Ludowe", "Polskie Stronnictwo Ludowe", 602L,
  "PL", "Solidarna Polska", "Solidarna Polska", 328L, # PF also has 9109 Suwerenna Polska (likely duplicate); used exact-name 328
  "PL", "Stronnictwo Konserwatywno-Ludowe-Ruch Nowej Polski", "Stronnictwo Konserwatywno-Ludowe", 8766L, # SKL-RNP = SKL after 2002 merger
  "PL", "Trzecia Droga", "Trzecia Droga [og]", NA, # PSL+PL2050 alliance; PF only technical alliance id
  "PL", "Twoj Ruch", "Ruch Palikota / Twój Ruch", 731L,
  "PL", "Wolnosc", "Koalicja Odnowy Rzeczypospolitej Wolność i Nadzieja KORWiN", 4629L, # KORWiN renamed Wolność 2016
  "PL", "Zjednoczona Lewica", "Zjednoczona Lewica", 9091L,
  "PT", "ADN - Alternativa Democrática Nacional", "Alternativa Democrática Nacional", 9817L,
  "PT", "BE - Bloco de Esquerda", "Bloco de Esquerda", 1310L,
  "PT", "Blank vote", "blank-invalid", NA,
  "PT", "Bloco de Esquerda (BE)", "Bloco de Esquerda", 1310L,
  "PT", "CDS-PP - CDS-Partido Popular", "Centro Democrático e Social / Partido Popular", 1308L,
  "PT", "CDU – Coligação Democrática Unitária (PCP-PEV)", "Coligação Democrática Unitária", 284L,
  "PT", "Centro Democrático Social - Partido Popular (CDS-PP)", "Centro Democrático e Social / Partido Popular", 1308L,
  "PT", "Coligação Democrática Unitária (CDU)", "Coligação Democrática Unitária", 284L,
  "PT", "JPP - Juntos pelo Povo", "JPP - Juntos pelo Povo [og]", NA,
  "PT", "L - LIVRE", "Livre — Liberdade, Esquerda, Europa e Ecologia", 3225L,
  "PT", "MAS - Movimento Alternativa Socialista", "MAS - Movimento Alternativa Socialista [og]", NA,
  "PT", "MPT - Partido da Terra", "Movimento Partido da Terra / Partido da Terra", 3226L,
  "PT", "Nova democracia (PND)", "Partido da Nova Democracia", 1095L,
  "PT", "Other", "other", NA,
  "PT", "Outro", "other", NA,
  "PT", "PCP-PEV - CDU - Coligação Democrática Unitária", "Coligação Democrática Unitária", 284L,
  "PT", "PCP-PEV - Partido Comunista Português - Partido Ecologista os Verdes", "Coligação Democrática Unitária", 284L, # PCP-PEV joint list runs as CDU
  "PT", "PCTP-MRPP - Partido Comunista dos Trabalhadores Portugueses - Movimento Reorganizativo do Proletariado Português", "Comunista dos Trabalhadores Portugueses / Movimento Reorganizativo do Partido do Proletariado", 857L,
  "PT", "PCTP/MRPP - Partido Comunista dos Trabalhadores Portugueses", "Comunista dos Trabalhadores Portugueses / Movimento Reorganizativo do Partido do Proletariado", 857L,
  "PT", "PH - Partido Humanista", "Partido Humanista", 8269L,
  "PT", "PNR - Partido Nacional Renovador", "PNR - Partido Nacional Renovador [og]", NA, # PF 6313 is the 1918-23 Partido Nacional Republicano, not PNR (2000-)
  "PT", "PPD/PSD - Partido Social Democrata", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "PT", "PPV/CDC - Partido Cidadania e Democracia Cristã", "PPV/CDC - Partido Cidadania e Democracia Cristã [og]", NA,
  "PT", "PURP - Partido Unido dos Reformados e Pensionistas", "PURP - Partido Unido dos Reformados e Pensionistas [og]", NA,
  "PT", "Partido Comunista dos Trabalhadores Port./Movimento Reorganizativo do Proletariado Port. (PCTP-MRPP)", "Comunista dos Trabalhadores Portugueses / Movimento Reorganizativo do Partido do Proletariado", 857L,
  "PT", "Partido Comunista dos Trabalhadores Portugueses/Movimento Reorganizativo do Partido do Proletariado (PCTP-MRPP)", "Comunista dos Trabalhadores Portugueses / Movimento Reorganizativo do Partido do Proletariado", 857L,
  "PT", "Partido Democrático do Atlântico (PDA)", "Partido Democrático do Atlântico (PDA) [og]", NA, # not the 1912-26 Partido Democrático
  "PT", "Partido Humanista (PH)", "Partido Humanista", 8269L,
  "PT", "Partido Nacional Renovador (PNR)", "Partido Nacional Renovador (PNR) [og]", NA, # PF 6313 is the 1918-23 Partido Nacional Republicano
  "PT", "Partido Social Democrata (PPD/PSD)", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "PT", "Partido Social Democrata (PSD)", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "PT", "Partido Socialista (PS)", "Partido Socialista", 655L,
  "PT", "R.I.R. - Reagir Incluir Reciclar", "R.I.R. - Reagir Incluir Reciclar [og]", NA,
  "PT", "VP – Volt Portugal", "VP – Volt Portugal [og]", NA,
  "PT", "Votou em branco / nulo", "blank-invalid", NA,
  "PT", "Votou em branco/ nulo", "blank-invalid", NA,
  "RO", "PPCD (PNTCD)", "Partidul Naţional Ţărănesc Creştin Democrat", 1750L, # PNŢCD briefly renamed PPCD (2005-07)
  "SE", "Annat parti", "other", NA,
  "SE", "Folkpartiet liberalerna", "Folkpartiet liberalerna / Liberalerna", 1274L,
  "SE", "Liberalerna", "Folkpartiet liberalerna / Liberalerna", 1274L,
  "SE", "Moderata samlingspartiet", "Allmänna valmansförbundet / Högern / Högerpartiet / Moderata samlingspartiet", 690L,
  "SE", "Other", "other", NA,
  "SE", "Socialdemokraterna", "Socialdemokraterna / Sveriges socialdemokratiska arbetareparti", 487L,
  "SE", "Vänsterpartiet", "Vänsterpartiet / Sveriges Komunistiska Parti", 830L,
  "SI", "Other", "other", NA,
  "SI", "SAB - Stranka Alenke Bratušek", "Zavezništvo Alenke Bratušek / Stranka Alenke Bratušek", 3114L,
  "SI", "SD - Socialni demokrati", "Socialni demokrati / Združena lista socialnih demokratov", 1403L,
  "SI", "SDS - Slovenska demokratska stranka", "Slovenska demokratska stranka", 472L,
  "SI", "SMC - Stranka modernega centra", "Stranka Mira Cerarja / Stranka modernega centra", 3098L,
  "SK", "Christian Democratic Movement (KDH)", "Kresťanskodemokratické hnutie", 63L,
  "SK", "Movement for a Democratic Slovakia (HZDS)", "Ľudová strana — Hnutie za demokratické Slovensko", 560L,
  "SK", "Other", "other", NA,
  "SK", "Party of Hungarian Coalition (SMK)", "Strana maďarskej koalície / Strana maďarskej komunity", 349L,
  "SK", "Pogresívne Slovensko - Spolu", "Pogresívne Slovensko - Spolu [og]", NA, # PS+Spolu 2019-20 alliance; PF only technical alliance id
  "SK", "SMER-Social Democracy", "Smer — sociálna demokracia", 311L,
  "SK", "Sloboda a Solidarita (SaS)", "Sloboda a Solidarita", 1386L,
  "SK", "Slovak Democratic Christian Union (SDKU)", "Slovenská demokratická a kresťanská únia — Demokratická strana", 1617L,
  "SK", "Slovenská národná strana (SNS)", "Slovenská národná strana", 730L,
  # names that matched a party of before 1950 by short name (see pf_historic())
  "ES", "PP", "Alianza-Partido Popular / Partido Popular", 441L, # not 5750, the Partido Progresista of 1834-74
  "SE", "Centern", "Bondeförbundet / Centerpartiet", 199L, # not 7415, "Centern" of 1873-82
  "RO", "PC", "Partidul Conservator / Partidul Umanist Român", 1443L,
  "PL", "Polska Partia Socjalistyczna", "Polska Partia Socjalistyczna [og]", NA # the post-1987 PPS; 4354 is the party of 1892-1948
) |>
  mutate(key = norm_key(label), source = "hand")

pf_check(map_hand, keys = c("cntry", "key"), core = core)
count(map_hand, kind = case_when(
  pf_name %in% pf_sentinels ~ pf_name,
  str_ends(pf_name, " \\[og\\]") ~ "[og]",
  .default = "Party Facts"
))
```

Forza Italia has two Party Facts ids, for the party of 1994–2009 and the one refounded in 2013, while its ESS label stayed the same.
The label alone cannot tell them apart, so it is coded by round.
(Solidarna Polska, renamed Suwerenna Polska in 2023, also has two ids, 328 and 9109, but they cover the same years and look like a duplicate; it keeps 328 throughout, as in the eu25games release.)

```{r}
#| label: era-map

map_era <- tribble(
  ~cntry, ~label, ~from, ~pf_name, ~pf_id,
  "IT", "Forza Italia", 1L, "Forza Italia", 1626L,
  "IT", "Forza Italia", 7L, "Forza Italia (2013)", 8058L
) |>
  mutate(key = norm_key(label), source = "hand")
```

# Crosswalk

Hand codes take precedence over name matches and links; round-specific codes take precedence over all.

```{r}
#| label: crosswalk

map_label <- bind_rows(
  mutate(map_hand, prio = 1L),
  mutate(map_link, prio = 2L),
  mutate(map_name, prio = 3L)
) |>
  slice_min(prio, n = 1, by = c(cntry, key)) |>
  select(-prio)

missing_codes <- tribble(
  ~code, ~pf_name,
  66L, NA, # not applicable
  77L, "refused",
  88L, "dont-know",
  99L, NA # no answer
)

crosswalk <- cells |>
  left_join(
    select(map_label, cntry, key, pf_name, pf_id, source),
    by = c("cntry", "key")
  ) |>
  # round-specific codes
  left_join(
    map_era |>
      select(cntry, key, from, era_name = pf_name, era_id = pf_id),
    by = join_by(cntry, key, closest(essround >= from))
  ) |>
  mutate(
    pf_name = coalesce(era_name, pf_name),
    pf_id = if_else(is.na(era_name), pf_id, era_id),
    source = if_else(is.na(era_name), source, "hand")
  ) |>
  select(-c(from, era_name, era_id)) |>
  # missing codes
  rows_update(
    cells |>
      filter(code %in% missing_codes$code) |>
      select(cntry, essround, variable, code) |>
      left_join(missing_codes, by = "code") |>
      mutate(pf_id = NA_integer_, source = "missing"),
    by = c("cntry", "essround", "variable", "code")
  ) |>
  arrange(cntry, variable, essround, code) |>
  select(cntry, essround, variable, code, label, pf_name, pf_id, source, n)

# Nothing unmapped: only 66 / 99 stay empty
unmapped <- crosswalk |>
  filter(is.na(pf_name), !code %in% c(66L, 99L))
unmapped
stopifnot(nrow(unmapped) == 0)

pf_check(crosswalk, keys = c("cntry", "essround", "variable", "code"), core = core)

crosswalk |>
  mutate(
    kind = case_when(
      is.na(pf_name) ~ "NA (66 / 99)",
      pf_name %in% pf_sentinels ~ "sentinel",
      str_ends(pf_name, " \\[og\\]") ~ "[og]",
      .default = "Party Facts"
    )
  ) |>
  count(kind, source, wt = n, name = "responses")

write_csv(
  select(crosswalk, -n),
  here("data", "02_processed", "ess_party_pf.csv"),
  na = ""
)
```

# Respondent File

```{r}
#| label: respondents

coded <- items |>
  filter(!is.na(stub)) |>
  left_join(
    select(crosswalk, cntry, essround, variable, code, pf_name, pf_id),
    by = c("cntry", "essround", "variable", "code")
  ) |>
  select(cntry, essround, name, idno, stub, pf_name, pf_id) |>
  pivot_wider(
    names_from = stub,
    values_from = c(pf_name, pf_id),
    names_glue = "ext_{stub}_{.value}"
  )

respondents <- items |>
  distinct(cntry, essround, name, idno) |>
  left_join(coded, by = c("cntry", "essround", "name", "idno"))

stopifnot(
  nrow(respondents) == nrow(distinct(respondents, cntry, essround, name, idno))
)

stub_order <- c("prtcl", "prtv", "prtv_de1", "prtv_lt2", "prtv_lt3")
respondents <- respondents |>
  select(
    cntry,
    essround,
    name,
    idno,
    all_of(paste0("ext_", rep(stub_order, each = 2), c("_pf_name", "_pf_id")))
  ) |>
  arrange(cntry, essround, name, idno)

write_parquet(
  respondents,
  here("data", "02_processed", "ess_party_pf.parquet"),
  compression = "zstd"
)

glimpse(respondents)
```

# Checks

Spot checks: CDU/CSU and AfD in Germany, the Five Star Movement in Italy, and the most frequent `[og]` answers.

```{r}
#| label: spot-checks

crosswalk |>
  filter(
    (cntry == "DE" & str_detect(label, "CDU|AFD")) |
      (cntry == "IT" & str_detect(label, "5 Stelle"))
  ) |>
  distinct(cntry, label, pf_name, pf_id)

crosswalk |>
  filter(str_ends(pf_name, " \\[og\\]")) |>
  count(cntry, pf_name, wt = n, sort = TRUE) |>
  head(10)
```

Among respondents who report both an attachment party and a main-vote party, the share whose two answers carry the same Party Facts id.
This is a check on the coding, not a result: a country far below the others would point to a party coded differently on the two lists.
The low shares (Hungary, Romania, Poland, France) come mostly from electoral alliances, which Party Facts keeps as entities of their own: attachment to Fidesz with a vote for the Fidesz–KDNP list, PO with KO, PSD with the PSD+PC alliance, SLD with SLD–UP.
The coding keeps that distinction here; notebook 1.9 adds the alliance membership, under which such a pair counts as the same party.

```{r}
#| label: same-party

respondents |>
  filter(!is.na(ext_prtcl_pf_id), !is.na(ext_prtv_pf_id)) |>
  summarise(
    pairs = n(),
    same = mean(ext_prtcl_pf_id == ext_prtv_pf_id),
    .by = cntry
  ) |>
  arrange(same) |>
  print(n = Inf)
```

# Known Quirks

What a user of the two output files should know, collected in one place:

- **Automatic links and name matches skip parties that ended before 1950** (`pf_historic()` in `pf_code.R`). Party Facts links Spain's Partido Popular in the CSES and several other datasets to the Partido Progresista of 1834–74 (5750, short name "PP"), and short names had matched labels to Sweden's "Centern" of 1873–82 and Germany's Deutschkonservative Partei (1876–1918); the affected labels are hand-coded. The cutoff is 1950 rather than 1990 because Party Facts' `year_last` for small parties is often their last year in parliament (PPM, UDP and PSP show 1989 but existed later).
- **Party Facts' ESS link ids are not a code map.**
  An id `DE-7-6-c` says that party 6 of Germany's attachment list first appears in round 7.
  When ESS renumbers a list later, Party Facts writes no new link, so taking "the link with the largest start round up to this round" would put the wrong party on about 800 labels (Austria's code 5 is LIF in rounds 1–2 and the Greens in round 3).
  Links are therefore used only at their exact position and carried to other rounds through the label; links that disagree with each other, or with the label, go to the hand map.
- **Small parties that Party Facts files under a technical id** ("1 percent", "other", "unknown", ...) are coded `"<label> [og]"`, not the `other` sentinel: they are named parties, only not in Party Facts.
  The sentinel `other` is kept for generic "other party" answers.
- **Forza Italia** is the one label coded by round: 1626 (1994–2009) up to round 6, 8058 (refounded 2013) from round 7.
- **Solidarna Polska** keeps id 328 in every round, although Party Facts also has 9109 (Suwerenna Polska) for the same years and links ESS round 10 there; 9109 looks like a duplicate, and 328 matches the eu25games release.
- **A few Party Facts links are wrong** and are overridden by hand: PSL's round-1 link points to the PSL of 1945–49 (6757, coded 602), and Lithuania's Nationalist Union to the interwar party (377, coded 9107).
- **Only the 25 democracies** are coded; other ESS countries are not in the output.
- **Codes 66 and 99** (not applicable, no answer) stay `NA`; 77 and 88 become `refused` and `dont-know`. Codes above 66 that are real answers ("Other", "Blank vote", Italy's "voted only in the majoritarian race") are coded by their label.
- **Electoral alliances** (Fidesz–KDNP, KO, PSD+PC, SLD–UP, ...) keep their own Party Facts id here, so the attachment/vote check above understates agreement; the alliance-aware comparison is in notebook 1.9.

# Session Info

```{r}
#| label: session-info

session_info()
```

# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()
rendering_time <- end_time - start_time
message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds."
))
```
