---
title: |
  CSAP Code Notebook 1.2
subtitle: |
  EES voter studies: reported party attachment, 2004-2024
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup
#| 
# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "haven",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load EES Data

```{r}
#| label: load-ees-data

ees_2004 <- read_dta(
  here(
    "data",
    "01_raw",
    "ees",
    "2004",
    "ZA4566.dta"
  )
)

ees_2009 <- read_sav(
  here(
    "data",
    "01_raw",
    "ees",
    "2009",
    "ZA5055_v1-1-1.sav"
  )
)

ees_2014 <- read_dta(
  here(
    "data",
    "01_raw",
    "ees",
    "2014",
    "ZA5160_v4-1-0.dta"
  )
)

ees_2019 <- read_dta(
  here(
    "data",
    "01_raw",
    "ees",
    "2019",
    "ZA7581_v2-0-1.dta"
  )
)

ees_2024 <- read_csv2(
  here(
    "data",
    "01_raw",
    "ees",
    "2024",
    "ZA8868_v1-0-0.csv"
  )
)
```

# Select PID and other key Variables

The PID item is the same question in every wave -- *do you consider yourself to be close to
any particular party?* -- but the archived coding differs wave by wave, and in 2004, 2014 and
2019 it also differs from the questionnaire reproduced in `data/01_raw/key-items.qmd`. The
codes used below were read off the files themselves.

Two of the variable choices are not the obvious ones. In 2019 the raw `Q25` codes *no, not
close to any party* with a country-specific value (`101` in Austria, `2801` in the UK), so
the harmonised `Q25n` is used instead. In 2024 the reverse holds: `q19n` and `q19_ees` do not
follow the 2019 scheme, so the raw `q19` is used.

Left--right self-placement -- *in political matters people talk of "the left" and "the right";
what is your position?* -- is carried through the same chunks, for the final section of this
notebook. Nothing between here and that section uses it. Its item number moves as much as the
PID item's does: `v134` in 2004, `q46` in 2009, `qpp13` in 2014, `Q11` in 2019 and `q10` in
2024. Two of those need more than a missing-code filter. The 2014 file stores the scale as
`1`--`11` for 0--10, so it is shifted down by one; and 2004 is the only wave fielded on
1--10 rather than 0--10, so it is stretched onto the common scale with `(x - 1) * 10 / 9`.
That stretch is linear and multiplies the 2004 spread by exactly 10/9, which is a decision
about comparability and not a finding -- it is written as its own line so that deleting it
restores the raw 1--10 and replacing it with `NA_real_` drops the wave, neither of which
touches anything else. The 20 zeroes in the 2004 Swedish file fall outside the fielded range
and are dropped with the missing codes.

```{r}
#| label: select-2004

# 2004: PID item is v211
ees_2004 <- ees_2004 |>
  # Item unfielded in BE, LT
  filter(!is.na(v211)) |>
  mutate(
    year = 2004,
    country_raw = as.character(as_factor(country)),
    # 94 no PID, 99 dk/na, 95 refused
    pid_cat = case_when(
      v211 == 94 ~ "No attachment",
      v211 == 99 ~ "Don't know",
      v211 == 95 ~ "Refused or no answer",
      .default = "Named a party"
    ),
    # LR self-placement is v134, fielded 1-10 with stray 0s in Sweden
    lr = if_else(as.numeric(v134) %in% 1:10, as.numeric(v134), NA_real_),
    # The only 1-10 wave, stretched onto the 0-10 of every later one
    lr = (lr - 1) * 10 / 9
  ) |>
  select(year, country_raw, pid_cat, lr)
```

```{r}
#| label: select-2009

# 2009: PID item is q87
ees_2009 <- ees_2009 |>
  mutate(
    year = 2009,
    country_raw = as.character(as_factor(t102)),
    # 0 no, 88 dk, 77 refused, 99 nap
    pid_cat = case_when(
      q87 == 0 ~ "No attachment",
      q87 == 88 ~ "Don't know",
      q87 %in% c(77, 99) ~ "Refused or no answer",
      .default = "Named a party"
    ),
    # LR self-placement is q46, 0-10 with 77, 88 and 98 missing
    lr = if_else(as.numeric(q46) %in% 0:10, as.numeric(q46), NA_real_)
  ) |>
  select(year, country_raw, pid_cat, lr)
```

```{r}
#| label: select-2014

# 2014: PID item is qpp21
ees_2014 <- ees_2014 |>
  mutate(
    year = 2014,
    country_raw = as.character(as_factor(countrycode)),
    # 1 no, -9 dk, -8 refused
    pid_cat = case_when(
      qpp21 == 1 ~ "No attachment",
      qpp21 == -9 ~ "Don't know",
      qpp21 == -8 ~ "Refused or no answer",
      .default = "Named a party"
    ),
    # LR self-placement is qpp13, stored 1-11 for 0-10
    lr = if_else(as.numeric(qpp13) %in% 1:11, as.numeric(qpp13) - 1, NA_real_)
  ) |>
  select(year, country_raw, pid_cat, lr)
```

```{r}
#| label: select-2019

# 2019: harmonised PID item is Q25n
ees_2019 <- ees_2019 |>
  mutate(
    year = 2019,
    country_raw = as.character(as_factor(countrycode)),
    # 0 no, 98 dk, no refusal
    pid_cat = case_when(
      Q25n == 0 ~ "No attachment",
      Q25n == 98 ~ "Don't know",
      .default = "Named a party"
    ),
    # LR self-placement is Q11, 0-10 with 97 and 98 missing
    lr = if_else(as.numeric(Q11) %in% 0:10, as.numeric(Q11), NA_real_)
  ) |>
  select(year, country_raw, pid_cat, lr)
```

```{r}
#| label: select-2024

# 2024: PID item is q19
ees_2024 <- ees_2024 |>
  mutate(
    year = 2024,
    country_raw = country_alpha2,
    # 1 no, 98 dk, no refusal
    pid_cat = case_when(
      q19 == 1 ~ "No attachment",
      q19 == 98 ~ "Don't know",
      .default = "Named a party"
    ),
    # LR self-placement is q10, 0-10 with 98 missing
    lr = if_else(as.numeric(q10) %in% 0:10, as.numeric(q10), NA_real_)
  ) |>
  select(year, country_raw, pid_cat, lr)
```

# Join to Long Format

Country identifiers are spelled differently in every wave -- lowercase names in 2004, `The
Netherlands` in 2009 and 2014, `Czech Rep.` in 2019, alpha-2 codes with `EL` for Greece in
2024 -- so all of them are mapped onto ISO2 before the waves are stacked. In 2004 Britain and
Northern Ireland are two separate units; both are mapped to `GB` so that the wave is
comparable with 2009--2019, which field the UK as one.

```{r}
#| label: bind

ees_pid <- bind_rows(ees_2004, ees_2009, ees_2014, ees_2019, ees_2024) |>
  mutate(
    # Map every spelling to ISO2
    country = recode_values(
      str_to_lower(country_raw),
      c("austria", "at") ~ "AT",
      c("belgium", "be") ~ "BE",
      c("bulgaria", "bg") ~ "BG",
      c("croatia", "hr") ~ "HR",
      c("cyprus", "cy") ~ "CY",
      c("czech republic", "czech rep.", "cz") ~ "CZ",
      c("denmark", "dk") ~ "DK",
      c("estonia", "ee") ~ "EE",
      c("finland", "fi") ~ "FI",
      c("france", "fr") ~ "FR",
      c("germany", "de") ~ "DE",
      c("greece", "el") ~ "GR",
      c("hungary", "hu") ~ "HU",
      c("ireland", "ie") ~ "IE",
      c("italy", "it") ~ "IT",
      c("latvia", "lv") ~ "LV",
      c("lithuania", "lt") ~ "LT",
      c("luxembourg", "lu") ~ "LU",
      c("malta", "mt") ~ "MT",
      c("netherlands", "the netherlands", "nl") ~ "NL",
      c("poland", "pl") ~ "PL",
      c("portugal", "pt") ~ "PT",
      c("romania", "ro") ~ "RO",
      c("slovakia", "sk") ~ "SK",
      c("slovenia", "si") ~ "SI",
      c("spain", "es") ~ "ES",
      c("sweden", "se") ~ "SE",
      c("britain", "northern ireland", "united kingdom") ~ "GB"
    ),
    # Fix category order for the legend
    pid_cat = factor(
      pid_cat,
      levels = c(
        "Named a party",
        "No attachment",
        "Don't know",
        "Refused or no answer"
      )
    )
  ) |>
  # Keep the project's 25 democracies
  filter(
    country %in%
      c(
        "AT",
        "BE",
        "BG",
        "CZ",
        "DE",
        "DK",
        "EE",
        "ES",
        "FI",
        "FR",
        "GB",
        "GR",
        "HR",
        "HU",
        "IE",
        "IT",
        "LT",
        "LV",
        "NL",
        "PL",
        "PT",
        "RO",
        "SE",
        "SI",
        "SK"
      )
  ) |>
  select(year, country, pid_cat, lr)

# Respondents and countries per wave
ees_pid |>
  summarise(
    respondents = n(),
    countries = n_distinct(country),
    .by = year
  )
```

# Plot Share of Respondents reporting PID (incl. other), denying PID and Don't Know

```{r}
#| label: shares

# Pooled share across the 25
shares_pooled <- ees_pid |>
  count(year, pid_cat) |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Share within each country
shares_country <- ees_pid |>
  count(year, country, pid_cat) |>
  group_by(year, country) |>
  mutate(share = n / sum(n)) |>
  ungroup()

shares_pooled
```

Three things are worth keeping in view when reading the figures.

First, **the survey mode changes between 2014 and 2019**. The 2004--2014 waves are
interviewer-administered, while 2019 and 2024 are online panels that offer no refusal option
at all. The *Refused or no answer* series therefore ends after 2014 by design rather than
because respondents stopped refusing, and the don't-know share rises across the same seam.
Movement over that boundary is not cleanly attributable to respondents.

Second, **the 2004 don't-know category is contaminated**. The archived `v211` codes `99` as
`d/k, n/a`, merging don't-know with no-answer, so the 2004 don't-know share is an upper bound
and its refusal share is correspondingly understated.

Third, **the set of countries is not fixed**. Bulgaria, Croatia and Romania are absent in 2004
(Croatia also in 2009), Belgium and Lithuania carry no PID data in 2004, and Great Britain is
absent in 2024. The pooled line is therefore not a balanced panel; the faceted figure below
shows where the gaps fall.

```{r}
#| label: plot-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled trend, four categories
ggplot(shares_pooled, aes(x = year, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(breaks = c(2004, 2009, 2014, 2019, 2024)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the EES, 2004-2024",
    subtitle = "Unweighted shares, pooled over 25 European democracies",
    x = NULL,
    y = "Share of respondents",
    colour = NULL,
    caption = "No refusal option was offered in 2019 and 2024."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-country
#| fig-width: 11
#| fig-height: 11

# Same trend, one panel per country
ggplot(shares_country, aes(x = year, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5) +
  scale_x_continuous(breaks = c(2004, 2014, 2024)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the EES by country, 2004-2024",
    subtitle = "Unweighted shares; gaps mark waves in which a country was not surveyed",
    x = NULL,
    y = "Share of respondents",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

# Plot Spread of Left-Right Self-Placement among Partisans and Deniers

The figures above count people. This one measures how spread out they are. The question is
whether respondents who name a party occupy a wider or a narrower band of the left--right
scale than respondents who say they feel close to none, and whether that difference moves
across the five waves.

Two things the figure is not. It is not the distance *between* the groups: two groups can
share a mean and differ only in spread, which is roughly what happens here. And a standard
deviation cannot tell a wide unimodal distribution from a bimodal one -- two distributions
that look nothing alike can share an SD -- so nothing below speaks to sorting, or to
polarisation in any of its senses. It is one descriptive cut.

Don't-know and refusal are dropped throughout this section; the contrast is between the two
substantive answers only.

Before the spread, the coverage, because it is the constraint everything else here sits
inside.

```{r}
#| label: lr-coverage

# Who answers the placement item at all, by wave and attachment
lr_coverage <- ees_pid |>
  filter(pid_cat %in% c("Named a party", "No attachment")) |>
  summarise(
    respondents = n(),
    placed = mean(!is.na(lr)),
    .by = c(year, pid_cat)
  ) |>
  arrange(year, pid_cat)

lr_coverage
```

Between 92.7% and 95.7% of respondents who name a party place themselves on the scale,
against 71.9% to 83.8% of those who name none -- a gap of ten to twenty-one points, present
in all five waves and widest in 2014. The standard deviations below are therefore computed on
a self-selected subset, and the selection is *different in the two groups*: the deniers who
do answer are the more engaged end of that group, which is exactly the end whose placements
should look most partisan-like. The direction of the resulting bias cannot be signed from the
data. Dropping the least engaged could narrow the denier spread, if what goes is arbitrary
mid-scale answers, or widen it, if what goes is default fives. The table is printed rather
than hidden so that the size of the problem is visible.

The mode break described above runs through this table too. The 2019 and 2024 online panels
offer a don't-know option but no interviewer to probe with, and the wave with the worst
denier coverage is 2014, the last interviewer-administered one -- so the coverage series does
not line up with the mode seam in any simple way, and neither does the spread series below.

```{r}
#| label: lr-select

# Partisans and deniers with a usable placement
ees_lr <- ees_pid |>
  filter(pid_cat %in% c("Named a party", "No attachment"), !is.na(lr)) |>
  mutate(
    group = factor(
      if_else(pid_cat == "Named a party", "Partisans", "Deniers"),
      levels = c("Partisans", "Deniers")
    ),
    # Keep all 25 panels in the faceted figure
    country = factor(country, levels = sort(unique(ees_pid$country)))
  )
```

The pooled figure takes the standard deviation over all respondents of a group in a wave,
pooling respondents the same way the share figures above do. The faceted figure takes it
within each country-wave. Cells of a single respondent are dropped because `sd()` is
undefined there, and the ten thinnest are printed so their size is visible. That table is
what this section uses in place of confidence bands: at these sample sizes the sampling error
of a standard deviation is small enough that the movement worth arguing about is
compositional, not statistical -- but two of the country-wave cells are genuinely thin, and
are named below.

```{r}
#| label: lr-sd

# Pooled spread within each wave
sd_pooled <- ees_lr |>
  summarise(
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(year, group)
  ) |>
  arrange(year, group)

# Spread within each country-wave
sd_country <- ees_lr |>
  summarise(
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(country, year, group)
  ) |>
  # sd() is undefined below two respondents
  filter(respondents >= 2)

sd_pooled

# Thinnest cells behind the faceted figure
sd_country |>
  slice_min(respondents, n = 10)
```

The pooled statistic is the one vulnerable to the unbalanced panel -- 20 countries in 2004,
24 in 2009, 25 in 2014 and 2019, 24 in 2024 -- so it is recomputed below over only the
nineteen countries present in all five waves. If the two versions move together, entry and
exit are not driving the trend; if they diverge, composition is.

```{r}
#| label: lr-balanced

# Countries surveyed in every wave
balanced <- ees_lr |>
  summarise(waves = n_distinct(year), .by = country) |>
  filter(waves == n_distinct(ees_lr$year)) |>
  pull(country)

# Same pooled statistic over that fixed set
ees_lr |>
  filter(country %in% balanced) |>
  summarise(
    countries = n_distinct(country),
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(year, group)
  ) |>
  arrange(year, group)
```

Four things are worth keeping in view when reading the figures.

First, **the 2004 point sits on a stretched scale**. That wave is the only one fielded on
1--10, and the linear stretch onto 0--10 multiplies its standard deviation by exactly 10/9:
the partisan figure of 2.88 is 2.59 on the raw scale, the denier figure of 2.25 is 2.03. The
stretch is an assumption about how respondents use a 1--10 scale, not a measurement, so the
2004 point should not be read on its own.

Second, **the groups are not fixed populations**. The share naming a party rises across the
five waves -- that is precisely what the figures above document -- so the marginal respondent
crosses between the two groups over time. A change in a group's spread can therefore be
compositional rather than attitudinal, and a standard deviation cannot tell the two readings
apart.

Third, **pooling mixes two variances**. The pooled standard deviation contains both the
spread within countries and the differences between countries in mean position, and the
country set is not fixed. The balanced-panel table above and the faceted figure below are the
two checks on that; neither is a correction.

Fourth, **two country-wave cells are too thin to read**. Estonia's 2004 deniers number nine
respondents and Italy's thirty-six; every other cell in the series runs into the hundreds.
Those two points are plotted because dropping them silently would be worse, but the Estonian
one in particular carries no information.

```{r}
#| label: plot-lr-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled spread, two groups
ggplot(sd_pooled, aes(x = year, y = sd_lr, colour = group)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(breaks = c(2004, 2009, 2014, 2019, 2024)) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Spread of left-right self-placement in the EES, 2004-2024",
    subtitle = "Unweighted standard deviation on a 0-10 scale, pooled over 25 democracies",
    x = NULL,
    y = "Standard deviation of self-placement",
    colour = NULL,
    caption = "The 2004 scale is stretched from 1-10 onto 0-10; non-placers are excluded."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-lr-country
#| fig-width: 11
#| fig-height: 11

# Same spread, one panel per country
ggplot(sd_country, aes(x = year, y = sd_lr, colour = group)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5, drop = FALSE) +
  scale_x_continuous(breaks = c(2004, 2014, 2024)) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Spread of left-right self-placement in the EES by country, 2004-2024",
    subtitle = "Unweighted standard deviations; gaps mark waves without a usable cell",
    x = NULL,
    y = "Standard deviation of self-placement",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

The two series do not cross. Respondents who name a party are the more dispersed group in
every wave, at a standard deviation between 2.79 and 2.98 against the deniers' 2.05 to 2.25,
a gap running from 0.59 scale points in 2024 to 0.84 in 2014. Neither series has much of a
trend of its own: the partisan one drifts from 2.88 in 2004 to 2.81 in 2024 with a peak in
2009, the denier one dips to 2.05 in 2014 and returns to 2.22. The balanced-panel table
above reproduces both series to within a tenth of a scale point at every wave, so the
country set entering and leaving is not what the pooled line is showing.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
