---
title: |
  CSAP Code Notebook 1.2
subtitle: |
  EES voter studies: reported party attachment, 2004-2024
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup
#| 
# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "haven",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load EES Data

```{r}
#| label: load-ees-data

ees_2004 <- read_dta(
  here(
    "data",
    "01_raw",
    "ees",
    "2004",
    "ZA4566.dta"
  )
)

ees_2009 <- read_sav(
  here(
    "data",
    "01_raw",
    "ees",
    "2009",
    "ZA5055_v1-1-1.sav"
  )
)

ees_2014 <- read_dta(
  here(
    "data",
    "01_raw",
    "ees",
    "2014",
    "ZA5160_v4-1-0.dta"
  )
)

ees_2019 <- read_dta(
  here(
    "data",
    "01_raw",
    "ees",
    "2019",
    "ZA7581_v2-0-1.dta"
  )
)

ees_2024 <- read_csv2(
  here(
    "data",
    "01_raw",
    "ees",
    "2024",
    "ZA8868_v1-0-0.csv"
  )
)
```

# Select PID and other key Variables

The PID item is the same question in every wave -- *do you consider yourself to be close to
any particular party?* -- but the archived coding differs wave by wave, and in 2004, 2014 and
2019 it also differs from the questionnaire reproduced in `data/01_raw/key-items.qmd`. The
codes used below were read off the files themselves.

Two of the variable choices are not the obvious ones. In 2019 the raw `Q25` codes *no, not
close to any party* with a country-specific value (`101` in Austria, `2801` in the UK), so
the harmonised `Q25n` is used instead. In 2024 the reverse holds: `q19n` and `q19_ees` do not
follow the 2019 scheme, so the raw `q19` is used.

```{r}
#| label: select-2004

# 2004: PID item is v211
ees_2004 <- ees_2004 |>
  # Item unfielded in BE, LT
  filter(!is.na(v211)) |>
  mutate(
    year = 2004,
    country_raw = as.character(as_factor(country)),
    # 94 no PID, 99 dk/na, 95 refused
    pid_cat = case_when(
      v211 == 94 ~ "No attachment",
      v211 == 99 ~ "Don't know",
      v211 == 95 ~ "Refused or no answer",
      .default = "Named a party"
    )
  ) |>
  select(year, country_raw, pid_cat)
```

```{r}
#| label: select-2009

# 2009: PID item is q87
ees_2009 <- ees_2009 |>
  mutate(
    year = 2009,
    country_raw = as.character(as_factor(t102)),
    # 0 no, 88 dk, 77 refused, 99 nap
    pid_cat = case_when(
      q87 == 0 ~ "No attachment",
      q87 == 88 ~ "Don't know",
      q87 %in% c(77, 99) ~ "Refused or no answer",
      .default = "Named a party"
    )
  ) |>
  select(year, country_raw, pid_cat)
```

```{r}
#| label: select-2014

# 2014: PID item is qpp21
ees_2014 <- ees_2014 |>
  mutate(
    year = 2014,
    country_raw = as.character(as_factor(countrycode)),
    # 1 no, -9 dk, -8 refused
    pid_cat = case_when(
      qpp21 == 1 ~ "No attachment",
      qpp21 == -9 ~ "Don't know",
      qpp21 == -8 ~ "Refused or no answer",
      .default = "Named a party"
    )
  ) |>
  select(year, country_raw, pid_cat)
```

```{r}
#| label: select-2019

# 2019: harmonised PID item is Q25n
ees_2019 <- ees_2019 |>
  mutate(
    year = 2019,
    country_raw = as.character(as_factor(countrycode)),
    # 0 no, 98 dk, no refusal
    pid_cat = case_when(
      Q25n == 0 ~ "No attachment",
      Q25n == 98 ~ "Don't know",
      .default = "Named a party"
    )
  ) |>
  select(year, country_raw, pid_cat)
```

```{r}
#| label: select-2024

# 2024: PID item is q19
ees_2024 <- ees_2024 |>
  mutate(
    year = 2024,
    country_raw = country_alpha2,
    # 1 no, 98 dk, no refusal
    pid_cat = case_when(
      q19 == 1 ~ "No attachment",
      q19 == 98 ~ "Don't know",
      .default = "Named a party"
    )
  ) |>
  select(year, country_raw, pid_cat)
```

# Join to Long Format

Country identifiers are spelled differently in every wave -- lowercase names in 2004, `The
Netherlands` in 2009 and 2014, `Czech Rep.` in 2019, alpha-2 codes with `EL` for Greece in
2024 -- so all of them are mapped onto ISO2 before the waves are stacked. In 2004 Britain and
Northern Ireland are two separate units; both are mapped to `GB` so that the wave is
comparable with 2009--2019, which field the UK as one.

```{r}
#| label: bind

ees_pid <- bind_rows(ees_2004, ees_2009, ees_2014, ees_2019, ees_2024) |>
  mutate(
    # Map every spelling to ISO2
    country = recode_values(
      str_to_lower(country_raw),
      c("austria", "at") ~ "AT",
      c("belgium", "be") ~ "BE",
      c("bulgaria", "bg") ~ "BG",
      c("croatia", "hr") ~ "HR",
      c("cyprus", "cy") ~ "CY",
      c("czech republic", "czech rep.", "cz") ~ "CZ",
      c("denmark", "dk") ~ "DK",
      c("estonia", "ee") ~ "EE",
      c("finland", "fi") ~ "FI",
      c("france", "fr") ~ "FR",
      c("germany", "de") ~ "DE",
      c("greece", "el") ~ "GR",
      c("hungary", "hu") ~ "HU",
      c("ireland", "ie") ~ "IE",
      c("italy", "it") ~ "IT",
      c("latvia", "lv") ~ "LV",
      c("lithuania", "lt") ~ "LT",
      c("luxembourg", "lu") ~ "LU",
      c("malta", "mt") ~ "MT",
      c("netherlands", "the netherlands", "nl") ~ "NL",
      c("poland", "pl") ~ "PL",
      c("portugal", "pt") ~ "PT",
      c("romania", "ro") ~ "RO",
      c("slovakia", "sk") ~ "SK",
      c("slovenia", "si") ~ "SI",
      c("spain", "es") ~ "ES",
      c("sweden", "se") ~ "SE",
      c("britain", "northern ireland", "united kingdom") ~ "GB"
    ),
    # Fix category order for the legend
    pid_cat = factor(
      pid_cat,
      levels = c(
        "Named a party",
        "No attachment",
        "Don't know",
        "Refused or no answer"
      )
    )
  ) |>
  # Keep the project's 25 democracies
  filter(
    country %in%
      c(
        "AT",
        "BE",
        "BG",
        "CZ",
        "DE",
        "DK",
        "EE",
        "ES",
        "FI",
        "FR",
        "GB",
        "GR",
        "HR",
        "HU",
        "IE",
        "IT",
        "LT",
        "LV",
        "NL",
        "PL",
        "PT",
        "RO",
        "SE",
        "SI",
        "SK"
      )
  ) |>
  select(year, country, pid_cat)

# Respondents and countries per wave
ees_pid |>
  summarise(
    respondents = n(),
    countries = n_distinct(country),
    .by = year
  )
```

# Plot Share of Respondents reporting PID (incl. other), denying PID and Don't Know

```{r}
#| label: shares

# Pooled share across the 25
shares_pooled <- ees_pid |>
  count(year, pid_cat) |>
  group_by(year) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Share within each country
shares_country <- ees_pid |>
  count(year, country, pid_cat) |>
  group_by(year, country) |>
  mutate(share = n / sum(n)) |>
  ungroup()

shares_pooled
```

Three things are worth keeping in view when reading the figures.

First, **the survey mode changes between 2014 and 2019**. The 2004--2014 waves are
interviewer-administered, while 2019 and 2024 are online panels that offer no refusal option
at all. The *Refused or no answer* series therefore ends after 2014 by design rather than
because respondents stopped refusing, and the don't-know share rises across the same seam.
Movement over that boundary is not cleanly attributable to respondents.

Second, **the 2004 don't-know category is contaminated**. The archived `v211` codes `99` as
`d/k, n/a`, merging don't-know with no-answer, so the 2004 don't-know share is an upper bound
and its refusal share is correspondingly understated.

Third, **the set of countries is not fixed**. Bulgaria, Croatia and Romania are absent in 2004
(Croatia also in 2009), Belgium and Lithuania carry no PID data in 2004, and Great Britain is
absent in 2024. The pooled line is therefore not a balanced panel; the faceted figure below
shows where the gaps fall.

```{r}
#| label: plot-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled trend, four categories
ggplot(shares_pooled, aes(x = year, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(breaks = c(2004, 2009, 2014, 2019, 2024)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the EES, 2004-2024",
    subtitle = "Unweighted shares, pooled over 25 European democracies",
    x = NULL,
    y = "Share of respondents",
    colour = NULL,
    caption = "No refusal option was offered in 2019 and 2024."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-country
#| fig-width: 11
#| fig-height: 11

# Same trend, one panel per country
ggplot(shares_country, aes(x = year, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5) +
  scale_x_continuous(breaks = c(2004, 2014, 2024)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the EES by country, 2004-2024",
    subtitle = "Unweighted shares; gaps mark waves in which a country was not surveyed",
    x = NULL,
    y = "Share of respondents",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
