---
title: |
  CSAP Code Notebook 1.3
subtitle: |
  CSES IMD and module 6: reported party attachment, 1996-2024
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "haven",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load CSES Data

Both releases are `.rdata` archives, so `load()` is called for its side effect: it creates the
object `cses_imd` (modules 1--5) and the object `cses6` (module 6) in the environment rather
than returning them. The module 6 archive also carries a stray `.Traceback` object, which is
ignored.

```{r}
#| label: load-cses-data

# Creates the object cses_imd
load(
  here(
    "data",
    "01_raw",
    "cses",
    "imd",
    "cses_imd.rdata"
  )
)

# Creates the object cses6
load(
  here(
    "data",
    "01_raw",
    "cses",
    "mod6",
    "cses6.rdata"
  )
)
```

# Select PID and other key Variables

The attachment item is the same question in both releases -- *do you usually think of yourself
as close to any particular party?* -- and shares one coding: `1` yes, `0` no, `8` don't know,
`7` refused and `9` missing. Only the variable names differ, and they differ completely: the
IMD uses upper-case `IMD`-prefixed names, module 6 lower-case `f`-prefixed ones.

CSES asks a follow-up probe of everyone who does not say yes (*do you feel yourself a little
closer to one of the political parties than the others?*, `IMD3005_2` / `f3023_2`). That probe
is deliberately **not** used here: the base item alone is the analogue of the EES question in
notebook 1.2, and the probe is unfielded in nine of the studies below.

```{r}
#| label: select-imd

# IMD: base PID item is IMD3005_1
cses_imd <- cses_imd |>
  mutate(
    # Module flags are study-level
    module = case_when(
      IMD1008_MOD_1 == 1 ~ 1,
      IMD1008_MOD_2 == 1 ~ 2,
      IMD1008_MOD_3 == 1 ~ 3,
      IMD1008_MOD_4 == 1 ~ 4,
      IMD1008_MOD_5 == 1 ~ 5
    ),
    study = IMD1004,
    country = IMD1006_UNALPHA2,
    year = IMD1008_YEAR,
    close = IMD3005_1,
    # 1 yes, 0 no, 8 dk, 7/9 refused
    pid_cat = case_when(
      close == 1 ~ "Close to a party",
      close == 0 ~ "No attachment",
      close == 8 ~ "Don't know",
      .default = "Refused or no answer"
    )
  ) |>
  select(module, study, country, year, close, pid_cat)
```

```{r}
#| label: select-mod6

# Module 6: base PID item is f3023_1
cses_mod6 <- cses6 |>
  mutate(
    module = 6,
    study = f1004,
    # Module 6 codes Slovenia SL
    country = if_else(f1006_unalpha2 == "SL", "SI", f1006_unalpha2),
    year = f1010_y,
    close = f3023_1,
    # 1 yes, 0 no, 8 dk, 7/9 refused
    pid_cat = case_when(
      close == 1 ~ "Close to a party",
      close == 0 ~ "No attachment",
      close == 8 ~ "Don't know",
      .default = "Refused or no answer"
    )
  ) |>
  select(module, study, country, year, close, pid_cat)
```

# Join to Long Format

The two releases do not overlap -- the IMD's latest election is 2021, module 6 begins in 2022
-- so no study is counted twice. Country identifiers are already alpha-2 in both files, so
unlike the EES in notebook 1.2 no name lookup is needed; the one exception is module 6's `SL`
for Slovenia, corrected above.

Two studies never fielded the item at all and are dropped whole rather than counted as
missing: `BELW1999` (Wallonia) and `LVA_2018`. Both are caught from the data, by dropping any
study in which every case is coded `9`.

```{r}
#| label: bind

cses_pid <- bind_rows(cses_imd, cses_mod6) |>
  # Drop studies never fielding the item
  filter(!all(close == 9), .by = study) |>
  # Keep the project's 25 democracies
  filter(
    country %in%
      c(
        "AT",
        "BE",
        "BG",
        "CZ",
        "DE",
        "DK",
        "EE",
        "ES",
        "FI",
        "FR",
        "GB",
        "GR",
        "HR",
        "HU",
        "IE",
        "IT",
        "LT",
        "LV",
        "NL",
        "PL",
        "PT",
        "RO",
        "SE",
        "SI",
        "SK"
      )
  ) |>
  mutate(
    # Fix category order for the legend
    pid_cat = factor(
      pid_cat,
      levels = c(
        "Close to a party",
        "No attachment",
        "Don't know",
        "Refused or no answer"
      )
    )
  ) |>
  select(module, study, country, year, pid_cat)

# Respondents, studies and countries per module
cses_pid |>
  summarise(
    respondents = n(),
    studies = n_distinct(study),
    countries = n_distinct(country),
    .by = module
  ) |>
  arrange(module)
```

# Plot Share of Respondents reporting PID (incl. other), denying PID and Don't Know

Unlike the EES, the CSES has no fixed fieldwork years: each study follows a national election,
so the 25 countries contribute 109 country-years scattered between 1996 and 2024. The pooled
figure therefore runs over the six CSES modules, each of which is a defined five-year window,
while the country panels keep the election year on the x-axis.

Three country-years are fielded as two studies each and are pooled into one point: Belgium 2019
(Flanders and Wallonia), Germany 2002, and the two Greek elections of 2015. Belgium 1999 is
not pooled, because the Walloon half of it never fielded the item.

```{r}
#| label: shares

# Pooled share within each module
shares_pooled <- cses_pid |>
  count(module, pid_cat) |>
  group_by(module) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Share within each country-year
shares_country <- cses_pid |>
  count(country, year, pid_cat) |>
  group_by(country, year) |>
  mutate(share = n / sum(n)) |>
  ungroup()

shares_pooled
```

Six things are worth keeping in view when reading the figures.

First, **close to a party is not the same as naming a party**. The CSES item is binary; the
party itself is asked at a separate variable (`IMD3005_3`). A respondent can be close to a
party here without ever naming one, so this series is not level-comparable with the EES series
in notebook 1.2 -- only its movement is.

Second, **the leaner probe is excluded**, and it is large. Respondents who say no to the base
item but yes to the follow-up probe are between 15% and 25% of all respondents depending on
the module. The *No attachment* band therefore contains a substantial layer of leaners, and
its level should not be read as the share with no partisan leaning whatsoever.

Third, **code `9` conflates missing with not asked**. The *Refused or no answer* band mixes
true item nonresponse with study-level gaps in administration, and only studies in which every
single case is `9` are dropped.

Fourth, **modules 1 and 2 used deviating wording**, asking about any particular *political*
party. That deviation sits exactly under the first two points of the pooled line.

Fifth, **CSES folds inconsistent answers into the binary** rather than dropping them: a "yes"
with no party identified is coded `1` and a "no" with a party identified is coded `0`, and
neither is corrected here.

Sixth, **the panel is unbalanced and the unit is an election, not a year**. Between 8 and 19 of
the 25 countries appear per module; Estonia and Croatia contribute a single study each, while
Czechia, Germany, Poland and Portugal contribute seven. The faceted figure exists to make this
visible.

```{r}
#| label: plot-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled trend, four categories
ggplot(shares_pooled, aes(x = module, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    breaks = 1:6,
    labels = c(
      "1\n1996-2001",
      "2\n2001-2006",
      "3\n2006-2011",
      "4\n2011-2016",
      "5\n2016-2021",
      "6\n2021-2026"
    )
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the CSES, modules 1-6",
    subtitle = "Unweighted shares, pooled over 25 European democracies",
    x = "CSES module",
    y = "Share of respondents",
    colour = NULL,
    caption = "The branching leaner probe is not used."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-country
#| fig-width: 11
#| fig-height: 11

# Same trend, one panel per country
ggplot(shares_country, aes(x = year, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5) +
  scale_x_continuous(breaks = c(2000, 2010, 2020)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the CSES by country, 1996-2024",
    subtitle = "Unweighted shares; each point is one national election study",
    x = NULL,
    y = "Share of respondents",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
