---
title: |
  CSAP Code Notebook 1.4
subtitle: |
  ESS rounds 1-11: reported party attachment over time
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "arrow",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load ESS Data

The source is the partitioned Parquet dataset written by notebook 1.1, not the 1.5 GB CSV it
was built from. The whole point of that conversion is that a reader fetches only the columns
it needs, so nothing here parses the extract again.

```{r}
#| label: load-ess-data

# Partitioned Parquet, never the CSV
ess <- open_dataset(
  here(
    "data",
    "01_raw",
    "ess",
    "parquet"
  ),
  partitioning = hive_partition(cntry = utf8(), essround = int64())
)

ess
```

# Select PID and other key Variables

The attachment item is `clsprty` -- *is there a particular political party you feel closer to
than all the other parties?* -- a core item asked of everyone, with wording and coding constant
across all eleven rounds: `1` yes, `2` no, `7` refusal, `8` don't know, `9` no answer. Unlike
the EES and CSES files it carries no `NA` at all, and no country-round cell is missing it, so
nothing has to be dropped.

The `name` column is kept because it is the only thing separating round 10's two stacked
samples, which the caveats below depend on.

```{r}
#| label: select

# Four columns of 430,581 rows
ess_pid <- ess |>
  select(cntry, essround, name, clsprty) |>
  collect() |>
  mutate(
    round = essround,
    country = cntry,
    # 1 yes, 2 no, 8 dk, 7/9 refused
    pid_cat = case_when(
      clsprty == 1 ~ "Close to a party",
      clsprty == 2 ~ "No attachment",
      clsprty == 8 ~ "Don't know",
      .default = "Refused or no answer"
    )
  ) |>
  select(round, country, name, pid_cat)
```

# Join to Long Format

There is nothing to join: the ESS arrives as one long file, the extract already holds exactly
the project's 25 democracies, and country identifiers are already alpha-2. The country filter
below is therefore a no-op, kept as a guard and for symmetry with notebooks 1.2 and 1.3.

```{r}
#| label: bind

ess_pid <- ess_pid |>
  # Extract already holds the 25
  filter(
    country %in%
      c(
        "AT",
        "BE",
        "BG",
        "CZ",
        "DE",
        "DK",
        "EE",
        "ES",
        "FI",
        "FR",
        "GB",
        "GR",
        "HR",
        "HU",
        "IE",
        "IT",
        "LT",
        "LV",
        "NL",
        "PL",
        "PT",
        "RO",
        "SE",
        "SI",
        "SK"
      )
  ) |>
  mutate(
    # Fix category order for the legend
    pid_cat = factor(
      pid_cat,
      levels = c(
        "Close to a party",
        "No attachment",
        "Don't know",
        "Refused or no answer"
      )
    )
  )

# Respondents and countries per round
ess_pid |>
  summarise(
    respondents = n(),
    countries = n_distinct(country),
    .by = round
  ) |>
  arrange(round)
```

# Plot Share of Respondents reporting PID (incl. other), denying PID and Don't Know

The x-axis is the ESS round rather than the fieldwork year. Year is derivable but not usable
as an axis: it lives in `inwyr` for rounds 1--2, `inwyys` for rounds 3--9 and `inwds` for
rounds 10--11, and the 18,386 respondents of round 10's self-completion sample carry no
interview date at all. The modal fieldwork year is therefore shown only as a label under each
round.

```{r}
#| label: shares

# Pooled share within each round
shares_pooled <- ess_pid |>
  count(round, pid_cat) |>
  group_by(round) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Share within each country-round
shares_country <- ess_pid |>
  count(country, round, pid_cat) |>
  group_by(country, round) |>
  mutate(share = n / sum(n)) |>
  ungroup()

shares_pooled
```

Four things are worth keeping in view when reading the figures.

First, **round 10 mixes two survey modes**. It is two stacked samples sharing `essround == 10`:
the face-to-face `ESS10e03_3` in 17 countries and the self-completion `ESS10SCe03_2` in Austria,
Germany, Spain, Latvia, Poland and Sweden. The self-completion half returns 51.2% close to a
party against the face-to-face half's 40.5%, and offers no refusal or don't-know code at all --
zero of each, against 703 in the face-to-face half -- which is why the round-10 don't-know share
is the lowest in the series. Because the two samples cover disjoint sets of countries, mode is
fully confounded with country composition there, and the gap cannot be attributed to mode alone.

Second, **`clsprty` is a filter item, not a party name**. The party actually named is recorded
separately in the country-specific `prtcl<cc>` family. As in notebook 1.3, this series is
therefore not level-comparable with the EES series in notebook 1.2 -- only its movement is.

Third, **the refusal band is not comparable across the three notebooks**. The ESS uses codes
`7` and `9` sparingly, 0.2% to 1.5% per round, against 4% to 7% in the EES. That difference is
coding practice, not respondents.

Fourth, **the panel is unbalanced**. Between 18 and 23 of the 25 countries appear per round.
Thirteen appear in all eleven rounds; Romania appears in one (round 4), Latvia in four and
Croatia in five. The faceted figure exists to make this visible.

```{r}
#| label: plot-pooled
#| fig-width: 9
#| fig-height: 5

# Pooled trend, four categories
ggplot(shares_pooled, aes(x = round, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    breaks = 1:11,
    labels = c(
      "1\n2002",
      "2\n2004",
      "3\n2006",
      "4\n2008",
      "5\n2011",
      "6\n2012",
      "7\n2014",
      "8\n2016",
      "9\n2018",
      "10\n2021",
      "11\n2023"
    )
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the ESS, rounds 1-11",
    subtitle = "Unweighted shares, pooled over 25 European democracies",
    x = "ESS round and modal fieldwork year",
    y = "Share of respondents",
    colour = NULL,
    caption = "Round 10 pools a face-to-face and a self-completion sample."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-country
#| fig-width: 11
#| fig-height: 11

# Same trend, one panel per country
ggplot(shares_country, aes(x = round, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5) +
  scale_x_continuous(breaks = c(1, 6, 11)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the ESS by country, rounds 1-11",
    subtitle = "Unweighted shares; gaps mark rounds in which a country was not surveyed",
    x = "ESS round",
    y = "Share of respondents",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

Read next to the other two notebooks, the ESS series declines from 50.9% to 43.1% close to a
party, moving in the same direction as the CSES series in notebook 1.3 (44.5% to 40.4%) and
against the EES series in notebook 1.2 (55.9% to 66.3%). Two of the three sources agree; the
EES is the one that does not, and its late waves are also the ones carrying the mode break
described there.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
