---
title: |
  Code Notebook 1.5
subtitle: |
  EU25 Games Data PID Distribution 
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "ggh4x",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load Data

```{r}
#| label: load-eu25-data

eu25games <- read_rds(
  here(
    "data",
    "01_raw",
    "eu25games2019.rds"
  )
)
```


# Select PID and other key Variables

The PID item is the same as used by the EES:
"Do you consider yourself to be close to any particular party? If so, which party do you feel close to?"

The central "No, I do not feel close to any political party" response is condensed to value "none" in the harmonized variable used below.

```{r}
#| label: select-vars

eu25games <- eu25games |>
  # only keep respondents who passed att check
  filter(
    der_att_check_3
  ) |>
  select(
    # Country
    meta_country,
    # party id matched to partyfacts
    ext_q_party_id_pf_name,
    # lr self
    q_lr_self,
    # eu self
    q_euint_self
  )
```

# PID Shares

At this point, treating "other" as reported attachement (just not of the listed parties).

```{r}
#| label: shares

# Pooled shares
shares_pooled <- eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  count(pid) |>
  mutate(
    share = n / sum(n),
    meta_country = "EU25"
  ) |>
  relocate(meta_country, .before = pid)

# Shares within country
shares_country <- eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  group_by(meta_country) |>
  count(pid) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# bind for plotting
toplot <- shares_pooled |>
  bind_rows(shares_country)
```

```{r}
#| label: fig-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled shares
ggplot(toplot, aes(y = meta_country, x = share, fill = pid)) +
  geom_col(position = "fill") +
  scale_colour_brewer(palette = "Dark2") +
  theme_minimal()
```


# LR Pos by PID Type

```{r}
#| label: lrpos-by-pid-pooled

# how many missings in lr self per type
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  group_by(pid) |>
  summarise(
    sum(is.na(q_lr_self)),
    sum(!is.na(q_lr_self))
  )

# summary stats per type
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  group_by(pid) |>
  summarise(
    mean_lr = mean(q_lr_self, na.rm = T),
    sd_lr = sd(q_lr_self, na.rm = T),
    nobs = n()
  )

# density per type
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  ggplot(
    aes(x = q_lr_self, fill = pid)
  ) +
  geom_histogram(position = "dodge")

# or alternatively
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  ggplot(
    aes(x = q_lr_self, fill = pid)
  ) +
  # geom_histogram() +
  geom_density() +
  facet_wrap(~pid)
```

```{r}
#| label: lrpos-by-pid-bycountry

# how many missings in lr self per type
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  group_by(meta_country, pid) |>
  summarise(
    sum(is.na(q_lr_self)),
    sum(!is.na(q_lr_self))
  ) |>
  ungroup()

# summary stats per type
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  group_by(meta_country, pid) |>
  summarise(
    mean_lr = mean(q_lr_self, na.rm = T),
    sd_lr = sd(q_lr_self, na.rm = T),
    nobs = n()
  ) |>
  ungroup()

# density per type
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  ggplot(
    aes(x = q_lr_self, fill = pid)
  ) +
  geom_histogram(position = "dodge") +
  facet_wrap(~meta_country)

# or alternatively
eu25games |>
  mutate(
    pid = case_when(
      ext_q_party_id_pf_name == "none" ~ "No attachement",
      is.na(ext_q_party_id_pf_name) ~ "Answer Missing",
      .default = "Named a party"
    )
  ) |>
  # rm missing
  filter(pid != "Answer Missing") |>
  ggplot(
    aes(x = q_lr_self, fill = pid)
  ) +
  # geom_histogram() +
  geom_density(alpha = 0.7) +
  facet_wrap(~meta_country)
```


# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```


  
