---
title: |
  CSAP Code Notebook 1.3
subtitle: |
  CSES IMD and module 6: reported party attachment, 1996-2024
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "haven",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load CSES Data

Both releases are `.rdata` archives, so `load()` is called for its side effect: it creates the
object `cses_imd` (modules 1--5) and the object `cses6` (module 6) in the environment rather
than returning them. The module 6 archive also carries a stray `.Traceback` object, which is
ignored.

```{r}
#| label: load-cses-data

# Creates the object cses_imd
load(
  here(
    "data",
    "01_raw",
    "cses",
    "imd",
    "cses_imd.rdata"
  )
)

# Creates the object cses6
load(
  here(
    "data",
    "01_raw",
    "cses",
    "mod6",
    "cses6.rdata"
  )
)
```

# Select PID and other key Variables

The attachment item is the same question in both releases -- *do you usually think of yourself
as close to any particular party?* -- and shares one coding: `1` yes, `0` no, `8` don't know,
`7` refused and `9` missing. Only the variable names differ, and they differ completely: the
IMD uses upper-case `IMD`-prefixed names, module 6 lower-case `f`-prefixed ones.

CSES asks a follow-up probe of everyone who does not say yes (*do you feel yourself a little
closer to one of the political parties than the others?*, `IMD3005_2` / `f3023_2`). That probe
is deliberately **not** used here: the base item alone is the analogue of the EES question in
notebook 1.2, and the probe is unfielded in nine of the studies below.

Left--right self-placement -- *where would you place yourself on a scale from 0 to 10, where
0 means the left and 10 means the right?* -- is carried through the same chunks, for the final
section of this notebook. Nothing between here and that section uses it. It is `IMD3006` in
the IMD and `f3020_r` in module 6, 0--10 in both, with `95` haven't heard of left-right, `97`
refused, `98` don't know where to place and `99` missing. Note that these are the *self*
items: the party-placement battery documented in `data/01_raw/key-items.qmd` is `IMD3007_A`
onwards, and is not used here.

```{r}
#| label: select-imd

# IMD: base PID item is IMD3005_1
cses_imd <- cses_imd |>
  mutate(
    # Module flags are study-level
    module = case_when(
      IMD1008_MOD_1 == 1 ~ 1,
      IMD1008_MOD_2 == 1 ~ 2,
      IMD1008_MOD_3 == 1 ~ 3,
      IMD1008_MOD_4 == 1 ~ 4,
      IMD1008_MOD_5 == 1 ~ 5
    ),
    study = IMD1004,
    country = IMD1006_UNALPHA2,
    year = IMD1008_YEAR,
    close = IMD3005_1,
    # 1 yes, 0 no, 8 dk, 7/9 refused
    pid_cat = case_when(
      close == 1 ~ "Close to a party",
      close == 0 ~ "No attachment",
      close == 8 ~ "Don't know",
      .default = "Refused or no answer"
    ),
    # Whitelisting the scale drops 95, 97, 98 and 99 in one go
    lr = if_else(as.numeric(IMD3006) %in% 0:10, as.numeric(IMD3006), NA_real_)
  ) |>
  select(module, study, country, year, close, pid_cat, lr)
```

```{r}
#| label: select-mod6

# Module 6: base PID item is f3023_1
cses_mod6 <- cses6 |>
  mutate(
    module = 6,
    study = f1004,
    # Module 6 codes Slovenia SL
    country = if_else(f1006_unalpha2 == "SL", "SI", f1006_unalpha2),
    year = f1010_y,
    close = f3023_1,
    # 1 yes, 0 no, 8 dk, 7/9 refused
    pid_cat = case_when(
      close == 1 ~ "Close to a party",
      close == 0 ~ "No attachment",
      close == 8 ~ "Don't know",
      .default = "Refused or no answer"
    ),
    # Same scale, lower-case name
    lr = if_else(as.numeric(f3020_r) %in% 0:10, as.numeric(f3020_r), NA_real_)
  ) |>
  select(module, study, country, year, close, pid_cat, lr)
```

# Join to Long Format

The two releases do not overlap -- the IMD's latest election is 2021, module 6 begins in 2022
-- so no study is counted twice. Country identifiers are already alpha-2 in both files, so
unlike the EES in notebook 1.2 no name lookup is needed; the one exception is module 6's `SL`
for Slovenia, corrected above.

Two studies never fielded the item at all and are dropped whole rather than counted as
missing: `BELW1999` (Wallonia) and `LVA_2018`. Both are caught from the data, by dropping any
study in which every case is coded `9`.

```{r}
#| label: bind

cses_pid <- bind_rows(cses_imd, cses_mod6) |>
  # Drop studies never fielding the item
  filter(!all(close == 9), .by = study) |>
  # Keep the project's 25 democracies
  filter(
    country %in%
      c(
        "AT",
        "BE",
        "BG",
        "CZ",
        "DE",
        "DK",
        "EE",
        "ES",
        "FI",
        "FR",
        "GB",
        "GR",
        "HR",
        "HU",
        "IE",
        "IT",
        "LT",
        "LV",
        "NL",
        "PL",
        "PT",
        "RO",
        "SE",
        "SI",
        "SK"
      )
  ) |>
  mutate(
    # Fix category order for the legend
    pid_cat = factor(
      pid_cat,
      levels = c(
        "Close to a party",
        "No attachment",
        "Don't know",
        "Refused or no answer"
      )
    )
  ) |>
  select(module, study, country, year, pid_cat, lr)

# Respondents, studies and countries per module
cses_pid |>
  summarise(
    respondents = n(),
    studies = n_distinct(study),
    countries = n_distinct(country),
    .by = module
  ) |>
  arrange(module)
```

# Plot Share of Respondents reporting PID (incl. other), denying PID and Don't Know

Unlike the EES, the CSES has no fixed fieldwork years: each study follows a national election,
so the 25 countries contribute 109 country-years scattered between 1996 and 2024. The pooled
figure therefore runs over the six CSES modules, each of which is a defined five-year window,
while the country panels keep the election year on the x-axis.

Three country-years are fielded as two studies each and are pooled into one point: Belgium 2019
(Flanders and Wallonia), Germany 2002, and the two Greek elections of 2015. Belgium 1999 is
not pooled, because the Walloon half of it never fielded the item.

```{r}
#| label: shares

# Pooled share within each module
shares_pooled <- cses_pid |>
  count(module, pid_cat) |>
  group_by(module) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Share within each country-year
shares_country <- cses_pid |>
  count(country, year, pid_cat) |>
  group_by(country, year) |>
  mutate(share = n / sum(n)) |>
  ungroup()

shares_pooled
```

Six things are worth keeping in view when reading the figures.

First, **close to a party is not the same as naming a party**. The CSES item is binary; the
party itself is asked at a separate variable (`IMD3005_3`). A respondent can be close to a
party here without ever naming one, so this series is not level-comparable with the EES series
in notebook 1.2 -- only its movement is.

Second, **the leaner probe is excluded**, and it is large. Respondents who say no to the base
item but yes to the follow-up probe are between 15% and 25% of all respondents depending on
the module. The *No attachment* band therefore contains a substantial layer of leaners, and
its level should not be read as the share with no partisan leaning whatsoever.

Third, **code `9` conflates missing with not asked**. The *Refused or no answer* band mixes
true item nonresponse with study-level gaps in administration, and only studies in which every
single case is `9` are dropped.

Fourth, **modules 1 and 2 used deviating wording**, asking about any particular *political*
party. That deviation sits exactly under the first two points of the pooled line.

Fifth, **CSES folds inconsistent answers into the binary** rather than dropping them: a "yes"
with no party identified is coded `1` and a "no" with a party identified is coded `0`, and
neither is corrected here.

Sixth, **the panel is unbalanced and the unit is an election, not a year**. Between 8 and 19 of
the 25 countries appear per module; Estonia and Croatia contribute a single study each, while
Czechia, Germany, Poland and Portugal contribute seven. The faceted figure exists to make this
visible.

```{r}
#| label: plot-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled trend, four categories
ggplot(shares_pooled, aes(x = module, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    breaks = 1:6,
    labels = c(
      "1\n1996-2001",
      "2\n2001-2006",
      "3\n2006-2011",
      "4\n2011-2016",
      "5\n2016-2021",
      "6\n2021-2026"
    )
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the CSES, modules 1-6",
    subtitle = "Unweighted shares, pooled over 25 European democracies",
    x = "CSES module",
    y = "Share of respondents",
    colour = NULL,
    caption = "The branching leaner probe is not used."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-country
#| fig-width: 11
#| fig-height: 11

# Same trend, one panel per country
ggplot(shares_country, aes(x = year, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5) +
  scale_x_continuous(breaks = c(2000, 2010, 2020)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the CSES by country, 1996-2024",
    subtitle = "Unweighted shares; each point is one national election study",
    x = NULL,
    y = "Share of respondents",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

# Plot Spread of Left-Right Self-Placement among Partisans and Deniers

The figures above count people. This one measures how spread out they are. The question is
whether respondents who say they feel close to a party occupy a wider or a narrower band of
the left--right scale than respondents who say they do not, and whether that difference moves
across the six modules.

Two things the figure is not. It is not the distance *between* the groups: two groups can
share a mean and differ only in spread, which is roughly what happens here. And a standard
deviation cannot tell a wide unimodal distribution from a bimodal one -- two distributions
that look nothing alike can share an SD -- so nothing below speaks to sorting, or to
polarisation in any of its senses. It is one descriptive cut.

Don't-know and refusal are dropped throughout this section; the contrast is between the two
substantive answers only.

Before the spread, the coverage, because it is the constraint everything else here sits
inside.

```{r}
#| label: lr-coverage

# Who answers the placement item at all, by module and attachment
lr_coverage <- cses_pid |>
  filter(pid_cat %in% c("Close to a party", "No attachment")) |>
  summarise(
    respondents = n(),
    placed = mean(!is.na(lr)),
    .by = c(module, pid_cat)
  ) |>
  arrange(module, pid_cat)

lr_coverage
```

Between 91.6% and 95.3% of respondents close to a party place themselves on the scale,
against 75.9% to 85.8% of those who are not -- a gap of ten to sixteen points, present in all
six modules. The standard deviations below are therefore computed on a self-selected subset,
and the selection is *different in the two groups*: the deniers who do answer are the more
engaged end of that group, which is exactly the end whose placements should look most
partisan-like. The direction of the resulting bias cannot be signed from the data. Dropping
the least engaged could narrow the denier spread, if what goes is arbitrary mid-scale
answers, or widen it, if what goes is default fives. The table is printed rather than hidden
so that the size of the problem is visible.

No study has to be dropped for this section. The two studies that never fielded the
attachment item are already gone from `cses_pid`; of those remaining, every one carries usable
placements, so the exclusion set here is exactly the item-level nonresponse above and nothing
more.

```{r}
#| label: lr-select

# Partisans and deniers with a usable placement
cses_lr <- cses_pid |>
  filter(pid_cat %in% c("Close to a party", "No attachment"), !is.na(lr)) |>
  mutate(
    group = factor(
      if_else(pid_cat == "Close to a party", "Partisans", "Deniers"),
      levels = c("Partisans", "Deniers")
    ),
    # Keep all 25 panels in the faceted figure
    country = factor(country, levels = sort(unique(cses_pid$country)))
  )
```

The pooled figure takes the standard deviation over all respondents of a group in a module,
pooling respondents the same way the share figure above does, and the faceted figure takes it
within each country-year -- which in the CSES means within each election study. Cells of a
single respondent are dropped because `sd()` is undefined there, and the ten thinnest are
printed so their size is visible. That table is what this section uses in place of confidence
bands: at these sample sizes the sampling error of a standard deviation is small enough that
the movement worth arguing about is compositional, not statistical.

```{r}
#| label: lr-sd

# Pooled spread within each module
sd_pooled <- cses_lr |>
  summarise(
    respondents = n(),
    studies = n_distinct(study),
    countries = n_distinct(country),
    sd_lr = sd(lr),
    .by = c(module, group)
  ) |>
  arrange(module, group)

# Spread within each country-year
sd_country <- cses_lr |>
  summarise(
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(country, year, group)
  ) |>
  # sd() is undefined below two respondents
  filter(respondents >= 2)

sd_pooled

# Thinnest cells behind the faceted figure
sd_country |>
  slice_min(respondents, n = 10)
```

The pooled statistic is the one vulnerable to the unbalanced panel, and the CSES panel is the
most unbalanced of the three sources: 14 countries in module 1, rising to 19 in modules 3 and
5, and only 8 in the advance release of module 6. The balanced check below is correspondingly
weak -- exactly three of the twenty-five countries appear in all six modules -- so it is
printed as a direction of travel rather than as a test.

```{r}
#| label: lr-balanced

# Countries appearing in every module
balanced <- cses_lr |>
  summarise(modules = n_distinct(module), .by = country) |>
  filter(modules == n_distinct(cses_lr$module)) |>
  pull(country)

balanced

# Same pooled statistic over that fixed set
cses_lr |>
  filter(country %in% balanced) |>
  summarise(
    countries = n_distinct(country),
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(module, group)
  ) |>
  arrange(module, group)
```

Four things are worth keeping in view when reading the figures.

First, **the deniers here include the leaners**. The branching probe is not used in this
notebook, so the 15% to 25% of respondents who say they feel a little closer to one party
than the others sit inside *No attachment*. Those are people with a real if weak partisan
leaning, counted on the denier side. The gap between the two lines below is therefore
narrower than the same gap in notebooks 1.2 and 1.4 would be for that reason alone, and the
three are not comparable in level.

Second, **placements are measured just after an election**. Every CSES study follows a
national election, so the scale is answered when left--right is at its most crystallised and
most campaign-primed. The pooled x-axis is the module window rather than a year, so a single
point averages elections up to five years apart as well as countries, and how polarising the
particular elections in a window happened to be is inside the point.

Third, **the groups are not fixed populations**. The share close to a party moves across the
modules -- that is what the figure above documents -- so the marginal respondent crosses
between the two groups over time, and a change in a group's spread can be compositional
rather than attitudinal.

Fourth, **pooling mixes two variances**. The pooled standard deviation contains both the
spread within countries and the differences between countries in mean position, and the
country set changes every module. The faceted figure below is the check on that; it is not a
correction.

```{r}
#| label: plot-lr-pooled
#| fig-width: 8
#| fig-height: 5

# Pooled spread, two groups
ggplot(sd_pooled, aes(x = module, y = sd_lr, colour = group)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    breaks = 1:6,
    labels = c(
      "1\n1996-2001",
      "2\n2001-2006",
      "3\n2006-2011",
      "4\n2011-2016",
      "5\n2016-2021",
      "6\n2021-2026"
    )
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Spread of left-right self-placement in the CSES, modules 1-6",
    subtitle = "Unweighted standard deviation on a 0-10 scale, pooled over 25 democracies",
    x = "CSES module",
    y = "Standard deviation of self-placement",
    colour = NULL,
    caption = "Deniers include the leaners; non-placers are excluded."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-lr-country
#| fig-width: 11
#| fig-height: 11

# Same spread, one panel per country
ggplot(sd_country, aes(x = year, y = sd_lr, colour = group)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5, drop = FALSE) +
  scale_x_continuous(breaks = c(2000, 2010, 2020)) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Spread of left-right self-placement in the CSES by country, 1996-2024",
    subtitle = "Unweighted standard deviations; each point is one national election study",
    x = NULL,
    y = "Standard deviation of self-placement",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

The two series do not cross. Respondents close to a party are the more dispersed group in
every module, at a standard deviation of 2.63 against the deniers' 2.03 in module 1 and 3.02
against 2.19 in module 6. Both series rise across the six modules and the partisan one rises
faster, so the gap widens from 0.60 to 0.83 scale points. The three-country balanced check
moves the same way, and the thinnest country-year cell in the series holds 78 respondents, so
neither the trend nor the gap is an artefact of a handful of small cells -- but module 6 is an
advance release covering eight countries, and its point should not be read as the others are.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
