---
title: |
  CSAP Code Notebook 1.4
subtitle: |
  ESS rounds 1-11: reported party attachment over time
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "arrow",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Load ESS Data

The source is the partitioned Parquet dataset written by notebook 1.1, not the 1.5 GB CSV it
was built from. The whole point of that conversion is that a reader fetches only the columns
it needs, so nothing here parses the extract again.

```{r}
#| label: load-ess-data

# Partitioned Parquet, never the CSV
ess <- open_dataset(
  here(
    "data",
    "01_raw",
    "ess",
    "parquet"
  ),
  partitioning = hive_partition(cntry = utf8(), essround = int64())
)

ess
```

# Select PID and other key Variables

The attachment item is `clsprty` -- *is there a particular political party you feel closer to
than all the other parties?* -- a core item asked of everyone, with wording and coding constant
across all eleven rounds: `1` yes, `2` no, `7` refusal, `8` don't know, `9` no answer. Unlike
the EES and CSES files it carries no `NA` at all, and no country-round cell is missing it, so
nothing has to be dropped.

The `name` column is kept because it is the only thing separating round 10's two stacked
samples, which the caveats below depend on.

`lrscale` -- *in politics people sometimes talk of "left" and "right"; where would you place
yourself on this scale?* -- is pulled in the same step, for the final section of this
notebook. It is a 0--10 self-placement asked in every round, with `77` refusal, `88` don't
know and `99` no answer. Nothing between here and that section uses it.

```{r}
#| label: select

# Five columns of 430,581 rows
ess_pid <- ess |>
  select(cntry, essround, name, clsprty, lrscale) |>
  collect() |>
  mutate(
    round = essround,
    country = cntry,
    # 1 yes, 2 no, 8 dk, 7/9 refused
    pid_cat = case_when(
      clsprty == 1 ~ "Close to a party",
      clsprty == 2 ~ "No attachment",
      clsprty == 8 ~ "Don't know",
      .default = "Refused or no answer"
    ),
    # Whitelisting the scale drops 77, 88 and 99 in one go
    lr = if_else(lrscale %in% 0:10, as.numeric(lrscale), NA_real_)
  ) |>
  select(round, country, name, pid_cat, lr)
```

# Join to Long Format

There is nothing to join: the ESS arrives as one long file, the extract already holds exactly
the project's 25 democracies, and country identifiers are already alpha-2. The country filter
below is therefore a no-op, kept as a guard and for symmetry with notebooks 1.2 and 1.3.

```{r}
#| label: bind

ess_pid <- ess_pid |>
  # Extract already holds the 25
  filter(
    country %in%
      c(
        "AT",
        "BE",
        "BG",
        "CZ",
        "DE",
        "DK",
        "EE",
        "ES",
        "FI",
        "FR",
        "GB",
        "GR",
        "HR",
        "HU",
        "IE",
        "IT",
        "LT",
        "LV",
        "NL",
        "PL",
        "PT",
        "RO",
        "SE",
        "SI",
        "SK"
      )
  ) |>
  mutate(
    # Fix category order for the legend
    pid_cat = factor(
      pid_cat,
      levels = c(
        "Close to a party",
        "No attachment",
        "Don't know",
        "Refused or no answer"
      )
    )
  )

# Respondents and countries per round
ess_pid |>
  summarise(
    respondents = n(),
    countries = n_distinct(country),
    .by = round
  ) |>
  arrange(round)
```

# Plot Share of Respondents reporting PID (incl. other), denying PID and Don't Know

The x-axis is the ESS round rather than the fieldwork year. Year is derivable but not usable
as an axis: it lives in `inwyr` for rounds 1--2, `inwyys` for rounds 3--9 and `inwds` for
rounds 10--11, and the 18,386 respondents of round 10's self-completion sample carry no
interview date at all. The modal fieldwork year is therefore shown only as a label under each
round.

```{r}
#| label: shares

# Pooled share within each round
shares_pooled <- ess_pid |>
  count(round, pid_cat) |>
  group_by(round) |>
  mutate(share = n / sum(n)) |>
  ungroup()

# Share within each country-round
shares_country <- ess_pid |>
  count(country, round, pid_cat) |>
  group_by(country, round) |>
  mutate(share = n / sum(n)) |>
  ungroup()

shares_pooled
```

Four things are worth keeping in view when reading the figures.

First, **round 10 mixes two survey modes**. It is two stacked samples sharing `essround == 10`:
the face-to-face `ESS10e03_3` in 17 countries and the self-completion `ESS10SCe03_2` in Austria,
Germany, Spain, Latvia, Poland and Sweden. The self-completion half returns 51.2% close to a
party against the face-to-face half's 40.5%, and offers no refusal or don't-know code at all --
zero of each, against 703 in the face-to-face half -- which is why the round-10 don't-know share
is the lowest in the series. Because the two samples cover disjoint sets of countries, mode is
fully confounded with country composition there, and the gap cannot be attributed to mode alone.

Second, **`clsprty` is a filter item, not a party name**. The party actually named is recorded
separately in the country-specific `prtcl<cc>` family. As in notebook 1.3, this series is
therefore not level-comparable with the EES series in notebook 1.2 -- only its movement is.

Third, **the refusal band is not comparable across the three notebooks**. The ESS uses codes
`7` and `9` sparingly, 0.2% to 1.5% per round, against 4% to 7% in the EES. That difference is
coding practice, not respondents.

Fourth, **the panel is unbalanced**. Between 18 and 23 of the 25 countries appear per round.
Thirteen appear in all eleven rounds; Romania appears in one (round 4), Latvia in four and
Croatia in five. The faceted figure exists to make this visible.

```{r}
#| label: plot-pooled
#| fig-width: 9
#| fig-height: 5

# Pooled trend, four categories
ggplot(shares_pooled, aes(x = round, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    breaks = 1:11,
    labels = c(
      "1\n2002",
      "2\n2004",
      "3\n2006",
      "4\n2008",
      "5\n2011",
      "6\n2012",
      "7\n2014",
      "8\n2016",
      "9\n2018",
      "10\n2021",
      "11\n2023"
    )
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the ESS, rounds 1-11",
    subtitle = "Unweighted shares, pooled over 25 European democracies",
    x = "ESS round and modal fieldwork year",
    y = "Share of respondents",
    colour = NULL,
    caption = "Round 10 pools a face-to-face and a self-completion sample."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-country
#| fig-width: 11
#| fig-height: 11

# Same trend, one panel per country
ggplot(shares_country, aes(x = round, y = share, colour = pid_cat)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5) +
  scale_x_continuous(breaks = c(1, 6, 11)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Reported party attachment in the ESS by country, rounds 1-11",
    subtitle = "Unweighted shares; gaps mark rounds in which a country was not surveyed",
    x = "ESS round",
    y = "Share of respondents",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

Read next to the other two notebooks, the ESS series declines from 50.9% to 43.1% close to a
party, moving in the same direction as the CSES series in notebook 1.3 (44.5% to 40.4%) and
against the EES series in notebook 1.2 (55.9% to 66.3%). Two of the three sources agree; the
EES is the one that does not, and its late waves are also the ones carrying the mode break
described there.

# Plot Spread of Left-Right Self-Placement among Partisans and Deniers

The figures above count people. This one measures how spread out they are. The question is
whether respondents who say they feel close to a party occupy a wider or a narrower band of
the left--right scale than respondents who say they do not, and whether that difference moves
across the eleven rounds.

Two things the figure is not. It is not the distance *between* the groups: two groups can
share a mean and differ only in spread, which is roughly what happens here. And a standard
deviation cannot tell a wide unimodal distribution from a bimodal one -- two distributions
that look nothing alike can share an SD -- so nothing below speaks to sorting, or to
polarisation in any of its senses. It is one descriptive cut.

Don't-know and refusal are dropped throughout this section; the contrast is between the two
substantive answers only.

Before the spread, the coverage, because it is the constraint everything else here sits
inside.

```{r}
#| label: lr-coverage

# Who answers the placement item at all, by round and attachment
lr_coverage <- ess_pid |>
  filter(pid_cat %in% c("Close to a party", "No attachment")) |>
  summarise(
    respondents = n(),
    placed = mean(!is.na(lr)),
    .by = c(round, pid_cat)
  ) |>
  arrange(round, pid_cat)

# Twenty-two rows, so past a tibble's default print
print(lr_coverage, n = Inf)

# Round 10 splits by sample as well as by attachment
ess_pid |>
  filter(round == 10, pid_cat %in% c("Close to a party", "No attachment")) |>
  summarise(
    respondents = n(),
    placed = mean(!is.na(lr)),
    .by = c(name, pid_cat)
  ) |>
  arrange(name, pid_cat)
```

Between 78% and 85% of deniers place themselves in a given round, against 94% to 96% of
partisans -- a gap of roughly fifteen points, present in every one of the eleven rounds. The
standard deviations below are therefore computed on a self-selected subset, and the selection
is *different in the two groups*: the deniers who do answer are the more engaged end of that
group, which is exactly the end whose placements should look most partisan-like. The
direction of the resulting bias cannot be signed from the data. Dropping the least engaged
could narrow the denier spread, if what goes is arbitrary mid-scale answers, or widen it, if
what goes is default fives. The table is printed rather than hidden so that the size of the
problem is visible.

The second table splits round 10 by sample. Both of its stacked samples carry `lrscale`, so
neither has to be set aside: coverage runs 93.1% among deniers and 97.2% among partisans in
the self-completion `ESS10SCe03_2`, against 81.5% and 95.1% in the face-to-face
`ESS10e03_3`. That is a diagnostic, not a mode effect -- the two samples cover disjoint sets
of countries, so mode and composition cannot be separated here any more than they could for
the attachment shares above.

```{r}
#| label: lr-select

# Partisans and deniers with a usable placement
ess_lr <- ess_pid |>
  filter(pid_cat %in% c("Close to a party", "No attachment"), !is.na(lr)) |>
  mutate(
    group = factor(
      if_else(pid_cat == "Close to a party", "Partisans", "Deniers"),
      levels = c("Partisans", "Deniers")
    ),
    # Keep all 25 panels in the faceted figure
    country = factor(country, levels = sort(unique(ess_pid$country)))
  )
```

The pooled figure takes the standard deviation over all respondents of a group in a round,
pooling respondents the same way the share figures above do. The faceted figure takes it
within each country-round. Cells of a single respondent are dropped because `sd()` is
undefined there; in practice none come close to that, and the ten thinnest are printed so
their size is visible. That table is what this section uses in place of confidence bands: at
these sample sizes the sampling error of a standard deviation is small enough that the
movement worth arguing about is compositional, not statistical.

```{r}
#| label: lr-sd

# Pooled spread within each round
sd_pooled <- ess_lr |>
  summarise(
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(round, group)
  ) |>
  arrange(round, group)

# Spread within each country-round
sd_country <- ess_lr |>
  summarise(
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(country, round, group)
  ) |>
  # sd() is undefined below two respondents
  filter(respondents >= 2)

print(sd_pooled, n = Inf)

# Thinnest cells behind the faceted figure
sd_country |>
  slice_min(respondents, n = 10)
```

The pooled statistic is the one vulnerable to the unbalanced panel, so it is recomputed below
over only the countries surveyed in every round. If the two versions move together, entry and
exit are not driving the trend; if they diverge, composition is.

```{r}
#| label: lr-balanced

# Countries surveyed in every round
balanced <- ess_lr |>
  summarise(rounds = n_distinct(round), .by = country) |>
  filter(rounds == n_distinct(ess_lr$round)) |>
  pull(country)

# Same pooled statistic over that fixed set
ess_lr |>
  filter(country %in% balanced) |>
  summarise(
    countries = n_distinct(country),
    respondents = n(),
    sd_lr = sd(lr),
    .by = c(round, group)
  ) |>
  arrange(round, group) |>
  print(n = Inf)
```

Three things are worth keeping in view when reading the figures.

First, **the groups are not fixed populations**. The share close to a party moves across the
rounds -- that is precisely what the figures above document -- so the marginal respondent
crosses between the two groups over time. A change in a group's spread can therefore be
compositional rather than attitudinal, and a standard deviation cannot tell the two readings
apart.

Second, **pooling mixes two variances**. The pooled standard deviation contains both the
spread within countries and the differences between countries in mean position, and the
country set is not fixed. The balanced-panel table above and the faceted figure below are the
two checks on that; neither is a correction.

Third, **the nonresponse gap is not an ESS peculiarity**. Measured on the same base, the
three sources land in the same place: 72% to 86% of deniers place themselves against 92% to
96% of partisans, in the EES of notebook 1.2 and the CSES of notebook 1.3 as much as here.
The gap is a property of the item, not of any one survey, so it is not something a change of
source would fix.

```{r}
#| label: plot-lr-pooled
#| fig-width: 9
#| fig-height: 5

# Pooled spread, two groups
ggplot(sd_pooled, aes(x = round, y = sd_lr, colour = group)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.2) +
  scale_x_continuous(
    breaks = 1:11,
    labels = c(
      "1\n2002",
      "2\n2004",
      "3\n2006",
      "4\n2008",
      "5\n2011",
      "6\n2012",
      "7\n2014",
      "8\n2016",
      "9\n2018",
      "10\n2021",
      "11\n2023"
    )
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Spread of left-right self-placement in the ESS, rounds 1-11",
    subtitle = "Unweighted standard deviation on a 0-10 scale, pooled over 25 democracies",
    x = "ESS round and modal fieldwork year",
    y = "Standard deviation of self-placement",
    colour = NULL,
    caption = "Respondents who did not place themselves are excluded."
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

```{r}
#| label: plot-lr-country
#| fig-width: 11
#| fig-height: 11

# Same spread, one panel per country
ggplot(sd_country, aes(x = round, y = sd_lr, colour = group)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.1) +
  facet_wrap(vars(country), ncol = 5, drop = FALSE) +
  scale_x_continuous(breaks = c(1, 6, 11)) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    title = "Spread of left-right self-placement in the ESS by country, rounds 1-11",
    subtitle = "Unweighted standard deviations; gaps mark rounds without a usable cell",
    x = "ESS round",
    y = "Standard deviation of self-placement",
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

The two series do not cross. Partisans are the more dispersed group in every round, at a
standard deviation of 2.39 against the deniers' 1.80 in round 1, and 2.59 against 1.84 in
round 11. Both drift upward across the eleven rounds, but not equally: the partisan series
gains 0.20 scale points over the period against the denier series' 0.04, so the gap between
them widens from 0.59 to 0.75. Both peak in round 10 -- the round whose two stacked samples
cover disjoint country sets, so that peak is the one point in the series that should not be
read on its own. Over the thirteen countries surveyed in every round the same two series run
2.39 to 2.59 and 1.79 to 1.84, within 0.05 of the full ones throughout, so entry and exit are
not what the pooled lines are showing.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
