---
title: |
  CSAP Code Notebook 1.9
subtitle: |
  Electoral alliances and the attachment-vote match
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

Party Facts keeps an electoral alliance, a joint list or a party union apart from its member parties: the CDU/CSU union (211) is not the CDU (1375) or the CSU (1731), the Fidesz–KDNP list (6366) is not Fidesz (1691), Koalicja Obywatelska (9122) is not Platforma Obywatelska (1117).
Comparing Party Facts names therefore scores a respondent who feels close to the CDU and votes for the CDU/CSU list as a *different* party.

This notebook lists the member parties of every alliance that occurs in the four sources (eu25games, ESS, CSES, EES) and uses the list to compare attachment and vote:

- **same**: the same party, or a party and an alliance that contains it, in either direction (close to the CSU, voted CDU/CSU; close to KO, voted PO);
- **different**: two different parties, including two members of the same alliance (CDU and CSU);
- **unmatchable**: either answer missing or a non-party answer (`none`, `other`, `independent`, ...).

The rule is implemented once, in `pf_match()` in `code/00_helper/pf_code.R`, and used here and in the composition notebooks 2.1–2.4.

Outputs, in `data/02_processed/`:

- `pf_alliances.csv`: one row per alliance and member (`cntry`, `alliance_name`, `member_name`, `member_id`), with names as they appear in the `ext_*_pf_name` variables (a Party Facts core name, or `"<label> [og]"`);
- `ess_party_match.parquet`, `cses_party_match.parquet`, `ees_party_match.parquet`: the respondent keys of notebooks 1.2–1.4 with one `ext_match_<vote item>` column per vote item, comparing it with the attachment item.
  The eu25games match is computed in notebook 2.1, which first resolves the release's open "other" answers.

The table is built from the crosswalks of notebooks 1.2–1.4, so this notebook runs after them; they do not depend on it.

# Setup

```{r}
#| label: setup

start_time <- Sys.time()
options(width = 80)

p_required <- c(
  "tidyverse",
  "here",
  "arrow",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

source(here("code", "00_helper", "pid_tree.R")) # countries_25
source(here("code", "00_helper", "pf_code.R"))

core <- pf_core()
```

# Party Answers in the Four Sources

```{r}
#| label: load

processed <- \(f) read_parquet(here("data", "02_processed", f))
ess <- processed("ess_party_pf.parquet")
cses <- processed("cses_party_pf.parquet")
ees <- processed("ees_party_pf.parquet")

eu25_iso2 <- c(
  Austria = "AT", Belgium = "BE", Bulgaria = "BG", Croatia = "HR",
  Czechia = "CZ", Denmark = "DK", Estonia = "EE", Finland = "FI",
  France = "FR", Germany = "DE", Greece = "GR", Hungary = "HU",
  Ireland = "IE", Italy = "IT", Latvia = "LV", Lithuania = "LT",
  Netherlands = "NL", Poland = "PL", Portugal = "PT", Romania = "RO",
  Slovakia = "SK", Slovenia = "SI", Spain = "ES", Sweden = "SE",
  `United Kingdom` = "GB"
)
eu25 <- read_rds(here("data", "01_raw", "eu25games2019.rds")) |>
  filter(der_att_check_3) |>
  transmute(
    cntry = eu25_iso2[meta_country],
    att = coalesce(ext_q_party_id_pf_name, ext_q_party_id_fra_pf_name),
    vote = coalesce(
      ext_q_vote_choice_intended_pf_name,
      ext_q_vote_choice_hypo_pf_name,
      ext_q_vote_choice_past_pf_name,
      ext_q_vote_choice_hypo_fra_pf_name,
      ext_q_vote_choice_past_fra_pf_name
    )
  )

# Attachment and vote pairs, one frame
pairs <- bind_rows(
  ess |>
    transmute(source = "ESS", cntry, att = ext_prtcl_pf_name, vote = coalesce(ext_prtv_pf_name, ext_prtv_de1_pf_name)),
  cses |>
    transmute(
      source = "CSES",
      cntry,
      att = ext_imd3005_3_pf_name,
      vote = coalesce(ext_imd3002_lh_pl_pf_name, ext_imd3002_lh_dc_pf_name)
    ),
  ees |>
    transmute(source = "EES", cntry, att = ext_pid_pf_name, vote = ext_vote_ep_pf_name),
  ees |>
    transmute(source = "EES", cntry, att = ext_pid_pf_name, vote = ext_vote_nat_pf_name),
  eu25 |>
    mutate(source = "eu25games")
) |>
  filter(!is.na(att), !is.na(vote), !att %in% pf_sentinels, !vote %in% pf_sentinels)
```

# Candidates

An alliance is a candidate for the table if its name reads as a joint list (a slash, a dash or a plus between names, or words such as *coalition*, *koalicja*, *alianța*, *blok*, *lista*, *spolu*), or if it is one side of an attachment/vote pair that disagrees at least five times in any source.
The candidates were classified by hand below; the list is printed so that a new alliance shows up when the sources change.

```{r}
#| label: candidates

disagree <- pairs |>
  filter(att != vote) |>
  count(cntry, att, vote, name = "n") |>
  filter(n >= 5)

joint_list <- regex(
  "/| — | – | - |\\+|&|coali|koali|alian|allian|bloc\\b|blok|lista |list\\b|spolu|together|zusammen|ensemble|nupes|union of|unión|unidas|uniti|ulivo|olive|united|frente|front de|sojusz|egység|összefog|, ",
  ignore_case = TRUE
)

entities <- pairs |>
  pivot_longer(c(att, vote), names_to = "item", values_to = "name") |>
  count(cntry, name, name = "n")

candidates <- entities |>
  mutate(
    by_name = str_detect(name, joint_list),
    in_pair = paste(cntry, name) %in% c(paste(disagree$cntry, disagree$att), paste(disagree$cntry, disagree$vote))
  ) |>
  filter(by_name | in_pair)

count(candidates, by_name, in_pair)
```

# Alliance Table

One row per alliance and member.
Members are named as in the data, so a member that appears under two names (a Party Facts name and an `[og]` spelling) has two rows; `member_id` is its Party Facts id where it has one.
Renamed parties are not alliances and are not listed.

```{r}
#| label: alliances

alliances <- tribble(
  ~cntry, ~alliance_name, ~member_name, ~member_id,
  "AT", "Communist Party of Austria and Platform PLUS - Open List [og]", "Kommunistische Partei Österreichs", 996L,
  "AT", "EUROPA Jetzt – Initiative Johannes Voggenhuber[og]", "Liste Peter Pilz", 6137L, # 2019 EP list backed by JETZT (Liste Pilz)
  "AT", "KPÖ Plus [og]", "Kommunistische Partei Österreichs", 996L,
  "AT", "KPÖ Plus – European Left, offene Liste[og]", "Kommunistische Partei Österreichs", 996L,
  "BE", "Christen-Democratisch & Vlaams / Nieuw-Vlaams Alliantie", "Christen-Democratisch en Vlaams", 604L, # CD&V/N-VA cartel 2004-08
  "BE", "Christen-Democratisch & Vlaams / Nieuw-Vlaams Alliantie", "Nieuw-Vlaamse Alliantie", 36L, # CD&V/N-VA cartel 2004-08
  "BE", "Open VLD + Vivant [og]", "Partij voor Vrijheid en Vooruitgang — Open Vlaamse Liberalen en Democraten", 49L,
  "BE", "Open VLD + Vivant [og]", "Vivant", 76L,
  "BE", "Socialistische Partij Anders / Sociaal-Liberale Partij / Spirit", "Sociaal-Liberale Partij", 7015L, # sp.a-Spirit cartel; Spirit later SLP
  "BE", "Socialistische Partij Anders / Sociaal-Liberale Partij / Spirit", "Socialistische Partij Anders — Vooruit", 1680L, # sp.a-Spirit cartel; Spirit later SLP
  "BG", "Balgarski Naroden Sajuz", "VMRO — Bulgarsko Natsionalno Dvizhenie", 2056L, # 2005 BZNS-NS+SSD+VMRO
  "BG", "Balgarski Naroden Sajuz", "Zemedelski Naroden Sǔyuz", 747L, # 2005 BZNS-NS+SSD+VMRO
  "BG", "Demokratichna Bulgaria", "Da Bulgaria", 5861L,
  "BG", "Demokratichna Bulgaria", "Demokrati za Silna Balgarija", 1195L,
  "BG", "Grazhdani za Evropeisko Razvitie Balgariya - Sayuz na demokratichnite sili (GERB- SDS) [og]", "Grazhdani za evropeysko razvitie na Balgariya", 760L,
  "BG", "Grazhdani za Evropeisko Razvitie Balgariya - Sayuz na demokratichnite sili (GERB- SDS) [og]", "Sayuz na Demokratichni Sili", 482L,
  "BG", "Grazhdani za evropeysko razvitie na Balgariya – Sayuz na demokratichnite sili (GERB - SDS) [og]", "Grazhdani za evropeysko razvitie na Balgariya", 760L,
  "BG", "Grazhdani za evropeysko razvitie na Balgariya – Sayuz na demokratichnite sili (GERB - SDS) [og]", "Sayuz na Demokratichni Sili", 482L,
  "BG", "Koalicija za Balgarija — BSP za Bŭlgariya", "Bălgarska Socialističeska Partija", 757L,
  "BG", "Koalitsiya ABV – Dvizhenie 21 [og]", "Alternativa za balgarsko vazrazhdane", 3187L,
  "BG", "Nova republika - DSB - Suyuz za Plovdiv - BDO [og]", "Demokrati za Silna Balgarija", 1195L,
  "BG", "Obedineni Demokratichni Sili", "Sayuz na Demokratichni Sili", 482L,
  "BG", "Obedineni Patrioti", "Natsionalen Sǎyuz Ataka", 1793L,
  "BG", "Obedineni Patrioti", "Natzionalen Front za Spasenie na Bulgaria", 2057L,
  "BG", "Obedineni Patrioti", "VMRO — Bulgarsko Natsionalno Dvizhenie", 2056L,
  "BG", "Prodalzhavame Promyanata – Demokratichna Balgariya (PP-DB) [og]", "Da Bulgaria", 5861L, # DB components included
  "BG", "Prodalzhavame Promyanata – Demokratichna Balgariya (PP-DB) [og]", "Demokrati za Silna Balgarija", 1195L, # DB components included
  "BG", "Prodalzhavame Promyanata – Demokratichna Balgariya (PP-DB) [og]", "Demokratichna Bulgaria", 9083L, # DB components included
  "BG", "Prodalzhavame Promyanata – Demokratichna Balgariya (PP-DB) [og]", "Prodalzhavame promyanata", 8994L, # DB components included
  "BG", "Prodalzhavame promyanata - Demokratichna Balgariya [og]", "Da Bulgaria", 5861L, # DB components included
  "BG", "Prodalzhavame promyanata - Demokratichna Balgariya [og]", "Demokrati za Silna Balgarija", 1195L, # DB components included
  "BG", "Prodalzhavame promyanata - Demokratichna Balgariya [og]", "Demokratichna Bulgaria", 9083L, # DB components included
  "BG", "Prodalzhavame promyanata - Demokratichna Balgariya [og]", "Prodalzhavame promyanata", 8994L, # DB components included
  "BG", "Reformatorski Blok", "Demokrati za Silna Balgarija", 1195L,
  "BG", "Reformatorski Blok", "Sayuz na Demokratichni Sili", 482L,
  "BG", "Sinyata koalitsia", "Demokrati za Silna Balgarija", 1195L,
  "BG", "Sinyata koalitsia", "Sayuz na Demokratichni Sili", 482L,
  "BG", "Демократична България (Да България! + Зелените + Демократи за силна България (ДСБ)) [og]", "Da Bulgaria", 5861L,
  "BG", "Демократична България (Да България! + Зелените + Демократи за силна България (ДСБ)) [og]", "Demokrati za Silna Balgarija", 1195L,
  "CZ", "Koalice PirSTAN (Piráti, STAN) [og]", "Ceska piratska strana", 2047L,
  "CZ", "Koalice PirSTAN (Piráti, STAN) [og]", "Starostové a nezávislí", 1707L,
  "CZ", "Koalice SPD + Trikolora [og]", "Svoboda a přímá demokracie", 6125L,
  "CZ", "Koalice Stačilo (KSČM+ČSNS, SD-SN) [og]", "Komunistická strana Cech a Moravy", 1728L, # ČSNS
  "CZ", "Koalice Stačilo (KSČM+ČSNS, SD-SN) [og]", "[ Liberální / Česká ] strana národně sociální", 1813L, # ČSNS
  "CZ", "Pirates and Mayors [og]", "Ceska piratska strana", 2047L, # = PirSTAN
  "CZ", "Pirates and Mayors [og]", "Starostové a nezávislí", 1707L, # = PirSTAN
  "CZ", "SPOLU (ODS+TOP 09+KDU-ČSL) [og]", "Křesťanská a demokratická unie / Československá strana lidová", 676L,
  "CZ", "SPOLU (ODS+TOP 09+KDU-ČSL) [og]", "Občanská demokratická strana", 466L,
  "CZ", "SPOLU (ODS+TOP 09+KDU-ČSL) [og]", "Tradice Odpovědnost Prosperita 09", 223L,
  "CZ", "Svobodní, Liberland a Radostné Česko - ODEJDEME BEZ PLACENÍ [og]", "Strana svobodných občanů", 254L,
  "CZ", "Together [og]", "Křesťanská a demokratická unie / Československá strana lidová", 676L, # = SPOLU
  "CZ", "Together [og]", "Občanská demokratická strana", 466L, # = SPOLU
  "CZ", "Together [og]", "Tradice Odpovědnost Prosperita 09", 223L, # = SPOLU
  "CZ", "ULD - Unie liberálních demokratů CZ (18), US-DEU (5), ODA (19) [only in v029 and v211 ] [og]", "Občanská demokratická aliance", 1008L, # per label: US-DEU + ODA
  "CZ", "ULD - Unie liberálních demokratů CZ (18), US-DEU (5), ODA (19) [only in v029 and v211 ] [og]", "Unie svobody", 905L, # per label: US-DEU + ODA
  "CZ", "ULD - Unie liberálních demokratů CZ (18), US-DEU (5), ODA (19) [only in v029 and v211 ] [og]", "Unie svobody / Demokratická unie", 104L, # per label: US-DEU + ODA
  "CZ", "Čtyřkoalice", "Křesťanská a demokratická unie / Československá strana lidová", 676L, # US and US-DEU both listed (separate PF ids)
  "CZ", "Čtyřkoalice", "Občanská demokratická aliance", 1008L, # US and US-DEU both listed (separate PF ids)
  "CZ", "Čtyřkoalice", "Unie svobody", 905L, # US and US-DEU both listed (separate PF ids)
  "CZ", "Čtyřkoalice", "Unie svobody / Demokratická unie", 104L, # US and US-DEU both listed (separate PF ids)
  "DE", "Christlich Demokratische Union / Christlich Soziale Union", "Christlich Demokratische Union", 1446L,
  "DE", "Christlich Demokratische Union / Christlich Soziale Union", "Christlich Soziale Union", 1731L,
  "ES", "A la valenciana", "Coalició Compromís", 5623L, # IU via EUPV
  "ES", "A la valenciana", "Izquierda Unida", 8128L, # IU via EUPV
  "ES", "A la valenciana", "Podemos", 3203L, # IU via EUPV
  "ES", "Ahora Repúblicas (Esquerra Republicana de Catalunya (ERC) & EH Bildu & Bloque Nationalista Galego (BNG)) [og]", "Bloque Nacionalista Galego", 1011L,
  "ES", "Ahora Repúblicas (Esquerra Republicana de Catalunya (ERC) & EH Bildu & Bloque Nationalista Galego (BNG)) [og]", "Esquerra Republicana de Catalunya", 848L,
  "ES", "Ahora Repúblicas (Esquerra Republicana de Catalunya (ERC) & EH Bildu & Bloque Nationalista Galego (BNG)) [og]", "Euskal Herria Bildu / Reunir Euskal Herria", 5668L,
  "ES", "Ahora Repúblicas [og]", "Bloque Nacionalista Galego", 1011L,
  "ES", "Ahora Repúblicas [og]", "Esquerra Republicana de Catalunya", 848L,
  "ES", "Ahora Repúblicas [og]", "Euskal Herria Bildu / Reunir Euskal Herria", 5668L,
  "ES", "CEU Coalicion por Europa [og]", "Coalición Canaria", 81L,
  "ES", "CEU Coalicion por Europa [og]", "Coalición Canaria [og]", NA,
  "ES", "CEU Coalicion por Europa [og]", "Convergència i Unió", 139L,
  "ES", "CEU Coalicion por Europa [og]", "Partido Nacionalista Vasco — Euzko Alderdi Jeltzalea", 1637L,
  "ES", "Coalición por una Europa Solidaria (CEUS (EAJ-PNV+CCa-PNC+CxG+El PI+DV+GBAI) [og]", "Coalición Canaria", 81L,
  "ES", "Coalición por una Europa Solidaria (CEUS (EAJ-PNV+CCa-PNC+CxG+El PI+DV+GBAI) [og]", "Coalición Canaria [og]", NA,
  "ES", "Coalición por una Europa Solidaria (CEUS (EAJ-PNV+CCa-PNC+CxG+El PI+DV+GBAI) [og]", "Partido Nacionalista Vasco — Euzko Alderdi Jeltzalea", 1637L,
  "ES", "Coalición por una Europa Solidaria [og]", "Coalición Canaria", 81L,
  "ES", "Coalición por una Europa Solidaria [og]", "Coalición Canaria [og]", NA,
  "ES", "Coalición por una Europa Solidaria [og]", "Partido Nacionalista Vasco — Euzko Alderdi Jeltzalea", 1637L,
  "ES", "Coalition for Europe (CiU (1724007) + EAJ-PNV (1724902) + CC (1724907) + others) [og]", "Coalición Canaria", 81L,
  "ES", "Coalition for Europe (CiU (1724007) + EAJ-PNV (1724902) + CC (1724907) + others) [og]", "Coalición Canaria [og]", NA,
  "ES", "Coalition for Europe (CiU (1724007) + EAJ-PNV (1724902) + CC (1724907) + others) [og]", "Convergència i Unió", 139L,
  "ES", "Coalition for Europe (CiU (1724007) + EAJ-PNV (1724902) + CC (1724907) + others) [og]", "Partido Nacionalista Vasco — Euzko Alderdi Jeltzalea", 1637L,
  "ES", "Compromís per Europa/Compromiso por Europa (CPE) [og]", "Coalició Compromís", 5623L,
  "ES", "IU (Izquierda Unida) - IC-V (Iniciativa per Catalunya - Verds) [og]", "Iniciativa per Catalunya Verds", 1260L,
  "ES", "IU (Izquierda Unida) - IC-V (Iniciativa per Catalunya - Verds) [og]", "Izquierda Unida", 8128L,
  "ES", "IU (Izquierda Unida) / IC-V (Iniciativa per Catalunya – Verds) [og]", "Iniciativa per Catalunya Verds", 1260L,
  "ES", "IU (Izquierda Unida) / IC-V (Iniciativa per Catalunya – Verds) [og]", "Izquierda Unida", 8128L,
  "ES", "Junts per Catalunya — Partit Demòcrata Europeu Català", "Partit Demòcrata Europeu Català", 6027L,
  "ES", "Left for The Right to Decide (Coalition ERC (1724905) + others) [og]", "Esquerra Republicana de Catalunya", 848L,
  "ES", "PP (Partido Popular) / UPN (Unión del Pueblo Navarro) [og]", "Alianza-Partido Popular / Partido Popular", 441L,
  "ES", "SUMAR [og]", "Coalició Compromís", 5623L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "SUMAR [og]", "Izquierda Unida", 8128L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "SUMAR [og]", "Más País", 8395L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "SUMAR [og]", "Podemos", 3203L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "Sumar [og]", "Coalició Compromís", 5623L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "Sumar [og]", "Izquierda Unida", 8128L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "Sumar [og]", "Más País", 8395L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "Sumar [og]", "Podemos", 3203L, # Podemos on list only in 2023 general, not 2024 EP
  "ES", "Unidas Podemos", "Izquierda Unida", 8128L,
  "ES", "Unidas Podemos", "Podemos", 3203L,
  "FR", "Alliance Rurale [og]", "Résistons !", 9876L, # 2024 EP list of Résistons! and LMR (ex-CPNT)
  "FR", "Ensemble ! (LREM, MODEM et Horizons) [og]", "Agir, la droite constructive", 9132L, # Agir in Ensemble 2022, not in label
  "FR", "Ensemble ! (LREM, MODEM et Horizons) [og]", "Horizons", 9134L,
  "FR", "Ensemble ! (LREM, MODEM et Horizons) [og]", "La République En Marche! / Renaissance", 5857L,
  "FR", "Ensemble ! (LREM, MODEM et Horizons) [og]", "Mouvement démocrate", 496L,
  "FR", "Ensemble - Renaissance, Mouvement Démocrate (MoDem) [og]", "Horizons", 9134L, # Horizons in Ensemble 2022/2024, not in label
  "FR", "Ensemble - Renaissance, Mouvement Démocrate (MoDem) [og]", "La République En Marche! / Renaissance", 5857L,
  "FR", "Ensemble - Renaissance, Mouvement Démocrate (MoDem) [og]", "Mouvement démocrate", 496L,
  "FR", "FN, MNR (Front national, Mouvement National Républicain) [og]", "Front National / Rassemblement national", 433L, # combined response code, not a joint list
  "FR", "FN, MNR (Front national, Mouvement National Républicain) [og]", "Mouvement national républicain", 1634L, # combined response code, not a joint list
  "FR", "Front de Gauche", "Parti communiste française", 1251L,
  "FR", "Front de Gauche", "Parti de Gauche", 8359L,
  "FR", "La liste Génération.s et Dème-Diem 25, menée par Benoît Hamon [og]", "Génération.s, le mouvement", 7342L,
  "FR", "Les Républicains et l’UDI [og]", "Les Républicains", 8168L,
  "FR", "Les Républicains et l’UDI [og]", "Union des démocrates et indépendants", 3229L,
  "FR", "MPF, RPF (Mouvement Pour la France / Rassemblement Pour la France [og]", "Mouvement pour la France", 364L, # combined response code
  "FR", "MPF, RPF (Mouvement Pour la France / Rassemblement Pour la France [og]", "RPF (Rassemblement du Peuple Français) [og]", NA, # label mis-expands Pasqua's Rassemblement pour la France
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", "Europe Écologie — Les Verts", 5650L,
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", "Génération.s, le mouvement", 7342L, # NUPES member, not in label
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", "La France Insoumise", 5858L,
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", "Les Verts", 1108L, # pre-2010 name of EELV
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", "Parti communiste française", 1251L,
  "FR", "NUPES (Nouvelle Union Populaire Ecologique et Sociale: FI, EELV, PCF, PS) [og]", "Parti socialiste", 1478L,
  "FR", "NUPES - La France Insoumise (LFI), Parti Socialiste (PS), Europe Écologie Les Verts (EELV) [og]", "Europe Écologie — Les Verts", 5650L,
  "FR", "NUPES - La France Insoumise (LFI), Parti Socialiste (PS), Europe Écologie Les Verts (EELV) [og]", "Génération.s, le mouvement", 7342L, # NUPES member, not in label
  "FR", "NUPES - La France Insoumise (LFI), Parti Socialiste (PS), Europe Écologie Les Verts (EELV) [og]", "La France Insoumise", 5858L,
  "FR", "NUPES - La France Insoumise (LFI), Parti Socialiste (PS), Europe Écologie Les Verts (EELV) [og]", "Les Verts", 1108L, # pre-2010 name of EELV
  "FR", "NUPES - La France Insoumise (LFI), Parti Socialiste (PS), Europe Écologie Les Verts (EELV) [og]", "Parti communiste française", 1251L, # NUPES member, not in label
  "FR", "NUPES - La France Insoumise (LFI), Parti Socialiste (PS), Europe Écologie Les Verts (EELV) [og]", "Parti socialiste", 1478L,
  "FR", "UMP + DVD [og]", "Union pour un mouvement populaire", 1595L,
  "FR", "UMP + DVD [og]", "divers droite", 1776L, # generic DVD category is literally part of the code
  "FR", "Union de la Droite et du Centre - Les Républicains et alliés [og]", "Les Républicains", 8168L,
  "FR", "Union de la Droite et du Centre - Les Républicains et alliés [og]", "Union des démocrates et indépendants", 3229L, # LR-UDI 2017 legislative alliance
  "FR", "l’Union pour la France (Debout la France, Les Patriotes et Génération Frexit) [og]", "Debout la France", 3207L,
  "GR", "Democratic Coalition [og]", "Dimokratiki Aristera", 1160L, # Dimokratiki Symparataxi PASOK-DIMAR
  "GR", "Democratic Coalition [og]", "Panellinio Sosialistikó Kínima", 1468L, # Dimokratiki Symparataxi PASOK-DIMAR
  "GR", "Front of the Greek Anticapitalist Left - Workers Revolutionary Party [og]", "Antikapitalistiki Aristeri Synergasia gia tin Anatropi", 95L, # ANTARSYA-EEK joint list
  "GR", "Kinima Allagis", "Kinima Dimokraton Sosialiston", 3212L, # KIDISO in KINAL from 2017
  "GR", "Kinima Allagis", "Panellinio Sosialistikó Kínima", 1468L, # KINAL electoral alliance 2017-2022
  "GR", "Olive Tree-Democratic Alignment (Coalition PASOK (1300313) + others) [og]", "Dimokratiki Aristera", 1160L, # Democratic Alignment 2015 = PASOK+DIMAR
  "GR", "Olive Tree-Democratic Alignment (Coalition PASOK (1300313) + others) [og]", "Panellinio Sosialistikó Kínima", 1468L,
  "GR", "Synaspismós Rizospastikís Aristerás", "Synaspismós tīs Aristerás", 516L, # SYN main component of SYRIZA coalition until 2013
  "GR", "Δημοκρατική Συμπαράταξη (ΔΗΣΥ) [og]", "Dimokratiki Aristera", 1160L,
  "GR", "Δημοκρατική Συμπαράταξη (ΔΗΣΥ) [og]", "Panellinio Sosialistikó Kínima", 1468L,
  "HR", "Amsterdam Coalition [og]", "Građansko-liberalni savez", 8178L,
  "HR", "Amsterdam Coalition [og]", "Istarski demokratski sabor", 513L,
  "HR", "BRUNA ESIH - ZLATKO HASANBEGOVIĆ: NEOVISNI ZA HRVATSKU + HRVATSKA STRANKA PRAV [og]", "Croatian Party of Rights [og]", NA, # label pools HSP and HSP-AS
  "HR", "BRUNA ESIH - ZLATKO HASANBEGOVIĆ: NEOVISNI ZA HRVATSKU + HRVATSKA STRANKA PRAV [og]", "Hrvatska stranka prava", 1128L,
  "HR", "BRUNA ESIH - ZLATKO HASANBEGOVIĆ: NEOVISNI ZA HRVATSKU + HRVATSKA STRANKA PRAV [og]", "Neovisni za Hrvatsku", 7422L,
  "HR", "Bandić Milan 365 i koalicijski partneri [og]", "Bandić Milan 365 — Stranka rada i solidarnosti", 5452L,
  "HR", "Coalition HDZ (1191511) + HSP AS (1191613)+HSS (1191810)+others [og]", "Croatian Party of Rights [og]", NA, # label pools HSP and HSP-AS
  "HR", "Coalition HDZ (1191511) + HSP AS (1191613)+HSS (1191810)+others [og]", "Hrvatska demokratska zajednica", 1431L, # Domoljubna koalicija 2015
  "HR", "Coalition HDZ (1191511) + HSP AS (1191613)+HSS (1191810)+others [og]", "Hrvatska seljačka stranka", 72L,
  "HR", "Coalition HDZ (1191511) + HSP AS (1191613)+HSS (1191810)+others [og]", "Hrvatska socijalno liberalna stranka", 214L, # HSLS in Domoljubna koalicija 2015
  "HR", "Coalition NF + HSLS (1191410)  + PGS + RI [og]", "Hrvatska socijalno liberalna stranka", 214L,
  "HR", "Coalition of the Croatian Peasant Party and the Croatian Social Liberal Party [og]", "Hrvatska seljačka stranka", 72L,
  "HR", "Coalition of the Croatian Peasant Party and the Croatian Social Liberal Party [og]", "Hrvatska socijalno liberalna stranka", 214L,
  "HR", "DP Miroslava Škore, Hrvatski suverenisti, Blok za Hrvatsku, HKS, Hrast, SU, Zelena Lista [og]", "Domovinski pokret", 8981L,
  "HR", "DP, HS, Blok za Hrvatsku, HKS, Hrast [og]", "Domovinski pokret", 8981L,
  "HR", "HDZ, HSLS [og]", "Hrvatska demokratska zajednica", 1431L,
  "HR", "HDZ, HSLS [og]", "Hrvatska socijalno liberalna stranka", 214L,
  "HR", "Hrvatska stranka prava (HSP) i koalicijski partneri [og]", "Croatian Party of Rights [og]", NA, # label pools HSP and HSP-AS
  "HR", "Hrvatska stranka prava (HSP) i koalicijski partneri [og]", "Hrvatska stranka prava", 1128L,
  "HR", "Koalicija IDS, PGS, LRI [og]", "Istarski demokratski sabor", 513L,
  "HR", "Koalicija Pametno i Za grad [og]", "Pametno", 7336L,
  "HR", "Koalicija Živi zid, Promijenimo Hrvatsku, Abeceda, HDSS [og]", "Živi zid", 4867L,
  "HR", "Kukuriku koalicija", "Hrvatska narodna stranka — Liberalni Demokrati", 799L,
  "HR", "Kukuriku koalicija", "Hrvatska stranka umirovljenika", 2062L,
  "HR", "Kukuriku koalicija", "Istarski demokratski sabor", 513L,
  "HR", "Kukuriku koalicija", "Socijaldemokratska partija Hrvatske", 1475L,
  "HR", "Možemo! - politička platforma, Zageb je naš, NL, RF, ORAH, Za Grad [og]", "Možemo! – politička platforma", 8980L,
  "HR", "Možemo, Zagreb je naš, NLJ, RF, Orah [og]", "Možemo! – politička platforma", 8980L,
  "HR", "Narodna stranka - Reformisti - Hrvatska seljačka stranka braće Radić, Stranka hrvatskih umirovljenika [og]", "Narodna stranka — Reformisti", 4866L,
  "HR", "PAMETNO + UNIJA KVARNERA [og]", "Pametno", 7336L,
  "HR", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS [og]", "Građansko-liberalni savez", 8178L,
  "HR", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS [og]", "Hrvatska seljačka stranka", 72L,
  "HR", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS [og]", "Hrvatska stranka umirovljenika", 2062L,
  "HR", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS [og]", "Istarski demokratski sabor", 513L,
  "HR", "Restart koalicija – SDP, HSS, HSU, SNAGA, GLAS, IDS, PGS [og]", "Socijaldemokratska partija Hrvatske", 1475L,
  "HR", "SDP, HSS, HSU [og]", "Hrvatska seljačka stranka", 72L,
  "HR", "SDP, HSS, HSU [og]", "Hrvatska stranka umirovljenika", 2062L,
  "HR", "SDP, HSS, HSU [og]", "Socijaldemokratska partija Hrvatske", 1475L,
  "HR", "SiP, Pametno, Fokus [og]", "Pametno", 7336L,
  "HU", "Alliance for Hungary - Center Party [og]", "Centrumpárt", 1779L, # 2002 Centrum list
  "HU", "Egységben Magyarországért [og]", "Demokratikus Koalíció", 469L,
  "HU", "Egységben Magyarországért [og]", "Jobbik Magyarországért Mozgalom", 42L,
  "HU", "Egységben Magyarországért [og]", "Lehet Más a Politika — LMP – Magyarország Zöld Pártja", 1650L,
  "HU", "Egységben Magyarországért [og]", "Magyar Szocialista Párt", 1408L,
  "HU", "Egységben Magyarországért [og]", "Momentum Mozgalom", 8249L,
  "HU", "Egységben Magyarországért [og]", "Párbeszéd Magyarországért", 3115L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", "Demokratikus Koalíció", 469L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", "Jobbik Magyarországért Mozgalom", 42L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", "Lehet Más a Politika — LMP – Magyarország Zöld Pártja", 1650L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", "Magyar Szocialista Párt", 1408L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", "Momentum Mozgalom", 8249L,
  "HU", "Ellenzéki Összefogás (DK-JOBBIK-LMP-MOMENTUM-MSZP-LMP-PÁRBESZÉD) [og]", "Párbeszéd Magyarországért", 3115L,
  "HU", "Fidesz — KDNP pártszövetség / Magyar Szolidaritás Szövetsége", "Fidesz — Magyar Polgári Szövetség", 1691L,
  "HU", "Fidesz — KDNP pártszövetség / Magyar Szolidaritás Szövetsége", "Kereszténydemokrata Néppárt", 1412L,
  "HU", "Fidesz — Magyar Demokrata Fórum", "Fidesz — Magyar Polgári Szövetség", 1691L,
  "HU", "Fidesz — Magyar Demokrata Fórum", "Magyar Demokrata Fórum", 1697L,
  "HU", "Hungarian Socialist Party-Dialogue for Hungary [og]", "Magyar Szocialista Párt", 1408L,
  "HU", "Hungarian Socialist Party-Dialogue for Hungary [og]", "Párbeszéd Magyarországért", 3115L,
  "HU", "MIÉP — Jobbik a Harmadik Út pártszövetség", "Jobbik Magyarországért Mozgalom", 42L,
  "HU", "MIÉP — Jobbik a Harmadik Út pártszövetség", "Magyar Igazság és Élet Pártja", 1597L,
  "HU", "Magyar Szocialista Párt - Párbeszéd (MSZP-P) [og]", "Magyar Szocialista Párt", 1408L,
  "HU", "Magyar Szocialista Párt - Párbeszéd (MSZP-P) [og]", "Párbeszéd Magyarországért", 3115L,
  "HU", "Together 2014 - Dialogue for Hungary [og]", "Együtt — A Korszakváltók Pártja", 2458L,
  "HU", "Together 2014 - Dialogue for Hungary [og]", "Párbeszéd Magyarországért", 3115L,
  "HU", "Összefogás", "Demokratikus Koalíció", 469L,
  "HU", "Összefogás", "Együtt — A Korszakváltók Pártja", 2458L,
  "HU", "Összefogás", "Magyar Szocialista Párt", 1408L, # Kormányváltás 2014
  "HU", "Összefogás", "Párbeszéd Magyarországért", 3115L,
  "IE", "Solidarity — People Before Profit", "Anti-Austerity Alliance / Solidarity", 8242L,
  "IE", "Solidarity — People Before Profit", "People Before Profit Alliance", 242L,
  "IE", "United Left Alliance", "People Before Profit Alliance", 242L,
  "IE", "United Left Alliance", "Páirtí Sóisialach", 1448L,
  "IT", "Alleanza Verdi e Sinistra [og]", "Europa Verde", 9110L,
  "IT", "Alleanza Verdi e Sinistra [og]", "Federazione dei Verdi", 851L, # pre-2021 name of Europa Verde
  "IT", "Alleanza Verdi e Sinistra [og]", "Sinistra Italiana", 1212L,
  "IT", "Azione - Italia Viva [og]", "Azione", 9082L,
  "IT", "Azione - Italia Viva [og]", "Italia Viva", 8641L,
  "IT", "Casa delle Libertà", "Alleanza Nazionale", 813L,
  "IT", "Casa delle Libertà", "Centro Cristiano Democratico", 1767L,
  "IT", "Casa delle Libertà", "Centro Cristiano Democratico  / Cristiani Democratici Uniti (Biancofiore)", 962L, # nested joint list
  "IT", "Casa delle Libertà", "Forza Italia", 1626L,
  "IT", "Casa delle Libertà", "Lega Nord", 1221L,
  "IT", "Casa delle Libertà", "Nuovo Partito Socialisti Italiani", 888L,
  "IT", "Casa delle Libertà", "Unione dei Democratici Cristiani e Democratici di Centro", 1758L,
  "IT", "Centro Cristiano Democratico  / Cristiani Democratici Uniti (Biancofiore)", "Centro Cristiano Democratico", 1767L,
  "IT", "Coalition NCD (1380633) + UDC (1380633) + PPI [og]", "Unione dei Democratici Cristiani e Democratici di Centro", 1758L,
  "IT", "Il Girasole", "Federazione dei Verdi", 851L,
  "IT", "Il Girasole", "Socialisti Democratici Italiani", 750L,
  "IT", "Il Popolo della Libertà", "Alleanza Nazionale", 813L, # 2008 joint list FI+AN; party 2009-13
  "IT", "Il Popolo della Libertà", "Forza Italia", 1626L, # 2008 joint list FI+AN; party 2009-13
  "IT", "L'Ulivo", "Democratici di Sinistra", 878L,
  "IT", "L'Ulivo", "Democrazia è Libertà — La Margherita", 279L,
  "IT", "L'Unione", "Democratici di Sinistra", 878L,
  "IT", "L'Unione", "Democrazia è Libertà — La Margherita", 279L,
  "IT", "L'Unione", "Federazione dei Verdi", 851L,
  "IT", "L'Unione", "L'Ulivo", 1737L, # nested: Ulivo list inside Unione coalition
  "IT", "L'Unione", "Lista di Pietro — Italia del Valori", 768L,
  "IT", "L'Unione", "Partito dei Comunisti Italiani", 1635L,
  "IT", "L'Unione", "Partito della Rifondazione Comunista", 1404L,
  "IT", "L'Unione", "Radicali Italiani", 8248L, # via Rosa nel Pugno
  "IT", "L'Unione", "Socialisti Democratici Italiani", 750L, # via Rosa nel Pugno
  "IT", "L'Unione", "Unione Democratici per l'Europa", 115L, # UDEUR
  "IT", "Liberi e Uguali", "Articolo Uno", 8361L,
  "IT", "Liberi e Uguali", "Sinistra Italiana", 1212L,
  "IT", "Noi con l'Italia - UDC [og]", "Unione dei Democratici Cristiani e Democratici di Centro", 1758L,
  "IT", "Partito della Rifondazione Comunista, PdCI-PRC-S2000-CU [og]", "Partito dei Comunisti Italiani", 1635L, # 2009 EP Lista Anticapitalista
  "IT", "Partito della Rifondazione Comunista, PdCI-PRC-S2000-CU [og]", "Partito della Rifondazione Comunista", 1404L, # 2009 EP Lista Anticapitalista
  "IT", "Sinistra (Rifondazione) [og]", "Partito della Rifondazione Comunista", 1404L, # 2019 EP La Sinistra
  "IT", "Sinistra (Rifondazione) [og]", "Sinistra Italiana", 1212L, # 2019 EP La Sinistra
  "IT", "Sinistra Arcobaleno [og]", "Federazione dei Verdi", 851L,
  "IT", "Sinistra Arcobaleno [og]", "Partito dei Comunisti Italiani", 1635L,
  "IT", "Sinistra Arcobaleno [og]", "Partito della Rifondazione Comunista", 1404L,
  "IT", "Stati Uniti d'Europa [og]", "Italia Viva", 8641L,
  "IT", "Stati Uniti d'Europa [og]", "Partito Socialista [og]", NA, # assumed to be PSI
  "IT", "Stati Uniti d'Europa [og]", "Più Europa", 6155L,
  "IT", "Stati Uniti d'Europa [og]", "Radicali Italiani", 8248L,
  "LT", "Anti-Corruption Coalition [og]", "Lietuvos Centro Partija", 5455L, # assumed = 2016 Puteikis-Krivickas coalition
  "LT", "Anti-corruption coalition of K. Krivickas and N. Puteikis (Lithuanian Centre Party, Lithuanian Pensioners' Party) (APKK) [og]", "Lietuvos Centro Partija", 5455L,
  "LT", "Antikorupcinė N. Puteikio ir K. Krivicko koalicija (Lietuvos centro partija, Li [og]", "Lietuvos Centro Partija", 5455L,
  "LT", "Coalition of LLRA and LRS [og]", "Lietuvos lenkų rinkimų akcija", 556L,
  "LT", "Koalicija Darbo partija + Jaunimas", "Darbo Partija", 102L,
  "LT", "Taikos koalicija (Lietuvos krikščioniškosios demokratijos partija, Žemaičių partija) [og]", "Žemaičių partija", 1291L,
  "LT", "„Valdemaro Tomaševskio blokas\" – Krikščioniškų šeimų sąjungos ir Rus [og]", "Lietuvos lenkų rinkimų akcija", 556L,
  "LV", "Attīstībai/Par!", "Kustība Par!", 7863L,
  "LV", "Attīstībai/Par!", "Latvijas attīstībai", 3213L,
  "LV", "Jaunā Vienotība", "Vienotība", 852L,
  "LV", "LSDSP/KDS/GKL [og]", "Kristīgi demokrātiskā savienība", 304L,
  "LV", "Latvijas Pirmā partija / Latvijas Ceļš", "Latvijas Ceļš", 1043L, # 2006 joint list, merged 2007
  "LV", "Latvijas Pirmā partija / Latvijas Ceļš", "Latvijas Pirmā partija", 1778L, # 2006 joint list, merged 2007
  "LV", "Nacionālā apvienība", "Tēvzemei un Brīvībai / LNNK", 1704L, # 2010 alliance, merged 2011
  "LV", "Nacionālā apvienība", "Visu Latvijai!", 506L, # 2010 alliance, merged 2011
  "LV", "Par Labu Latviju!", "Latvijas Ceļš", 1043L, # via LPP/LC
  "LV", "Par Labu Latviju!", "Latvijas Pirmā partija", 1778L, # via LPP/LC
  "LV", "Par Labu Latviju!", "Latvijas Pirmā partija / Latvijas Ceļš", 461L,
  "LV", "Par Labu Latviju!", "Tautas Partija", 1755L,
  "LV", "Saskaņas Centrs", "Sociāldemokrātiskā Partija \"Saskaņa\"", 5976L,
  "LV", "Vienotība", "Jaunais Laiks", 1531L, # 2010 alliance, merged 2011
  "LV", "Vienotība", "Pilsoniskā savienība", 415L, # 2010 alliance, merged 2011
  "NL", "ChristenUnie — Staatkundig Gereformeerde Partij", "ChristenUnie", 1459L, # EP joint list
  "NL", "ChristenUnie — Staatkundig Gereformeerde Partij", "Staatkundig Gereformeerde Partij", 1178L, # EP joint list
  "NL", "Coalition GL - PvdA (GroenLinks,PvdA) [og]", "GroenLinks", 1537L,
  "NL", "Coalition GL - PvdA (GroenLinks,PvdA) [og]", "Partij van de Arbeid", 1234L,
  "NL", "Coalition GL - PvdA (GroenLinks,PvdA) [og]", "Partij van de Arbeid (PvdA)/Europese Sociaaldemocraten [og]", NA,
  "PL", "Akcja Wyborcza Solidarność Prawicy", "Zjednoczenie Chrześcijańsko-Narodowe", 70L, # AWS founding member
  "PL", "Europa Plus Social Movement [og]", "Ruch Palikota / Twój Ruch", 731L, # 2014 EP list
  "PL", "Europa Plus Social Movement [og]", "Socjaldemokracja Polska", 1328L, # 2014 EP list
  "PL", "Koalicja Europejska PO PSL SLD .N Zieloni (KE ) [og]", "Nowoczesna", 4630L,
  "PL", "Koalicja Europejska PO PSL SLD .N Zieloni (KE ) [og]", "Platforma Obywatelska", 1117L,
  "PL", "Koalicja Europejska PO PSL SLD .N Zieloni (KE ) [og]", "Polskie Stronnictwo Ludowe", 6757L,
  "PL", "Koalicja Europejska PO PSL SLD .N Zieloni (KE ) [og]", "Sojusz Lewicy Demokratycznej", 57L,
  "PL", "Koalicja Obywatelska", "Nowoczesna", 4630L,
  "PL", "Koalicja Obywatelska", "Platforma Obywatelska", 1117L,
  "PL", "Koalicyjny Komitet Wyborczy Porozumienie dla Przyszłości - CentroLewica (PD+SDPL+Zieloni 2004) [og]", "Socjaldemokracja Polska", 1328L,
  "PL", "Konfederacja Wolność i Niepodległość", "Koalicja Odnowy Rzeczypospolitej Wolność i Nadzieja KORWiN", 4629L,
  "PL", "Konfederacja Wolność i Niepodległość", "Ruch Narodowy", 3224L,
  "PL", "Lewica [og]", "Nowa Lewica", 9349L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Lewica [og]", "Polska Partia Socjalistyczna, PPS [og]", NA, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Lewica [og]", "Sojusz Lewicy Demokratycznej", 57L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Lewica [og]", "Spring", NA, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Lewica [og]", "Wiosna", 7356L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Lewica [og]", "[Lewica] Razem", 4871L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Lewica i Demokraci", "Socjaldemokracja Polska", 1328L,
  "PL", "Lewica i Demokraci", "Sojusz Lewicy Demokratycznej", 57L,
  "PL", "Lewica i Demokraci", "Unia Pracy", 1104L,
  "PL", "Nowa Lewica", "Polska Partia Socjalistyczna, PPS [og]", NA, # 2021 merger SLD+Wiosna; 2023 list incl. Razem, PPS
  "PL", "Nowa Lewica", "Sojusz Lewicy Demokratycznej", 57L, # 2021 merger SLD+Wiosna; 2023 list incl. Razem, PPS
  "PL", "Nowa Lewica", "Spring", NA, # 2021 merger SLD+Wiosna; 2023 list incl. Razem, PPS
  "PL", "Nowa Lewica", "Wiosna", 7356L, # 2021 merger SLD+Wiosna; 2023 list incl. Razem, PPS
  "PL", "Nowa Lewica", "[Lewica] Razem", 4871L, # 2021 merger SLD+Wiosna; 2023 list incl. Razem, PPS
  "PL", "Ogólnopolska Koalicja Obywatelska [og]", "Nowoczesna", 4630L,
  "PL", "Ogólnopolska Koalicja Obywatelska [og]", "Platforma Obywatelska", 1117L,
  "PL", "Senate 2001 coalition [og]", "Akcja Wyborcza Solidarność Prawicy", 863L, # Blok Senat 2001
  "PL", "Senate 2001 coalition [og]", "Platforma Obywatelska", 1117L, # Blok Senat 2001
  "PL", "Senate 2001 coalition [og]", "Prawo i Sprawiedliwość", 1565L, # Blok Senat 2001
  "PL", "Senate 2001 coalition [og]", "Unia Wolności", 8268L, # Blok Senat 2001
  "PL", "Sojusz Lewicy Demokratycznej", "Socjaldemokracja Rzeczypospolitej Polskiej", 8224L, # SdRP core of SLD coalition, merged 1999
  "PL", "Sojusz Lewicy Demokratycznej — Unia Pracy", "Sojusz Lewicy Demokratycznej", 57L,
  "PL", "Sojusz Lewicy Demokratycznej — Unia Pracy", "Unia Pracy", 1104L,
  "PL", "The Left [og]", "Nowa Lewica", 9349L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "The Left [og]", "Polska Partia Socjalistyczna, PPS [og]", NA, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "The Left [og]", "Sojusz Lewicy Demokratycznej", 57L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "The Left [og]", "Spring", NA, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "The Left [og]", "Wiosna", 7356L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "The Left [og]", "[Lewica] Razem", 4871L, # 2019/2023 Lewica list; Spring = Wiosna
  "PL", "Trzecia Droga [og]", "Polska 2050", 9027L,
  "PL", "Trzecia Droga [og]", "Polskie Stronnictwo Ludowe", 6757L,
  "PL", "United Right [og]", "Prawo i Sprawiedliwość", 1565L, # United Poland = Solidarna Polska
  "PL", "United Right [og]", "Solidarna Polska", 328L, # United Poland = Solidarna Polska
  "PL", "United Right [og]", "United Poland", NA, # United Poland = Solidarna Polska
  "PL", "Zjednoczona Lewica", "Polska Partia Socjalistyczna, PPS [og]", NA,
  "PL", "Zjednoczona Lewica", "Ruch Palikota / Twój Ruch", 731L,
  "PL", "Zjednoczona Lewica", "Sojusz Lewicy Demokratycznej", 57L,
  "PL", "Zjednoczona Lewica", "Unia Pracy", 1104L,
  "PT", "Aliança Democrática", "Centro Democrático e Social / Partido Popular", 1308L,
  "PT", "Aliança Democrática", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "PT", "Coligação Democrática Unitária", "Partido Comunista Português", 310L,
  "PT", "Coligação Democrática Unitária", "Partido Ecologista — Os Verdes", 886L,
  "PT", "Coligação: PSD (6), CDS - PP (2), Força Portugal [only in v112] [og]", "Centro Democrático e Social / Partido Popular", 1308L,
  "PT", "Coligação: PSD (6), CDS - PP (2), Força Portugal [only in v112] [og]", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "PT", "Numa coligação PSD/CDS [only in v114] [og]", "Centro Democrático e Social / Partido Popular", 1308L,
  "PT", "Numa coligação PSD/CDS [only in v114] [og]", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "PT", "Portugal à Frente", "Centro Democrático e Social / Partido Popular", 1308L,
  "PT", "Portugal à Frente", "Partido Popular Democrático / Partido Social Democrata", 1359L,
  "RO", "Alianţa Dreptate şi Adevăr", "Partidul Democrat", 1715L,
  "RO", "Alianţa Dreptate şi Adevăr", "Partidul Naţional Liberal", 481L,
  "RO", "Alianţa România Dreaptă", "Partidul Democrat-Liberal", 660L,
  "RO", "Alianţa România Dreaptă", "Partidul Naţional Ţărănesc Creştin Democrat", 1750L,
  "RO", "Alianța 2020 USR PLUS (2020 USR PLUS) [og]", "Partidul Libertate, Unitate și Solidaritate (PLUS) [og]", NA,
  "RO", "Alianța 2020 USR PLUS (2020 USR PLUS) [og]", "Partidul Libertății Unității și Solidarității (PLUS) [og]", NA,
  "RO", "Alianța 2020 USR PLUS (2020 USR PLUS) [og]", "Uniunea Salvați România", 5969L,
  "RO", "Alianța Dreapta Unită [og]", "Partidul Mișcarea Populară", 3210L,
  "RO", "Alianța Dreapta Unită [og]", "Uniunea Salvați România", 5969L,
  "RO", "Alianța Electorală PSD-PNL [og]", "Partidul Naţional Liberal", 481L,
  "RO", "Alianța Electorală PSD-PNL [og]", "Partidul Social Democrat", 2555L,
  "RO", "Alianța PSD+PC", "Partidul Conservator / Partidul Umanist Român", 1443L,
  "RO", "Alianța PSD+PC", "Partidul Social Democrat", 2555L,
  "RO", "Alianța USR-PLUS [og]", "Partidul Libertate, Unitate și Solidaritate (PLUS) [og]", NA,
  "RO", "Alianța USR-PLUS [og]", "Partidul Libertății Unității și Solidarității (PLUS) [og]", NA,
  "RO", "Alianța USR-PLUS [og]", "Uniunea Salvați România", 5969L,
  "RO", "Christian-Liberal Alliance [og]", "Partidul Democrat-Liberal", 660L, # ACL 2014
  "RO", "Christian-Liberal Alliance [og]", "Partidul Naţional Liberal", 481L, # ACL 2014
  "RO", "Convenţia Democrată Română", "Partidul Alianţa Civică", 645L,
  "RO", "Convenţia Democrată Română", "Partidul Naţional Liberal", 481L, # CDR member 1992-2000
  "RO", "Convenţia Democrată Română", "Partidul Naţional Ţărănesc Creştin Democrat", 1750L,
  "RO", "Electoral Alliance PSD (1642300) + UNPR + PC (1642600) [og]", "Partidul Conservator / Partidul Umanist Român", 1443L,
  "RO", "Electoral Alliance PSD (1642300) + UNPR + PC (1642600) [og]", "Partidul Social Democrat", 2555L,
  "RO", "PSD-UNPR-PC Electoral Alliance [og]", "Partidul Conservator / Partidul Umanist Român", 1443L,
  "RO", "PSD-UNPR-PC Electoral Alliance [og]", "Partidul Social Democrat", 2555L,
  "RO", "PSD-UNPR-PC Electoral Alliance [og]", "Partidul Social Democrat Român", 1034L, # label used for PSD in this source
  "RO", "Uniunea Naţională PSD+PUR", "Partidul Conservator / Partidul Umanist Român", 1443L,
  "RO", "Uniunea Naţională PSD+PUR", "Partidul Social Democrat", 2555L,
  "RO", "Uniunea Social Liberală", "Partidul Conservator / Partidul Umanist Român", 1443L,
  "RO", "Uniunea Social Liberală", "Partidul Naţional Liberal", 481L,
  "RO", "Uniunea Social Liberală", "Partidul Social Democrat", 2555L,
  "SI", "Coalition New Slovenia + Slovenian People's Party (NSi (1705522) + SLS (1705521)) [og]", "Nova Slovenija — Krščanska ljudska stranka", 1618L,
  "SI", "Coalition New Slovenia + Slovenian People's Party (NSi (1705522) + SLS (1705521)) [og]", "Slovenska ljudska stranka", 6740L,
  "SI", "LDS (2) + DESUS (1) [only in v112] [og]", "Demokratična stranka upokojencev Slovenije", 467L,
  "SI", "LDS (2) + DESUS (1) [only in v112] [og]", "Liberalna demokracija Slovenije", 975L,
  "SI", "Slovenska demokratska stranka + Slovenska ljudska stranka [og]", "Slovenska demokratska stranka", 472L,
  "SI", "Slovenska demokratska stranka + Slovenska ljudska stranka [og]", "Slovenska ljudska stranka", 6740L,
  "SI", "Slovenska demokratska stranka in Slovenska ljudska stranka(SDS+SLS) [og]", "Slovenska demokratska stranka", 472L,
  "SI", "Slovenska demokratska stranka in Slovenska ljudska stranka(SDS+SLS) [og]", "Slovenska ljudska stranka", 6740L,
  "SK", "Hungarian Community Togetherness [og]", "Strana maďarskej koalície / Strana maďarskej komunity", 349L, # 2020 MKÖ joint list
  "SK", "Koalícia Komunistická Strana Slovenska, Vzdor - Strana Práce [og]", "Komunistická strana Slovenska", 6798L,
  "SK", "Koalícia Komunistická Strana Slovenska, Vzdor - Strana Práce [og]", "VZDOR - strana práce [og]", NA,
  "SK", "Obyčajní Ľudia a Nezávislé Osobnosti", "Nová väčšina", 6686L, # joint lists 2016/2020
  "SK", "Obyčajní Ľudia a Nezávislé Osobnosti", "Za ľudí", 8394L, # 2023 joint list
  "SK", "Pogresívne Slovensko - Spolu [og]", "Progresívne Slovensko", 7361L,
  "SK", "Pogresívne Slovensko - Spolu [og]", "Spolu — občianska demokracia", 8177L,
  "SK", "Progresivne Slovensko-SPOLU [og]", "Progresívne Slovensko", 7361L,
  "SK", "Progresivne Slovensko-SPOLU [og]", "Spolu — občianska demokracia", 8177L,
  "SK", "Progressive Slovakia - TOGETHER-Civic Democracy [og]", "Progresívne Slovensko", 7361L,
  "SK", "Progressive Slovakia - TOGETHER-Civic Democracy [og]", "Spolu — občianska demokracia", 8177L,
  "SK", "Progressive Slovakia and Together - Civic Democracy [og]", "Progresívne Slovensko", 7361L,
  "SK", "Progressive Slovakia and Together - Civic Democracy [og]", "Spolu — občianska demokracia", 8177L,
  "SK", "Progresívne Slovensko/ SPOLU [og]", "Progresívne Slovensko", 7361L,
  "SK", "Progresívne Slovensko/ SPOLU [og]", "Spolu — občianska demokracia", 8177L,
  "SK", "SNS (Slovenská národná strana) / Pravá SNS [og]", "Slovenská národná strana", 730L, # combined answer category
  "SK", "hungarian community togetherness [og]", "Strana maďarskej koalície / Strana maďarskej komunity", 349L # 2020 MKÖ joint list
)

pf_check_alliances(alliances, core = core)

alliances |>
  summarise(members = n(), .by = c(cntry, alliance_name)) |>
  count(cntry, name = "alliances") |>
  print(n = Inf)

write_csv(alliances, here("data", "02_processed", "pf_alliances.csv"), na = "")
```

# Before and After

The share of *same party* among attached voters with two real party answers, compared by Party Facts name only and with the alliance rule.

```{r}
#| label: before-after

pairs |>
  mutate(
    by_name = att == vote,
    with_alliances = pf_match(cntry, att, vote, alliances) == "same"
  ) |>
  summarise(
    pairs = n(),
    by_name = mean(by_name),
    with_alliances = mean(with_alliances),
    .by = c(source, cntry)
  ) |>
  mutate(gain = with_alliances - by_name) |>
  arrange(desc(gain)) |>
  print(n = 40)
```

Spot checks: attachment to the CDU and a vote for the CDU/CSU union match, attachment to the CDU and a vote for the CSU do not; the same for Fidesz and Fidesz–KDNP, and for PO and KO.

```{r}
#| label: spot-checks

tribble(
  ~cntry, ~att, ~vote,
  "DE", "Christlich Demokratische Union", "Christlich Demokratische Union / Christlich Soziale Union",
  "DE", "Christlich Soziale Union", "Christlich Demokratische Union / Christlich Soziale Union",
  "DE", "Christlich Demokratische Union", "Christlich Soziale Union",
  "HU", "Fidesz — Magyar Polgári Szövetség", "Fidesz — KDNP pártszövetség / Magyar Szolidaritás Szövetsége",
  "PL", "Platforma Obywatelska", "Koalicja Obywatelska"
) |>
  mutate(match = pf_match(cntry, att, vote, alliances))
```

The largest pairs still scored *different*: these should be genuine switches, not alliances missing from the table.
Two kinds of pair are *different* by construction, not by oversight:

- **Renamed or refounded parties with a new Party Facts id** (UMP and Les Républicains, PD and PDL in Romania, SDĽ and Smer): the table lists alliances only, not successors or later mergers.
- **Suspect answers in the sources**, e.g. EES 2004 Italy, where attachment to the Margherita or the PdCI pairs with a vote for Il Girasole (Verdi + SDI) hundreds of times, which points to a mislabelled vote code upstream rather than to an alliance.

```{r}
#| label: remaining

pairs |>
  filter(pf_match(cntry, att, vote, alliances) == "different") |>
  count(cntry, att, vote, sort = TRUE) |>
  mutate(across(c(att, vote), \(x) str_trunc(x, 32))) |>
  head(30)
```

# Match Files

For every vote item, the match with the attachment item of the same respondent.

```{r}
#| label: match-files

add_matches <- function(df, att, votes) {
  for (v in votes) {
    df[[paste0("ext_match_", v)]] <- pf_match(
      df$cntry,
      df[[paste0("ext_", att, "_pf_name")]],
      df[[paste0("ext_", v, "_pf_name")]],
      alliances
    )
    # no vote item for this respondent
    df[[paste0("ext_match_", v)]][is.na(df[[paste0("ext_", v, "_pf_name")]])] <- NA
  }
  df
}

vote_stubs <- function(df, pattern) {
  str_subset(names(df), pattern) |>
    str_remove("^ext_") |>
    str_remove("_pf_name$")
}

ess_match <- ess |>
  add_matches("prtcl", vote_stubs(ess, "^ext_prtv.*_pf_name$")) |>
  select(cntry, essround, name, idno, starts_with("ext_match_"))

cses_match <- cses |>
  add_matches("imd3005_3", vote_stubs(cses, "^ext_(imd3002|f3016)_.*_pf_name$")) |>
  select(release, study, respondent, cntry, year, starts_with("ext_match_"))

ees_match <- ees |>
  add_matches("pid", vote_stubs(ees, "^ext_vote_.*_pf_name$")) |>
  select(wave, row, respondent, cntry, region, starts_with("ext_match_"))

write_parquet(ess_match, here("data", "02_processed", "ess_party_match.parquet"), compression = "zstd")
write_parquet(cses_match, here("data", "02_processed", "cses_party_match.parquet"), compression = "zstd")
write_parquet(ees_match, here("data", "02_processed", "ees_party_match.parquet"), compression = "zstd")

map(
  list(ESS = ess_match, CSES = cses_match, EES = ees_match),
  \(d) d |>
    select(starts_with("ext_match_")) |>
    pivot_longer(everything()) |>
    count(name, value) |>
    pivot_wider(names_from = value, values_from = n, values_fill = 0)
)
```

# Session Info

```{r}
#| label: session-info

session_info()
```

# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()
rendering_time <- end_time - start_time
message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds."
))
```
