---
title: |
  Code Notebook 4.1
subtitle: |
  Bayesian Multilevel Models of Attitudinal and Behavioral Affective Polarization
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

This notebook implements the specification stated in `index.qmd`, @sec-model-spec and
@sec-appendix-model: one backbone estimated twice, once with the token allocation as the
outcome and once with the rescaled thermometer rating, differing only in the levels $R$
can take and in the grouping structure wrapped around it.

$$
Y_{riac} = \alpha + \beta_1 R_{ri} + \beta_2 T_i + \beta_3 (R_{ri} \times T_i) + \boldsymbol{X}_{ri}'\boldsymbol{\gamma}
         + u_{i0}
         + v_{a0} + v_{aR} R_{ri} + v_{aT} T_i
         + w_{c0} + w_{cR} R_{ri} + w_{cT} T_i + w_{c,RT}(R_{ri} \times T_i)
         + \varepsilon_{riac}
$$

Nothing here is read off a coefficient table. The reported quantities are marginal means
and contrasts between them,

$$
\mathrm{AP}_t = \mu_t(\mathrm{Co}) - \mu_t(\mathrm{Out}),
\qquad
\Delta = \mathrm{AP}_1 - \mathrm{AP}_0,
$$

computed per posterior draw and summarized afterwards.

Robustness fits — the location-scale attitudinal model, the ordinal likelihood, the
no-anchor-slopes refit, the defector recode and the full-sample attitudinal refit — belong
to notebook 4.2 and are deliberately not run here.

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)

# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "brms",
  "posterior",
  "tidybayes",
  "bayesplot",
  "ggpubr",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Four chains, run in parallel.
options(mc.cores = 4)

# Fitted model objects live outside the tracked tree (see .gitignore).
model_dir <- here("models")
dir.create(model_dir, showWarnings = FALSE)
```

`smoke_test` is the switch between a fast structural check and the specification the
manuscript commits to. The cache file name carries the flag, because `file_refit =
"on_change"` compares the formula, data and priors -- not the sampler settings -- and
would otherwise hand back a 500-iteration fit when a 4,000-iteration one was asked for.

Where `cmdstanr` is available we use it, with within-chain threading. The bottleneck in
both models is a Gaussian likelihood evaluated over tens of thousands of rows, which is
precisely what `reduce_sum` splits across threads, so four chains of four threads occupy
sixteen cores rather than four and cut the wall clock by roughly a factor of five. The
backend is detected rather than hard-coded, and falls back to `rstan` with no threading if
CmdStan is not installed, so the notebook runs either way.

```{r}
#| label: sampler-settings

smoke_test <- TRUE # TRUE -> 2 chains x 500 iterations, for a structural check

# prefer cmdstanr + threading; fall back to rstan if CmdStan is not installed
has_cmdstan <- requireNamespace("cmdstanr", quietly = TRUE) &&
  !is.null(tryCatch(cmdstanr::cmdstan_path(), error = function(e) NULL))

n_threads <- 4

sampler <- list(
  chains = if (smoke_test) 2 else 4,
  cores = if (smoke_test) 2 else 4,
  iter = if (smoke_test) 500 else 4000,
  warmup = if (smoke_test) 250 else 2000,
  seed = 20260810,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  backend = if (has_cmdstan) "cmdstanr" else "rstan",
  threads = if (has_cmdstan) threading(n_threads) else NULL
)

# cache path; the suffix keeps smoke and full fits from overwriting one another,
# and the backend from silently reusing a fit compiled the other way
model_file <- function(name) {
  file.path(
    model_dir,
    paste0(name, if (smoke_test) "_smoke", if (has_cmdstan) "_mt")
  )
}

list(
  backend = sampler$backend,
  chains = sampler$chains,
  threads_per_chain = if (has_cmdstan) n_threads else 1,
  cores_busy = sampler$cores * if (has_cmdstan) n_threads else 1,
  iter = sampler$iter,
  warmup = sampler$warmup
)
```

# Load data

```{r}
#| label: load-data

eu25games2019 <- readRDS(
  here(
    "data",
    "03_final",
    "eu25games2019.rds"
  )
)

thermo_long <- readRDS(
  here(
    "data",
    "03_final",
    "thermo_long.rds"
  )
)
```

# Analysis frames

Both frames are reduced to the columns the models use and given identical names for the
three framework variables (`R`, `type`) and the three grouping factors (`pid`, `anchor`,
`country`), so that every downstream function works on either.

Out-partisan is the reference level of $R$ and vote-anchored the reference level of $T$,
which makes $\beta_1^{\mathrm{Co}}$ the co-partisan effect among vote-anchored respondents
and $\beta_3^{\mathrm{Co}}$ the difference in that effect between the types.

## Behavioral frame

The unit is a respondent-round. Rows enter when the respondent has a partisan type and the
round has a defined relationship category, which after the correction to
`der_partisan_relationship` in notebook 2.1 means: the two party-showing treatments, plus
the co-national arm of `4_somenat` as the no-cue control.

Two coding decisions deserve a note.

*EU stance.* The treatment branch shows the EU-stance attribute only in
`1_conat_eustance_partisan` and `3_eunat_eustance`. Since treatment 3 shows no party and
so contributes no rows here, the retained rounds split into treatment 1 (party and EU
stance), treatment 2 (party only) and the control (neither). Coding `eupos` with an
explicit `not_shown` level identifies the EU-stance effect from the treatment 1 versus
treatment 2 comparison, so the `None` contrast is adjusted for the absent attribute rather
than confounded with it. Left uncoded, $\mu_t(\mathrm{None})$ would not be a clean no-cue
baseline.

*What is left out.* `cj_nationality_shown` is `own_country` in every retained row and so
carries no information. Respondent-level covariates are omitted from both models on
purpose: they are collinear with the respondent-level varying intercept and would
contribute nothing to a within-person contrast. Partner age enters as a factor rather than
a number, since it is a randomized attribute with five levels and nothing requires its
effect to be linear.

```{r}
#| label: frame-behavioral

dat_behav <- eu25games2019 |>
  filter(
    !is.na(der_partisan_type),
    !is.na(der_partisan_relationship)
  ) |>
  transmute(
    tokens = cj_pl2,
    # relationship category R, out-partisan as reference
    R = fct_relevel(factor(as.character(der_partisan_relationship)), "Out"),
    # partisan type T, vote-anchored as reference
    type = factor(
      der_partisan_type,
      levels = c(0, 1),
      labels = c("vote_anchored", "explicit")
    ),
    # grouping factors
    pid = factor(meta_pid),
    anchor = factor(der_partisan_anchor),
    country = factor(meta_country),
    # session structure: game, and round position within game
    game = fct_relevel(factor(meta_game), "dict"),
    round = factor(meta_round),
    # remaining design attributes
    eupos = factor(
      if_else(
        cj_treatment %in% c("1_conat_eustance_partisan", "3_eunat_eustance"),
        cj_eupos,
        "not_shown"
      ),
      levels = c("not_shown", "eu_citizen", "not_eu_citizen")
    ),
    # hyphen replaced so the level survives into a readable coefficient name
    cj_rel = factor(str_replace_all(cj_rel, "-", "_")),
    cj_class = factor(cj_class),
    cj_sex = factor(cj_sex),
    cj_age = factor(cj_age)
  )

dat_behav |>
  summarise(
    rows = n(),
    respondents = n_distinct(pid),
    anchors = n_distinct(anchor),
    countries = n_distinct(country)
  )
```

## Attitudinal frame

The unit is a respondent-party rating. Ratings are already rescaled from 0--100 to the
0--10 range of the token allocations in notebook 3.2, so coefficients from the two models
are in the same units.

Two restrictions apply, both recorded in the checks below rather than asserted here. The
United Kingdom is dropped, because its thermometer battery carries scores for one party
only. And the primary sample keeps respondents who supply at least one co-partisan and at
least one out-partisan rating, which is what it takes for the within-person contrast to be
identified: a respondent with no rating of their own anchor contributes nothing to
$\mathrm{AP}_t$ once they carry a varying intercept, but does pull $\mu_t(\mathrm{Out})$
toward a population that $\mu_t(\mathrm{Co})$ is not averaged over. The full-sample refit
is deferred to notebook 4.2.

```{r}
#| label: frame-attitudinal

dat_attit_all <- thermo_long |>
  filter(
    !is.na(der_partisan_type),
    !is.na(thermo_score),
    meta_country != "United Kingdom"
  ) |>
  transmute(
    thermo = thermo_score,
    R = fct_relevel(
      factor(if_else(thermo_party == der_partisan_anchor, "Co", "Out")),
      "Out"
    ),
    type = factor(
      der_partisan_type,
      levels = c(0, 1),
      labels = c("vote_anchored", "explicit")
    ),
    pid = factor(meta_pid),
    anchor = factor(der_partisan_anchor),
    country = factor(meta_country),
    # rated party, crossed with the anchor grouping; also absorbs display
    # position, which is fixed per party within a country
    rated_party = factor(thermo_party)
  )

# primary sample: the within-person contrast must be identified
dat_attit <- dat_attit_all |>
  group_by(pid) |>
  filter(any(R == "Co"), any(R == "Out")) |>
  ungroup() |>
  mutate(across(c(anchor, rated_party), droplevels))

bind_rows(
  dat_attit_all |> mutate(sample = "all rated"),
  dat_attit |> mutate(sample = "primary (>=1 Co & >=1 Out)")
) |>
  group_by(sample, type) |>
  summarise(
    rows = n(),
    respondents = n_distinct(pid),
    .groups = "drop"
  )
```

# Pre-estimation checks

@sec-model-spec lists these under *To check before estimation*. Each decides something,
and the ones that decide the most are run first.

## 1. The no-cue baseline, $\mu_1(\mathrm{None}) - \mu_0(\mathrm{None})$

Cheap and load-bearing. It decides how $\Delta \approx 0$ may be read. If the two types
allocate alike when no partisan information is displayed at all, the type variable carries
no general behavioral signature, and a small $\Delta$ cannot be attributed to two
dissimilar populations happening to converge. If they do not, that level difference has to
be interpreted before anything else.

```{r}
#| label: check-baseline-nocue

baseline_nocue <- dat_behav |>
  filter(R == "None") |>
  group_by(type) |>
  summarise(
    mean_tokens = mean(tokens),
    sd = sd(tokens),
    n = n(),
    .groups = "drop"
  )

baseline_nocue

# raw difference and its standard error
with(
  baseline_nocue,
  tibble(
    baseline_gap = mean_tokens[type == "explicit"] -
      mean_tokens[type == "vote_anchored"],
    se = sqrt(sum(sd^2 / n))
  )
)
```

## 2. Cell counts, and where the evidence on $\beta_3$ actually comes from

The conjoint party cue was constructed around the attachment item: among self-identified
partisans, roughly half of all party-showing rounds display their own attachment party,
whereas for vote-anchored respondents a match arises only by chance. The cue is still
randomized *within* each type, so the co-partisan contrast remains a causal effect of the
cue inside each type, but the assignment probability differs by type, and the
vote-anchored respondents who do see a co-partisan profile are disproportionately anchored
to large parties. That is what the anchor-level varying slope on $R$ is in the model to
absorb.

The practical consequence is that $\Delta$ rests on the co-partisan rounds of the
vote-anchored group, and there are far fewer of them.

```{r}
#| label: check-cell-counts

# how often the displayed party matched the respondent's own anchor
eu25games2019 |>
  filter(
    str_detect(cj_treatment, "partisan"),
    !is.na(der_partisan_type),
    !is.na(ext_cj_party_pf_name)
  ) |>
  group_by(type = if_else(der_partisan_type == 1, "explicit", "vote_anchored")) |>
  summarise(
    party_cue_rounds = n(),
    share_matching_anchor = mean(ext_cj_party_pf_name == der_partisan_anchor),
    .groups = "drop"
  )

# rows and respondents behind each cell
dat_behav |>
  group_by(type, R) |>
  summarise(
    rows = n(),
    respondents = n_distinct(pid),
    mean_tokens = mean(tokens),
    .groups = "drop"
  ) |>
  arrange(type, R)
```

## 3. Can the anchor level carry a varying slope on $R \times T$?

An identification question, not a modelling preference: it requires both partisan types to
be present in workable numbers *within* each party. The binding requirement is co-partisan
rounds from both types in the same anchor.

```{r}
#| label: check-anchor-interaction

anchor_feasibility <- dat_behav |>
  filter(R == "Co") |>
  count(country, anchor, type) |>
  pivot_wider(
    names_from = type,
    values_from = n,
    values_fill = 0
  )

anchor_feasibility |>
  summarise(
    anchors_with_co_rounds = n(),
    both_types_present = sum(vote_anchored > 0 & explicit > 0),
    both_types_ge_10 = sum(vote_anchored >= 10 & explicit >= 10)
  )
```

Roughly half the anchors carry co-partisan rounds from one type only, so the interaction
stays at the country level, exactly as written in @eq-multilevel. The party-level type gap
is thereby assumed homogeneous, and that assumption is stated rather than hidden.

## 4. Thermometer battery coverage

The battery holds six parties per country, so respondents anchored to a party outside it
supply no co-partisan rating. This is the restriction behind the primary attitudinal
sample, and it does not fall evenly.

```{r}
#| label: check-battery-country

battery_coverage <- dat_attit_all |>
  distinct(country, pid, type, anchor) |>
  left_join(
    dat_attit_all |>
      filter(R == "Co") |>
      distinct(pid) |>
      mutate(has_co = TRUE),
    by = "pid"
  ) |>
  mutate(has_co = coalesce(has_co, FALSE)) |>
  group_by(country, type) |>
  summarise(
    respondents = n(),
    pct_no_co = 100 * mean(!has_co),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = type,
    values_from = c(respondents, pct_no_co)
  ) |>
  mutate(
    type_gap = pct_no_co_vote_anchored - pct_no_co_explicit
  ) |>
  arrange(desc(pct_no_co_vote_anchored))

battery_coverage |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))
```

```{r}
#| label: check-battery-party

# which anchors fall outside their country's battery, by respondents lost
battery <- dat_attit_all |>
  distinct(country, rated_party)

dat_attit_all |>
  distinct(country, pid, type, anchor) |>
  anti_join(battery, by = c("country", "anchor" = "rated_party")) |>
  count(country, anchor, sort = TRUE) |>
  head(20)
```

The loss is concentrated in a handful of party systems, and within them it is far heavier
for vote-anchored respondents than for self-identified partisans, because the battery
covers the large established parties while vote-anchored respondents are more often
anchored to newer or smaller ones. The country-level attitudinal estimates for the worst
affected systems therefore rest on a narrowed and differentially selected base, and must be
reported as such in @sec-attitudinal.

## 5. Item non-response on the thermometer, by type

Non-response on a party correlates with disliking it, so if the rates differ by type,
listwise deletion selects in the direction that widens the attitudinal gap and the estimate
would have to be reported as a bound.

```{r}
#| label: check-thermo-nonresponse

eu25games2019 |>
  distinct(meta_pid, .keep_all = TRUE) |>
  filter(!is.na(der_partisan_type)) |>
  select(der_partisan_type, meta_country, num_range("q_thermo_", 1:6)) |>
  pivot_longer(num_range("q_thermo_", 1:6)) |>
  group_by(
    country_set = if_else(
      meta_country == "United Kingdom",
      "United Kingdom",
      "other 24"
    ),
    type = if_else(der_partisan_type == 1, "explicit", "vote_anchored")
  ) |>
  summarise(
    ratings = n(),
    pct_missing = 100 * mean(is.na(value)),
    .groups = "drop"
  )
```

## 6. Number of rated parties per country, and battery order

The rated-party varying intercept is relied on to absorb display position, which holds only
if the order of the battery really is identical within each country. It also matters how
many parties each country contributes, since respondents in fragmented systems supply more
out-partisan rows and so more weight in the pooled contrast.

```{r}
#| label: check-battery-structure

# parties per country, and ratings per respondent
thermo_long |>
  group_by(meta_country) |>
  summarise(
    parties = n_distinct(thermo_party),
    ratings_per_respondent = mean(table(meta_pid)),
    .groups = "drop"
  ) |>
  arrange(parties)

# each battery slot maps to exactly one party within a country
eu25games2019 |>
  distinct(meta_pid, .keep_all = TRUE) |>
  select(meta_country, matches("^ext_q_thermo_[0-9]+_pf_name$")) |>
  pivot_longer(-meta_country, names_to = "slot", values_to = "party") |>
  group_by(meta_country, slot) |>
  summarise(parties_in_slot = n_distinct(party, na.rm = TRUE), .groups = "drop") |>
  count(parties_in_slot)
```

The United Kingdom is the exception that forces its own exclusion: its six battery slots
are named, but only the first carries any scores, in both waves and already in the raw
data. UK respondents supply one rating each and no within-person contrast, so the
attitudinal model runs on 24 countries while the behavioral model keeps all 25.

## 7. Reference levels

The mapping from coefficients to estimands in @sec-model-spec assumes out-partisan is the
reference level of $R$ and vote-anchored the reference level of $T$. Asserted rather than
eyeballed.

```{r}
#| label: check-reference-levels

stopifnot(
  levels(dat_behav$R)[1] == "Out",
  levels(dat_attit$R)[1] == "Out",
  levels(dat_behav$type)[1] == "vote_anchored",
  levels(dat_attit$type)[1] == "vote_anchored",
  # three relationship levels in the games, two in the ratings
  nlevels(dat_behav$R) == 3,
  nlevels(dat_attit$R) == 2
)

list(R_behav = levels(dat_behav$R), R_attit = levels(dat_attit$R))
```

## 8. Covariate balance

The remaining conjoint attributes are randomized independently of the partisan cue, so
omitting them would not bias $\mathrm{AP}_t$. They are kept because @sec-behavioral
benchmarks the partisan attribute against EU position, class and religion within the same
conjoint. Balance is reported rather than assumed.

```{r}
#| label: check-balance

dat_behav |>
  select(type, R, eupos, cj_rel, cj_class, cj_sex, cj_age) |>
  pivot_longer(-c(type, R), names_to = "attribute", values_to = "level") |>
  count(type, R, attribute, level) |>
  group_by(type, R, attribute) |>
  mutate(share = round(n / sum(n), 3)) |>
  ungroup() |>
  select(-n) |>
  pivot_wider(names_from = c(type, R), values_from = share) |>
  arrange(attribute, level)
```

# Model specification

## Priors

Weakly informative and scaled to the 0--10 outcome, exactly as tabulated in
@sec-appendix-model. They rule out effects larger than the outcome range itself while
remaining far more diffuse than the likelihood.

```{r}
#| label: priors

priors <- c(
  # centred at the midpoint of the outcome range
  prior(normal(5, 2), class = Intercept),
  # centred at no effect; ~95% of prior mass within +/- 4 units
  prior(normal(0, 2), class = b),
  # positive support with heavy tails, shrinking toward zero
  prior(student_t(3, 0, 1), class = sd),
  prior(student_t(3, 0, 2), class = sigma),
  # mildly favours weaker correlations, regularizing the covariance blocks
  prior(lkj(2), class = cor)
)

priors
```

## Formulas

Respondents receive a varying intercept only. Anchors receive an intercept and slopes on
$R$ and on $T$ -- the slope on $R$ being the design's answer to the composition objection,
since anchors are parties *within* countries and letting the slope vary by anchor estimates
the type contrast inside party-country cells. Countries receive an intercept and slopes on
$R$, on $T$ and on their interaction; that last term is not a nuisance but the source of
every comparative quantity reported.

Neither model carries a respondent-level varying slope on $R$. In the games that is a
matter of power, with roughly two co-partisan rounds of six per respondent. In the ratings
the term is not thinly identified but unidentified: each respondent supplies exactly one
co-partisan rating, so a person-specific slope would absorb that single observation as a
residual and leave the contrast determined by the prior.

```{r}
#| label: formulas

# Behavioral: R has three levels; game is interacted with R and T so that the
# agreement between the two instruments is a quantity with a posterior rather
# than a visual comparison of two tables. Round position enters separately, so
# that practice within an instrument stays visible.
f_behav <- bf(
  tokens ~ R * type * game + round +
    eupos + cj_rel + cj_class + cj_sex + cj_age +
    (1 | pid) +
    (1 + R + type | anchor) +
    (1 + R * type | country)
)

# Attitudinal: R is binary, there are no design attributes, and the rated party
# enters as a further varying intercept crossed with the anchor grouping.
f_attit <- bf(
  thermo ~ R * type +
    (1 | pid) +
    (1 + R + type | anchor) +
    (1 + R * type | country) +
    (1 | rated_party)
)

list(behavioral = f_behav, attitudinal = f_attit)
```

Both outcomes are bounded on $[0, 10]$ and discrete, and a Gaussian likelihood is assumed
anyway. The choice is a deliberate trade, argued in @sec-appendix-model: the quantities of
interest are differences in conditional means, and under a Normal likelihood the
coefficients are exactly those differences on the original token scale, with no link
function or set of cutpoints standing between the estimate and its interpretation. Because
this is a convenience and not a claim, notebook 4.2 re-estimates both models with an
ordinal likelihood.

## Prior predictive check

Confirms that draws from the prior stay inside the admissible range before any data are
used. Cheap, since no likelihood is evaluated.

```{r}
#| label: prior-predictive
#| fig-cap: "Prior predictive draws, behavioral model."

m_prior <- brm(
  f_behav,
  data = dat_behav,
  family = gaussian(),
  prior = priors,
  sample_prior = "only",
  chains = 2,
  cores = 2,
  iter = 500,
  warmup = 250,
  seed = sampler$seed,
  backend = sampler$backend,
  file = file.path(
    model_dir,
    paste0("m_prior_behavioral", if (has_cmdstan) "_cs")
  ),
  file_refit = "on_change"
)

pp_check(m_prior, ndraws = 50) +
  coord_cartesian(xlim = c(-15, 25)) +
  geom_vline(xintercept = c(0, 10), linetype = "dashed") +
  labs(
    subtitle = "Dashed lines mark the admissible range [0, 10]"
  ) +
  theme_pubr()
```

# Estimation

Hamiltonian Monte Carlo in Stan through `brms`, four chains run in parallel and each
threaded across four cores where CmdStan is available. Fits are cached to `models/`, which
is untracked: each object is large, and regenerating it is a matter of re-running this
notebook.

```{r}
#| label: fit-behavioral

m_behav <- brm(
  f_behav,
  data = dat_behav,
  family = gaussian(),
  prior = priors,
  chains = sampler$chains,
  cores = sampler$cores,
  iter = sampler$iter,
  warmup = sampler$warmup,
  seed = sampler$seed,
  control = sampler$control,
  backend = sampler$backend,
  threads = sampler$threads,
  file = model_file("m_behavioral"),
  file_refit = "on_change"
)
```

```{r}
#| label: fit-attitudinal

m_attit <- brm(
  f_attit,
  data = dat_attit,
  family = gaussian(),
  prior = priors,
  chains = sampler$chains,
  cores = sampler$cores,
  iter = sampler$iter,
  warmup = sampler$warmup,
  seed = sampler$seed,
  control = sampler$control,
  backend = sampler$backend,
  threads = sampler$threads,
  file = model_file("m_attitudinal"),
  file_refit = "on_change"
)
```

# Convergence diagnostics

Rank-normalized $\hat{R} < 1.01$ and bulk and tail effective sample sizes above 1,000, as
committed to in @sec-model-spec. Reported as a summary that flags violations, rather than
as a wall of output.

```{r}
#| label: diagnostics-table

convergence_summary <- function(fit, label) {
  d <- as_draws_df(fit) |>
    summarise_draws("rhat", "ess_bulk", "ess_tail") |>
    filter(!is.na(rhat))

  np <- nuts_params(fit)
  n_divergent <- sum(np$Value[np$Parameter == "divergent__"])

  tibble(
    model = label,
    parameters = nrow(d),
    max_rhat = max(d$rhat),
    n_rhat_above_1.01 = sum(d$rhat > 1.01),
    min_ess_bulk = min(d$ess_bulk),
    n_ess_bulk_below_1000 = sum(d$ess_bulk < 1000),
    min_ess_tail = min(d$ess_tail),
    n_ess_tail_below_1000 = sum(d$ess_tail < 1000),
    divergences = n_divergent
  )
}

diagnostics <- bind_rows(
  convergence_summary(m_behav, "behavioral"),
  convergence_summary(m_attit, "attitudinal")
)

diagnostics
```

```{r}
#| label: diagnostics-worst

# the parameters that come closest to failing, per model
worst_parameters <- function(fit, label) {
  as_draws_df(fit) |>
    summarise_draws("rhat", "ess_bulk", "ess_tail") |>
    filter(!is.na(rhat)) |>
    slice_min(ess_bulk, n = 10) |>
    mutate(model = label, .before = 1)
}

bind_rows(
  worst_parameters(m_behav, "behavioral"),
  worst_parameters(m_attit, "attitudinal")
)
```

```{r}
#| label: fig-trace-behavioral
#| fig-cap: "Trace plots, population-level parameters carrying the estimands (behavioral model)."

mcmc_plot(
  m_behav,
  type = "trace",
  variable = "^b_(Intercept|RCo|RNone|typeexplicit)",
  regex = TRUE
)
```

```{r}
#| label: fig-trace-attitudinal
#| fig-cap: "Trace plots, population-level parameters carrying the estimands (attitudinal model)."

mcmc_plot(
  m_attit,
  type = "trace",
  variable = "^b_(Intercept|RCo|typeexplicit)",
  regex = TRUE
)
```

```{r}
#| label: fig-ppcheck
#| fig-cap: "Posterior predictive checks. Both outcomes are bounded and discrete; the Gaussian likelihood is a convenience, and notebook 4.2 re-estimates both with an ordinal likelihood."

ggarrange(
  pp_check(m_behav, ndraws = 50) + labs(subtitle = "Behavioral (tokens)"),
  pp_check(m_attit, ndraws = 50) + labs(subtitle = "Attitudinal (thermometer)"),
  ncol = 2,
  common.legend = TRUE
)
```

# Derived quantities

## How they are computed

Two properties of the specification make this cheap and exact, and both are worth stating
because they are the reason a small prediction grid suffices where G-computation over
seventy-odd thousand rows would otherwise be needed.

First, the model is linear with an identity link, so the population-level prediction *is*
the mean marginalized over the Normal varying effects. `re_formula = NA` therefore gives
the pooled quantity of @eq-marginal exactly, not an approximation to it.

Second, every additive covariate cancels from a contrast taken within a type. $\mathrm{AP}_t$,
$\Delta$, $\Delta_c$ and the in-group/out-group decomposition are consequently invariant to
the values at which the conjoint attributes are held. Only the *levels* $\mu_t(R)$ depend
on them, and those are reported at the reference level of each attribute, which -- because
`eupos` has `not_shown` as its reference -- is a combination that occurs in the data rather
than an extrapolation beyond it.

The game is the one covariate that does not cancel, since it is interacted with $R$ and
$T$. It is marginalized explicitly, with equal weights across the six game-by-round cells:
every respondent played each cell exactly once, so uniform weights are the design
distribution.

```{r}
#| label: prediction-grids

# reference values for the attributes that cancel from every contrast,
# kept as factors with the levels the model was fitted on
behav_reference <- dat_behav |>
  slice(1) |>
  select(eupos, cj_rel, cj_class, cj_sex, cj_age) |>
  mutate(across(everything(), \(x) factor(levels(x)[1], levels = levels(x))))

grid_behav <- expand_grid(
  R = factor(levels(dat_behav$R), levels = levels(dat_behav$R)),
  type = factor(levels(dat_behav$type), levels = levels(dat_behav$type)),
  game = factor(levels(dat_behav$game), levels = levels(dat_behav$game)),
  round = factor(levels(dat_behav$round), levels = levels(dat_behav$round))
) |>
  bind_cols(behav_reference)

grid_attit <- expand_grid(
  R = factor(levels(dat_attit$R), levels = levels(dat_attit$R)),
  type = factor(levels(dat_attit$type), levels = levels(dat_attit$type))
)

nrow(grid_behav)
nrow(grid_attit)
```

`ap_draws()` turns a matrix of posterior expected values into the per-draw quantities of
@eq-estimand and @eq-decomposition. Contrasts are formed on the matrix and only then
reshaped, which keeps memory proportional to draws times cells rather than draws times
rows.

```{r}
#| label: helper-ap-draws

# ep   : draws x grid-rows matrix of posterior expected values
# grid : the newdata that produced it
# by   : optionally one grid column, to compute the quantities separately per level
ap_draws <- function(ep, grid, by = NULL) {
  grid <- mutate(grid, .col = row_number())
  pieces <- if (is.null(by)) list(pooled = grid) else split(grid, grid[[by]])

  imap(pieces, function(g, key) {
    # marginal mean at relationship r for type t, averaged over the design cells
    mu <- function(r, t) {
      cols <- g$.col[g$R == r & g$type == t]
      rowMeans(ep[, cols, drop = FALSE])
    }

    out <- tibble(
      .draw = seq_len(nrow(ep)),
      mu_co_0 = mu("Co", "vote_anchored"),
      mu_out_0 = mu("Out", "vote_anchored"),
      mu_co_1 = mu("Co", "explicit"),
      mu_out_1 = mu("Out", "explicit")
    ) |>
      mutate(
        ap_0 = mu_co_0 - mu_out_0,
        ap_1 = mu_co_1 - mu_out_1,
        delta = ap_1 - ap_0
      )

    # the no-cue control exists in the games only
    if (any(g$R == "None")) {
      out <- out |>
        mutate(
          mu_none_0 = mu("None", "vote_anchored"),
          mu_none_1 = mu("None", "explicit"),
          baseline_gap = mu_none_1 - mu_none_0,
          ingroup_0 = mu_co_0 - mu_none_0,
          ingroup_1 = mu_co_1 - mu_none_1,
          outgroup_0 = mu_out_0 - mu_none_0,
          outgroup_1 = mu_out_1 - mu_none_1
        )
    }

    if (!is.null(by)) out[[by]] <- key
    out
  }) |>
    list_rbind()
}

# posterior median, 95% credible interval, and the probability of a positive value
summarise_quantities <- function(draws, by = NULL) {
  draws |>
    pivot_longer(
      -c(.draw, all_of(by)),
      names_to = "quantity",
      values_to = "value"
    ) |>
    group_by(across(all_of(c(by, "quantity")))) |>
    summarise(
      median = median(value),
      lower = quantile(value, 0.025),
      upper = quantile(value, 0.975),
      p_gt_0 = mean(value > 0),
      .groups = "drop"
    )
}
```

## Pooled quantities

@sec-behavioral fixes the reporting order, and it is not negotiable: the magnitude of
$\mathrm{AP}_t$ comes first, for both types, and only then the contrast between them. Both
partisan types discriminate substantially in favour of co-partisans; the question the paper
turns on is whether that discrimination differs by type.

```{r}
#| label: pooled-behavioral

ep_behav <- posterior_epred(
  m_behav,
  newdata = grid_behav,
  re_formula = NA
)

draws_behav <- ap_draws(ep_behav, grid_behav)

summarise_quantities(draws_behav) |>
  filter(quantity %in% c("mu_co_0", "mu_out_0", "mu_co_1", "mu_out_1", "ap_0", "ap_1"))
```

```{r}
#| label: pooled-behavioral-delta

# the type contrast, and the diagnostic that licenses its reading
summarise_quantities(draws_behav) |>
  filter(quantity %in% c("delta", "baseline_gap"))
```

```{r}
#| label: pooled-behavioral-decomposition

# what the no-cue control buys: in-group and out-group components of AP
summarise_quantities(draws_behav) |>
  filter(str_detect(quantity, "^(ingroup|outgroup)_"))
```

```{r}
#| label: pooled-attitudinal

ep_attit <- posterior_epred(
  m_attit,
  newdata = grid_attit,
  re_formula = NA
)

draws_attit <- ap_draws(ep_attit, grid_attit)

summarise_quantities(draws_attit)
```

## Stability across the two games

The order of the games was fixed rather than randomized: no respondent played the trust
game first, so the game factor and position in the session are perfectly confounded. The
game-specific estimates are therefore a stability check, not a contrast between allocation
regimes. An estimate that does not vary across the two games is evidence that the result
holds across two different allocation instruments and across sequence position at once. A
difference could not be attributed to the strategic element of the trust game rather than
to practice or fatigue, and is not interpreted as such.

```{r}
#| label: by-game

draws_by_game <- ap_draws(ep_behav, grid_behav, by = "game")

summarise_quantities(draws_by_game, by = "game") |>
  filter(quantity %in% c("ap_0", "ap_1", "delta"))
```

## Country-specific quantities

The country-level slope on $R \times T$ is what carries these: no other country-level term
does, since the intercept and the slope on $T$ enter both conditional means and cancel from
their difference, while the slope on $R$ shifts the level of polarization without touching
how it varies by type.

```{r}
#| label: country-behavioral

grid_behav_country <- expand_grid(
  country = factor(levels(dat_behav$country), levels = levels(dat_behav$country)),
  grid_behav
)

ep_behav_country <- posterior_epred(
  m_behav,
  newdata = grid_behav_country,
  re_formula = ~ (1 + R * type | country),
  allow_new_levels = FALSE
)

country_behav <- ap_draws(ep_behav_country, grid_behav_country, by = "country") |>
  summarise_quantities(by = "country") |>
  mutate(measurement = "Behavioral", .before = 1)

country_behav |>
  filter(quantity == "delta") |>
  arrange(desc(median))
```

```{r}
#| label: country-attitudinal

grid_attit_country <- expand_grid(
  country = factor(levels(dat_attit$country), levels = levels(dat_attit$country)),
  grid_attit
)

ep_attit_country <- posterior_epred(
  m_attit,
  newdata = grid_attit_country,
  re_formula = ~ (1 + R * type | country),
  allow_new_levels = FALSE
)

country_attit <- ap_draws(ep_attit_country, grid_attit_country, by = "country") |>
  summarise_quantities(by = "country") |>
  mutate(measurement = "Attitudinal", .before = 1)

country_attit |>
  filter(quantity == "delta") |>
  arrange(desc(median))
```

```{r}
#| label: fig-country-delta
#| fig-cap: "Country-specific type gap $\\Delta_c$, both measures. Points are posterior medians, bars 95% credible intervals."
#| fig-dpi: 300

bind_rows(country_behav, country_attit) |>
  filter(quantity == "delta") |>
  mutate(country = fct_reorder(country, median)) |>
  ggplot(aes(x = median, y = country, colour = measurement)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_pointrange(
    aes(xmin = lower, xmax = upper),
    position = position_dodge(width = 0.5),
    size = 0.3
  ) +
  labs(
    x = expression(Delta[c]),
    y = NULL,
    colour = "Measurement"
  ) +
  theme_pubr()
```

## Party-specific quantities

The same construction at the anchor level. Note what it can and cannot show: the anchor
grouping carries slopes on $R$ and on $T$ but not on their interaction, so $\mathrm{AP}_{a,t}$
varies by party while $\Delta_a$ does not. That is the homogeneity assumption recorded in
check 3, made visible rather than left implicit.

```{r}
#| label: anchor-behavioral

# round cancels from every contrast, so one round level suffices here
grid_behav_anchor <- expand_grid(
  anchor = factor(levels(dat_behav$anchor), levels = levels(dat_behav$anchor)),
  grid_behav |> filter(round == levels(dat_behav$round)[1])
)

ep_behav_anchor <- posterior_epred(
  m_behav,
  newdata = grid_behav_anchor,
  re_formula = ~ (1 + R + type | anchor),
  allow_new_levels = FALSE,
  ndraws = min(2000, ndraws(m_behav))
)

anchor_behav <- ap_draws(ep_behav_anchor, grid_behav_anchor, by = "anchor") |>
  summarise_quantities(by = "anchor") |>
  mutate(measurement = "Behavioral", .before = 1)

anchor_behav |>
  filter(quantity %in% c("ap_0", "ap_1")) |>
  slice_max(median, n = 10)
```

```{r}
#| label: anchor-attitudinal

grid_attit_anchor <- expand_grid(
  anchor = factor(levels(dat_attit$anchor), levels = levels(dat_attit$anchor)),
  grid_attit
)

ep_attit_anchor <- posterior_epred(
  m_attit,
  newdata = grid_attit_anchor,
  re_formula = ~ (1 + R + type | anchor),
  allow_new_levels = FALSE,
  ndraws = min(2000, ndraws(m_attit))
)

anchor_attit <- ap_draws(ep_attit_anchor, grid_attit_anchor, by = "anchor") |>
  summarise_quantities(by = "anchor") |>
  mutate(measurement = "Attitudinal", .before = 1)

anchor_attit |>
  filter(quantity %in% c("ap_0", "ap_1")) |>
  slice_max(median, n = 10)
```

## Variance decomposition

The share of variance attributable to respondents, anchors and countries, as promised in
@sec-appendix-model.

```{r}
#| label: variance-decomposition

variance_shares <- function(fit, label) {
  vc <- VarCorr(fit)

  # intercept SD per grouping level; the residual block carries a single
  # unnamed row rather than an "Intercept" one
  intercept_sd <- map_dbl(vc, function(g) {
    if (is.null(g$sd)) {
      return(NA_real_)
    }
    if ("Intercept" %in% rownames(g$sd)) g$sd["Intercept", "Estimate"] else g$sd[1, "Estimate"]
  })

  tibble(
    model = label,
    component = if_else(names(intercept_sd) == "residual__", "residual", names(intercept_sd)),
    variance = as.numeric(intercept_sd)^2
  ) |>
    filter(!is.na(variance)) |>
    mutate(share = variance / sum(variance)) |>
    arrange(desc(share))
}

bind_rows(
  variance_shares(m_behav, "behavioral"),
  variance_shares(m_attit, "attitudinal")
)
```

# Save derived quantities

The fits stay in the untracked `models/` directory; the summarized quantities are written
to `data/03_final/`, so that the figures in @sec-attitudinal and @sec-behavioral, and
notebook 4.2, can be built without refitting.

```{r}
#| label: save-quantities

model_quantities <- list(
  meta = list(
    fitted_at = Sys.time(),
    smoke_test = smoke_test,
    sampler = sampler,
    n_behavioral = nrow(dat_behav),
    n_attitudinal = nrow(dat_attit),
    countries_behavioral = nlevels(dat_behav$country),
    countries_attitudinal = nlevels(dat_attit$country)
  ),
  diagnostics = diagnostics,
  pooled = bind_rows(
    summarise_quantities(draws_behav) |> mutate(measurement = "Behavioral"),
    summarise_quantities(draws_attit) |> mutate(measurement = "Attitudinal")
  ),
  by_game = summarise_quantities(draws_by_game, by = "game"),
  by_country = bind_rows(country_behav, country_attit),
  by_anchor = bind_rows(anchor_behav, anchor_attit),
  checks = list(
    baseline_nocue = baseline_nocue,
    anchor_feasibility = anchor_feasibility,
    battery_coverage = battery_coverage
  )
)

saveRDS(
  model_quantities,
  file = here(
    "data",
    "03_final",
    "model_quantities.rds"
  )
)
```

# Session Info

```{r}
#| label: session-info

session_info()
```

# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
