---
title: |
  Comparative AP Literature Summary
subtitle: |
  How does reported partisan attachement influence affective polarization?: A comparative study of 25 European democracies
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
bibliography: ../references.bib
biblio-style: apsr
---

# Comparative Affective Polarization Studies: Overview Table

## Coding Guide

Each study is coded on four dimensions. Definitions are fixed below — apply them consistently across all entries and use the **Notes** column only for genuine ambiguities that cannot be resolved by the rules given here.

### 1. AP Target
*What is the object of polarized affect — parties/elites, or fellow citizens?*

| Code | Definition | Typical measure |
|:-----|:-----------|:----------------|
| **Vertical** | Affect directed at **parties or political elites** | Feeling thermometers toward parties or leaders |
| **Horizontal** | Affect or behavior directed at **fellow citizens** with different partisan identities | Social distance, discrimination tasks, allocation games |
| **Both** | Study **explicitly reports both** vertical and horizontal measures as distinct dimensions | Combined survey + behavioral experiment |

> **Key distinction:** Feeling thermometers toward *parties* = Vertical. Feeling thermometers toward *party supporters* or behavioral tasks involving *citizens* = Horizontal. If a study uses party thermometers and treats them as a proxy for citizen-level polarization without measuring it directly, code as **Vertical** and flag in Notes.


### 2. AP Manifestation
*How is AP expressed and operationalized?*

| Code | Definition | Typical measure |
|:-----|:-----------|:----------------|
| **Attitudinal** | **Self-reported evaluations** of parties, leaders, or partisan outgroups | Feeling thermometers, like-dislike scales, social distance items |
| **Behavioral** | **Actions** in incentivized or experimental settings | Dictator games, trust games, conjoint discrimination tasks |
| **Both** | Study reports and analyzes **both** attitudinal and behavioral measures | Survey thermometers + allocation experiment in same study |


### 3. Sample
*Who is included in the analysis as a "partisan"?*

This is the dimension most consequential for the paper's argument. Apply codes strictly.

| Code | Definition |
|:-----|:-----------|
| **Identifiers** | Restricted to respondents who **self-report psychological attachment** to a party (explicit partisans only) |
| **Voters** | Restricted to respondents who **report a vote choice**, regardless of felt attachment; includes implicit partisans |
| **Mixed** | Combines identifiers and voters under a shared partisan anchor, or uses a strategy encompassing both groups |
| **All** | **No partisan restriction** — non-partisans included in the analysis |

**Decision rules:**

- Reiljan's API on CSES → **Identifiers** (restricted to respondents naming a liked party)
- Wagner's spread measure on full sample → **All**
- Any anchor that accepts vote choice without requiring attachment → **Voters** or **Mixed**
- When unclear, note the exact operationalization in Notes


### 4. Data Source
*What dataset does the study rely on?*

Record the primary dataset by its common name. This column reveals which studies draw on the **same underlying data** — a key part of the paper's framing regarding the field's dependence on a small number of datasets.

Common sources: `CSES`, `ESS`, `Original survey`, `Multiple`


## Table

| Study  | AP Target | AP Manifestation | Sample | N Countries | Data Source | Notes |
|:------:|:---------:|:----------------:|:------:|:-----------:|:------------|:------|
| @iyengar2012affect | Both | Attitudinal | Identifiers | 2 | Multiple (ANES, YouGov, @almond_verba_1959_civic_culture, Blair Center Election Study, AP-Yahoo! News 2008 election study) | Thermometers toward parties (V) + social trust/discrimination items (H); US & UK |
| @Westwood2018tie | Horizontal | Both | Voters | 4 | Original survey | Trust games; US, UK, Belgium, Spain |
| @reiljan2020fear  | Vertical | Attitudinal | Mixed | 23 | CSES | Introduces API |
| @gidron2020american  | Vertical | Attitudinal | Identifiers | 20 | Original survey | US, UK, Israel; party thermometers |
| @wagner2021affective  | Vertical | Attitudinal | All | 51 | CSES | Spread-of-scores & distance measures; includes non-partisans |
| @Helbling2020social | Horizontal | Attitudinal | Voters | 2 | Original survey | Social avoidance; Austria & Germany |
| @kekkonen2021affective  | Vertical | Attitudinal | Mixed | 1 | Original survey | Finland; citizen-targeted affect |
| @Harteveld2024consequences  | Both | Both | Mixed | 9 | Original Survey (CONAP) | Operationalize in- and out-group via most- and least liked party supporters in like/dislike items |
| @areal2024verticalvshorizontal  | Both | Attitudinal | Voters | 9 | Multiple | Explicit V vs H comparison; correlation ~0.4–0.7 by country |
| @reiljan2024patterns  | Vertical | Attitudinal | Voters | 40 | CSES | Distinguish vertical AP into LAP and PAP; analyses based on identifiers only as robustness |
| @hahm2024divided  | Horizontal | Behavioral | Identifiers | 25 | Original Survey | Conjoint dictator + trust games |
| @Hernandez2021affective | Vertical | Attitudinal | All  | 42 | CSES |  |
| @boxell2024cross | Vertical | Attitudinal | Identifiers  | 12 | Multiple |  |
| @lauka2018mass | Vertical | Attitudinal | Voters  | 48 | CSES |  |
| @bradley2022ethnic | Vertical | Attitudinal | All  | 48 | CSES | Investigate identifiers with API and all with Wagner |
| @tichelbaecker2023we | Both | Attitudinal | Mixed | 10 | Original Survey | |
| @garzia2023affective | Both | Attitudinal | All | 18 | Original Survey | Investigate identifiers with API and all with Wagner |
| @Gidron2023who | Vertical | Attitudinal | Identifiers | 20 | CSES | |
| @Horne2023way | Vertical | Attitudinal | Identifiers | 19 | CSES | |
| @harteveld2021ticking | Vertical | Attitudinal | Identifiers | 40 | CSES | |
| @Comellas2023ideological | Horizontal | Attitudinal | Identifiers | 5 | TRI-POL | Study bloc AP |
| @Fuller2022populism | Both | Both | Attitudinal | 9 | Original Survey | Main analysis uses scaling method on thermo items, further analyses use identifiers with vertical/attitudinal (thermo) measures and horizontal/attitudinal social distance items |
| @harteveld2022affective | Vertical | Attitudinal | Mixed | 28 | CSES | Includes a second analysis with longitudinal data from Netherlands (LISS) |
| @Torcal2022affective | Both | Attitudinal | All | 48 | CSES, CNEP, E-DEM | Code most liked party as in-group |
| @adams2023can | Vertical | Attitudinal | Identifiers | 20 | CSES | | 
| @renstrom2021intergroup | Both | Attitudinal | All | 1 | Original Survey | Code most liked party as in-group |
| @renstrom2022protecting | Both | Attitudinal | All  | 1 | Original Survey | Code most liked party as in-group |
| @hahm2023divided | Horizontal | Behavioral | Identifiers | 25 | Original Survey | Study AP around EU attachement | 
| @kekkonen2022puzzles | Both | Attitudinal | Identifiers | 1 | Original Survey | Compare single and multiple ingroup operationalizations |
| @bettarelli2023regional | Vertical | Attitudinal | All | 30 | CSES | Compare wagner spread aggregated across country and NUTS regions |
| @ryan2023exploring | Vertical | Attitudinal | All | 5 | CSES | Nordic countries |
| @reiljan2021ideological | Vertical | Attitudinal | All | 1 | SOM | Code most liked party as in-group |
| @Knudsen2021affective | Both | Attitudinal | Voters | 2 | Original Survey | Calls for differential treatment of vertical and horizontal AP |
| @segovia2022affective | Vertical | Attitudinal | All | 1 | CEP | Chile case study |
| @Orhan2022relationship | Vertical | Attitudinal | Identifiers | 53 | CSES | |
| @berntzen2023consequencesofaffective | Both | Attitudinal | All | 2 | NCP, YouGov | |
| @Torcal2023social | Vertical | Attitudinal  | ALL | 1 | CIUPANEL, E-DEM | Spain |
| @Martini2019trust | Vertical | Behavioral | Identifiers | 2 | Original Survey | notes |
| @ruggeri2021general | trgt | mfst | smpl | n | data | notes | 
| @Ward2019partisan | Vertical | Attitudinal | All | 34 | CSES | notes | 

continue w/ nir 2012

to-do table:
  - browse "top" journals (APSR, AJPS, CPS, BJPS, JOP) for (comparative/multiparty/Europe) AP papers
  - change sample column to ingroup strategy -> highlight identity/attachement primacy in literature
  - ingroup via most liked party as robustness (or correlation with anchors)


## To-Do / Studies Still to Code

- [ ] Boxell et al. (2024) — cross-national AP trends
- [ ] Lauka, Mann & Frees (2018) — early European cross-national AP
- [ ] Ward & Tavits (2019) — party cues and AP
- [ ] Knudsen (2020) — Norway, horizontal AP
- [ ] Renström et al. (2022) — Sweden
- [ ] Bradley & Chauchard (2022) — cross-national measure correlations
- [ ] Hobolt, Leeper & Tilley (2020) — Brexit & AP, UK
- [ ] Hernández et al. (2021) — electoral cycle effects on AP
- [ ] Garzia et al. (2023) — affective dealignment, multiparty context
- [ ] Torcal & Comellas (2025) — operationalization in multiparty systems




## Collection of potentially useful quotes

> There are nevertheless key aspects that still generate debate. How important are partisan identities for affective polarization? What is relationship between vertical (leader, party) and horizontal (citizen) affective polarization? [@IyengarWagner2025Conceptualizing, S.47]

@kekkonen2022puzzles high correlation between single ingroup and multiple ingroup

@Torcal2023social for recall maximizing ingroup strategy 






---
title: "Partisanship, Partisan Identity and De-/Realignment in Europe"
subtitle: "A structured bibliography of the post-1945 debate (EU-27 excl. MT/CY/LU, plus the UK)"
format: html
---

## How to read this list

Entries are grouped by **phase of the debate** and then by **country/region**, so the list doubles as a
map of the argument rather than a flat reference dump.

**DOI column convention.** DOIs marked as links were verified during compilation. Entries with `—` are
canonical references whose bibliographic details are given from standard citation practice but whose DOI
was **not** individually verified — check these against Crossref or your library before citing. Books
generally have no DOI and are marked `[book]`.

---

## 1. Foundations: importing the Michigan model into Europe (1950s–1970s)

The founding question: is "party identification", developed for a two-party US context, a meaningful
concept in European multiparty systems with strong cleavage structures and disciplined parties?

| Reference | Scope | Contribution to the debate | DOI |
|---|---|---|---|
| Campbell, A., Converse, P. E., Miller, W. E. & Stokes, D. E. (1960). *The American Voter.* Wiley. | US (baseline) | Establishes party identification as an enduring affective attachment and the "funnel of causality". | [book] |
| Lipset, S. M. & Rokkan, S. (1967). Cleavage structures, party systems, and voter alignments. In *Party Systems and Voter Alignments.* Free Press. | Western Europe | The freezing hypothesis: party systems reflect frozen social cleavages of the 1920s. The reference point for all later de-/realignment claims. | [book] |
| Butler, D. & Stokes, D. (1969). *Political Change in Britain.* Macmillan. | UK | First systematic European application of the Michigan model; introduces cohort/socialisation reasoning to British electoral change. | [book] |
| Converse, P. E. (1969). Of time and partisan stability. *Comparative Political Studies*, 2(2), 139–171. | Comparative | Learning model: partisanship strengthens with democratic experience; foundational for later CEE work. | — |
| Budge, I., Crewe, I. & Farlie, D. (eds.) (1976). *Party Identification and Beyond.* Wiley. | Comparative | The first collective stress-test of the concept outside the US; still the canonical statement of the sceptical position. | [book] |
| Thomassen, J. (1976). Party identification as a cross-national concept: Its meaning in the Netherlands. In Budge et al. (eds.), *Party Identification and Beyond.* | Netherlands | The classic negative finding: in the Dutch case party ID appears to follow rather than cause vote choice. | [book] |
| Rose, R. & Mishler, W. (1998). Negative and positive party identification in post-communist countries. *Electoral Studies*, 17(2), 217–234. | CEE | Introduces negative party identification as a distinct orientation in new democracies. | — |

---

## 2. The dealignment thesis (1980s–1990s)

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Dalton, R. J. (1984). Cognitive mobilization and partisan dealignment in advanced industrial democracies. *The Journal of Politics*, 46(1), 264–284. | 8 nations | The cognitive-mobilisation mechanism: rising skills and resources reduce the need for partisan cues, producing "apartisans". | [10.2307/2130444](https://doi.org/10.2307/2130444) |
| Dalton, R. J., Flanagan, S. C. & Beck, P. A. (eds.) (1984). *Electoral Change in Advanced Industrial Democracies: Realignment or Dealignment?* Princeton UP. | Comparative | Frames the de- vs. realignment dichotomy that organises the field to this day. | [book] |
| Crewe, I. & Denver, D. (eds.) (1985). *Electoral Change in Western Democracies: Patterns and Sources of Electoral Volatility.* Croom Helm. | Western Europe | Early systematic treatment of volatility as the behavioural signature of dealignment. | [book] |
| Schmitt, H. (1989). On party attachment in Western Europe and the utility of Eurobarometer data. *West European Politics*, 12(2), 122–139. | EC member states | Establishes the Eurobarometer partisanship series and its measurement pitfalls. | — |
| Richardson, B. M. (1991). European party loyalties revisited. *American Political Science Review*, 85(3), 751–775. | Western Europe | Rehabilitates partisanship in Europe against the sceptics; distinguishes party loyalty from vote. | — |
| Franklin, M., Mackie, T. & Valen, H. (eds.) (1992). *Electoral Change: Responses to Evolving Social and Attitudinal Structures in Western Countries.* Cambridge UP. | 16 countries | The definitive test of declining cleavage voting; shows social-structural effects weakening across Western democracies. | [book] |
| Schmitt, H. & Holmberg, S. (1995). Political parties in decline? In Klingemann, H.-D. & Fuchs, D. (eds.), *Citizens and the State.* Oxford UP. | Western Europe | Nuanced verdict: attachment declines, but unevenly and less than the "crisis of parties" narrative implies. | [book] |
| Pedersen, M. N. (1979). The dynamics of European party systems: Changing patterns of electoral volatility. *European Journal of Political Research*, 7(1), 1–26. | Western Europe | The Pedersen index; the standard aggregate measure in all subsequent volatility work. | — |

---

## 3. Consolidation and contestation (2000s)

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Dalton, R. J. & Wattenberg, M. P. (eds.) (2000). *Parties without Partisans: Political Change in Advanced Industrial Democracies.* Oxford UP. | Comparative | The canonical statement of the dealignment consensus; Dalton's chapter documents cross-national decline in party ID. | [book] |
| Thomassen, J. (ed.) (2005). *The European Voter: A Comparative Study of Modern Democracies.* Oxford UP. | 6 countries, long series | Long-run national election-study evidence; finds decline weaker and more country-specific than assumed. | [book] |
| Berglund, F., Holmberg, S., Schmitt, H. & Thomassen, J. (2005). Party identification and party choice. In Thomassen (ed.), *The European Voter.* | 6 countries | Directly addresses whether party ID is causally prior to vote in European systems. | [book] |
| Holmberg, S. (2007). Partisanship reconsidered. In Dalton, R. J. & Klingemann, H.-D. (eds.), *The Oxford Handbook of Political Behavior.* | Comparative | Balance sheet of the concept's travels; defends its utility. | [book] |
| Dalton, R. J. & Weldon, S. (2007). Partisanship and party system institutionalization. *Party Politics*, 13(2), 179–196. | Comparative incl. CEE | Links individual partisanship to system-level institutionalisation. | — |
| Bartle, J. & Bellucci, P. (eds.) (2009). *Political Parties and Partisanship: Social Identity and Individual Attitudes.* Routledge. | Comparative | The turning point toward a social-identity reading of European partisanship. | [book] |
| Kroh, M. & Selb, P. (2009). Inheritance and the dynamics of party identification. *Political Behavior*, 31(4), 559–574. | Germany (SOEP) | Reconciles traditionalist and revisionist views: inherited attachments behave traditionally, acquired ones revisionist. | [10.1007/s11109-009-9084-2](https://doi.org/10.1007/s11109-009-9084-2) |
| Schmitt-Beck, R., Weick, S. & Christoph, B. (2006). Shaky attachments: Individual-level stability and change of partisanship among West German voters, 1984–2001. *European Journal of Political Research*, 45(4), 581–608. | Germany (SOEP panel) | Panel evidence on the instability of German partisanship. | — |
| Marsh, M. (2006). Party identification in Ireland: An insecure anchor for a floating party system. *Electoral Studies*, 25(3), 489–508. | Ireland | Key national case: weak party ID in a system without a classic left–right cleavage. | — |
| Neundorf, A., Stegmueller, D. & Scotto, T. J. (2011). The individual-level dynamics of bounded partisanship. *Public Opinion Quarterly*, 75(3), 458–482. | UK, Germany | "Bounded partisanship": stability within blocs, movement across parties. | — |
| van Biezen, I., Mair, P. & Poguntke, T. (2012). Going, going, ... gone? The decline of party membership in contemporary Europe. *European Journal of Political Research*, 51(1), 24–56. | Europe-wide | The organisational counterpart to attitudinal dealignment. | — |
| Mair, P. (2013). *Ruling the Void: The Hollowing of Western Democracy.* Verso. | Western Europe | The influential normative synthesis of disengagement and party decline. | [book] |

---

## 4. Volatility, dealignment and its consequences (2010s–today)

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Dassonneville, R. (2013). Questioning generational replacement: An age, period and cohort analysis of electoral volatility in the Netherlands, 1971–2010. *Electoral Studies*, 32(1), 37–47. | Netherlands | APC decomposition: volatility is a period rather than purely cohort phenomenon. | — |
| Dassonneville, R., Hooghe, M. & Vanhoutte, B. (2012). Age, period and cohort effects in the decline of party identification in Germany, 1992–2009. *German Politics*, 21(2), 209–227. | Germany | Central APC contribution to the German dealignment dispute (with Dalton's rejoinder). | [10.1080/09644008.2012.679659](https://doi.org/10.1080/09644008.2012.679659) |
| Dassonneville, R. & Hooghe, M. (2018). Indifference and alienation: Diverging dimensions of electoral dealignment in Europe. *Acta Politica*, 53(1), 1–23. | Europe | Distinguishes two behavioural routes out of alignment. | — |
| Dassonneville, R. & Hooghe, M. (2017). Economic indicators and electoral volatility: Economic effects on electoral volatility in Western Europe, 1950–2013. *Comparative European Politics*, 15(6), 919–943. | 16 countries, 1950–2013 | Macro-economic drivers of aggregate dealignment. | — |
| Dejaeghere, Y. & Dassonneville, R. (2017). A comparative investigation into the effects of party-system variables on party switching using individual-level data. *Party Politics*, 23(2), 110–123. | Comparative | System-level moderators of individual switching. | [10.1177/1354068815576294](https://doi.org/10.1177/1354068815576294) |
| Dassonneville, R., Fournier, P. & Somer-Topcu, Z. (2023). Partisan attachments in a multidimensional space. *West European Politics*, 46(4). | Western Europe | Supply-side account: ideologically consistent parties across dimensions generate clearer brands and more party ID. | [10.1080/01402382.2022.2087387](https://doi.org/10.1080/01402382.2022.2087387) |
| Garzia, D., Ferreira da Silva, F. & De Angelis, A. (2022). Partisan dealignment and the personalisation of politics in West European parliamentary democracies, 1961–2018. *West European Politics*, 45(2), 311–334. | 14 countries, 109 election surveys | The key demonstration that dealignment causes rising leader effects. | [10.1080/01402382.2020.1845941](https://doi.org/10.1080/01402382.2020.1845941) |
| Chiaramonte, A. & Emanuele, V. (2017). Party system volatility, regeneration and de-institutionalization in Western Europe (1945–2015). *Party Politics*, 23(4), 376–388. | 20 countries, 1945–2015 | The definitive long-run aggregate volatility series for Western Europe. | — |
| van der Meer, T. W. G., van Elsas, E., Lubbe, R. & van der Brug, W. (2015). Are volatile voters erratic, whimsical or seriously picky? *Party Politics*, 21(1), 100–114. | Netherlands | Disaggregates "the volatile voter"; volatility is largely intra-bloc and reasoned. | — |
| Fieldhouse, E., Green, J., Evans, G., Mellon, J., Prosser, C., Schmitt, H. & van der Eijk, C. (2020). *Electoral Shocks: The Volatile Voter in a Turbulent World.* Oxford UP. | UK | Ties record volatility to a sequence of political shocks in a dealigned electorate. | [book] |
| Kroh, M. (2014). *Documentation of Sample Sizes and Panel Attrition in the SOEP* / related SOEP-based partisanship work. | Germany | Data infrastructure underpinning much German panel work on partisanship. | — |

---

## 5. Realignment: new cleavages, education and the sociocultural divide

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Kriesi, H., Grande, E., Lachat, R., Dolezal, M., Bornschier, S. & Frey, T. (2006). Globalization and the transformation of the national political space: Six European countries compared. *European Journal of Political Research*, 45(6), 921–956. | AT, CH, DE, FR, NL, UK | The integration–demarcation divide; the foundational realignment statement. | — |
| Kriesi, H. et al. (2008). *West European Politics in the Age of Globalization.* Cambridge UP. | 6 countries | Book-length elaboration with party-supply and mass-demand evidence. | [book] |
| Bornschier, S. (2010). *Cleavage Politics and the Populist Right.* Temple UP. | Western Europe | Links the new cultural divide to radical-right party formation. | [book] |
| Stubager, R. (2009). Education-based group identity and consciousness in the authoritarian–libertarian value conflict. *European Journal of Political Research*, 48(2), 204–233. | Denmark | Early demonstration that education functions as a *group identity*, not just a predictor. | — |
| Oesch, D. & Rennwald, L. (2018). Electoral competition in Europe's new tripolar political space: Class voting for the left, centre-right and radical right. *European Journal of Political Research*, 57(4), 783–807. | Western Europe | Tripolar class realignment; the standard class-schema reference. | — |
| Hooghe, L. & Marks, G. (2018). Cleavage theory meets Europe's crises: Lipset, Rokkan, and the transnational cleavage. *Journal of European Public Policy*, 25(1), 109–135. | Europe | The transnational cleavage thesis. | — |
| Marks, G., Attewell, D., Hooghe, L., Rovny, J. & Steenbergen, M. (2023). The social bases of political parties: A new measure and survey. *British Journal of Political Science*, 53(1), 249–260. | Europe | New party-structuration measure enabling inter-temporal comparison of cleavage strength. | — |
| Bornschier, S., Häusermann, S., Zollinger, D. & Colombo, C. (2021). How "us" and "them" relates to voting behavior: Group identities and voting for the new left and radical right. *Comparative Political Studies*, 54(12), 2087–2122. | Switzerland/Western Europe | Brings explicit group-identity measurement into cleavage research. | — |
| Zollinger, D. (2024). Cleavage identities in voters' own words: Harnessing open-ended survey responses. *American Journal of Political Science*, 68(1), 139–159. | Switzerland | Inductive identity measurement; a methodological bridge between cleavage and identity literatures. | — |
| Hooghe, L. & Marks, G. (2026). How does the education cleavage stack up against the classic cleavages of the past? *West European Politics*, 49(3). | Europe + US | Argues GAL/TAN parties are as socially structured as classic cleavage parties were in the 1950s. | [10.1080/01402382.2025.2452789](https://doi.org/10.1080/01402382.2025.2452789) |
| Zollinger, D. & Attewell, D. (2025). Updating cleavage theory for the twenty-first century. *West European Politics*. | Western democracies | Introduction to the current state-of-the-art special issue on cleavage politics. | [10.1080/01402382.2025.2560781](https://doi.org/10.1080/01402382.2025.2560781) |
| Ford, R. & Jennings, W. (2020). The changing cleavage politics of Western Europe. *Annual Review of Political Science*, 23, 295–314. | Western Europe | The standard review linking education expansion, migration, ageing and geography to realignment. | [10.1146/annurev-polisci-052217-104957](https://doi.org/10.1146/annurev-polisci-052217-104957) |
| Emanuele, V. & Chiaramonte, A. (2018). A growing impact of new parties: Myth or reality? *Party Politics*, 24(5), 475–487. | Western Europe | Distinguishes realignment through new parties from mere volatility. | — |

---

## 6. The identity turn: partisanship as social identity

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Bankert, A., Huddy, L. & Rosema, M. (2017). Measuring partisanship as a social identity in multi-party systems. *Political Behavior*, 39(1), 103–132. | NL, SE, UK | The eight-item partisan identity scale; establishes measurement equivalence across multiparty systems. | [10.1007/s11109-016-9349-5](https://doi.org/10.1007/s11109-016-9349-5) |
| Huddy, L., Bankert, A. & Davies, C. (2018). Expressive versus instrumental partisanship in multiparty European systems. *Political Psychology*, 39(S1), 173–199. | UK, SE, NL, IT | Shows expressive partisanship outperforms ideological intensity in Europe. | [10.1111/pops.12482](https://doi.org/10.1111/pops.12482) |
| Bakker, B. N., Hopmann, D. N. & Persson, M. (2015). Personality traits and party identification over time. *European Journal of Political Research*, 54(1), 197–215. | Germany (panel) | Dispositional origins of the direction, strength and stability of party ID. | [10.1111/1475-6765.12070](https://doi.org/10.1111/1475-6765.12070) |
| Westwood, S. J., Iyengar, S., Walgrave, S., Leonisio, R., Miller, L. & Strijbis, O. (2018). The tie that divides: Cross-national evidence of the primacy of partyism. *European Journal of Political Research*, 57(2), 333–354. | UK, BE, ES (+US) | Partisan discrimination exceeds religious, linguistic, ethnic and regional discrimination. | [10.1111/1475-6765.12228](https://doi.org/10.1111/1475-6765.12228) |
| Hobolt, S. B., Leeper, T. J. & Tilley, J. (2021). Divided by the vote: Affective polarization in the wake of the Brexit referendum. *British Journal of Political Science*, 51(4), 1476–1493. | UK | Opinion-based identities can generate polarisation as intense as partisanship. | [10.1017/S0007123420000125](https://doi.org/10.1017/S0007123420000125) |
| Rekker, R., Keijsers, L., Branje, S. & Meeus, W. (2019). The formation of party preference in adolescence and early adulthood. *Young*, 27(1). | Netherlands | Cohort-sequential panel pinpointing the formative window for party preference. | [10.1177/1103308818757037](https://doi.org/10.1177/1103308818757037) |
| Slothuus, R. & Bisgaard, M. (2021). How political parties shape public opinion in the real world. *American Journal of Political Science*, 65(4), 896–911. | Denmark | Quasi-experimental panel around two real party position reversals; strong evidence for partisan-led opinion. | [10.1111/ajps.12550](https://doi.org/10.1111/ajps.12550) |
| Bisgaard, M. & Slothuus, R. (2018). Partisan elites as culprits? How party cues shape partisan perceptual gaps. *American Journal of Political Science*, 62(2), 456–469. | Denmark | Elite cues as the mechanism behind partisan factual disagreement. | — |

### 6a. Negative partisanship

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Medeiros, M. & Noël, A. (2014). The forgotten side of partisanship: Negative party identification in four Anglo-American democracies. *Comparative Political Studies*, 47(7), 1022–1046. | incl. UK | Re-establishes negative party ID as a distinct construct. | — |
| Mayer, S. J. (2017). How negative partisanship affects voting behavior in Europe. *Research & Politics*, 4(1). | 17 European PR systems (CSES) | Negative party ID independently raises turnout (~9 pp) and shapes vote choice. | [10.1177/2053168016686636](https://doi.org/10.1177/2053168016686636) |
| Anderson, C. D., McGregor, R. M. & Stephenson, L. B. (2022). Us versus them: Do the rules of the game encourage negative partisanship? *European Journal of Political Research*, 61(4), 1060–1079. | CSES | Institutional origins of negative partisanship, moderated by sophistication. | [10.1111/1475-6765.12485](https://doi.org/10.1111/1475-6765.12485) |
| Meléndez, C. & Rovira Kaltwasser, C. (2021). Negative partisanship towards the populist radical right and democratic resilience in Western Europe. *Democratization*, 28(5), 949–969. | Western Europe | Negative partisanship as a possible brake on democratic erosion. | [10.1080/13510347.2021.1883002](https://doi.org/10.1080/13510347.2021.1883002) |
| Mayer, S. J. & Russo, L. (2024). What one is not: A new scale to measure negative party identity in multi-party systems. *Quality & Quantity*, 58, 2887–2906. | DE, IT | Measurement instrument for negative party identity. | — |
| Areal, J. (2022). "Them" without "us": Negative identities and affective polarization in Brazil. *Politics, Groups, and Identities* / *Political Research Exchange*. | Brazil (comparative relevance) | Shows negative identities can operate without positive attachment. | [10.1080/2474736X.2022.2117635](https://doi.org/10.1080/2474736X.2022.2117635) |

---

## 7. Affective polarisation in European multiparty systems

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Reiljan, A. (2020). "Fear and loathing across party lines" (also) in Europe. *European Journal of Political Research*, 59(2), 376–396. | 22 European democracies | The Affective Polarisation Index; CEE and Southern Europe exceed the US. | [10.1111/1475-6765.12351](https://doi.org/10.1111/1475-6765.12351) |
| Wagner, M. (2021). Affective polarization in multiparty systems. *Electoral Studies*, 69, 102199. | 51 countries, 166 elections | The standard individual-level measurement toolkit. | [10.1016/j.electstud.2020.102199](https://doi.org/10.1016/j.electstud.2020.102199) |
| Gidron, N., Adams, J. & Horne, W. (2020). *American Affective Polarization in Comparative Perspective.* Cambridge UP. | 20 Western publics | Macro-correlates: unemployment, inequality, elite cultural conflict, majoritarian institutions. | [10.1017/9781108914123](https://doi.org/10.1017/9781108914123) |
| Gidron, N., Adams, J. & Horne, W. (2023). Who dislikes whom? *British Journal of Political Science*, 53(3), 997–1015. | 20 Western democracies | Dyadic approach; elite policy conflict and coalition partnership drive party-pair affect. | [10.1017/S0007123422000394](https://doi.org/10.1017/S0007123422000394) |
| Horne, W., Adams, J. & Gidron, N. (2023). The way we were: How histories of co-governance alleviate partisan hostility. *Comparative Political Studies*, 56(3), 299–325. | Western democracies | Coalition history as a structural depolarising force. | [10.1177/00104140221100197](https://doi.org/10.1177/00104140221100197) |
| Harteveld, E. (2021). Ticking all the boxes? A comparative study of social sorting and affective polarization. *Electoral Studies*, 72, 102337. | 40 countries, 119 elections | Social sorting travels beyond the US. | [10.1016/j.electstud.2021.102337](https://doi.org/10.1016/j.electstud.2021.102337) |
| Harteveld, E. (2021). Fragmented foes: Affective polarization in the multiparty context of the Netherlands. *Electoral Studies*, 71, 102332. | Netherlands | Shows out-party affect is bloc-structured, not undifferentiated. | [10.1016/j.electstud.2021.102332](https://doi.org/10.1016/j.electstud.2021.102332) |
| Hernández, E., Anduiza, E. & Rico, G. (2021). Affective polarization and the salience of elections. *Electoral Studies*, 69, 102203. | 42 countries, 99 surveys | Election proximity as an exogenous driver of measured polarisation. | [10.1016/j.electstud.2020.102203](https://doi.org/10.1016/j.electstud.2020.102203) |
| Knudsen, E. (2021). Affective polarization in multiparty systems? *Scandinavian Political Studies*, 44(1), 34–44. | Norway vs US | Target of measurement (parties vs voters) changes cross-national rankings. | [10.1111/1467-9477.12186](https://doi.org/10.1111/1467-9477.12186) |
| Kekkonen, A. & Ylä-Anttila, T. (2021). Affective blocs: Understanding affective polarization in multiparty systems. *Electoral Studies*, 72, 102367. | Finland | Bloc-based conceptualisation for fragmented systems. | [10.1016/j.electstud.2021.102367](https://doi.org/10.1016/j.electstud.2021.102367) |
| Reiljan, A. & Ryan, A. (2021). Ideological tripolarization, partisan tribalism and institutional trust. *Scandinavian Political Studies*, 44(2), 195–219. | Sweden | Foundations of affective polarisation in a tripolar system. | — |
| Reiljan, A., Garzia, D., Ferreira da Silva, F. & Trechsel, A. H. (2024). Patterns of affective polarization toward parties and leaders across the democratic world. *American Political Science Review*, 118(2), 654–670. | CSES | Separates leader- from party-directed polarisation. | [10.1017/S0003055423000485](https://doi.org/10.1017/S0003055423000485) |
| Garzia, D., Ferreira da Silva, F. & Maye, S. (2023). Affective polarization in comparative and longitudinal perspective. *Public Opinion Quarterly*, 87(1), 219–231. | Parliamentary democracies | Long-run trends outside the US. | [10.1093/poq/nfad004](https://doi.org/10.1093/poq/nfad004) |
| Boxell, L., Gentzkow, M. & Shapiro, J. M. (2024). Cross-country trends in affective polarization. *Review of Economics and Statistics*, 106(2), 557–565. | 12 OECD democracies | Divergent trends; rise is not universal. | [10.1162/rest_a_01160](https://doi.org/10.1162/rest_a_01160) |
| Wagner, M. (2024). Affective polarization in Europe. *European Political Science Review*, 16(3), 378–392. | Europe | The state-of-the-field review; essential for conceptual and measurement decisions. | [10.1017/S1755773923000383](https://doi.org/10.1017/S1755773923000383) |
| Russo, L., Wagner, M. & Harteveld, E. (2023). Introduction: Affective polarization in multiparty systems. *Electoral Studies*, 86, 102691. | Europe | Special-issue framing of conceptualisation, causes and consequences. | [10.1016/j.electstud.2023.102691](https://doi.org/10.1016/j.electstud.2023.102691) |
| Gidron, N., Sheffer, L. & Mor, G. (2022). Validating the feeling thermometer as a measure of partisan affect in multi-party systems. *Electoral Studies*, 80, 102542. | Multiparty | Measurement validation. | [10.1016/j.electstud.2022.102542](https://doi.org/10.1016/j.electstud.2022.102542) |
| Areal, J. & Harteveld, E. (2024). Vertical vs horizontal affective polarization: Disentangling feelings towards elites and voters. *Electoral Studies*, 90, 102814. | Europe | Separates elite- from voter-directed animosity. | [10.1016/j.electstud.2024.102814](https://doi.org/10.1016/j.electstud.2024.102814) |
| Praprotnik, K. & Wagner, M. (2023). Affective polarization and coalition signals. *Political Science Research and Methods.* | Austria | Coalition signalling as a short-term depolarising cue. | [10.1017/psrm.2023.33](https://doi.org/10.1017/psrm.2023.33) |
| Wagner, M. & Harteveld, E. (2025). Elite cooperation and affective polarization: Evidence from German coalitions. *Political Studies.* | Germany, 20 years monthly data | Coalition formation reduces affective distance, conditional on surprise and recency. | [10.1177/00323217241300993](https://doi.org/10.1177/00323217241300993) |
| Hahm, H., Hilpert, D. & König, T. (2023). Divided by Europe: Affective polarisation in the context of European elections. *West European Politics*, 46(4), 705–731. | EU | EU integration as an axis of partisan affect. | — |
| Harteveld, E., Mendoza, P. & Rooduijn, M. (2022). Affective polarization and the populist radical right: Creating the hating? *Government and Opposition*, 57(4), 703–727. | 28 countries, 103 elections | Tests whether PRR parties generate or merely reflect animosity. | — |
| Harteveld, E. & Wagner, M. (2023). Does affective polarisation increase turnout? *West European Politics*, 46(4), 732–759. | DE, NL, ES | Mobilising consequences of polarisation. | [10.1080/01402382.2022.2087395](https://doi.org/10.1080/01402382.2022.2087395) |
| Berntzen, L. E., Kelsall, H. & Harteveld, E. (2024). Consequences of affective polarization. *European Journal of Political Research*, 63(3), 927–949. | UK, Norway | Avoidance and intolerance yes; support for violence largely not. | [10.1111/1475-6765.12623](https://doi.org/10.1111/1475-6765.12623) |
| Harteveld, E., Berntzen, L. E., Kokkonen, A., Kelsall, H., Linde, J. & Dahlberg, S. (2026). The (alleged) consequences of affective polarization: A survey experiment in nine democracies. *European Journal of Political Research.* | 9 countries, N≈18,000 | The main causal test outside the US. | [10.1017/S1475676526101273](https://doi.org/10.1017/S1475676526101273) |
| Torcal, M. & Magalhães, P. C. (2022). Ideological extremism, perceived party system polarization, and support for democracy. *European Political Science Review*, 14(2), 188–205. | Europe | Curvilinear relationship between perceived polarisation and democratic support. | [10.1017/S1755773922000066](https://doi.org/10.1017/S1755773922000066) |
| Janssen, L. & Turkenburg, E. (2025). Breaking free from linear assumptions. *European Journal of Political Research.* | Europe | Reconciles mixed findings via an inverted-U specification. | [10.1111/1475-6765.12725](https://doi.org/10.1111/1475-6765.12725) |
| Torcal, M. & Harteveld, E. (eds.) (2025). *Handbook of Affective Polarization.* Edward Elgar. | Global/European | Current reference work; chapters on measurement in multiparty systems, negative partisanship, party competition. | [book] |

---

## 8. Central and Eastern Europe

| Reference | Scope | Contribution | DOI |
|---|---|---|---|
| Whitefield, S. (2002). Political cleavages and post-communist politics. *Annual Review of Political Science*, 5, 181–200. | CEE | The framing review of cleavage formation in new democracies. | — |
| Tavits, M. (2005). The development of stable party support: Electoral dynamics in post-communist Europe. *American Journal of Political Science*, 49(2), 283–298. | CEE | Whether Converse-style learning produces stabilising attachments. | — |
| Tavits, M. (2008). On the linkage between electoral volatility and party system instability in Central and Eastern Europe. *European Journal of Political Research*, 47(5), 537–555. | CEE | Disentangles voter-level from supply-side instability. | — |
| Powell, E. N. & Tucker, J. A. (2014). Revisiting electoral volatility in post-communist countries: New data, new results and new approaches. *British Journal of Political Science*, 44(1), 123–147. | CEE | Separates Type A (party entry/exit) from Type B (vote-switching) volatility. | — |
| Rovny, J. & Polk, J. (2019). New wine in old bottles: Explaining the dimensional structure of East European party systems. *Party Politics*, 25(1), 12–24. | CEE | Dimensional structure and its implications for stable alignments. | — |
| Markowski, R. (2016). The Polish parliamentary election of 2015: A free and fair election that results in unfair political consequences. *West European Politics*, 39(6), 1311–1322. | Poland | Entry point to the Polish realignment/polarisation literature. | — |
| Tworzecki, H. (2019). Poland: A case of top-down polarization. *The ANNALS of the AAPSS*, 681(1), 97–119. | Poland | Elite-driven polarisation in a dealigned electorate. | — |
| Enyedi, Z. (2005). The role of agency in cleavage formation. *European Journal of Political Research*, 44(5), 697–720. | Hungary/CEE | Agency-based account of how elites construct alignments. | — |
| Gherghina, S. (2014). *Party Organization and Electoral Volatility in Central and Eastern Europe.* Routledge. | CEE | Organisational sources of (in)stability. | [book] |
| Sikk, A. (2005). How unstable? Volatility and the genuinely new parties in Eastern Europe. *European Journal of Political Research*, 44(3), 391–412. | CEE/Baltics | New-party entry as the driver of measured volatility. | — |
| Haughton, T. & Deegan-Krause, K. (2020). *The New Party Challenge: Changing Cycles of Party Birth and Death in Central Europe and Beyond.* Oxford UP. | CEE | Explains recurrent party turnover in weakly aligned electorates. | [book] |

---

## 9. Country-level literatures (Western and Southern Europe)

### United Kingdom

| Reference | Contribution | DOI |
|---|---|---|
| Crewe, I., Särlvik, B. & Alt, J. (1977). Partisan dealignment in Britain 1964–1974. *British Journal of Political Science*, 7(2), 129–190. | The paper that named British dealignment. | — |
| Särlvik, B. & Crewe, I. (1983). *Decade of Dealignment.* Cambridge UP. | Book-length statement. | [book] |
| Clarke, H., Sanders, D., Stewart, M. & Whiteley, P. (2004). *Political Choice in Britain.* Oxford UP. | Valence politics as the successor to cleavage-based alignment. | [book] |
| Evans, G. & Tilley, J. (2012). How parties shape class politics: Explaining the decline of the class basis of party support. *British Journal of Political Science*, 42(1), 137–161. | Supply-side explanation: class dealignment is party-induced. | — |
| Evans, G. & Tilley, J. (2017). *The New Politics of Class.* Oxford UP. | The full argument, with long-run BES data. | [book] |
| Fieldhouse, E. et al. (2020). *Electoral Shocks.* Oxford UP. | Volatility, shocks and the Brexit realignment. | [book] |
| Hobolt, S. B., Leeper, T. J. & Tilley, J. (2021). Divided by the vote. *BJPS*, 51(4). | Brexit identities as functional substitutes for partisanship. | [10.1017/S0007123420000125](https://doi.org/10.1017/S0007123420000125) |

### Germany

| Reference | Contribution | DOI |
|---|---|---|
| Baker, K. L., Dalton, R. J. & Hildebrandt, K. (1981). *Germany Transformed.* Harvard UP. | Classic account of value change and partisan realignment. | [book] |
| Arzheimer, K. (2006). "Dead men walking?" Party identification in Germany, 1977–2002. *Electoral Studies*, 25(4), 791–807. | Systematic test of the German decline claim. | — |
| Schmitt-Beck, R., Weick, S. & Christoph, B. (2006). Shaky attachments. *EJPR*, 45(4). | Panel evidence on instability. | — |
| Dassonneville, R., Hooghe, M. & Vanhoutte, B. (2012). Age, period and cohort effects... *German Politics*, 21(2). | APC decomposition of German dealignment. | [10.1080/09644008.2012.679659](https://doi.org/10.1080/09644008.2012.679659) |
| Kroh, M. & Selb, P. (2009). Inheritance and the dynamics of party identification. *Political Behavior*, 31(4). | Socialisation vs. running tally, SOEP. | [10.1007/s11109-009-9084-2](https://doi.org/10.1007/s11109-009-9084-2) |
| Hudde, A. (2022). Partisan affect in times of fractionalization: Visualizing who likes whom in Germany, 1977 to 2020. *Socius*, 8. | Long-run descriptive mapping of German partisan affect. | — |
| Wagner, M. & Harteveld, E. (2025). Elite cooperation and affective polarization. *Political Studies.* | Coalition effects on partisan affect. | [10.1177/00323217241300993](https://doi.org/10.1177/00323217241300993) |

### Netherlands & Belgium

| Reference | Contribution | DOI |
|---|---|---|
| Thomassen, J. (1976). Party identification as a cross-national concept. In Budge et al. | The founding Dutch sceptical case. | [book] |
| Thomassen, J. & Rosema, M. (2009). Party identification revisited. In Bartle & Bellucci (eds.). | Reassessment after three decades. | [book] |
| Dassonneville, R. (2013). Questioning generational replacement. *Electoral Studies*, 32(1). | APC analysis of Dutch volatility. | — |
| van der Meer, T. et al. (2015). Are volatile voters erratic, whimsical or seriously picky? *Party Politics*, 21(1). | Reasoned intra-bloc volatility. | — |
| Aarts, K. & van der Kolk, H. / de Lange, S. et al. (2020). Dealignment, realignment and generational differences in the Netherlands. *West European Politics*, 43(4). | Decomposes dealignment vs. cleavage realignment by cohort. | [10.1080/01402382.2020.1774203](https://doi.org/10.1080/01402382.2020.1774203) |
| Harteveld, E. (2021). Fragmented foes. *Electoral Studies*, 71, 102332. | Bloc structure of Dutch partisan affect. | [10.1016/j.electstud.2021.102332](https://doi.org/10.1016/j.electstud.2021.102332) |
| Rekker, R. et al. (2019). The formation of party preference in adolescence and early adulthood. *Young*, 27(1). | Formative-window evidence. | [10.1177/1103308818757037](https://doi.org/10.1177/1103308818757037) |
| Walgrave, S. & Westwood, S. J. et al. (2018). Partyism in Belgium — see Westwood et al. (2018). | Belgian evidence on partyism vs. linguistic cleavage. | [10.1111/1475-6765.12228](https://doi.org/10.1111/1475-6765.12228) |

### Nordics (DK, SE, FI)

| Reference | Contribution | DOI |
|---|---|---|
| Borre, O. & Andersen, J. G. (1997). *Voting and Political Attitudes in Denmark.* Aarhus UP. | Long-run Danish alignment change. | [book] |
| Stubager, R. (2009). Education-based group identity and consciousness. *EJPR*, 48(2). | Education as identity in Denmark. | — |
| Slothuus, R. & Bisgaard, M. (2021). How political parties shape public opinion in the real world. *AJPS*, 65(4). | Danish quasi-experiment on partisan following. | [10.1111/ajps.12550](https://doi.org/10.1111/ajps.12550) |
| Hansen, K. M. & Kosiara-Pedersen, K. (2017). How campaigns polarize the electorate. *Party Politics*, 23(3). | Danish bloc-based polarisation. | — |
| Oscarsson, H. & Holmberg, S. (2020). *Research on Swedish Voting Behaviour* / *Swedish Voting Behavior*. | The reference synthesis for Sweden. | [book] |
| Reiljan, A. & Ryan, A. (2021). Ideological tripolarization, partisan tribalism and institutional trust. *Scandinavian Political Studies*, 44(2). | Swedish tripolar structure. | — |
| Oskarson, M. & Demker, M. (2015). Room for realignment: The working-class sympathy for Sweden Democrats. *Government and Opposition*, 50(4), 629–651. | Swedish working-class realignment. | — |
| Kekkonen, A. & Ylä-Anttila, T. (2021). Affective blocs. *Electoral Studies*, 72, 102367. | Finnish bloc structure. | [10.1016/j.electstud.2021.102367](https://doi.org/10.1016/j.electstud.2021.102367) |
| Westinen, J. (2015). *Cleavages in Contemporary Finland.* Åbo Akademi. | Finnish cleavage change. | [book] |

### Southern Europe (ES, IT, PT, GR)

| Reference | Contribution | DOI |
|---|---|---|
| Barnes, S. H., McDonough, P. & López Pina, A. (1985). The development of partisanship in new democracies: The case of Spain. *American Journal of Political Science*, 29(4), 695–720. | Classic study of partisanship formation after transition. | — |
| Torcal, M. & Comellas, J. M. (2022). Affective polarisation in times of political instability and conflict: Spain from a comparative perspective. *South European Society and Politics*, 27(1). | Spanish trends with comparative benchmarking. | [10.1080/13608746.2022.2044236](https://doi.org/10.1080/13608746.2022.2044236) |
| Orriols, L. & León, S. (2020). Looking for affective polarisation in Spain: PSOE and Podemos from conflict to coalition. *South European Society and Politics*, 25(3–4). | Intra-bloc polarisation and coalition effects. | — |
| Rodon, T. (2022). Affective and territorial polarisation: The impact on vote choice in Spain. *South European Society and Politics*, 27(1), 147–169. | Territorial identity as a second polarisation axis. | — |
| Bellucci, P., Maraffi, M. & Segatti, P. (2000). *PCI, PDS, DS: La trasformazione dell'identità politica della sinistra di governo.* Donzelli. | Party-identity transformation in Italy. | [book] |
| Chiaramonte, A. & Emanuele, V. (2017). Party system volatility, regeneration and de-institutionalization. *Party Politics*, 23(4). | Italy in comparative volatility perspective. | — |
| Lisi, M. (2015). *Party Change, Recent Democracies, and Portugal.* Lexington. | Portuguese party-voter linkage. | [book] |
| Teperoglou, E. & Tsatsanis, E. (2014). Dealignment, de-legitimation and the implosion of the two-party system in Greece. *Journal of Elections, Public Opinion and Parties*, 24(2), 222–242. | The paradigmatic case of rapid dealignment under crisis. | — |
| Tsatsanis, E., Teperoglou, E. & Seriatos, A. (2020). Two-party system collapse and party–voter realignment in times of crisis. *South European Society and Politics*, 25(3–4). | Greek realignment after 2012. | — |

### France, Austria, Ireland

| Reference | Contribution | DOI |
|---|---|---|
| Converse, P. E. & Dupeux, G. (1962). Politicization of the electorate in France and the United States. *Public Opinion Quarterly*, 26(1), 1–23. | The founding comparative statement of French partisan weakness. | — |
| Haegel, F. (1993). Partisan ties. In Boy, D. & Mayer, N. (eds.), *The French Voter Decides.* | French partisanship in comparative perspective. | [book] |
| Tiberj, V. (2017). Running to stand still: Cohort replacement and political change in France. | Generational account of French realignment. | — |
| Plasser, F. & Ulram, P. A. (2006). *Wechselwahl und Wählerdynamik in Österreich.* | Austrian dealignment and volatility. | [book] |
| Praprotnik, K. & Wagner, M. (2023). Affective polarization and coalition signals. *PSRM.* | Austrian coalition signalling experiment. | [10.1017/psrm.2023.33](https://doi.org/10.1017/psrm.2023.33) |
| Marsh, M. (2006). Party identification in Ireland: An insecure anchor for a floating party system. *Electoral Studies*, 25(3), 489–508. | Irish partisanship without a class cleavage. | — |
| Marsh, M., Sinnott, R., Garry, J. & Kennedy, F. (2008). *The Irish Voter.* Manchester UP. | Comprehensive Irish account. | [book] |

---

## 10. Coverage gaps you will need to fill

Countries with comparatively thin English-language journal literature on partisanship specifically
(as opposed to volatility or party systems): **Bulgaria, Romania, Croatia, Slovenia, Slovakia, the Baltic
states**. For these, the productive entry points are the CEE volatility and cleavage literatures in §8,
national election-study reports, and *East European Politics*, *Problems of Post-Communism* and
*Europe-Asia Studies* rather than the mainstream comparative journals.

Data infrastructures that anchor almost all of this work, and which are worth citing directly:
CSES (Modules 1–5), the European Election Studies / Eurobarometer partisanship series, the ESS,
national election studies (BES, GLES, DPES/NKO, SNES, Danish NES, CIS/CNEP, ITANES), and long-run
household panels (SOEP, LISS, Understanding Society).



# EU Partisanship Survey Projects and Datasets

## Cross-national Programs

| Project | Design | Span | Countries | Partisanship item | Link |
|---|---|---|---|---|---|
| European Social Survey (ESS) | Repeated cross-section, biennial | 2002– | All 25 at some point; BG, HR, EL, LV, RO, SK patchy by round | "Is there a particular political party you feel closer to than all the other parties?" → which → how close | [europeansocialsurvey.org](https://www.europeansocialsurvey.org/) |
| Comparative Study of Electoral Systems (CSES) | Post-election modules appended to national election studies (Modules 1–6) | 1996– | Nearly all 25; continuous for DE, NL, SE, PL, CZ, IE, PT, ES | "Do you usually think of yourself as close to any particular party?" + degree of closeness | [cses.org](https://cses.org/) |
| European Election Studies (EES) | Post-EP-election voter studies; panel components 1989, 2009 | 1979– | All EU members at time of fielding; UK to 2019 | Closeness battery, CSES-compatible | [europeanelectionstudies.net](https://europeanelectionstudies.net/) |
| Eurobarometer (Standard EB) | Repeated cross-section, ~2×/year | 1973– | All EU members; UK to 2019 | Party attachment asked intermittently; thin after ~2000 | [GESIS EB Data Service](https://www.gesis.org/en/eurobarometer-data-service) |
| Mannheim Eurobarometer Trend File | Harmonised EB trend file (ZA3521) | 1970–2002 | EU-15 | Harmonised party-closeness item across early waves | [search.gesis.org](https://search.gesis.org/research_data/ZA3521) |
| ISSP | Annual repeated cross-section | 1985– | Most of the 25, varies by module | Party affiliation as standard harmonised background variable | [issp.org](https://issp.org/) |
| European Values Study (EVS) | Repeated cross-section | 1981, 1990, 1999, 2008, 2017 | All 25 | Closeness / vote-propensity items; weaker and less consistent than ESS/CSES | [europeanvaluesstudy.eu](https://europeanvaluesstudy.eu/) |
| CRONOS / CRONOS-2 | Probability-based cross-national online panel on ESS sample; 6 waves each | 2016–18 (EE, UK, SI); Nov 2021–Mar 2023 (AT, BE, CZ, FI, FR, HU, IS, IT, PT, SI, SE, UK) | 3 + 12 | ESS closeness item repeated across waves | [europeansocialsurvey.org](https://www.europeansocialsurvey.org/) |
| Comparative National Elections Project (CNEP) | Election-based cross-sections | 1990– | DE, EL, ES, HU, IT, PT among EU cases | Party identification plus social-network transmission items | [u.osu.edu/cnep](https://u.osu.edu/cnep/) |
| True European Voter | *Ex post* harmonisation of national election studies | 1950s–2010s | ~15 W/C European countries | Harmonised party-ID variable; best route to long back-series comparability | [CESSDA catalogue](https://datacatalogue.cessda.eu/) |

## National Studies

| Country | Project | Design | Span | Link |
|---|---|---|---|---|
| Austria | AUTNES (incl. Online Panel Study); predecessors 1986–2008 | Rolling cross-section + multi-wave panels | 2013– | [autnes.at](https://www.autnes.at/) |
| Belgium | RepResent (EOS) | Multi-wave election panel | 2019– | [represent-project.be](https://www.represent-project.be/) |
| Belgium | PartiRep Election Study; Belgian NES (ISPO/PIOP) | Cross-sections, panel waves 2014 | 1991– | [partirep.eu](https://www.partirep.eu/) |
| Bulgaria | Bulgarian Election Study (CSES vehicle) | Election-triggered cross-sections; no continuous series | ~2001, 2014, 2021 | [cses.org](https://cses.org/) |
| Croatia | Croatian Election Studies (FPZG Zagreb), harmonised file | 11 pre/post-election cycles | 1990–2024 | [CROSSDA](https://data.crossda.hr/dataset.xhtml?persistentId=doi:10.23669/IWHYLM) |
| Czechia | Czech NES; *Naše společnost* continuous survey | Post-election cross-sections; monthly series | 1990–; 2002– | [cvvm.soc.cas.cz](https://cvvm.soc.cas.cz/en) |
| Czechia | Czech Social Science Data Archive (CSDA) | Archive for CNES + CHPS | — | [archiv.soc.cas.cz](https://archiv.soc.cas.cz/en) |
| Denmark | Danish National Election Study (DNES / Valgprojektet) | Cross-sections; 2-wave panel 2019→2022 | 1971– | [valgprojektet.dk](https://www.valgprojektet.dk/default.asp?l=eng) |
| Estonia | Estonian National Election Study (Johan Skytte Institute, Tartu) | Post-election cross-sections | 2011, 2015, 2019, 2023 | [skytte.ut.ee](https://skytte.ut.ee/en) |
| Finland | Finnish National Election Study (FNES) | Post-election cross-sections | 2003– (roots to 1948) | [Finnish Social Science Data Archive](https://www.fsd.tuni.fi/en/) |
| France | CEVIPOF *Enquête Électorale Française*; *Baromètre de la confiance politique* | Large multi-wave election panels; annual barometer | 2002–; 2009– | [sciencespo.fr/cevipof](https://www.sciencespo.fr/cevipof/en.html) |
| France | ELIPSS | Probability-based online panel | 2012– | [elipss.fr](https://www.elipss.fr/) |
| Germany | GLES | Rolling cross-section + short- and long-term panels | 2009– | [gles-studie.de](https://gles-studie.de/) |
| Germany | SOEP | Annual household panel; party ID since 1984 | 1984– | [diw.de/soep](https://www.diw.de/soep) |
| Germany | ALLBUS | Biennial cross-section | 1980– | [gesis.org/allbus](https://www.gesis.org/en/allbus/allbus-home) |
| Germany | GESIS Panel | Probability-based mixed-mode panel | 2013– | [gesis.org/gesis-panel](https://www.gesis.org/en/gesis-panel) |
| Germany | Politbarometer | Monthly time series | 1977– | [forschungsgruppe.de](https://www.forschungsgruppe.de/) |
| Greece | Hellenic/Greek National Election Studies (CSES vehicle) | Election-triggered; discontinuous | 2009, 2012, 2015 | [cses.org](https://cses.org/) |
| Hungary | Hungarian Election Study (TK PTI); Election Panel Study 2018 | Cross-sections + 2018 multi-wave panel | 2014, 2018, 2022 | [politikatudomany.tk.hu](https://politikatudomany.tk.hu/en) |
| Ireland | Irish National Election Study (INES) | 2002–07 panel; later cross-sections | 2002– | [ISSDA, UCD](https://www.ucd.ie/issda/) |
| Italy | ITANES | Election cross-sections with panel components | 1990– (back-series to 1968) | [itanes.org](https://www.itanes.org/) |
| Italy | CISE Osservatorio Politico | Repeated election surveys | 2000s– | [cise.luiss.it](https://cise.luiss.it/) |
| Latvia | No continuous academic series; CSES Latvia | Election-triggered | 2010/11, 2014 | [cses.org](https://cses.org/) |
| Latvia | SKDS omnibus / Latvijas Barometrs | Commercial time series | 1990s– | [skds.lv](https://www.skds.lv/) |
| Lithuania | Lithuanian National Election Study (VU TSPMI) | Post-election cross-sections, CSES-linked | 1997– | [tspmi.vu.lt](https://www.tspmi.vu.lt/en/) |
| Netherlands | Dutch Parliamentary Election Study (DPES/NKO, SKON) | Pre/post panel design; recent waves via LISS + I&O | 1971–2025 | [surveydata.nl](https://surveydata.nl/dutch-parliamentary-election-studies) |
| Netherlands | DPES data deposits | Archive access | 1971– | [DANS Data Station SSH](https://ssh.datastations.nl/) |
| Netherlands | LISS Panel | Probability online panel; annual Politics & Values module | 2007– | [lissdata.nl](https://www.lissdata.nl/) |
| Poland | Polish National Election Study (PGSW) | Election cross-sections | 1997– | [pgsw.pl](https://www.pgsw.pl/) |
| Poland | POLPAN | True panel, 5-year waves | 1988– | [polpan.org](https://polpan.org/) |
| Poland | CBOS *Aktualne problemy i wydarzenia* | Monthly time series | 1982– | [cbos.pl](https://www.cbos.pl/EN/home_en/cbos_en.php) |
| Portugal | Portuguese Election Study; *Barómetro da Qualidade da Democracia* | Election cross-sections; annual barometer | 2002–; 2008– | [ics.ulisboa.pt](https://www.ics.ulisboa.pt/) |
| Romania | Romanian Election Studies (RES); CSES Romania | Election cross-sections | 1996– | [cses.org](https://cses.org/) |
| Slovakia | CSES Slovakia (Comenius University) | Post-election cross-sections | 2010, 2016, 2020, 2023 | [cses.org](https://cses.org/) |
| Slovenia | Slovenian Public Opinion Survey (SJM), hosts national election studies | Near-annual repeated cross-section | 1968– | [ADP, Ljubljana](https://www.adp.fdv.uni-lj.si/eng/) |
| Spain | CIS *Barómetro* + pre/post-election studies | Monthly time series; election panels | 1979– | [cis.es](https://www.cis.es/) |
| Sweden | Swedish National Election Study (Valundersökningen) | Rotating panel across consecutive elections | 1956– | [gu.se/en/snes](https://www.gu.se/en/snes) |
| Sweden | SOM Institute national survey | Annual cross-section | 1986– | [gu.se/en/som-institute](https://www.gu.se/en/som-institute) |
| Sweden | Medborgarpanelen (LORE) | Online panel | 2010– | [gu.se/en/lore](https://www.gu.se/en/lore) |
| UK | British Election Study + BES Internet Panel | Election cross-sections; continuous multi-wave panel | 1964–; 2014– | [britishelectionstudy.com](https://www.britishelectionstudy.com/) |
| UK | Understanding Society (UKHLS), incl. BHPS | Annual household panel; party support items | 1991– | [understandingsociety.ac.uk](https://www.understandingsociety.ac.uk/) |
| UK | British Social Attitudes | Annual cross-section | 1983– | [natcen.ac.uk](https://natcen.ac.uk/british-social-attitudes) |

## Panel Datasets with a Partisanship Item

Filtered view of the two tables above: **only designs that re-interview the same respondents**,
and only where a partisanship item (closeness, party ID, *Parteineigung*, or an equivalent
party-attachment battery) is fielded. Repeated cross-sections, monthly time series and data
archives are excluded even where the parent programme appears above — ESS, Eurobarometer, EVS,
ISSP, SOM, ALLBUS, Politbarometer, CBOS, CIS *Barómetro*, SJM and *Naše společnost* are all out
on that rule. Scope is the 25 countries of the analysis frame (EU-27 less CY/MT/LU, plus the UK).

The table exists to make the section 7 choice and its coverage cost readable in one place: which
of the 25 can support a within-person test of attachment movement, and which cannot.

**Access marks follow the legend in `data/01_raw/external-data.md`:** ✅ on disk in this repo,
⬜ identified but not downloaded, 🔒 registration / data-use agreement required.
**⚠ in the last column** flags a cell not verified against a source — the note says what to check.

### Cross-national panels

| Project | Countries (of the 25) | Panel type | Waves / span | Partisanship item | Access | Check |
|---|---|---|---|---|---|---|
| **eu25games2019** (Hahm, Hilpert & König) — see [inventory §A](../data/01_raw/external-data.md) | All 25 | Election panel (EP 2019) | 3 waves, 2019 | Attachment item + vote choice (intended / hypothetical / past) | ✅ | — |
| TRI-POL — see [inventory §B](../data/01_raw/external-data.md) | ES, IT, PT (+ AR, CL outside frame) | Online panel (opt-in Netquest, no weights) | 3 waves, Sept 2021 – Apr 2022 | Attachment item in all three waves; party/leader thermometers; incentivized trust game | ✅ | — |
| [CRONOS-2](https://www.europeansocialsurvey.org/sites/default/files/2023-07/CRONOS2_user_guide_e01_0.pdf) | AT, BE, CZ, FI, FR, HU, IT, PT, SE, SI, UK — 11 (+ IS outside frame) | Probability online panel, recruited off ESS Round 10 | 6 waves, Nov 2021 – Mar 2023; HU completed waves 1–2 only | ESS closeness item carried from the R10 interview | ⬜ | ⚠ which panel waves repeat the closeness item |
| [CRONOS-1](https://www.europeansocialsurvey.org/methodology/methodological-research/modes-data-collection/cronos) | EE, SI, UK | Probability online panel, recruited off ESS Round 8 | 6 waves, 2016–18 | ESS closeness item | ⬜ | ⚠ same |
| [EES 2014 online panel](https://www.gesis.org/en/services/finding-and-accessing-data/international-survey-programs/ees-2014-study) | AT, DE, EL, ES, IT, PL, SE, UK — 8 | Election panel | EP wave (May 2014) matched to the subsequent first-order national election | Closeness battery, CSES-compatible | ⬜ 🔒 | ⚠ item presence in the panel waves specifically |

The EES 1989 and 2009 panel components (listed under *Cross-national Programs* above) are older
and thinner; the 2014 online panel is the one worth using, and it is Greece's only panel of any
kind. The **EES 2024 voter study (ZA8868) is a post-election cross-section, not a panel**, and is
excluded here despite appearing in the inventory.

### National panels

| Country | Project | Panel type | Waves / span | Partisanship item | Access | Check |
|---|---|---|---|---|---|---|
| Austria | [AUTNES Online Panel 2017–2024](https://doi.org/10.11587/HNUFCC) | Rolling online panel | Multi-wave, 2017–2024 | Party ID with strength, repeated as a core variable | ⬜ 🔒 | — |
| Belgium | [RepResent Belgian Panel](https://link.springer.com/article/10.1057/s41304-023-00430-z) | Election panel | 4 waves: pre-2019 (N=7,351), post-2019 (N=3,909), +1 yr Apr 2020 (N=1,996), +2 yr Jun 2021 (N=1,119) | Party attachment within a ~250-variable representation battery | ⬜ | ⚠ exact closeness wording |
| Belgium | [PartiRep Election Study](https://www.partirep.eu/) | Pre/post 2-wave | 2014 cycle | Party ID | ⬜ | ⚠ wave structure and item |
| Bulgaria | — none identified — | — | — | — | — | ⚠ searched Aug 2026; the CSES vehicle is cross-sectional |
| Croatia | — none identified — | — | — | — | — | ⚠ [Croatian Election Studies](https://data.crossda.hr/dataset.xhtml?persistentId=doi:10.23669/IWHYLM) are pre/post cross-sections, not a linked panel |
| Czechia | [Czech Household Panel Survey (CHPS)](https://www.soc.cas.cz/en/publications/the-czech-household-panel-survey) | Household | 5 annual waves 2015–2019, 2 COVID collections 2020–21, wave 6 2023–24 | Party preference in the political-participation module; [analysed as panel data](https://journals.sagepub.com/doi/full/10.1177/14789299251410829) | ⬜ | ⚠ whether the item is closeness or vote intention, and its wave coverage |
| Denmark | [DNES / Valgprojektet](https://www.valgprojektet.dk/default.asp?l=eng) | Pre/post 2-wave | 2019 → 2022 | Closeness with strength | ⬜ | ⚠ panel linkage across the two elections |
| Estonia | — none identified — | — | — | — | — | covered only by CRONOS-1 above |
| Finland | [Finnish online panel (Kekkonen & Ylä-Anttila)](https://www.frontiersin.org/journals/political-science/articles/10.3389/fpos.2022.920567/full) | Online panel | Multi-wave, around 2019 | Party like–dislike and partisan social distance | ⬜ | ⚠ wave count and whether released for secondary use |
| France | [CEVIPOF *Enquête Électorale Française*](https://www.sciencespo.fr/cevipof/en.html) | Election panel | Large multi-wave panels per cycle, 2002– | Party closeness | ⬜ | ⚠ item constant across waves |
| France | [ELIPSS](https://www.elipss.fr/) | Probability online panel | 2012– | Party closeness in the political modules | ⬜ 🔒 | ⚠ which modules carry it |
| Germany | [GLES Panel 2016–2021](https://doi.org/10.4232/1.14114) | Rolling online panel | Multi-wave, 2016–2021 | *Parteineigung* with strength; party and leader scalometers | ⬜ 🔒 | ⚠ wave count |
| Germany | [SOEP](https://www.diw.de/soep) | Household | Annual, 1984– | *Parteineigung* with strength, asked since 1984 | ⬜ 🔒 | — |
| Germany | [GESIS Panel](https://www.gesis.org/en/gesis-panel) | Probability online panel, mixed-mode | 2013– | Party ID in the recurring political battery | ⬜ 🔒 | ⚠ item periodicity |
| Greece | [Hellenic Panel — EES 2014](https://elnes.gr/node/1483) | Election panel | EP 2014 → subsequent national election | Closeness battery | ⬜ | ⚠ Greece's only panel; all other ELNES studies are cross-sections |
| Hungary | [Election Panel Study 2018](https://politikatudomany.tk.hu/en) | Election panel | Multi-wave, 2018 cycle | Party preference | ⬜ | ⚠ wave count and item |
| Ireland | [INES 2002–2007](https://www.ucd.ie/issda/) | Election panel | 2002 → 2007 | Party ID | ⬜ 🔒 | ⚠ item and linkage; later INES rounds are cross-sections |
| Italy | [ITANES](https://www.itanes.org/) | Election panel (selected cycles) | Panel components within the 1990– series | Party ID | ⬜ | ⚠ which cycles carry a panel component |
| Latvia | — none identified — | — | — | — | — | ⚠ searched Aug 2026; New Baltic Barometer is repeated cross-section |
| Lithuania | — none identified — | — | — | — | — | ⚠ searched Aug 2026; LNES rounds are post-election cross-sections |
| Netherlands | [LISS](https://www.lissdata.nl) | Probability online panel | Annual *Politics & Values* module, 2007– | Closeness; also like–dislike toward party *supporters* | ⬜ 🔒 | — |
| Netherlands | [DPES / NKO](https://surveydata.nl/dutch-parliamentary-election-studies) | Pre/post 2-wave | 1971–2025 | Closeness with strength | ⬜ | ⚠ inter-election linkage |
| Poland | [POLPAN](https://polpan.org/) | True panel, 5-year waves | 1988–2018 | Political-behaviour battery; a partisanship item is **not** confirmed for every wave | ⬜ | ⚠ verify the item wave by wave before relying on it |
| Portugal | — none identified — | — | — | — | — | covered only by CRONOS-2 and TRI-POL above |
| Romania | — none identified — | — | — | — | — | ⚠ searched Aug 2026; RES rounds are election cross-sections |
| Slovakia | — none identified — | — | — | — | — | ⚠ searched Aug 2026; CSES Slovakia is cross-sectional, Volebný Kompas is VAA data |
| Slovenia | — none identified — | — | — | — | — | covered only by CRONOS-1 and CRONOS-2 above |
| Spain | [POLAT Panel](https://doi.org/10.34810/DATA1486) | Online panel | **12 waves, 2010–2020**, 4,633 respondents, 657 variables | Partisanship, explicitly among the fielded political variables | ⬜ | — |
| Spain | [E-DEM](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7451797/) | Election panel | 2018–2019 | Partisanship plus party thermometers; built for affective polarisation | ⬜ | ⚠ wave count; already cited here via `@Torcal2022affective` and `@Torcal2023social` |
| Spain | [CIS pre/post-election studies](https://www.cis.es/) | Pre/post 2-wave | 1979– | Party ID / *simpatía* | ⬜ | ⚠ which cycles are genuinely panelled |
| Sweden | [SNES / Valundersökningen](https://www.gu.se/en/snes) | Rotating (consecutive elections) | 1956– | Party ID with strength | ⬜ | ⚠ rotation half-sample linkage |
| Sweden | [Medborgarpanelen (LORE)](https://www.gu.se/en/lore) | Online panel | 2010– | Party sympathy / party ID | ⬜ 🔒 | ⚠ item wording and periodicity |
| UK | [BES Internet Panel](https://doi.org/10.5255/UKDA-SN-8202-2) | Rolling online panel | Waves 1–29, 2014–, ~30k per wave | Party ID with strength; paradata enables satisficing checks | ⬜ 🔒 | — |
| UK | [Understanding Society (UKHLS), incl. BHPS](https://www.understandingsociety.ac.uk/) | Household | Annual, 1991– | Party support and closeness items | ⬜ 🔒 | ⚠ item continuity across the BHPS → UKHLS transition |

### Coverage

Nineteen of the 25 have a national panel carrying a partisanship item, though depth varies by an
order of magnitude: DE, UK, ES and NL have long multi-wave series (SOEP, BES IP, POLAT, LISS),
while IE, EL and HU have a single historical panel each. **EE, PT and SI have no national panel**
and are reachable only through CRONOS-1, CRONOS-2 or TRI-POL. **BG, HR, LV, LT, RO and SK have
none at all** — the search returning nothing is itself the result, and it reproduces exactly the
CEE and Baltic thinness noted in §10 above.

For section 7 this narrows the field sharply. Only three sources put a repeated attachment measure
and an affective-polarisation outcome on the same respondents — TRI-POL, E-DEM and
`eu25games2019` itself — and only `eu25games2019` covers the full frame. Status, access and the
transition-matrix gate for each candidate live in `data/01_raw/external-data.md` §C.1; this table
is the literature-side map of what exists, not a record of what is on disk.


# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Code 1

```{r}
#| label: code-1

# start your code here

```


# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```


  
