---
title: |
  Comparative AP Literature Summary
subtitle: |
  How does reported partisan attachement influence affective polarization?: A comparative study of 25 European democracies
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
bibliography: ../references.bib
biblio-style: apsr
---

# Comparative Affective Polarization Studies: Overview Table

## Coding Guide

Each study is coded on four dimensions. Definitions are fixed below — apply them consistently across all entries and use the **Notes** column only for genuine ambiguities that cannot be resolved by the rules given here.

### 1. AP Target
*What is the object of polarized affect — parties/elites, or fellow citizens?*

| Code | Definition | Typical measure |
|:-----|:-----------|:----------------|
| **Vertical** | Affect directed at **parties or political elites** | Feeling thermometers toward parties or leaders |
| **Horizontal** | Affect or behavior directed at **fellow citizens** with different partisan identities | Social distance, discrimination tasks, allocation games |
| **Both** | Study **explicitly reports both** vertical and horizontal measures as distinct dimensions | Combined survey + behavioral experiment |

> **Key distinction:** Feeling thermometers toward *parties* = Vertical. Feeling thermometers toward *party supporters* or behavioral tasks involving *citizens* = Horizontal. If a study uses party thermometers and treats them as a proxy for citizen-level polarization without measuring it directly, code as **Vertical** and flag in Notes.


### 2. AP Manifestation
*How is AP expressed and operationalized?*

| Code | Definition | Typical measure |
|:-----|:-----------|:----------------|
| **Attitudinal** | **Self-reported evaluations** of parties, leaders, or partisan outgroups | Feeling thermometers, like-dislike scales, social distance items |
| **Behavioral** | **Actions** in incentivized or experimental settings | Dictator games, trust games, conjoint discrimination tasks |
| **Both** | Study reports and analyzes **both** attitudinal and behavioral measures | Survey thermometers + allocation experiment in same study |


### 3. Sample
*Who is included in the analysis as a "partisan"?*

This is the dimension most consequential for the paper's argument. Apply codes strictly.

| Code | Definition |
|:-----|:-----------|
| **Identifiers** | Restricted to respondents who **self-report psychological attachment** to a party (explicit partisans only) |
| **Voters** | Restricted to respondents who **report a vote choice**, regardless of felt attachment; includes implicit partisans |
| **Mixed** | Combines identifiers and voters under a shared partisan anchor, or uses a strategy encompassing both groups |
| **All** | **No partisan restriction** — non-partisans included in the analysis |

**Decision rules:**

- Reiljan's API on CSES → **Identifiers** (restricted to respondents naming a liked party)
- Wagner's spread measure on full sample → **All**
- Any anchor that accepts vote choice without requiring attachment → **Voters** or **Mixed**
- When unclear, note the exact operationalization in Notes


### 4. Data Source
*What dataset does the study rely on?*

Record the primary dataset by its common name. This column reveals which studies draw on the **same underlying data** — a key part of the paper's framing regarding the field's dependence on a small number of datasets.

Common sources: `CSES`, `ESS`, `Original survey`, `Multiple`


## Table

| Study  | AP Target | AP Manifestation | Sample | N Countries | Data Source | Notes |
|:------:|:---------:|:----------------:|:------:|:-----------:|:------------|:------|
| @iyengar2012affect | Both | Attitudinal | Identifiers | 2 | Multiple (ANES, YouGov, @almond_verba_1959_civic_culture, Blair Center Election Study, AP-Yahoo! News 2008 election study) | Thermometers toward parties (V) + social trust/discrimination items (H); US & UK |
| @Westwood2018tie | Horizontal | Both | Voters | 4 | Original survey | Trust games; US, UK, Belgium, Spain |
| @reiljan2020fear  | Vertical | Attitudinal | Mixed | 23 | CSES | Introduces API |
| @gidron2020american  | Vertical | Attitudinal | Identifiers | 20 | Original survey | US, UK, Israel; party thermometers |
| @wagner2021affective  | Vertical | Attitudinal | All | 51 | CSES | Spread-of-scores & distance measures; includes non-partisans |
| @Helbling2020social | Horizontal | Attitudinal | Voters | 2 | Original survey | Social avoidance; Austria & Germany |
| @kekkonen2021affective  | Vertical | Attitudinal | Mixed | 1 | Original survey | Finland; citizen-targeted affect |
| @Harteveld2024consequences  | Both | Both | Mixed | 9 | Original Survey (CONAP) | Operationalize in- and out-group via most- and least liked party supporters in like/dislike items |
| @areal2024verticalvshorizontal  | Both | Attitudinal | Voters | 9 | Multiple | Explicit V vs H comparison; correlation ~0.4–0.7 by country |
| @reiljan2024patterns  | Vertical | Attitudinal | Voters | 40 | CSES | Distinguish vertical AP into LAP and PAP; analyses based on identifiers only as robustness |
| @hahm2024divided  | Horizontal | Behavioral | Identifiers | 25 | Original Survey | Conjoint dictator + trust games |
| @Hernandez2021affective | Vertical | Attitudinal | All  | 42 | CSES |  |
| @boxell2024cross | Vertical | Attitudinal | Identifiers  | 12 | Multiple |  |
| @lauka2018mass | Vertical | Attitudinal | Voters  | 48 | CSES |  |
| @bradley2022ethnic | Vertical | Attitudinal | All  | 48 | CSES | Investigate identifiers with API and all with Wagner |
| @tichelbaecker2023we | Both | Attitudinal | Mixed | 10 | Original Survey | |
| @garzia2023affective | Both | Attitudinal | All | 18 | Original Survey | Investigate identifiers with API and all with Wagner |
| @Gidron2023who | Vertical | Attitudinal | Identifiers | 20 | CSES | |
| @Horne2023way | Vertical | Attitudinal | Identifiers | 19 | CSES | |
| @harteveld2021ticking | Vertical | Attitudinal | Identifiers | 40 | CSES | |
| @Comellas2023ideological | Horizontal | Attitudinal | Identifiers | 5 | TRI-POL | Study bloc AP |
| @Fuller2022populism | Both | Both | Attitudinal | 9 | Original Survey | Main analysis uses scaling method on thermo items, further analyses use identifiers with vertical/attitudinal (thermo) measures and horizontal/attitudinal social distance items |
| @harteveld2022affective | Vertical | Attitudinal | Mixed | 28 | CSES | Includes a second analysis with longitudinal data from Netherlands (LISS) |
| @Torcal2022affective | Both | Attitudinal | All | 48 | CSES, CNEP, E-DEM | Code most liked party as in-group |
| @adams2023can | Vertical | Attitudinal | Identifiers | 20 | CSES | | 
| @renstrom2021intergroup | Both | Attitudinal | All | 1 | Original Survey | Code most liked party as in-group |
| @renstrom2022protecting | Both | Attitudinal | All  | 1 | Original Survey | Code most liked party as in-group |
| @hahm2023divided | Horizontal | Behavioral | Identifiers | 25 | Original Survey | Study AP around EU attachement | 
| @kekkonen2022puzzles | Both | Attitudinal | Identifiers | 1 | Original Survey | Compare single and multiple ingroup operationalizations |
| @bettarelli2023regional | Vertical | Attitudinal | All | 30 | CSES | Compare wagner spread aggregated across country and NUTS regions |
| @ryan2023exploring | Vertical | Attitudinal | All | 5 | CSES | Nordic countries |
| @reiljan2021ideological | Vertical | Attitudinal | All | 1 | SOM | Code most liked party as in-group |
| @Knudsen2021affective | Both | Attitudinal | Voters | 2 | Original Survey | Calls for differential treatment of vertical and horizontal AP |
| @segovia2022affective | Vertical | Attitudinal | All | 1 | CEP | notes |


## To-Do / Studies Still to Code

- [ ] Boxell et al. (2024) — cross-national AP trends
- [ ] Lauka, Mann & Frees (2018) — early European cross-national AP
- [ ] Ward & Tavits (2019) — party cues and AP
- [ ] Knudsen (2020) — Norway, horizontal AP
- [ ] Renström et al. (2022) — Sweden
- [ ] Bradley & Chauchard (2022) — cross-national measure correlations
- [ ] Hobolt, Leeper & Tilley (2020) — Brexit & AP, UK
- [ ] Hernández et al. (2021) — electoral cycle effects on AP
- [ ] Garzia et al. (2023) — affective dealignment, multiparty context
- [ ] Torcal & Comellas (2025) — operationalization in multiparty systems




## Collection of potentially useful quotes

> There are nevertheless key aspects that still generate debate. How important are partisan identities for affective polarization? What is relationship between vertical (leader, party) and horizontal (citizen) affective polarization? [@IyengarWagner2025Conceptualizing, S.47]

@kekkonen2022puzzles high correlation between single ingroup and multiple ingroup

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

# Code 1

```{r}
#| label: code-1

# start your code here

```


# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```


  
