---
title: |
  CSAP Code Notebook 2.1
subtitle: |
  eu25games 2019: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts in this notebook are Mermaid diagrams written by R code.
The trees visible when the page opens are drawn by Quarto; those inside tabs are drawn by a small script the first time their tab is opened, because a diagram in a hidden tab cannot be laid out (see `code/00_helper/pid_tree.R`).

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# The Tree

Every composition notebook (2.1 to 2.4) sorts respondents into the same sequential tree, following the typology in the notes of `index.qmd`.

1. **Attachment.** Does the respondent report feeling close to a party?
   The answer is *yes* (a party is named, including "other"), *no*, or *don't know / refused / missing*.
   The last group gets its own grey box so that the shares at every level add up to 100%.
2. **Vote.** Does the respondent report a party vote?
   *No vote reported* collects non-voters, blank or invalid ballots, don't know, refusals and missing answers.
   Where a survey wave never asked the vote item, the branch reads *vote item not fielded* instead, so that "not asked" is not confused with "did not answer".
3. **Match.** Only for respondents who report both, the two parties are compared: *same party*, *different party*, or *not matchable*, when either side is "other" or an unlisted party.

Two leaves are named after the typology: *no attachment, vote reported* is the **vote-anchored** group (the data files call it *implicit*), and *no attachment, no vote* is the non-partisan group.
Leaner probes are deliberately **not** part of the tree.
Where a survey carries one, a dashed note hangs off the *No attachment* box so that it can be picked up later.

Every box gives the number of respondents, and every arrow gives the share of the box above it.

# Load Data

The source is the published harmonised release, not the old analysis frame in `data/03_final/`.
As in notebook 1.5, only respondents who passed the attention check (`der_att_check_3`) are kept.

```{r}
#| label: load-eu25-data

eu25games <- read_rds(
  here(
    "data",
    "01_raw",
    "eu25games2019.rds"
  )
) |>
  # Attention-check passers, as in 1.5
  filter(der_att_check_3)
```

# Select Attachment and Vote Items

The attachment item is the EES wording -- *do you consider yourself to be close to any particular political party? If so, which party do you feel close to?* -- asked in all three waves.
French respondents in wave 2 answered a separate French version of every party item (`q_party_id_fra` and friends), so those columns are coalesced into the main ones.

The vote item concerns the **European Parliament election of May 2019** and exists only in waves 2 and 3.
Wave 2 was fielded around the election and asks for the intended vote, the vote already cast, or a hypothetical vote; wave 3 asks for the vote cast or a hypothetical vote.
They are coalesced in the order intention, hypothetical, past -- the recall-maximising rule of the old pipeline (`code/02_key_variables/2.1_key_variables.qmd`, commit `4651e2c`).
Wave 1 carries no vote item at all.

Every party answer comes with a Party Facts name (`ext_*_pf_name`).
Non-party answers carry a literal sentinel string instead, and parties missing from Party Facts are marked `[og]` -- those are real parties and are kept as such.
So the sentinel strings are matched literally rather than by testing for a missing Party Facts ID.

```{r}
#| label: sentinels

# Non-party answers in Party Facts
pf_no_party <- c(
  "none",
  "dont-know",
  "refused",
  "did-not-vote",
  "blank-invalid",
  "not-eligible"
)

# Named, but not a codeable party
pf_other <- c("other", "other-country", "independent")

# Sentinels actually present
eu25games |>
  transmute(
    pid = coalesce(ext_q_party_id_pf_name, ext_q_party_id_fra_pf_name)
  ) |>
  filter(pid %in% c(pf_no_party, pf_other) | is.na(pid)) |>
  count(pid)
```

The attachment item has no don't-know code.
A respondent either names a party, says "none", or leaves it empty; the empty answers are what the grey *DK / missing* box counts.

```{r}
#| label: select-vars

eu25_tree <- eu25games |>
  transmute(
    wave = meta_wave,
    country_name = meta_country,
    # French wave-2 items folded in
    pid = coalesce(ext_q_party_id_pf_name, ext_q_party_id_fra_pf_name),
    # Intention, hypothetical, past
    vote_party = coalesce(
      ext_q_vote_choice_intended_pf_name,
      ext_q_vote_choice_hypo_pf_name,
      ext_q_vote_choice_past_pf_name,
      ext_q_vote_choice_hypo_fra_pf_name,
      ext_q_vote_choice_past_fra_pf_name
    )
  )
```

# Harmonise to the Tree

```{r}
#| label: harmonise

eu25_tree <- eu25_tree |>
  mutate(
    # Country names to ISO2
    country = recode_values(
      country_name,
      "Austria" ~ "AT",
      "Belgium" ~ "BE",
      "Bulgaria" ~ "BG",
      "Croatia" ~ "HR",
      "Czechia" ~ "CZ",
      "Denmark" ~ "DK",
      "Estonia" ~ "EE",
      "Finland" ~ "FI",
      "France" ~ "FR",
      "Germany" ~ "DE",
      "Greece" ~ "GR",
      "Hungary" ~ "HU",
      "Ireland" ~ "IE",
      "Italy" ~ "IT",
      "Latvia" ~ "LV",
      "Lithuania" ~ "LT",
      "Netherlands" ~ "NL",
      "Poland" ~ "PL",
      "Portugal" ~ "PT",
      "Romania" ~ "RO",
      "Slovakia" ~ "SK",
      "Slovenia" ~ "SI",
      "Spain" ~ "ES",
      "Sweden" ~ "SE",
      "United Kingdom" ~ "GB"
    ),
    # Level 1: attachment
    att = case_when(
      is.na(pid) | pid %in% c("dont-know", "refused") ~ "dk",
      pid %in% pf_no_party ~ "no",
      .default = "yes"
    ),
    # Level 2: vote, unasked in wave 1
    vote = case_when(
      wave == 1 ~ "not_fielded",
      is.na(vote_party) | vote_party %in% pf_no_party ~ "no",
      .default = "yes"
    ),
    # Level 3: same Party Facts name?
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      pid %in% pf_other | vote_party %in% pf_other ~ "unmatchable",
      pid == vote_party ~ "same",
      .default = "different"
    ),
    wave = factor(paste("Wave", wave))
  ) |>
  select(wave, country, att, vote, match, pid, vote_party)

# Every country mapped
stopifnot(all(eu25_tree$country %in% countries_25))
```

# Checks

The raw answers behind each tree level, printed so that any unexpected code is visible.
The first table crosses the three levels by wave, the second lists the most frequent attachment/vote pairs that end up as *different party*.

```{r}
#| label: checks

# Levels by wave
eu25_tree |>
  count(wave, att, vote, match) |>
  print(n = Inf)

# Most common mismatches
eu25_tree |>
  filter(match == "different") |>
  count(country, pid, vote_party, sort = TRUE) |>
  slice_head(n = 15)
```

The unit here is the **respondent-wave**, not the person.
The panel re-interviews part of the sample, so a respondent who took part in two waves is counted twice in the pooled tree, and the attachment item may be answered differently each time.
Attention-check passers also include a few within-wave duplicates of the same panelist, which the published frame does not remove.

```{r}
#| label: panel-structure

# Waves per panelist
eu25games |>
  count(meta_pid, name = "rows") |>
  count(rows, name = "panelists")
```

# Pooled Tree

All 25 countries and all three waves.
Because wave 1 never asked the vote item, a third of the pooled sample sits in the grey *vote item not fielded* boxes; the per-wave trees below separate this out.

```{r}
#| label: tree-pooled
#| output: asis

eu25_tree |>
  tree_counts() |>
  mermaid_tree() |>
  emit_mermaid()
```

# Trees by Wave

```{r}
#| label: tree-wave
#| output: asis

eu25_tree |>
  tree_counts(wave) |>
  emit_tree_tabset(by = "wave", level = 2)
```

# Trees by Country

All three waves pooled within each country.

```{r}
#| label: tree-country
#| output: asis

eu25_tree |>
  tree_counts(country) |>
  emit_tree_tabset(by = "country", level = 2)
```

# Country by Wave

The figure compares the leaves of the tree across countries and waves.
Each bar is one country-wave, and each colour one leaf: blues are respondents who report an attachment, oranges those who report none, and greys the DK and not-fielded groups.
The Mermaid trees below it give the exact counts for every cell.

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-wave
eu25_tree |>
  mutate(wave = fct_relabel(wave, \(x) str_remove(x, "Wave "))) |>
  plot_leaf_shares(x = wave, facet = country) +
  labs(
    title = "Partisan composition in eu25games by country and wave",
    subtitle = "Unweighted shares of attention-check passers",
    x = "Wave"
  )
```

```{r}
#| label: tree-country-wave
#| output: asis

eu25_tree |>
  tree_counts(country, wave) |>
  emit_nested_tabset(outer = "country", inner = "wave", level = 2)
```

# Leaner Probes

The eu25games questionnaire carries no leaner probe.
The only follow-up to the attachment item is its strength (`q_party_closeness`: very close, fairly close, merely a sympathiser), asked of those who name a party and only in wave 1.
So no tree in this notebook has a dashed note.

| Item | Waves | Leaner probe |
|---|---|---|
| `q_party_id` (attachment) | 1, 2, 3 | none |
| `q_party_closeness` (strength) | 1 | not a probe: asked of those who named a party |

# Caveats

- **The vote is a European Parliament vote**, and in wave 2 mostly an intended or hypothetical one.
  Comparing it with a national-election recall in the CSES and ESS notebooks means comparing different quantities.
- **Wave 1 has no vote.** Its trees stop at the attachment level.
- **Respondent-waves, not persons.** See the panel table above.
- **Matching on Party Facts names** treats an electoral alliance and its member party as different parties when Party Facts codes them separately.
  The mismatch table above is dominated by such pairs: Fidesz against the Fidesz–KDNP list in Hungary, Civic Platform against the European Coalition in Poland, Podemos against Unidas Podemos in Spain, PLUS against the USR-PLUS alliance in Romania.
  A member-to-list crosswalk would move these into *same party*; genuine switches (e.g. Conservative attachment, Brexit Party vote in Britain) would stay.
- **Panel duplicates.** A handful of panelists appear many times over (one 29 times), which the attention-check filter does not remove.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
