---
title: |
  CSAP Code Notebook 2.1
subtitle: |
  eu25games 2019: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts in this notebook are Mermaid diagrams written by R code.
The trees visible when the page opens are drawn by Quarto; those inside tabs are drawn by a small script the first time their tab is opened, because a diagram in a hidden tab cannot be laid out (see `code/00_helper/pid_tree.R`).

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# The Tree

Every composition notebook (2.1 to 2.4) sorts respondents into the same sequential tree, following the typology in the notes of `index.qmd`.

1. **Attachment.** Does the respondent report feeling close to a party?
   The answer is *yes* (a party is named), *no*, or *don't know / refused / missing*.
   The last group gets its own grey box so that the shares at every level add up to 100%.
   An "other" answer counts as *yes* only if its open text names a real party (see [Other Answers](#other-answers)).
2. **Vote.** Does the respondent report a party vote?
   *No vote reported* collects non-voters, blank or invalid ballots, don't know, refusals and missing answers, and "other" votes whose open text names no party.
   Where a survey wave never asked the vote item, the branch reads *vote item not fielded* instead, so that "not asked" is not confused with "did not answer".
3. **Match.** Only for respondents who report both, the two parties are compared: *same party* or *different party*.
   Because both sides are always a named party, no respondent in this notebook is *not matchable*.

Two leaves are named after the typology: *no attachment, vote reported* is the **vote-anchored** group (the data files call it *implicit*), and *no attachment, no vote* is the non-partisan group.
Leaner probes are deliberately **not** part of the tree.
Where a survey carries one, a dashed note hangs off the *No attachment* box so that it can be picked up later.

Every box gives the number of respondents, and every arrow gives the share of the box above it.

# Load Data

The source is the published harmonised release, not the old analysis frame in `data/03_final/`.
As in notebook 1.5, only respondents who passed the attention check (`der_att_check_3`) are kept.

```{r}
#| label: load-eu25-data

eu25games <- read_rds(
  here(
    "data",
    "01_raw",
    "eu25games2019.rds"
  )
) |>
  # Attention-check passers, as in 1.5
  filter(der_att_check_3)
```

# Select Attachment and Vote Items

The attachment item is the EES wording -- *do you consider yourself to be close to any particular political party? If so, which party do you feel close to?* -- asked in all three waves.
French respondents in wave 2 answered a separate French version of every party item (`q_party_id_fra` and friends), so those columns are coalesced into the main ones.

The vote item concerns the **European Parliament election of May 2019** and exists only in waves 2 and 3.
Wave 2 was fielded around the election and asks for the intended vote, the vote already cast, or a hypothetical vote; wave 3 asks for the vote cast or a hypothetical vote.
They are coalesced in the order intention, hypothetical, past -- the recall-maximising rule of the old pipeline (`code/02_key_variables/2.1_key_variables.qmd`, commit `4651e2c`).
Wave 1 carries no vote item at all.

Every party answer comes with a Party Facts name (`ext_*_pf_name`).
Non-party answers carry a literal sentinel string instead, and parties missing from Party Facts are marked `[og]` -- those are real parties and are kept as such.
So the sentinel strings are matched literally rather than by testing for a missing Party Facts ID.

```{r}
#| label: sentinels

# Non-party answers in Party Facts
pf_no_party <- c(
  "none",
  "dont-know",
  "refused",
  "did-not-vote",
  "blank-invalid",
  "not-eligible"
)

# Named, but not a codeable party
pf_other <- c("other", "other-country", "independent")

# Sentinels actually present
eu25games |>
  transmute(
    pid = coalesce(ext_q_party_id_pf_name, ext_q_party_id_fra_pf_name)
  ) |>
  filter(pid %in% c(pf_no_party, pf_other) | is.na(pid)) |>
  count(pid)
```

The attachment item has no don't-know code.
A respondent either names a party, says "none", or leaves it empty; the empty answers are what the grey *DK / missing* box counts.

## Other Answers {#other-answers}

Every party item offers an "other" option with an open text field (`q_*_other_raw`), and the release codes that text to Party Facts as well (`ext_q_*_other_pf_name`).
An "other" answer is replaced by what its open text says, so that it counts as attachment or vote only if the text names a real party:

- **A real party:** the party is used, for the tree level and for the match.
  This includes parties Party Facts does not list, which the release marks `[og]`.
- **A clear denial:** *no attachment*, and *no vote reported* on the vote item.
- **Anything else** (don't know, refusal, several parties, a political camp, an independent candidate, junk): *DK / missing* on the attachment item, *no vote reported* on the vote item.

The release codes most open texts, but not all: about half of the open attachment answers and most open vote answers carry no code at all, among them plain party names ("Aontu", "Mislav Kolakušić", "Δημιουργία Ξανά").
These are coded by hand below, from a reading of every uncoded text.

- `other_text_parties` maps a text to a party.
  Where the release already uses a name for that party in that country, the same name is used, so that an attachment and a vote for the same party match.
  A party that appears nowhere else gets a new name ending in `[hand]`.
  A text counts only if it names one party: a party's name or abbreviation (also misspelt or transliterated), its leader, or a candidate on its list.
  Ambiguous abbreviations are left out.
- `other_denials` lists the attachment texts that say the respondent is close to no party.
  Complaints about parties, "a bit of several" and undecided answers do not count.
  The vote item needs no such list, because a denial and anything else both end in *no vote reported*.

```{r}
#| label: other-denials

# Uncoded open attachment texts that deny any attachment
other_denials <- c(
  # AT
  "Jetzt stehe ich keiner politischen Partei mehr nahe",
  # BG
  "ne",
  "не",
  "Никоя партия. Трябва да се премахнат",
  "безпартиен",
  "не съм близък с никоя партия",
  "не чувствам никоя партия близка",
  "нямам симпатии",
  # CZ
  "nyní mne žádná strana svými programy neoslovuje",
  # DK
  "nej",
  "Ikke et bestemt parti. Flere partier kan komme på tale. Min stemmeafgivelse var på en bestemt kandidat fra Venstre.",
  # EE
  "mitte ükski",
  "mitte päris",
  "Olen apoliitilne",
  # FI
  "en mitaan suomalaista",
  # FR
  "plus en ce moment",
  # GR
  "κανενα δεν μου ταιριαζει",
  # HR
  "Nijednoj",
  # HU
  "Egyik politikai párthoz sem érzem magam közel",
  "Egyikhez sem, de a Fidesz politikájával értek egyet",
  "Egyikhez sem.",
  "Nem érzem közel",
  # IE
  "no, i do not feel close to any party",
  "none",
  "I did feal close to fine Geal but now i am not happy with any party",
  # LT
  "Neesu artimas jokiai partijai",
  # LV
  "Neesmu tva nevienai partijai",
  "Nejūtos tuva nevienai partijai",
  "nejūtos tuva nevienai partijai",
  "Partijai nē, bet tuvas sociāldemokrātiskās vērtības",
  # NL
  "geen verbondenheid",
  "niet echt maar ik stem eigenlijk altijd op het CDA al zo lang ik stem stemrecht heb.",
  # PT
  "nenhum tudo a mesma merda",
  "Sou de esquerda mas não me identifico",
  # RO
  "apolitic",
  # SE
  "Inget speciellt parti, men är överlag mer vänster än höger.",
  "Känner inte direkt tillhörighet till något valet är en kompromiss",
  # SI
  "Ne maram nobenega",
  "Ne verjamem več nobeni stranki, ker lažejo kot cigani.",
  "nisem blizu",
  "nobeni",
  "nobeni od teh",
  "za nobeno ker nisem volila",
  # SK
  "vôbec k žiadnej politiku moc nesledujem lebo mám z toho len nervy",
  # ES
  "No me considero proxima a ninguno pero si tubiera que decir uno diria ERC",
  "No me identifico con ninguno, pero prefiero y voto a PSOE"
)

```

```{r}
#| label: other-parties

# Uncoded open texts that name one party
other_text_parties <- tribble(
  ~country, ~party, ~text,
  # AT
  "Austria", "Österreichische Volkspartei",
  c("Türkis", "heißt das, ob ich ein Parteibuch habe? Dann nein. Ansonsten ÖVP"),
  "Austria", "Sozialdemokratische Partei Österreichs",
  "Für keine Partei, aber für einen Kandidaten (SPÖ), obwohl ich diese Partei noch nie gewählt habe",
  "Austria", "KPÖ Plus – European Left, offene Liste[og]", "KPÖ Plus ausProtest",
  "Austria", "Die Weißen [hand]", "Die Weißen",
  "Austria", "GILT [hand]", "GILT",
  "Austria", "Männerpartei [hand]", "Männerpartei",
  # BE
  "Belgium", "Partij van de Arbeid van België — Parti du Travail de Belgique", "PvDa",
  "Belgium", "Liste Destexhe [og]", "destexhe",
  "Belgium", "DierAnimal [og]", "partij van de dieren",
  "Belgium", "Christen-Democratisch en Vlaams",
  "verschillend: CD&V voor Europa, NVA voor Vlaanderen",
  "Belgium", "Lijst Dedecker [hand]", "LIjst Dedecker",
  "Belgium", "Parti des pensionnés [hand]", "Parti des pensionnés",
  "Belgium", "Pirate Party Belgium [hand]", "Pirate Party Belgium",
  "Belgium", "Beveren2020 [hand]", "Beveren2020",
  # BG
  "Bulgaria", "Vazrazhdane", c("\"Възраждане\"", "vazrajdane"),
  "Bulgaria", "Glas Naroden [og]", "glas naroden",
  "Bulgaria", "Pat na mladite [og]",
  c(
    "Движение на младите",
    "Партия на младите",
    "Път за младите",
    "за младите",
    "младите"
  ),
  "Bulgaria", "Koalicija za Balgarija — BSP za Bŭlgariya",
  c(
    "Коалиция за България",
    "гласувах не за партия а за личност - Елена Йончева"
  ),
  "Bulgaria", "Grazhdani za evropeysko razvitie na Balgariya",
  "десислава атанасова",
  "Bulgaria", "Sayuz na Demokratichnite Sili [og]", "преференция сдс",
  "Bulgaria", "Zelena Partija", "партията на зелените",
  "Bulgaria", "Българска пролет [hand]",
  c("Българската пролет", "българската пролет"),
  "Bulgaria", "БНРП [hand]", "БНРП",
  "Bulgaria", "България на труда и разума [hand]", "България на труда и разума",
  "Bulgaria", "Български демократичен форум [hand]", "Български демократичен форум",
  "Bulgaria", "Екогласност [hand]", "Екогласност",
  "Bulgaria", "НДСВ [hand]", "НДСВ",
  "Bulgaria", "Съюз за Пловдив [hand]", "ПП Съюз за Пловдив",
  # HR
  "Croatia", "Lista Mislava Kolakušića [hand]",
  c(
    "Lista Kolakušić",
    "Mislav Kolakušić",
    "Mislav kolakušić",
    "Nezavisna lista Kolakušić",
    "15 Mislav Kolakušić",
    "KANDIDACIJSKA LISTA GRUPE BIRAČA Nositelj kandidacijske liste: MISLAV KOLAKUŠIĆ",
    "KOLAKUŠIĆ",
    "Kalakušić",
    "Kolakušić",
    "Kolakušič",
    "Mislav Kolakušić čak i anketa ide u korist partije",
    "Mislav Kolakušič",
    "Mislav Tiolušić",
    "Mislava Kolakušića",
    "kolakušić",
    "korlakušić",
    "mislav kolakušić",
    "nezavisna stranka Mislava Kolakušića"
  ),
  "Croatia", "Možemo! [og]",
  c(
    "Možemo",
    "MOŽEMO",
    "Možemo! politička platforma",
    "možemo",
    "MOŽEMO! - POLITIČKA PLATFORMA NOVA LJEVICA ODRŽIVI RAZVOJ HRVATSKE - ORaH",
    "Nova ljevica (lista Možemo!, Nova ljevica, ORAH)",
    "Rada Borić"
  ),
  "Croatia", "Nezavisna lista Marijane Petir [og]",
  c(
    "Marijana Petir",
    "Neovisna lista Marijana Petir",
    "Nezavisna lista-Marijana Petir",
    "marijana petir"
  ),
  "Croatia", "Hrvatska konzervativna stranka", c("Ruža Tomašić", "RUŽA TOMAŠIĆ"),
  "Croatia", "Hrvatski suverenisti [og]", c("suverenisti", "Suverenisti"),
  "Croatia", "Radnička fronta [og]", "Radnička dronta",
  "Croatia", "Hrvatska demokršćanska stranka [og]", c("HDS", "hds"),
  "Croatia", "Hrvatska stranka umirovljenika [og]", "HSU",
  "Croatia", "Održivi razvoj Hrvatske", "orah",
  "Croatia", "Reformisti [og]", "Radimir Čačić",
  "Croatia", "Samostalna demokratska srpska stranka",
  "SDSS- Samostalna demokratska srpska stranka",
  "Croatia", "Hrvatski laburisti — Stranka rada", "Stranka Hrvatskih Laburista",
  "Croatia", "Pametno", c("Stranka Pametno", "pametno - dalia orešković"),
  "Croatia", "Snaga [og]", "Stranka snaga",
  "Croatia", "HSP AS [hand]", "HSP AS",
  "Croatia", "Generacija obnove [hand]", "Generacija obnove",
  "Croatia", "Mreža nezavisnih lista [hand]", "MREŽA NEZAVISNIH LISTA",
  "Croatia", "Primorsko-goranski savez [hand]", "PGS",
  "Croatia", "Slavonsko-baranjska hrvatska stranka [hand]",
  c("SBHS - prestala sa djelovanjem", "Slavonsko baranjska hrvatska stranka"),
  "Croatia", "Snaga Slavonije i Baranje [hand]",
  "Snaga Slavonije i Baranje - u osnivanju",
  "Croatia", "Zagreb je naš [hand]", c("Zageb je naš", "Zagreb je naš"),
  "Croatia", "Stranka Ivana Pernara [hand]", "stranka Ivana Pernara",
  # CZ
  "Czechia", "Dělnická strana [og]", "DSSS",
  "Czechia", "Strana zelených", "zeleni",
  "Czechia", "Alternativa pro Českou republiku [hand]",
  c("Alternativa pro Českou republiku", "APČR", "Alternativa pro ČR"),
  "Czechia", "Cesta odpovědné společnosti [hand]",
  c("Cesta odpovědné společnosti", "CESTA"),
  "Czechia", "Patrioti ČR [hand]", "Patrioti ČR",
  "Czechia", "Trikolóra [hand]", c("TRIKOLORA", "Trikolora", "Trikolóra"),
  "Czechia", "Agrární demokratická strana [hand]",
  "agrární demokratická společnost",
  "Czechia", "ANO, vytrollíme europarlament [hand]",
  c(
    "ano vytrollíme europarlament",
    "ANO, vytrollíme europarlament",
    "Ano, vytrolíme Evropský parlament",
    "trolly"
  ),
  "Czechia", "Strana nezávislosti České republiky [hand]",
  "Strana nezávislosti České republiky",
  "Czechia", "Vědci pro Českou republiku [hand]", c("Vědci", "Vědci..."),
  "Czechia", "Strana za práva zvířat [hand]", "Za práva zvířat",
  # DK
  "Denmark", "Folkebevægelsen mod EU",
  c("Folkebhevægelsen mod EU", "mod EU", "mod eu"),
  "Denmark", "Venstre",
  c(
    "Jeg er langt fra enig i alt med mit parti som erVenstre",
    "stemmer kun personligt her en fra partiet venster"
  ),
  "Denmark", "Socialistisk Folkeparti", c("SF", "Personligt på Margrethe Auken"),
  "Denmark", "Stram Kurs", "stram Kurs1",
  "Denmark", "Klaus Riskær Pedersen [hand]", "Claus riskjær",
  "Denmark", "Retsforbundet [hand]", "retsforbundet",
  # EE
  "Estonia", "Eesti Konservatiivne Rahvaerakond",
  c("EKRE", "Toetan EKRE põhimõtteid", "OLeksin hääletanud EKRE poolt", "ekre"),
  "Estonia", "Elurikkuse Erakond", "ERE",
  "Estonia", "Savisaare valimisliit [hand]", "Savisaare valimsliit",
  # FI
  "Finland", "Liberaalipuolue - Vapaus valita [og]",
  c("Liberaali", "Liberaalipuolue", "Liberaalit"),
  "Finland", "Piraattipuolue / Piratpartiet",
  c("Piraatti", "Piraatti Puolue", "Piraatti puolue", "Pirattipuolue"),
  "Finland", "Seitsemän tähden liike [og]",
  c("7 tähden puolue", "Seitsemän tähden liike", "Seitsämän tähden liike"),
  "Finland", "Eläinoikeuspuolue [og]", "Eläioikeuspuolue",
  "Finland", "Suomen Kommunistinen Puolue", "Suomen kommunistinen puolue",
  "Finland", "Feministinen puolue [og]", "feministi puolue",
  # FR
  "France", "Parti animaliste [og]",
  c(
    "Animalis",
    "Liste animaliste",
    "PARTI POUR LES ANIMAUX",
    "parti animlaliste",
    "pour les animaux"
  ),
  "France", "Gilets jaunes [og]", c("liste gilets jaunes", "gillet jaune", "les jaunes"),
  "France", "Alliance jaune [og]", "Allaince jaune",
  "France", "À voix égales [hand]", "A voix égales",
  # DE
  "Germany", "Bündnis Grundeinkommen [og]",
  c(
    "Bündnnis für ein bedingungsloses Grundeinkommen",
    "Bündnis für bedingungslose Grundeinkommen"
  ),
  "Germany", "Partei der Humanisten [og]", c("Die Humanisten", "die Humanisten"),
  "Germany", "Bündnis C – Christen für Deutschland [og]",
  "Bündnis C - Christen für Deutschland",
  "Germany", "DiEM25 [og]", c("DIEM", "DIEM25", "diem"),
  "Germany", "Familien-Partei Deutschlands [og]",
  "Familien-Partei Deutschlands (FAMILIE)",
  "Germany", "Tierschutzpartei",
  c(
    "Partei Mensch Umwelt Tierschutz",
    "Tierschutz, Mensch,Umwelt",
    "partei mensch umwelt tierschutz tierschutzpartei",
    "deutsche tierschützer"
  ),
  "Germany", "Die Grauen [og]", "graue panter",
  "Germany", "Liberal-Konservative Reformer [hand]", c("LKR", "Lucke Partei"),
  "Germany", "Bayernpartei [hand]", "Bayern Partei",
  "Germany", "Demokratie Direkt [hand]", "Demokratie Direkt",
  # GR
  "Greece", "Métopo Evropaikís Realistikís Anypakoís",
  c(
    "DIEM 25",
    "Diem",
    "Diem25",
    "¨ΜΕΡΑ 25",
    "ΜΕΡΑ",
    "ΜέΡΑ25",
    "Μέρα25",
    "Μέρα 25",
    "ΜΕΡΑ24"
  ),
  "Greece", "Dimiourgia Xana [og]",
  c(
    "ΔΗ ΞΑ",
    "ΔΗ.Ξα",
    "ΔΗΜΙΟΥΡΓΙΑ, ΞΑΝΑ",
    "ΔΗΞΑ",
    "Δημιουργία Ξανά",
    "Δημιουργία ξανά",
    "Δημιουργία, ξανά",
    "δηξα",
    "δημιουργία ξανά",
    "δημιουργία ξανά11"
  ),
  "Greece", "Ellada o Allos Dromos [og]",
  c(
    "ΕΛΛΑΔΑ- Ο ΑΛΛΟΣ ΔΡΟΜΟΣ",
    "Ελλάδα, ο Άλλος Δρόμος",
    "ΕΛΛΑΔΑ Ο ΑΛΛΟΣ ΔΡΟΜΟΣ - ΝΟΤΗΣ ΜΑΡΙΑΣ",
    "Ο Άλλος Δρόμος",
    "άλλο δρόμο",
    "μΑΡΙΩς"
  ),
  "Greece", "Antarsya [og]", c("ανταρσία", "Ανταρσια"),
  "Greece", "Agrotiko Ktinotrofiko Komma Ellados [og]", "ΚΑΕΠ",
  "Greece", "Oikologoi Prasinoi", c("Οικολόγοι Πράσινοι", "οικολόγοι πράσινοι"),
  "Greece", "Laïkós Sýndesmos / Chrysí Avgí", "Χρυσή Αυγή",
  "Greece", "Δράση [hand]", "ΔΡΑΣΗ",
  "Greece", "Δραχμή [hand]", "ΔΡΑΧΜΗ",
  "Greece", "ΕΠΑΜ [hand]", "Ε.ΠΑ.Μ.",
  "Greece", "Ελεύθερη Πατρίδα [hand]", c("ΕΛΕΥΘΕΡΗ ΠΑΤΡΙΔΑ", "Ελεύθερη Πατρίδα"),
  "Greece", "Κοινωνία [hand]", "Κοινωνία Συνεχιστές Ιωάννη Καποδίστρια",
  "Greece", "Φιλελεύθερη Συμμαχία [hand]", "Φιλελεύθερη Συμμαχία",
  "Greece", "Ελλήνων Συνέλευσις [hand]", c("Ελλήνων Συνέλευσις", "ελληνων συνελευσι"),
  "Greece", "Όρθια Ελλάδα [hand]", c("ΟΡΘΙΑ ΕΛΛΑΔΑ", "ορθια ελλαδα"),
  "Greece", "Πολίτες [hand]", c("ΠΟΛΙΤΕΣ", "Πολίτες", "Πολίτες (Ψινάκης)"),
  "Greece", "Κόμμα Πειρατών [hand]", "πειρατες",
  # HU
  "Hungary", "A Haza Pártja (A Haza Nem Eladó Mozgalom Párt, HNEM)[og]", "A Haza Pártja",
  "Hungary", "Fidesz — Magyar Polgári Szövetség",
  "Ezek közül, az összes közül kizárólag a Fidesz-...-nak van normális hozzáállása!!! A többi egy sunyi kalap szar!!!",
  "Hungary", "Jobbik Magyarországért Mozgalom",
  "JOBBIK-ra szavaztam, de NAGYON megbántam. Legszívesebben egyikre se szavaznék.",
  "Hungary", "Balpárt [hand]", "Balpárt",
  "Hungary", "MIÉP [hand]", "Miép",
  # IE
  "Ireland", "Aontú [og]", "Aontu",
  "Ireland", "Fianna Fáil", "Fianna Fail",
  # IT
  "Italy", "CasaPound [og]", c("CASAPOUND", "casapound"),
  "Italy", "Partito della Rifondazione Comunista", "Rifondazione Comunista",
  "Italy", "Fratelli d'Italia", "meloni",
  "Italy", "Forza Nuova [hand]", "forza nuova",
  "Italy", "Partito Animalista [hand]", "Partito animalista",
  # LV
  "Latvia", "Attīstībai/Par!", c("Attistibai Par!", "Attīstībai Par", "Atīstībai PAR"),
  "Latvia", "Kam pieder valsts? — Par cilvēcīgu Latviju",
  c("Gobzemam", "kplv", "KPLV", "KPV.LV", "lkpv.lv"),
  "Latvia", "Kristīgi demokrātiskā savienība", c("KDS", "Kristīgo demokrātu savienība"),
  "Latvia", "Progresīvie",
  c(
    "Partija Progresīvie",
    "Progresivie",
    "Progresīvie1",
    "par progresīvajiem",
    "progresivie"
  ),
  "Latvia", "Vienotība", "Visdrīzāk par \"Vienotība\"",
  "Latvia", "Latvijas Krievu savienība / Русский союз Латвии", "par Ždanoku",
  "Latvia", "Reformu partija [hand]", "Reformu partija",
  # LT
  "Lithuania", "Lietuvos Laisvės Sąjunga (Liberalai)", "LLS",
  "Lithuania", "Lietuvos Respublikos Liberalų sąjūdis", "LRLS",
  "Lithuania", "Visuomeninis rinkimų komitetas „Lemiamas šuolis“ [og]", "Lemiamas šuolis",
  "Lithuania", "Visuomeninis rinkimų komitetas „Prezidento Rolando Pakso judėjimas“ [og]",
  "Prezidento Rolando Pakso judėjimas",
  "Lithuania", "Lietuvos sąrašas [hand]", "Lietuvos sąrašas",
  "Lithuania", "Lietuvos liaudies partija [hand]", "lietuvos liaudies partija",
  # NL
  "Netherlands", "Piratenpartij [og]", "Piratenpartij/regio",
  "Netherlands", "Volt [og]", "Volt Nederland",
  "Netherlands", "De Groenen [hand]", "de groenen",
  # PL
  "Poland", "Konfederacja Wolność i Niepodległość", "Konfederacja Korwin",
  "Poland", "Polska Fair Play [og]", "Gwiazdowski Fair Play",
  "Poland", "Federacja dla Rzeczypospolitej [hand]",
  "Federacja na rzecz Rzeczpospolitej Marka Jakubiaka",
  # PT
  "Portugal", "Basta! [og]", "CHEGA",
  "Portugal", "Nós Cidadãos [og]",
  c("Nós, Cidadãos", "Nos Cidadãos", "Nós, Cidadãos!", "nós, cidadãos", "Nós", "nos"),
  "Portugal", "Partido Ecologista — Os Verdes", "Partido ecologista \"os verdes\"",
  "Portugal", "Livre — Liberdade, Esquerda, Europa e Ecologia", "Partido Livre",
  "Portugal", "Partido pelos Animais e pela Natureza", "pan",
  "Portugal", "Juntos pelo Povo [hand]", "JPP",
  # RO
  "Romania", "Uniunea Salvați România",
  c("Alianta 2020", "USR PLUS", "USR plus", "alianta usr plus"),
  "Romania", "Partidul Social Democrat", "pesede",
  "Romania", "Partidul Naţional Ţărănesc Creştin Democrat", "PNTCD",
  "Romania", "Partidul România Unită [og]", "PRU Partidul Romania Unita",
  "Romania", "Prodemo [og]", "Pro Demo",
  "Romania", "UNPR [hand]", "UNPR",
  "Romania", "Partida Romilor [hand]", "PARTIDA ROMILOR",
  # SK
  "Slovakia", "Komunistická strana Slovenska", "Komunistická strana Slovenska",
  "Slovakia", "Strana zelených", "Strana zelenych",
  "Slovakia", "Slovenská Ľudová Strana Andreja Hlinku [og]",
  "slovenska ludova strana ANDREA HLINKU",
  "Slovakia", "Spolu — občianska demokracia", c("SPOLU", "Spolu"),
  "Slovakia", "Ľudová strana Naše Slovensko", "Volil som človeka na kandidátke ĽSNS",
  "Slovakia", "Sloboda a Solidarita", "sas",
  "Slovakia", "Za ľudí [og]", "za ľudi",
  "Slovakia", "Korektúra [hand]", c("Korektura", "Korektúra"),
  "Slovakia", "Starostovia a nezávislí [hand]", "Strana starostov a nezávislých",
  # SI
  "Slovenia", "Solidarnost", "Solidarnosti",
  "Slovenia", "Za zdravo družbo [og]", "Za zdravo družbo.",
  "Slovenia", "Dobra Drzava", "za dobro državo- Bojan Dobovšek",
  "Slovenia", "Zavezništvo Alenke Bratušek / Stranka Alenke Bratušek",
  "ne za stranko, ampak za dr. Angeliko Milinar, pa čeprav je bila za stranko SAB",
  "Slovenia", "Lista novinarja Bojana Požarja [hand]",
  c("Lista novinarja Bojana Požarja", "lista bojana požarja"),
  "Slovenia", "Stranka mladih Slovenije [hand]", "sms stranka mladih slovenije",
  # ES
  "Spain", "Compromís [og]", c("COMPROMIS", "compromis"),
  "Spain", "Izquierda Unida", "IU, pero lo fuí más en el pasado",
  "Spain", "Unidas Podemos", "Unidos Podemos",
  "Spain", "Volt [og]", "VOLT europa",
  "Spain", "Esquerra Republicana de Catalunya", "esquerra repubicana",
  "Spain", "Actúa [hand]",
  c("ACTUA", "Actúa (para Varufakis)", "Actúa (por Iannis Varufakis)"),
  "Spain", "Falange Española de las JONS [hand]", c("FE de las JONS", "Falange"),
  "Spain", "Unión del Pueblo Navarro [hand]", "UPN",
  "Spain", "En Marea [hand]", "en marea",
  "Spain", "Iniciativa Feminista [hand]", "Iniciativa feminista",
  "Spain", "Recortes Cero [hand]", "recortes cero",
  # SE
  "Sweden", "Socialdemokraterna / Sveriges socialdemokratiska arbetareparti",
  c("socialdemokraterna", "SAP men sviktande efter Januariöverenskommelsen", "s"),
  "Sweden", "Miljöpartiet de gröna",
  "MP-Står nära dess grundsten men inte dess politiker och byreokrati.",
  "Sweden", "Medborgerlig samling [og]", "Medborglig samling",
  "Sweden", "Alternativ för Sverige [og]", "altenativ för sverige",
  "Sweden", "Partiet de Fria [hand]",
  c("Partiet de Fria", "Partiet de fria", "partiet de fria"),
  "Sweden", "Klassiskt liberala partiet [hand]", "Klassiskt liberala partiet",
  # GB
  "United Kingdom", "Ulster Unionist Party", "Ulster Unionist Party",
  "United Kingdom", "Libertarian Party [hand]", "Libertarian Party UK",
  "United Kingdom", "Veterans and People's Party [hand]", "VPPP",
  "United Kingdom", "Animal Welfare Party [hand]", "Animal Welfare Party",
  "United Kingdom", "Scottish Green Party [hand]",
  c("Scottish Green Party", "Scottish Greens"),
  "United Kingdom", "Yorkshire Party [hand]", c("The Yorkshire Party", "Yorkshire party")
) |>
  unnest_longer(text)

# One party per text
stopifnot(!anyDuplicated(other_text_parties[c("country", "text")]))

# Replace an "other" answer by what its open text says
resolve_other <- function(answer, other_pf, text, country, denials = character()) {
  text <- str_squish(text)
  hand <- other_text_parties$party[match(
    paste(country, text),
    paste(other_text_parties$country, other_text_parties$text)
  )]
  case_when(
    !answer %in% pf_other ~ answer,
    # Coded upstream to a party or non-party answer
    !is.na(other_pf) & !other_pf %in% pf_other ~ other_pf,
    !is.na(hand) ~ hand,
    text %in% denials ~ "none",
    # Anything else stays "other"
    .default = "other"
  )
}
```

```{r}
#| label: select-vars

eu25_tree <- eu25games |>
  transmute(
    wave = meta_wave,
    country_name = meta_country,
    # French wave-2 items folded in
    pid_answer = coalesce(ext_q_party_id_pf_name, ext_q_party_id_fra_pf_name),
    pid_other_pf = coalesce(
      ext_q_party_id_other_pf_name,
      ext_q_party_id_other_fra_pf_name
    ),
    pid_other_text = coalesce(q_party_id_other_raw, q_party_id_other_fra_raw),
    pid = resolve_other(
      pid_answer,
      pid_other_pf,
      pid_other_text,
      meta_country,
      denials = other_denials
    ),
    # The "other" answer behind the vote
    vote_answer = coalesce(
      ext_q_vote_choice_intended_pf_name,
      ext_q_vote_choice_hypo_pf_name,
      ext_q_vote_choice_past_pf_name,
      ext_q_vote_choice_hypo_fra_pf_name,
      ext_q_vote_choice_past_fra_pf_name
    ),
    # Intention, hypothetical, past
    vote_party = coalesce(
      resolve_other(
        ext_q_vote_choice_intended_pf_name,
        ext_q_vote_choice_intended_other_pf_name,
        q_vote_choice_intended_other_raw,
        meta_country
      ),
      resolve_other(
        ext_q_vote_choice_hypo_pf_name,
        ext_q_vote_choice_hypo_other_pf_name,
        q_vote_choice_hypo_other_raw,
        meta_country
      ),
      resolve_other(
        ext_q_vote_choice_past_pf_name,
        ext_q_vote_choice_past_other_pf_name,
        q_vote_choice_past_other_raw,
        meta_country
      ),
      resolve_other(
        ext_q_vote_choice_hypo_fra_pf_name,
        ext_q_vote_choice_hypo_other_fra_pf_name,
        q_vote_choice_hypo_other_fra_raw,
        meta_country
      ),
      resolve_other(
        ext_q_vote_choice_past_fra_pf_name,
        ext_q_vote_choice_past_other_fra_pf_name,
        q_vote_choice_past_other_fra_raw,
        meta_country
      )
    )
  )

# Every hand-coded text found
open_texts <- eu25games |>
  select(
    country_name = meta_country,
    matches("^q_(party_id|vote_choice_(intended|hypo|past))_other(_fra)?_raw$")
  ) |>
  pivot_longer(-country_name, values_to = "text") |>
  mutate(text = str_squish(text)) |>
  filter(!is.na(text))
stopifnot(
  all(other_denials %in% open_texts$text),
  all(
    paste(other_text_parties$country, other_text_parties$text) %in%
      paste(open_texts$country_name, open_texts$text)
  )
)

# Names reused from the release exist in it
pf_names <- eu25games |>
  select(matches("^ext_q_(party_id|vote_choice).*_pf_name$")) |>
  unlist(use.names = FALSE) |>
  unique()
reused <- other_text_parties |>
  filter(!str_ends(party, fixed("[hand]")))
stopifnot(all(reused$party %in% pf_names))

# How "other" answers resolve
resolution <- function(resolved, other_pf) {
  case_when(
    resolved %in% c(pf_no_party, pf_other) ~ resolved,
    !is.na(other_pf) & !other_pf %in% pf_other ~ "party (coded)",
    .default = "party (by hand)"
  )
}
eu25_tree |>
  filter(pid_answer %in% pf_other) |>
  count(attachment = resolution(pid, pid_other_pf))
eu25_tree |>
  filter(vote_answer %in% pf_other) |>
  count(vote = case_when(
    vote_party %in% c(pf_no_party, pf_other) ~ vote_party,
    .default = "party"
  ))
```

# Harmonise to the Tree

```{r}
#| label: harmonise

eu25_tree <- eu25_tree |>
  mutate(
    # Country names to ISO2
    country = recode_values(
      country_name,
      "Austria" ~ "AT",
      "Belgium" ~ "BE",
      "Bulgaria" ~ "BG",
      "Croatia" ~ "HR",
      "Czechia" ~ "CZ",
      "Denmark" ~ "DK",
      "Estonia" ~ "EE",
      "Finland" ~ "FI",
      "France" ~ "FR",
      "Germany" ~ "DE",
      "Greece" ~ "GR",
      "Hungary" ~ "HU",
      "Ireland" ~ "IE",
      "Italy" ~ "IT",
      "Latvia" ~ "LV",
      "Lithuania" ~ "LT",
      "Netherlands" ~ "NL",
      "Poland" ~ "PL",
      "Portugal" ~ "PT",
      "Romania" ~ "RO",
      "Slovakia" ~ "SK",
      "Slovenia" ~ "SI",
      "Spain" ~ "ES",
      "Sweden" ~ "SE",
      "United Kingdom" ~ "GB"
    ),
    # Level 1: attachment, uncoded other as DK
    att = case_when(
      is.na(pid) | pid %in% c("dont-know", "refused", pf_other) ~ "dk",
      pid %in% pf_no_party ~ "no",
      .default = "yes"
    ),
    # Level 2: vote, unasked in wave 1
    vote = case_when(
      wave == 1 ~ "not_fielded",
      is.na(vote_party) | vote_party %in% c(pf_no_party, pf_other) ~ "no",
      .default = "yes"
    ),
    # Level 3: same Party Facts name?
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      pid == vote_party ~ "same",
      .default = "different"
    ),
    wave = factor(paste("Wave", wave))
  ) |>
  select(wave, country, att, vote, match, pid, vote_party)

# Every country mapped
stopifnot(all(eu25_tree$country %in% countries_25))
```

# Checks

The raw answers behind each tree level, printed so that any unexpected code is visible.
The first table crosses the three levels by wave, the second lists the most frequent attachment/vote pairs that end up as *different party*.

```{r}
#| label: checks

# Levels by wave
eu25_tree |>
  count(wave, att, vote, match) |>
  print(n = Inf)

# Most common mismatches
eu25_tree |>
  filter(match == "different") |>
  count(country, pid, vote_party, sort = TRUE) |>
  slice_head(n = 15)
```

The unit here is the **respondent-wave**, not the person.
The panel re-interviews part of the sample, so a respondent who took part in two waves is counted twice in the pooled tree, and the attachment item may be answered differently each time.
Attention-check passers also include a few within-wave duplicates of the same panelist, which the published frame does not remove.

```{r}
#| label: panel-structure

# Waves per panelist
eu25games |>
  count(meta_pid, name = "rows") |>
  count(rows, name = "panelists")
```

# Pooled Tree

All 25 countries and all three waves.
Because wave 1 never asked the vote item, a third of the pooled sample sits in the grey *vote item not fielded* boxes; the per-wave trees below separate this out.

```{r}
#| label: tree-pooled
#| output: asis

eu25_tree |>
  tree_counts() |>
  mermaid_tree() |>
  emit_mermaid()
```

# Trees by Wave

```{r}
#| label: tree-wave
#| output: asis

eu25_tree |>
  tree_counts(wave) |>
  emit_tree_tabset(by = "wave", level = 2)
```

# Trees by Country

All three waves pooled within each country.

```{r}
#| label: tree-country
#| output: asis

eu25_tree |>
  tree_counts(country) |>
  emit_tree_tabset(by = "country", level = 2)
```

# Country by Wave

The figure compares the leaves of the tree across countries and waves.
Each bar is one country-wave, and each colour one leaf: blues are respondents who report an attachment, oranges those who report none, and greys the DK and not-fielded groups.
The Mermaid trees below it give the exact counts for every cell.

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-wave
eu25_tree |>
  mutate(wave = fct_relabel(wave, \(x) str_remove(x, "Wave "))) |>
  plot_leaf_shares(x = wave, facet = country) +
  labs(
    title = "Partisan composition in eu25games by country and wave",
    subtitle = "Unweighted shares of attention-check passers",
    x = "Wave"
  )
```

```{r}
#| label: tree-country-wave
#| output: asis

eu25_tree |>
  tree_counts(country, wave) |>
  emit_nested_tabset(outer = "country", inner = "wave", level = 2)
```

# Leaner Probes

The eu25games questionnaire carries no leaner probe.
The only follow-up to the attachment item is its strength (`q_party_closeness`: very close, fairly close, merely a sympathiser), asked of those who name a party and only in wave 1.
So no tree in this notebook has a dashed note.

| Item | Waves | Leaner probe |
|---|---|---|
| `q_party_id` (attachment) | 1, 2, 3 | none |
| `q_party_closeness` (strength) | 1 | not a probe: asked of those who named a party |

# Caveats

- **The vote is a European Parliament vote**, and in wave 2 mostly an intended or hypothetical one.
  Comparing it with a national-election recall in the CSES and ESS notebooks means comparing different quantities.
- **Wave 1 has no vote.** Its trees stop at the attachment level.
- **Hand-coded "other" answers.** Open texts the release left uncoded are coded by hand (see [Other Answers](#other-answers)), which is a judgement call for short or misspelt texts.
  A party that appears only in hand-coded texts carries a `[hand]` name, so an attachment to it can only match a vote that was hand-coded as well.
  Votes for independent candidates count as *no vote reported*, and attachment to them as *DK / missing*.
- **Respondent-waves, not persons.** See the panel table above.
- **Matching on Party Facts names** treats an electoral alliance and its member party as different parties when Party Facts codes them separately.
  The mismatch table above is dominated by such pairs: Fidesz against the Fidesz–KDNP list in Hungary, Civic Platform against the European Coalition in Poland, Podemos against Unidas Podemos in Spain, PLUS against the USR-PLUS alliance in Romania.
  A member-to-list crosswalk would move these into *same party*; genuine switches (e.g. Conservative attachment, Brexit Party vote in Britain) would stay.
- **Panel duplicates.** A handful of panelists appear many times over (one 29 times), which the attention-check filter does not remove.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
