---
title: |
  CSAP Code Notebook 2.2
subtitle: |
  EES voter studies 2004-2024: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts are Mermaid diagrams written by R code; trees inside tabs are drawn the first time their tab is opened.
Notebook 2.1 explains the tree itself; it is the same in all four composition notebooks.

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "haven",
  "arrow",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# Load EES Data

The same five files as in notebook 1.5.

```{r}
#| label: load-ees-data

ees_2004 <- read_dta(
  here("data", "01_raw", "ees", "2004", "ZA4566.dta")
)

ees_2009 <- read_sav(
  here("data", "01_raw", "ees", "2009", "ZA5055_v1-1-1.sav")
)

ees_2014 <- read_dta(
  here("data", "01_raw", "ees", "2014", "ZA5160_v4-1-0.dta")
)

ees_2019 <- read_dta(
  here("data", "01_raw", "ees", "2019", "ZA7581_v2-0-1.dta")
)

ees_2024 <- read_csv2(
  here("data", "01_raw", "ees", "2024", "ZA8868_v1-0-0.csv")
)
```

# Select Attachment and Vote Items

The attachment item is the same question in every wave -- *do you consider yourself to be close to any particular party? If so, which party do you feel close to?* -- and it names the party in the same step.
The vote item is the **European Parliament vote** of the election the study accompanies, since the EES is fielded right after each European election.
Every wave also carries a recall of the last **national** general election, which is not used here but could replace the EP vote in a later variant:

| Wave | Attachment | EP vote (used) | National recall (not used) | Leaner probe |
|---|---|---|---|---|
| 2004 | `v211` | `v112` | `v113` | none |
| 2009 | `q87` | `q25` | `q27` | `q89` / `q90` |
| 2014 | `qpp21_ees` | `qp2_ees` | `qpp5_ees` | none |
| 2019 | `Q25n` | `Q7n` | `Q9n` | none |
| 2024 | `q19` (party: `q19n`) | `q6` (party: `q6n`) | `q8` | none |

The items above only *classify* attachment and vote (party, no party, don't know; party vote, no vote).
The party comparison uses the Party Facts coding of notebook 1.4 instead of the release's own codes, which do not line up across the two items in every wave (2014 and 2024 raw codes are shifted against each other, 2019 raw codes follow separate lists):
the attachment party and the EP vote are compared by Party Facts party, with the rule of notebook 1.9, so that a party and an electoral alliance or joint list containing it count as the *same party*, and so do a party and the party it continued in through a rename, refoundation or merger (from the year of the succession on).
The match comes from `data/02_processed/ees_party_match.parquet`, joined on wave and the respondent's position in the wave's file.

The attachment categories follow notebook 1.5, with one difference: don't know, refusal and missing are pooled into the tree's grey *DK* box.

```{r}
#| label: select-2004

# 2004: v211 attachment, v112 EP vote
ees_2004 <- ees_2004 |>
  # Position in the file, the key of notebook 1.4
  mutate(row = row_number()) |>
  # Item unfielded in BE, LT
  filter(!is.na(v211)) |>
  transmute(
    year = 2004,
    row,
    country_raw = as.character(as_factor(country)),
    a = as.numeric(v211),
    v = as.numeric(v112),
    # 94 no, 95 refused, 99 dk/na
    att = case_when(
      a == 94 ~ "no",
      a %in% c(95, 99) ~ "dk",
      .default = "yes"
    ),
    # 1-89 parties, 90 other
    vote = if_else(v %in% 1:90, "yes", "no"),
    att_party = if_else(a %in% 1:89, as.character(a), NA),
    vote_party = if_else(v %in% 1:89, as.character(v), NA)
  )
```

```{r}
#| label: select-2009

# 2009: q87 attachment, q25 EP vote
ees_2009 <- ees_2009 |>
  transmute(
    year = 2009,
    row = row_number(),
    country_raw = as.character(as_factor(t102)),
    a = as.numeric(q87),
    v = as.numeric(q25),
    # 0 no, 77/99 refused, 88 dk
    att = case_when(
      a == 0 ~ "no",
      is.na(a) | a %in% c(77, 88, 99) ~ "dk",
      .default = "yes"
    ),
    # Seven-digit parties, 20/98 other
    vote = if_else(v > 1000 | v %in% c(20, 98), "yes", "no", missing = "no"),
    att_party = if_else(a > 1000, as.character(a), NA),
    vote_party = if_else(v > 1000, as.character(v), NA)
  )
```

```{r}
#| label: select-2014

# 2014: harmonised qpp21_ees, qp2_ees
ees_2014 <- ees_2014 |>
  transmute(
    year = 2014,
    row = row_number(),
    country_raw = as.character(as_factor(countrycode)),
    a = as.numeric(qpp21_ees),
    v = as.numeric(qp2_ees),
    # 0 no, 77 refused, 88 dk
    att = case_when(
      a == 0 ~ "no",
      is.na(a) | a %in% c(77, 88, 99) ~ "dk",
      .default = "yes"
    ),
    # Seven-digit parties, 20/98 other
    vote = if_else(v > 1000 | v %in% c(20, 98), "yes", "no", missing = "no"),
    att_party = if_else(a > 1000, as.character(a), NA),
    vote_party = if_else(v > 1000, as.character(v), NA)
  )
```

```{r}
#| label: select-2019

# 2019: unified Q25n, Q7n
ees_2019 <- ees_2019 |>
  transmute(
    year = 2019,
    row = row_number(),
    country_raw = as.character(as_factor(countrycode)),
    a = as.numeric(Q25n),
    v = as.numeric(Q7n),
    # 0 no, 98 dk, no refusal
    att = case_when(
      a == 0 ~ "no",
      is.na(a) | a == 98 ~ "dk",
      .default = "yes"
    ),
    # 0 no vote, 96 blank, 98/99 missing
    vote = if_else(v %in% c(1:19, 20, 90), "yes", "no", missing = "no"),
    att_party = if_else(a %in% 1:19, as.character(a), NA),
    vote_party = if_else(v %in% 1:19, as.character(v), NA)
  )
```

```{r}
#| label: select-2024

# 2024: raw q19/q6, unified for parties
ees_2024 <- ees_2024 |>
  transmute(
    year = 2024,
    row = row_number(),
    country_raw = country_alpha2,
    a = q19,
    v = q6,
    # 1 no, 98 dk, no refusal
    att = case_when(
      a == 1 ~ "no",
      is.na(a) | a == 98 ~ "dk",
      .default = "yes"
    ),
    # 91 blank, 96 no vote, 98 dk
    vote = if_else(v > 1000 | v == 90, "yes", "no", missing = "no"),
    att_party = if_else(q19n %in% 1:87, as.character(q19n), NA),
    vote_party = if_else(q6n %in% 1:87, as.character(q6n), NA)
  )
```

# Harmonise to the Tree

Country names are mapped onto ISO2 exactly as in notebook 1.5, with Britain and Northern Ireland pooled into `GB` in 2004.

```{r}
#| label: bind

ees_tree <- bind_rows(ees_2004, ees_2009, ees_2014, ees_2019, ees_2024) |>
  mutate(
    # Map every spelling to ISO2
    country = recode_values(
      str_to_lower(country_raw),
      c("austria", "at") ~ "AT",
      c("belgium", "be") ~ "BE",
      c("bulgaria", "bg") ~ "BG",
      c("croatia", "hr") ~ "HR",
      c("cyprus", "cy") ~ "CY",
      c("czech republic", "czech rep.", "cz") ~ "CZ",
      c("denmark", "dk") ~ "DK",
      c("estonia", "ee") ~ "EE",
      c("finland", "fi") ~ "FI",
      c("france", "fr") ~ "FR",
      c("germany", "de") ~ "DE",
      c("greece", "el") ~ "GR",
      c("hungary", "hu") ~ "HU",
      c("ireland", "ie") ~ "IE",
      c("italy", "it") ~ "IT",
      c("latvia", "lv") ~ "LV",
      c("lithuania", "lt") ~ "LT",
      c("luxembourg", "lu") ~ "LU",
      c("malta", "mt") ~ "MT",
      c("netherlands", "the netherlands", "nl") ~ "NL",
      c("poland", "pl") ~ "PL",
      c("portugal", "pt") ~ "PT",
      c("romania", "ro") ~ "RO",
      c("slovakia", "sk") ~ "SK",
      c("slovenia", "si") ~ "SI",
      c("spain", "es") ~ "ES",
      c("sweden", "se") ~ "SE",
      c("britain", "northern ireland", "united kingdom") ~ "GB"
    )
  ) |>
  # Keep the project's 25 democracies
  filter(country %in% countries_25) |>
  mutate(
    year = factor(year)
  )

# Level 3: Party Facts match with the EP vote, with alliances and lineage
# (notebook 1.9); `route` says which table made a pair the same party
ees_match <- read_parquet(here("data", "02_processed", "ees_party_match.parquet")) |>
  transmute(year = factor(wave, levels = levels(ees_tree$year)), row, match_cntry = cntry, pf_match = ext_match_vote_ep, route = ext_route_vote_ep)
ees_pf <- read_parquet(here("data", "02_processed", "ees_party_pf.parquet")) |>
  transmute(year = factor(wave, levels = levels(ees_tree$year)), row, att_pf = ext_pid_pf_name, vote_pf = ext_vote_ep_pf_name)

n_before <- nrow(ees_tree)
ees_tree <- ees_tree |>
  left_join(ees_match, by = c("year", "row")) |>
  left_join(ees_pf, by = c("year", "row")) |>
  # the (wave, row) key lines up with notebook 1.4: same country on every row
  (\(d) {
    stopifnot(!anyNA(d$match_cntry), all(d$country == d$match_cntry))
    d
  })() |>
  mutate(
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      .default = coalesce(pf_match, "unmatchable")
    )
  ) |>
  mutate(route = if_else(match == "same", route, NA)) |>
  select(year, country, att, vote, match, route, a, v, att_pf, vote_pf)
stopifnot(nrow(ees_tree) == n_before)

# Respondents and countries per wave
ees_tree |>
  summarise(
    respondents = n(),
    countries = n_distinct(country),
    .by = year
  )
```

# Checks

The first table crosses the tree levels by wave.
The second is the alignment check described above: among respondents who report both an attachment and a comparable vote, the share naming the *same* party, by country and wave.
A cell far below its neighbours points to a coding problem rather than to voters; the ones that remain low are discussed under the caveats.
The third lists the pairs that are the *same party* only through an alliance or lineage (the audit trail of the tables of notebook 1.9).

```{r}
#| label: checks

# Levels by wave
ees_tree |>
  count(year, att, vote, match) |>
  pivot_wider(names_from = year, values_from = n) |>
  print(n = Inf)

# Same-party share among comparable pairs
ees_tree |>
  filter(match %in% c("same", "different")) |>
  summarise(same = round(mean(match == "same"), 2), .by = c(country, year)) |>
  pivot_wider(names_from = year, values_from = same) |>
  arrange(country) |>
  print(n = Inf)

# Same party through a table, by route
ees_tree |>
  filter(route %in% c("alias", "alliance", "lineage")) |>
  count(route, country, year, att_pf, vote_pf, sort = TRUE) |>
  mutate(across(c(att_pf, vote_pf), \(x) str_trunc(x, 40))) |>
  print(n = 40)
```

# Pooled Tree

All 25 countries and all five waves.
The panel of countries is unbalanced (see notebook 1.5), so this tree mixes different country sets.

```{r}
#| label: tree-pooled
#| output: asis

ees_tree |>
  tree_counts() |>
  mermaid_tree(leaner = "q89, 2009 wave only") |>
  emit_mermaid()
```

# Trees by Wave

```{r}
#| label: tree-wave
#| output: asis

# Probe only fielded in 2009
lean_2009 <- function(counts) {
  if (as.character(counts$year[1]) == "2009") "q89 (then q90 for the party)"
}

ees_tree |>
  tree_counts(year) |>
  emit_tree_tabset(by = "year", level = 2, leaner = lean_2009)
```

# Trees by Country

All five waves pooled within each country.

```{r}
#| label: tree-country
#| output: asis

ees_tree |>
  tree_counts(country) |>
  emit_tree_tabset(by = "country", level = 2, leaner = "q89, 2009 wave only")
```

# Country by Wave

Each bar is one country-wave; blues report an attachment, oranges report none, greys are DK.
Missing bars are waves in which a country was not surveyed, or (Belgium and Lithuania 2004) did not field the attachment item.

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-wave
ees_tree |>
  plot_leaf_shares(x = year, facet = country) +
  scale_x_discrete(labels = \(x) str_sub(x, 3, 4)) +
  labs(
    title = "Partisan composition in the EES by country and wave, 2004-2024",
    subtitle = "Unweighted shares; the vote is the European Parliament vote",
    x = "Wave (20..)"
  )
```

```{r}
#| label: tree-country-wave
#| output: asis

ees_tree |>
  tree_counts(country, year) |>
  emit_nested_tabset(
    outer = "country",
    inner = "year",
    level = 2,
    leaner = lean_2009
  )
```

# Leaner Probes

Only the 2009 wave follows the attachment item with a leaner probe: `q89` (*do you feel yourself a little closer to one of the political parties than others?*), asked of everyone who said no, don't know or refused at `q87`, and `q90` for the party.
The other four waves ask only the strength of attachment of those who named a party (`v212`, `q88`, `qpp22`, `Q26`), which is not a probe.
In the trees, the 2009 *No attachment* box therefore still contains the respondents who would have been classed as leaners.

# Caveats

- **The vote is an EP vote.** European elections are second-order: turnout is low and sincere voting for small parties is common, so *no vote reported* and *different party* are both larger than they would be with a national recall.
  The national recall is listed in the table above for a later comparison.
- **Mode changes between 2014 and 2019**, from interviewer-administered to online panels with no refusal option (see notebook 1.5).
- **Electoral alliances count as the same party as their members** (notebook 1.9): an attachment to HDZ and a vote for the HDZ-led coalition list, or to PO and a vote for the European Coalition, are *same party*.
  So do a party and the party it continued in through a rename, refoundation or merger, from the year of the succession on (the UMP and Les Républicains, the PD and the PDL).
  Two members of one alliance, two predecessors of one successor, and a breakaway and its parent stay *different party*.
  Italy 2004 stays low for a different reason: attachment to the Margherita or the PdCI pairs with a vote for Il Girasole hundreds of times, which points to a mislabelled vote code in the release.
- **The 2004 DK category** also contains no-answers, as noted in notebook 1.5.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
