---
title: |
  CSAP Code Notebook 2.2
subtitle: |
  EES voter studies 2004-2024: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts are Mermaid diagrams written by R code; trees inside tabs are drawn the first time their tab is opened.
Notebook 2.1 explains the tree itself; it is the same in all four composition notebooks.

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "haven",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# Load EES Data

The same five files as in notebook 1.2.

```{r}
#| label: load-ees-data

ees_2004 <- read_dta(
  here("data", "01_raw", "ees", "2004", "ZA4566.dta")
)

ees_2009 <- read_sav(
  here("data", "01_raw", "ees", "2009", "ZA5055_v1-1-1.sav")
)

ees_2014 <- read_dta(
  here("data", "01_raw", "ees", "2014", "ZA5160_v4-1-0.dta")
)

ees_2019 <- read_dta(
  here("data", "01_raw", "ees", "2019", "ZA7581_v2-0-1.dta")
)

ees_2024 <- read_csv2(
  here("data", "01_raw", "ees", "2024", "ZA8868_v1-0-0.csv")
)
```

# Select Attachment and Vote Items

The attachment item is the same question in every wave -- *do you consider yourself to be close to any particular party? If so, which party do you feel close to?* -- and it names the party in the same step.
The vote item is the **European Parliament vote** of the election the study accompanies, since the EES is fielded right after each European election.
Every wave also carries a recall of the last **national** general election, which is not used here but could replace the EP vote in a later variant:

| Wave | Attachment | EP vote (used) | National recall (not used) | Leaner probe |
|---|---|---|---|---|
| 2004 | `v211` | `v112` | `v113` | none |
| 2009 | `q87` | `q25` | `q27` | `q89` / `q90` |
| 2014 | `qpp21_ees` | `qp2_ees` | `qpp5_ees` | none |
| 2019 | `Q25n` | `Q7n` | `Q9n` | none |
| 2024 | `q19` (party: `q19n`) | `q6` (party: `q6n`) | `q8` | none |

The difficult part is the party comparison, because the party codes of the two items do not line up in every file.
Which code to compare was decided by checking, wave by wave and country by country, how often attached voters named the same code on both items.

- **2004** codes both items on one country-specific list (`1`, `2`, ...), so the raw codes compare directly.
- **2009** codes both items with the same seven-digit EES party codes.
- **2014** raw codes are shifted against each other (`qpp21` counts parties from `2`, because `1` is "no"; `qp2` counts from `1`), so the harmonised `_ees` versions are used, which share the seven-digit codes.
- **2019** raw codes do not line up at all; the unified `n` versions (`Q25n`, `Q7n`) do, and match better than the `_EES` codes in Estonia, Lithuania, Slovenia and Sweden.
- **2024** raw codes are shifted in the same way as 2014.
  The unified `q19n` and `q6n` line up, but they code some named parties as "other" (`90`) where the raw item has a party.
  So attachment and vote are *classified* from the raw `q19` and `q6`, as in notebook 1.2, and the unified codes are used only for the *comparison*.
  A party that the unified coding lacks becomes *not matchable*.

The attachment categories follow notebook 1.2, with one difference: don't know, refusal and missing are pooled into the tree's grey *DK* box.

```{r}
#| label: select-2004

# 2004: v211 attachment, v112 EP vote
ees_2004 <- ees_2004 |>
  # Item unfielded in BE, LT
  filter(!is.na(v211)) |>
  transmute(
    year = 2004,
    country_raw = as.character(as_factor(country)),
    a = as.numeric(v211),
    v = as.numeric(v112),
    # 94 no, 95 refused, 99 dk/na
    att = case_when(
      a == 94 ~ "no",
      a %in% c(95, 99) ~ "dk",
      .default = "yes"
    ),
    # 1-89 parties, 90 other
    vote = if_else(v %in% 1:90, "yes", "no"),
    att_party = if_else(a %in% 1:89, as.character(a), NA),
    vote_party = if_else(v %in% 1:89, as.character(v), NA)
  )
```

```{r}
#| label: select-2009

# 2009: q87 attachment, q25 EP vote
ees_2009 <- ees_2009 |>
  transmute(
    year = 2009,
    country_raw = as.character(as_factor(t102)),
    a = as.numeric(q87),
    v = as.numeric(q25),
    # 0 no, 77/99 refused, 88 dk
    att = case_when(
      a == 0 ~ "no",
      is.na(a) | a %in% c(77, 88, 99) ~ "dk",
      .default = "yes"
    ),
    # Seven-digit parties, 20/98 other
    vote = if_else(v > 1000 | v %in% c(20, 98), "yes", "no", missing = "no"),
    att_party = if_else(a > 1000, as.character(a), NA),
    vote_party = if_else(v > 1000, as.character(v), NA)
  )
```

```{r}
#| label: select-2014

# 2014: harmonised qpp21_ees, qp2_ees
ees_2014 <- ees_2014 |>
  transmute(
    year = 2014,
    country_raw = as.character(as_factor(countrycode)),
    a = as.numeric(qpp21_ees),
    v = as.numeric(qp2_ees),
    # 0 no, 77 refused, 88 dk
    att = case_when(
      a == 0 ~ "no",
      is.na(a) | a %in% c(77, 88, 99) ~ "dk",
      .default = "yes"
    ),
    # Seven-digit parties, 20/98 other
    vote = if_else(v > 1000 | v %in% c(20, 98), "yes", "no", missing = "no"),
    att_party = if_else(a > 1000, as.character(a), NA),
    vote_party = if_else(v > 1000, as.character(v), NA)
  )
```

```{r}
#| label: select-2019

# 2019: unified Q25n, Q7n
ees_2019 <- ees_2019 |>
  transmute(
    year = 2019,
    country_raw = as.character(as_factor(countrycode)),
    a = as.numeric(Q25n),
    v = as.numeric(Q7n),
    # 0 no, 98 dk, no refusal
    att = case_when(
      a == 0 ~ "no",
      is.na(a) | a == 98 ~ "dk",
      .default = "yes"
    ),
    # 0 no vote, 96 blank, 98/99 missing
    vote = if_else(v %in% c(1:19, 20, 90), "yes", "no", missing = "no"),
    att_party = if_else(a %in% 1:19, as.character(a), NA),
    vote_party = if_else(v %in% 1:19, as.character(v), NA)
  )
```

```{r}
#| label: select-2024

# 2024: raw q19/q6, unified for parties
ees_2024 <- ees_2024 |>
  transmute(
    year = 2024,
    country_raw = country_alpha2,
    a = q19,
    v = q6,
    # 1 no, 98 dk, no refusal
    att = case_when(
      a == 1 ~ "no",
      is.na(a) | a == 98 ~ "dk",
      .default = "yes"
    ),
    # 91 blank, 96 no vote, 98 dk
    vote = if_else(v > 1000 | v == 90, "yes", "no", missing = "no"),
    att_party = if_else(q19n %in% 1:87, as.character(q19n), NA),
    vote_party = if_else(q6n %in% 1:87, as.character(q6n), NA)
  )
```

# Harmonise to the Tree

Country names are mapped onto ISO2 exactly as in notebook 1.2, with Britain and Northern Ireland pooled into `GB` in 2004.
Party codes are only ever compared within one respondent, so it does not matter that the unified codes of 2019 and 2024 are country-specific indices rather than unique party codes.

```{r}
#| label: bind

ees_tree <- bind_rows(ees_2004, ees_2009, ees_2014, ees_2019, ees_2024) |>
  mutate(
    # Map every spelling to ISO2
    country = recode_values(
      str_to_lower(country_raw),
      c("austria", "at") ~ "AT",
      c("belgium", "be") ~ "BE",
      c("bulgaria", "bg") ~ "BG",
      c("croatia", "hr") ~ "HR",
      c("cyprus", "cy") ~ "CY",
      c("czech republic", "czech rep.", "cz") ~ "CZ",
      c("denmark", "dk") ~ "DK",
      c("estonia", "ee") ~ "EE",
      c("finland", "fi") ~ "FI",
      c("france", "fr") ~ "FR",
      c("germany", "de") ~ "DE",
      c("greece", "el") ~ "GR",
      c("hungary", "hu") ~ "HU",
      c("ireland", "ie") ~ "IE",
      c("italy", "it") ~ "IT",
      c("latvia", "lv") ~ "LV",
      c("lithuania", "lt") ~ "LT",
      c("luxembourg", "lu") ~ "LU",
      c("malta", "mt") ~ "MT",
      c("netherlands", "the netherlands", "nl") ~ "NL",
      c("poland", "pl") ~ "PL",
      c("portugal", "pt") ~ "PT",
      c("romania", "ro") ~ "RO",
      c("slovakia", "sk") ~ "SK",
      c("slovenia", "si") ~ "SI",
      c("spain", "es") ~ "ES",
      c("sweden", "se") ~ "SE",
      c("britain", "northern ireland", "united kingdom") ~ "GB"
    )
  ) |>
  # Keep the project's 25 democracies
  filter(country %in% countries_25) |>
  mutate(
    # Level 3: compare within country
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      is.na(att_party) | is.na(vote_party) ~ "unmatchable",
      att_party == vote_party ~ "same",
      .default = "different"
    ),
    year = factor(year)
  ) |>
  select(year, country, att, vote, match, a, v)

# Respondents and countries per wave
ees_tree |>
  summarise(
    respondents = n(),
    countries = n_distinct(country),
    .by = year
  )
```

# Checks

The first table crosses the tree levels by wave.
The second is the alignment check described above: among respondents who report both an attachment and a comparable vote, the share naming the *same* party, by country and wave.
A cell far below its neighbours points to a coding problem rather than to voters; the ones that remain low are discussed under the caveats.

```{r}
#| label: checks

# Levels by wave
ees_tree |>
  count(year, att, vote, match) |>
  pivot_wider(names_from = year, values_from = n) |>
  print(n = Inf)

# Same-party share among comparable pairs
ees_tree |>
  filter(match %in% c("same", "different")) |>
  summarise(same = round(mean(match == "same"), 2), .by = c(country, year)) |>
  pivot_wider(names_from = year, values_from = same) |>
  arrange(country) |>
  print(n = Inf)
```

# Pooled Tree

All 25 countries and all five waves.
The panel of countries is unbalanced (see notebook 1.2), so this tree mixes different country sets.

```{r}
#| label: tree-pooled
#| output: asis

ees_tree |>
  tree_counts() |>
  mermaid_tree(leaner = "q89, 2009 wave only") |>
  emit_mermaid()
```

# Trees by Wave

```{r}
#| label: tree-wave
#| output: asis

# Probe only fielded in 2009
lean_2009 <- function(counts) {
  if (as.character(counts$year[1]) == "2009") "q89 (then q90 for the party)"
}

ees_tree |>
  tree_counts(year) |>
  emit_tree_tabset(by = "year", level = 2, leaner = lean_2009)
```

# Trees by Country

All five waves pooled within each country.

```{r}
#| label: tree-country
#| output: asis

ees_tree |>
  tree_counts(country) |>
  emit_tree_tabset(by = "country", level = 2, leaner = "q89, 2009 wave only")
```

# Country by Wave

Each bar is one country-wave; blues report an attachment, oranges report none, greys are DK.
Missing bars are waves in which a country was not surveyed, or (Belgium and Lithuania 2004) did not field the attachment item.

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-wave
ees_tree |>
  plot_leaf_shares(x = year, facet = country) +
  scale_x_discrete(labels = \(x) str_sub(x, 3, 4)) +
  labs(
    title = "Partisan composition in the EES by country and wave, 2004-2024",
    subtitle = "Unweighted shares; the vote is the European Parliament vote",
    x = "Wave (20..)"
  )
```

```{r}
#| label: tree-country-wave
#| output: asis

ees_tree |>
  tree_counts(country, year) |>
  emit_nested_tabset(
    outer = "country",
    inner = "year",
    level = 2,
    leaner = lean_2009
  )
```

# Leaner Probes

Only the 2009 wave follows the attachment item with a leaner probe: `q89` (*do you feel yourself a little closer to one of the political parties than others?*), asked of everyone who said no, don't know or refused at `q87`, and `q90` for the party.
The other four waves ask only the strength of attachment of those who named a party (`v212`, `q88`, `qpp22`, `Q26`), which is not a probe.
In the trees, the 2009 *No attachment* box therefore still contains the respondents who would have been classed as leaners.

# Caveats

- **The vote is an EP vote.** European elections are second-order: turnout is low and sincere voting for small parties is common, so *no vote reported* and *different party* are both larger than they would be with a national recall.
  The national recall is listed in the table above for a later comparison.
- **Mode changes between 2014 and 2019**, from interviewer-administered to online panels with no refusal option (see notebook 1.2).
- **Electoral alliances count as different parties.** Where a party runs inside a joint list, an attachment to the party and a vote for the list do not share a code.
  This is visible in the alignment table: Croatia 2014 (HDZ against the HDZ-led coalition list), Romania 2014 and 2019, Poland 2019 (the European Coalition list), Portugal 2014, and Italy and Portugal 2004 sit far below the other cells.
  These are codings, not voters, and a crosswalk from members to lists would move them into *same party*.
- **The 2004 DK category** also contains no-answers, as noted in notebook 1.2.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
