---
title: |
  CSAP Code Notebook 2.3
subtitle: |
  CSES IMD and module 6: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts are Mermaid diagrams written by R code; trees inside tabs are drawn the first time their tab is opened.
Notebook 2.1 explains the tree itself; it is the same in all four composition notebooks.

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# Load CSES Data

As in notebook 1.3, `load()` creates `cses_imd` (modules 1--5) and `cses6` (module 6) as a side effect.

```{r}
#| label: load-cses-data

# Creates the object cses_imd
load(here("data", "01_raw", "cses", "imd", "cses_imd.rdata"))

# Creates the object cses6
load(here("data", "01_raw", "cses", "mod6", "cses6.rdata"))
```

# Select Attachment and Vote Items

**Attachment** is the base item -- *do you usually think of yourself as close to any particular party?* -- coded `1` yes, `0` no, `7` refused, `8` don't know, `9` missing.
The party is asked separately (*which party do you feel closest to?*), so a respondent can be close to a party without a usable party code; such respondents land in *not matchable* at the last level.

**Vote** is the vote in the **current lower-house election**, the election the study follows.
The party-list vote is used where the system has one; in systems without a list vote (code `9999995`, e.g. Britain and France) the district-candidate vote replaces it.
Where the study followed no lower-house election at all (code `9999996`, e.g. the French presidential study of 2022), the vote branch reads *vote item not fielded*.
Blank or invalid ballots, refusals, don't know and abstention count as *no vote reported*.
Abstention and missing share the code `9999999`, so they cannot be told apart.

**Match.** Attachment party and vote are recorded on the same harmonised CSES party codes, so the codes are compared directly.
Codes `9999988`--`9999992` (none of the parties, independent, other left/right/unspecified) name something that is not a codeable party and count as *not matchable*.

The branching leaner probe (`IMD3005_2` / `f3023_2`) is carried along only to flag where it exists.

```{r}
#| label: select-imd

# IMD codes are seven digits
cses_imd_tree <- cses_imd |>
  transmute(
    # Module flags are study-level
    module = case_when(
      IMD1008_MOD_1 == 1 ~ 1,
      IMD1008_MOD_2 == 1 ~ 2,
      IMD1008_MOD_3 == 1 ~ 3,
      IMD1008_MOD_4 == 1 ~ 4,
      IMD1008_MOD_5 == 1 ~ 5
    ),
    study = IMD1004,
    country = IMD1006_UNALPHA2,
    year = IMD1008_YEAR,
    close = IMD3005_1,
    probe = IMD3005_2,
    party = IMD3005_3,
    vote_pl = IMD3002_LH_PL,
    vote_dc = IMD3002_LH_DC,
    other_from = 9999988
  ) |>
  # Drop archive label tables
  as.list() |>
  as_tibble()
```

```{r}
#| label: select-mod6

# Module 6 codes are six digits
cses_mod6_tree <- cses6 |>
  transmute(
    module = 6,
    study = f1004,
    # Module 6 codes Slovenia SL
    country = if_else(f1006_unalpha2 == "SL", "SI", f1006_unalpha2),
    year = f1010_y,
    close = f3023_1,
    probe = f3023_2,
    party = f3023_3,
    vote_pl = f3011_lh_pl,
    vote_dc = f3011_lh_dc,
    other_from = 999988
  ) |>
  # Drop archive label tables
  as.list() |>
  as_tibble()
```

# Harmonise to the Tree

Both releases share one coding scheme, differing only in the number of digits, so they are stacked first and classified once.
The offset `other_from` holds the first non-party code of each release.
As in notebook 1.3, the two studies that never fielded the attachment item (`BELW1999`, `LVA_2018`) are dropped whole.

```{r}
#| label: harmonise

cses_tree <- bind_rows(cses_imd_tree, cses_mod6_tree) |>
  # Drop studies never fielding the item
  filter(!all(close == 9), .by = study) |>
  # Keep the project's 25 democracies
  filter(country %in% countries_25) |>
  mutate(
    # List vote, else district vote
    vote_code = if_else(vote_pl == other_from + 7, vote_dc, vote_pl),
    # Level 1: attachment
    att = case_when(
      close == 1 ~ "yes",
      close == 0 ~ "no",
      .default = "dk"
    ),
    # Level 2: vote, 8 = no election
    vote = case_when(
      vote_code == other_from + 8 ~ "not_fielded",
      vote_code == other_from + 7 ~ "not_fielded",
      vote_code < other_from + 5 ~ "yes",
      .default = "no"
    ),
    # Level 3: shared party codes
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      party >= other_from | vote_code >= other_from ~ "unmatchable",
      party == vote_code ~ "same",
      .default = "different"
    ),
    module = factor(module),
    year = factor(year)
  ) |>
  select(
    module,
    study,
    country,
    year,
    att,
    vote,
    match,
    probe,
    party,
    vote_code
  )

# Respondents, studies and countries per module
cses_tree |>
  summarise(
    respondents = n(),
    studies = n_distinct(study),
    countries = n_distinct(country),
    .by = module
  ) |>
  arrange(module)
```

# Checks

The first table crosses the tree levels by module.
The second gives, per election study, the share naming the *same* party among attached voters whose two codes are both real parties, sorted from the lowest.
The low end of that list is dominated by German studies, where the two most frequent mismatches -- `2760002` against `2760001`, and `2760003` against `2760001` -- are together several times larger than all other German mismatch pairs listed.
That pattern fits a vote coded for the CDU/CSU union against an attachment coded for CDU or CSU separately, and should be checked against part 3 of the CSES codebook before it is corrected.
Romanian studies show the same pattern around electoral alliances.

```{r}
#| label: checks

# Levels by module
cses_tree |>
  count(module, att, vote, match) |>
  pivot_wider(names_from = module, values_from = n) |>
  print(n = Inf)

# Same-party share per study
cses_tree |>
  filter(match %in% c("same", "different")) |>
  summarise(
    pairs = n(),
    same = round(mean(match == "same"), 2),
    .by = study
  ) |>
  arrange(same) |>
  print(n = 15)

# Most frequent German mismatches
cses_tree |>
  filter(country == "DE", match == "different") |>
  count(party, vote_code, sort = TRUE) |>
  slice_head(n = 6)
```

# Pooled Tree

All 25 countries and all six modules.
The unit is the respondent, and each study contributes its own sample size, so countries with more or larger studies weigh more.

```{r}
#| label: tree-pooled
#| output: asis

cses_tree |>
  tree_counts() |>
  mermaid_tree(leaner = "IMD3005_2 / f3023_2 (not in every study)") |>
  emit_mermaid()
```

# Trees by Module

Each module is a five-year window of national elections, so the module is the natural time unit for a pooled series.

```{r}
#| label: tree-module
#| output: asis

cses_tree |>
  mutate(module = fct_relabel(module, \(x) paste("Module", x))) |>
  tree_counts(module) |>
  emit_tree_tabset(
    by = "module",
    level = 2,
    leaner = "IMD3005_2 / f3023_2 (not in every study)"
  )
```

# Trees by Country

All studies of a country pooled.

```{r}
#| label: tree-country
#| output: asis

cses_tree |>
  tree_counts(country) |>
  emit_tree_tabset(
    by = "country",
    level = 2,
    leaner = "IMD3005_2 / f3023_2 (not in every study)"
  )
```

# Country by Election Study

In the CSES the natural cell is the election study, not a calendar year.
Three country-years hold two studies each (Belgium 2019 as Flanders and Wallonia, Germany 2002, Greece 2015); the bars pool them, the trees keep them apart.

```{r}
#| label: probe-studies

# Studies where the probe was asked
probe_studies <- cses_tree |>
  filter(att == "no") |>
  summarise(fielded = any(probe %in% c(0, 1)), .by = study) |>
  filter(fielded) |>
  pull(study)

# Probe note only where fielded
lean_study <- function(counts) {
  if (counts$study[1] %in% probe_studies) "IMD3005_2 / f3023_2"
}
```

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-year
cses_tree |>
  plot_leaf_shares(x = year, facet = country) +
  scale_x_discrete(
    breaks = as.character(seq(1996, 2024, 4)),
    labels = \(x) str_sub(x, 3, 4)
  ) +
  labs(
    title = "Partisan composition in the CSES by country and election, 1996-2024",
    subtitle = "Unweighted shares; the vote is the current lower-house vote",
    x = "Election year (19.. / 20..)"
  )
```

```{r}
#| label: tree-country-study
#| output: asis

cses_tree |>
  tree_counts(country, study) |>
  emit_nested_tabset(
    outer = "country",
    inner = "study",
    level = 2,
    leaner = lean_study
  )
```

# Leaner Probes

The CSES is the one source here with a leaner probe in almost every study: *do you feel yourself a little closer to one of the political parties than the others?* (`IMD3005_2`, `f3023_2`), asked of everyone who does not say yes to the base item.
The table lists the studies in which the probe was **not** fielded for the *no attachment* group.
Among the rest, a large share of the *No attachment* box are leaners -- notebook 1.3 puts them at 15% to 25% of all respondents -- which a later version of the tree can split off.

```{r}
#| label: probe-table

# Studies without the probe
cses_tree |>
  distinct(module, study) |>
  filter(!study %in% probe_studies) |>
  arrange(module, study)
```

# Caveats

- **The vote is a national lower-house vote**, measured right after the election; the EES and eu25games notebooks use a European Parliament vote.
- **Abstention and missing share code `9999999`**, so *no vote reported* mixes both.
- **Close is not naming.** Respondents close to a party with no usable party code are in *not matchable*.
- **CSES folds inconsistent answers into the binary** (notebook 1.3): a "yes" with no party is kept as yes, a "no" with a party as no.
- **Union and alliance codes** push German and Romanian studies towards *different party*, see the checks above.
- **Module 6 is an advance release** covering eight of the 25 countries.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
