---
title: |
  CSAP Code Notebook 2.4
subtitle: |
  ESS rounds 1-11: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts are Mermaid diagrams written by R code; trees inside tabs are drawn the first time their tab is opened.
Notebook 2.1 explains the tree itself; it is the same in all four composition notebooks.

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "arrow",
  "rvest",
  "stringi",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# Load ESS Data

The partitioned Parquet dataset from notebook 1.1, as in notebook 1.4.

```{r}
#| label: load-ess-data

# Partitioned Parquet, never the CSV
ess <- open_dataset(
  here("data", "01_raw", "ess", "parquet"),
  partitioning = hive_partition(cntry = utf8(), essround = int64())
)
```

# The Party Problem

The ESS asks both items as a filter followed by a party:

- **Attachment:** `clsprty` -- *is there a particular political party you feel closer to than all the other parties?* -- then *which one?* in `prtcl<cc>`.
- **Vote:** `vote` -- *did you vote in the last national election?* -- then *which party did you vote for?* in `prtv<cc>`.

The party variables are **country-specific**, and re-lettered whenever a country's party list changes (`prtclde`, `prtclade`, `prtclbde`, ...).
More importantly, the attachment list and the vote list are **numbered independently**: the same party can be `3` on one and `5` on the other.
The Parquet file carries only the numbers.
So the numbers have to be turned back into party names, and the names compared.

The names come from the value labels in the ESS codebook (`data/01_raw/ess/Datafile-subset codebook.html`).
Each `prtcl*` and `prtv*` variable is one block there, with a table of codes and labels.

Two countries record more than one vote.
Germany has a first (constituency) and a second (list) vote, and the **second** vote (`prtv*de2`) is used, as it decides the seat distribution.
Lithuania has a multi-member list vote and two single-member constituency rounds, and the **list** vote (`prtv*lt1`) is used.

```{r}
#| label: codebook-labels

# One block per party variable
codebook <- read_html(
  here("data", "01_raw", "ess", "Datafile-subset codebook.html")
)
blocks <- html_elements(
  codebook,
  xpath = "//h3[starts-with(@id,'prtcl') or starts-with(@id,'prtv')]/.."
)

# Code-label pairs per variable
party_labels <- blocks |>
  map(\(b) {
    tibble(
      variable = html_attr(html_element(b, "h3"), "id"),
      value = as.integer(html_text2(html_elements(b, "td.nowrap"))),
      label = html_text2(html_elements(b, "td:not(.nowrap)"))
    )
  }) |>
  list_rbind() |>
  # Missing codes carry no party
  filter(value < 66)

party_labels |>
  summarise(variables = n_distinct(variable), labels = n())
```

Names are compared after normalising them: lower case, accents removed, punctuation turned into spaces.
Two families of labels are not parties.
*Other* answers ("Other", "Andet - other", "Independent candidate", ...) name something, but nothing that can be matched, so they count as *not matchable*.
*Non-votes* on the vote list ("Blank paper", "Invalid", "Did not vote", ...) count as *no vote reported*.
Both are caught by patterns over the normalised labels, and the full list of what the patterns catch is printed so it can be checked by eye.

```{r}
#| label: label-classes

# Lower case, no accents or punctuation
normalise_party <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |>
    str_to_lower() |>
    str_replace_all("[^a-z0-9]+", " ") |>
    str_squish()
}

party_labels <- party_labels |>
  mutate(
    norm = normalise_party(label),
    # Named, but not a party
    is_other = str_detect(
      norm,
      "^(other|andet|altro|druga|98 other)\\b|^independents?( candidate| nir| self nominated)?$"
    ),
    # Ballots that are not a vote
    is_novote = str_detect(
      norm,
      "blank|^invalid|invalid (vote|ballot)|^nul+$|spoil|did not vote|don t know if|branco|nulo|none of them"
    )
  )

# What the patterns catch
party_labels |>
  filter(is_other | is_novote) |>
  count(is_other, is_novote, label, sort = TRUE) |>
  print(n = Inf)
```

# Select Attachment and Vote Items

For every country, the one attachment variable and the one vote variable that apply to a respondent are the only non-empty ones among the country's party variables.
So the country's variables are stacked long, the empty cells are dropped, and the variable name is kept to look the label up.
This runs country by country, to keep the Parquet reads narrow.

```{r}
#| label: select

# One country at a time
read_country <- function(cc) {
  lc <- str_to_lower(cc)

  # Vote list per country
  vote_pattern <- case_when(
    cc == "DE" ~ paste0("^prtv[a-z]?", lc, "2$"),
    cc == "LT" ~ paste0("^prtv[a-z]?", lc, "1$"),
    .default = paste0("^prtvt[a-z]?", lc, "$")
  )

  raw <- ess |>
    filter(cntry == cc) |>
    select(
      cntry,
      essround,
      name,
      idno,
      clsprty,
      vote,
      matches(paste0("^prtcl[a-z]?", lc, "$")),
      matches(vote_pattern)
    ) |>
    collect() |>
    mutate(
      row = row_number(),
      # Some lists are stored as text
      across(starts_with("prt"), as.integer)
    )

  # Attachment party, long
  att <- raw |>
    select(row, starts_with("prtcl")) |>
    pivot_longer(
      -row,
      names_to = "att_var",
      values_to = "att_value",
      values_drop_na = TRUE
    )

  # Vote party, long
  vot <- raw |>
    select(row, starts_with("prtv")) |>
    pivot_longer(
      -row,
      names_to = "vote_var",
      values_to = "vote_value",
      values_drop_na = TRUE
    )

  raw |>
    select(row, cntry, essround, name, idno, clsprty, vote) |>
    left_join(att, by = "row") |>
    left_join(vot, by = "row")
}

ess_raw <- map(countries_25, read_country) |>
  list_rbind()

# One row per respondent
stopifnot(nrow(ess_raw) == nrow(ess |> filter(cntry %in% countries_25)))
```

# Harmonise to the Tree

**Attachment** is `clsprty`: `1` yes, `2` no, `7`/`8`/`9` DK.
**Vote** is reported if `vote` is `1` and the party answer is a party or an *other* answer; a non-vote label, `66`--`99`, or no voting at all (`vote` `2`, not eligible `3`, or missing) means *no vote reported*.

```{r}
#| label: harmonise

# Label lookups for both items
att_lookup <- party_labels |>
  select(
    att_var = variable,
    att_value = value,
    att_label = label,
    att_norm = norm,
    att_other = is_other
  )
vote_lookup <- party_labels |>
  select(
    vote_var = variable,
    vote_value = value,
    vote_label = label,
    vote_norm = norm,
    vote_other = is_other,
    vote_novote = is_novote
  )

ess_tree <- ess_raw |>
  left_join(att_lookup, by = c("att_var", "att_value")) |>
  left_join(vote_lookup, by = c("vote_var", "vote_value")) |>
  mutate(
    country = cntry,
    # Level 1: attachment
    att = case_when(
      clsprty == 1 ~ "yes",
      clsprty == 2 ~ "no",
      .default = "dk"
    ),
    # Level 2: a party or other vote
    vote = if_else(
      vote == 1 & !is.na(vote_label) & !vote_novote,
      "yes",
      "no",
      missing = "no"
    )
  )
```

# Matching Party Names

Comparing the names exactly is far too strict.
The two lists of one country and round often spell the same party differently -- "PP" against "Partido Popular (PP)", "SMER-Social Democracy" against "Smer – SD", "Fidesz (Fidesz Magyar Polgári Párt)" against "Fidesz - KDNP" -- and an exact comparison scores all of these as *different party*.
So a pair of names counts as the same party if one of four rules holds, tried in this order:

1. **Exact:** the normalised names are identical.
2. **Acronym:** an acronym written in capitals in one label ("PP", "SMER", "HDZ") occurs as a word in the other.
3. **First word:** both names start with the same distinctive word of at least four letters ("fidesz", "sojusz"), after dropping generic words such as *party*, *social*, *democratic* or a country adjective.
4. **Contained:** a name that is a single word occurs as a word in the other ("podemos" in "unidos podemos").

Rules 2 to 4 credit a vote for an **electoral alliance** that carries the party's name or acronym ("Fidesz - KDNP", "HDZ, HSLS") as a vote for the party.
The EES, CSES and eu25games notebooks compare codes and do not do this, so the ESS *same party* share is not directly comparable with theirs; see the caveats.

Every pair matched by rules 2 to 4 is printed below, so that false matches can be spotted.
The `name_overrides` table then forces individual pairs either way.

```{r}
#| label: match-rules

# Words that identify no party
generic_words <- c(
  "the",
  "of",
  "and",
  "for",
  "a",
  "i",
  "na",
  "za",
  "de",
  "la",
  "le",
  "les",
  "del",
  "di",
  "per",
  "des",
  "du",
  "van",
  "voor",
  "fur",
  "und",
  "party",
  "parti",
  "partido",
  "partij",
  "partei",
  "partia",
  "partiya",
  "partidul",
  "part",
  "strana",
  "stranka",
  "puolue",
  "partiet",
  "parties",
  "social",
  "socialist",
  "democratic",
  "democrats",
  "democracy",
  "christian",
  "liberal",
  "liberals",
  "national",
  "people",
  "peoples",
  "popular",
  "union",
  "alliance",
  "coalition",
  "movement",
  "list",
  "new",
  "green",
  "greens",
  "left",
  "right",
  "centre",
  "center",
  "citizens",
  "civic",
  "united",
  "hrvatska",
  "hrvatski",
  "slovenska",
  "ceska",
  "polska",
  "polskie",
  "magyar",
  "magyarorszag",
  "balgarska",
  "bulgarian",
  "lietuvos",
  "latvijas",
  "eesti",
  "suomen",
  "svenska",
  "dansk",
  "danish",
  "irish",
  "movimento",
  "bloco",
  "coligacao",
  "koalicja",
  "koalitsia",
  "koalice",
  "alianta",
  "narodna",
  "demokratska",
  "socijaldemokratska",
  "erakond",
  "eestimaa",
  "jauna",
  "lithuania",
  "folkeparti",
  "labour",
  "lithuanian",
  "slovak",
  "slovenian",
  "czech",
  "hungarian",
  "croatian",
  "polish",
  "estonian",
  "latvian",
  "finnish",
  "swedish",
  "dutch",
  "german",
  "french",
  "spanish",
  "portuguese",
  "italian",
  "greek",
  "british",
  "belgian",
  "austrian",
  "romanian"
)

# Capitalised acronyms in a label
label_acronyms <- function(label) {
  label |>
    str_extract_all("\\b\\p{Lu}[\\p{Lu}\\d]+\\b") |>
    map(normalise_party)
}

# Distinctive words of a name
name_words <- function(norm) {
  str_split(norm, " ") |>
    map(\(w) w[!w %in% generic_words])
}

# One row per distinct name pair
name_pairs <- ess_tree |>
  filter(att == "yes", vote == "yes", !is.na(att_norm)) |>
  filter(!att_other, !vote_other) |>
  distinct(cntry, att_label, att_norm, vote_label, vote_norm) |>
  mutate(
    att_acr = label_acronyms(att_label),
    vote_acr = label_acronyms(vote_label),
    att_all = str_split(att_norm, " "),
    vote_all = str_split(vote_norm, " "),
    att_words = name_words(att_norm),
    vote_words = name_words(vote_norm),
    rule = case_when(
      att_norm == vote_norm ~ "exact",
      map2_lgl(att_acr, vote_all, \(a, b) any(a %in% b)) |
        map2_lgl(vote_acr, att_all, \(a, b) any(a %in% b)) ~ "acronym",
      map2_lgl(
        att_words,
        vote_words,
        \(a, b) {
          length(a) > 0 && length(b) > 0 && a[1] == b[1] && nchar(a[1]) >= 4
        }
      ) ~ "first word",
      !str_detect(att_norm, " ") &
        nchar(att_norm) >= 4 &
        map2_lgl(att_norm, vote_all, \(a, b) a %in% b) ~ "contained",
      .default = "none"
    )
  ) |>
  select(cntry, att_label, att_norm, vote_label, vote_norm, rule)

# How pairs were matched
name_pairs |>
  count(rule)
```

The table lists every pair matched by a rule other than *exact*, with the number of respondents behind it.
This is the audit trail for the matching: a pair that names two different parties belongs in `name_overrides` with `same = FALSE`.

```{r}
#| label: fuzzy-audit

# Non-exact matches, by weight
ess_tree |>
  inner_join(
    name_pairs |> filter(!rule %in% c("exact", "none")),
    by = c("cntry", "att_label", "att_norm", "vote_label", "vote_norm")
  ) |>
  count(cntry, rule, att_label, vote_label, sort = TRUE) |>
  print(n = Inf, width = Inf)
```

```{r}
#| label: overrides

# Force individual pairs either way
name_overrides <- tribble(
  ~cntry , ~att_norm                                 , ~vote_norm                              , ~same ,
  # "PP" is only the party prefix
  "BG"   , "pp red zakonnost i spravedlivost rzs"    , "pp ataka"                              , FALSE ,
  # Venstre and Radikale Venstre differ
  "DK"   , "venstre"                                 , "det radikale venstre"                  , FALSE ,
  # Two different freedom parties
  "LT"   , "party freedom and justice lt"            , "freedom party lp"                      , FALSE ,
  # Same party, other transliteration
  "BG"   , "dvizhenie za prava i svobodi dps"        , "dvijenie za prava i svobodi"           , TRUE  ,
  # Party inside its own alliance
  "BG"   , "balgarska sotsialisticheska partiya bsp" , "koalitziia za balgariia"               , TRUE  ,
  "BG"   , "pp ataka"                                , "koalitziia ataka"                      , TRUE  ,
  "BG"   , "sds"                                     , "koalitziia ods"                        , TRUE  ,
  "PL"   , "platforma obywatelska"                   , "koalicja obywatelska"                  , TRUE  ,
  "CZ"   , "kdu csl christian democratic p"          , "koalice christian dem freedom union"   , TRUE  ,
  "FR"   , "lr les republicains"                     , "les republicains et l udi"             , TRUE  ,
  "FR"   , "renaissance ex la republique en marche"  , "ensemble lrem modem et horizons"       , TRUE  ,
  "IT"   , "forza italia"                            , "popolo delle liberta pdl"              , TRUE  ,
  # Renamed party, earlier name
  "FR"   , "lr les republicains"                     , "ump union pour un mouvement populaire" , TRUE  ,
  "BE"   , "cd v"                                    , "cvp"                                   , TRUE  ,
  "BE"   , "mouvement reformateur mr"                , "prl fdf"                               , TRUE
)

name_pairs <- name_pairs |>
  left_join(name_overrides, by = c("cntry", "att_norm", "vote_norm")) |>
  mutate(same = coalesce(same, rule != "none"))

# Level 3: rule-based name match
ess_tree <- ess_tree |>
  left_join(
    name_pairs |> select(cntry, att_label, vote_label, same),
    by = c("cntry", "att_label", "vote_label")
  ) |>
  mutate(
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      is.na(att_norm) | att_other | vote_other ~ "unmatchable",
      same ~ "same",
      .default = "different"
    )
  )
```

The names that remain unmatched are the other half of the audit.
For every attachment name, the table gives the vote name its holders most often report when no rule links the two.
Most rows are genuine switches or parties that did not stand; a row where both names plainly describe one party (a renamed party, a spelling no rule catches) belongs in `name_overrides` with `same = TRUE`.

```{r}
#| label: unmatched-names

# Most common unmatched pair per name
ess_tree |>
  filter(match == "different") |>
  count(cntry, essround, att_label, vote_label, sort = TRUE) |>
  slice_max(n, n = 1, by = c(cntry, essround, att_label), with_ties = FALSE) |>
  arrange(desc(n)) |>
  print(n = 40, width = Inf)
```

```{r}
#| label: crosswalk

# Pair-level crosswalk, stable order
name_pairs |>
  arrange(cntry, att_norm, vote_norm, att_label, vote_label) |>
  write_csv(here("data", "02_processed", "ess_party_crosswalk.csv"))
```

The pair-level crosswalk is written to `data/02_processed/ess_party_crosswalk.csv`: every attachment/vote name pair that occurs among attached voters, with the rule that matched it (or `none`) and the final decision.

```{r}
#| label: round-labels

ess_tree <- ess_tree |>
  mutate(
    # Round with modal fieldwork year
    round = factor(
      essround,
      levels = 1:11,
      labels = c(
        "R1 2002",
        "R2 2004",
        "R3 2006",
        "R4 2008",
        "R5 2011",
        "R6 2012",
        "R7 2014",
        "R8 2016",
        "R9 2018",
        "R10 2021",
        "R11 2023"
      )
    )
  ) |>
  select(
    round,
    essround,
    country,
    name,
    att,
    vote,
    match,
    att_label,
    vote_label
  )
```

# Checks

```{r}
#| label: checks

# Levels by round
ess_tree |>
  count(essround, att, vote, match) |>
  pivot_wider(names_from = essround, values_from = n) |>
  print(n = Inf)

# Same-party share per country-round
ess_tree |>
  filter(match %in% c("same", "different")) |>
  summarise(
    same = round(mean(match == "same"), 2),
    .by = c(country, essround)
  ) |>
  pivot_wider(names_from = essround, values_from = same, names_sort = TRUE) |>
  arrange(country) |>
  print(n = Inf)
```

# Pooled Tree

All 25 countries and all eleven rounds.

```{r}
#| label: tree-pooled
#| output: asis

ess_tree |>
  tree_counts() |>
  mermaid_tree() |>
  emit_mermaid()
```

# Trees by Round

```{r}
#| label: tree-round
#| output: asis

ess_tree |>
  tree_counts(round) |>
  emit_tree_tabset(by = "round", level = 2)
```

# Trees by Country

All rounds pooled within each country.

```{r}
#| label: tree-country
#| output: asis

ess_tree |>
  tree_counts(country) |>
  emit_tree_tabset(by = "country", level = 2)
```

# Country by Round

Each bar is one country-round; blues report an attachment, oranges report none, greys are DK.
Missing bars are rounds in which a country was not surveyed.

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-round
ess_tree |>
  mutate(essround = factor(essround, levels = 1:11)) |>
  plot_leaf_shares(x = essround, facet = country) +
  scale_x_discrete(breaks = c("1", "6", "11"), drop = FALSE) +
  labs(
    title = "Partisan composition in the ESS by country and round, 2002-2023",
    subtitle = "Unweighted shares; the vote is the last national election",
    x = "ESS round"
  )
```

```{r}
#| label: tree-country-round
#| output: asis

ess_tree |>
  tree_counts(country, round) |>
  emit_nested_tabset(outer = "country", inner = "round", level = 2)
```

# Leaner Probes

The ESS has no leaner probe.
`clsprty` already asks for a party one feels *closer to than all the other parties*, a comparative wording that other surveys use for the probe itself, not for the base item.
Its only follow-up is the strength of closeness (`prtdgcl`), asked of those who named a party.
So the ESS *Close to a party* box is closer to "attachment or leaning" than the EES or CSES base items are, and no tree here carries a dashed note.

# Caveats

- **The vote is the last national election**, which can lie up to four or five years before the interview.
- **Round 10 pools two modes** (face-to-face and self-completion) on disjoint country sets, see notebook 1.4.
- **Matching is by name, not by code.** The rules catch spelling variants but are heuristics; the two audit tables above are the record of what they decided, and `name_overrides` corrects individual pairs.
  Renamed parties are not linked by any rule; the largest cases (UMP and LR, CVP and CD&V, PRL-FDF and MR) are linked by hand in `name_overrides`, smaller ones remain *different party*.
- **Alliances count as the same party here**, when the list carries the party's name or acronym (Fidesz and Fidesz-KDNP, the SDP inside Croatia's Restart coalition, the French parties inside NUPES).
  The EES, CSES and eu25games notebooks compare party codes and count such votes as *different party*, so the ESS *same party* share runs higher for this reason alone.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
