---
title: |
  CSAP Code Notebook 2.4
subtitle: |
  ESS rounds 1-11: partisan composition trees
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

The flowcharts are Mermaid diagrams written by R code; trees inside tabs are drawn the first time their tab is opened.
Notebook 2.1 explains the tree itself; it is the same in all four composition notebooks.

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)


# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "arrow",
  "rvest",
  "stringi",
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)

# Shared tree helpers
source(here("code", "00_helper", "pid_tree.R"))
```

# Load ESS Data

The partitioned Parquet dataset from notebook 1.1, as in notebook 1.7.

```{r}
#| label: load-ess-data

# Partitioned Parquet, never the CSV
ess <- open_dataset(
  here("data", "01_raw", "ess", "parquet"),
  partitioning = hive_partition(cntry = utf8(), essround = int64())
)
```

# The Party Problem

The ESS asks both items as a filter followed by a party:

- **Attachment:** `clsprty` -- *is there a particular political party you feel closer to than all the other parties?* -- then *which one?* in `prtcl<cc>`.
- **Vote:** `vote` -- *did you vote in the last national election?* -- then *which party did you vote for?* in `prtv<cc>`.

The party variables are **country-specific**, and re-lettered whenever a country's party list changes (`prtclde`, `prtclade`, `prtclbde`, ...).
More importantly, the attachment list and the vote list are **numbered independently**: the same party can be `3` on one and `5` on the other.
The Parquet file carries only the numbers.
So the numbers have to be turned back into parties before the two answers can be compared.
Notebook 1.2 does that, coding every code of both lists to Party Facts; this notebook uses the codebook labels only to classify the vote (party, other, non-vote) and compares the parties through the Party Facts coding, with the alliance rule of notebook 1.9.

The names come from the value labels in the ESS codebook (`data/01_raw/ess/Datafile-subset codebook.html`).
Each `prtcl*` and `prtv*` variable is one block there, with a table of codes and labels.

Two countries record more than one vote.
Germany has a first (constituency) and a second (list) vote, and the **second** vote (`prtv*de2`) is used, as it decides the seat distribution.
Lithuania has a multi-member list vote and two single-member constituency rounds, and the **list** vote (`prtv*lt1`) is used.

```{r}
#| label: codebook-labels

# One block per party variable
codebook <- read_html(
  here("data", "01_raw", "ess", "Datafile-subset codebook.html")
)
blocks <- html_elements(
  codebook,
  xpath = "//h3[starts-with(@id,'prtcl') or starts-with(@id,'prtv')]/.."
)

# Code-label pairs per variable
party_labels <- blocks |>
  map(\(b) {
    tibble(
      variable = html_attr(html_element(b, "h3"), "id"),
      value = as.integer(html_text2(html_elements(b, "td.nowrap"))),
      label = html_text2(html_elements(b, "td:not(.nowrap)"))
    )
  }) |>
  list_rbind() |>
  # Missing codes carry no party
  filter(value < 66)

party_labels |>
  summarise(variables = n_distinct(variable), labels = n())
```

Names are compared after normalising them: lower case, accents removed, punctuation turned into spaces.
Two families of labels are not parties.
*Other* answers ("Other", "Andet - other", "Independent candidate", ...) name something, but nothing that can be matched, so they count as *not matchable*.
*Non-votes* on the vote list ("Blank paper", "Invalid", "Did not vote", ...) count as *no vote reported*.
Both are caught by patterns over the normalised labels, and the full list of what the patterns catch is printed so it can be checked by eye.

```{r}
#| label: label-classes

# Lower case, no accents or punctuation
normalise_party <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |>
    str_to_lower() |>
    str_replace_all("[^a-z0-9]+", " ") |>
    str_squish()
}

party_labels <- party_labels |>
  mutate(
    norm = normalise_party(label),
    # Named, but not a party
    is_other = str_detect(
      norm,
      "^(other|andet|altro|druga|98 other)\\b|^independents?( candidate| nir| self nominated)?$"
    ),
    # Ballots that are not a vote
    is_novote = str_detect(
      norm,
      "blank|^invalid|invalid (vote|ballot)|^nul+$|spoil|did not vote|don t know if|branco|nulo|none of them"
    )
  )

# What the patterns catch
party_labels |>
  filter(is_other | is_novote) |>
  count(is_other, is_novote, label, sort = TRUE) |>
  print(n = Inf)
```

# Select Attachment and Vote Items

For every country, the one attachment variable and the one vote variable that apply to a respondent are the only non-empty ones among the country's party variables.
So the country's variables are stacked long, the empty cells are dropped, and the variable name is kept to look the label up.
This runs country by country, to keep the Parquet reads narrow.

```{r}
#| label: select

# One country at a time
read_country <- function(cc) {
  lc <- str_to_lower(cc)

  # Vote list per country
  vote_pattern <- case_when(
    cc == "DE" ~ paste0("^prtv[a-z]?", lc, "2$"),
    cc == "LT" ~ paste0("^prtv[a-z]?", lc, "1$"),
    .default = paste0("^prtvt[a-z]?", lc, "$")
  )

  raw <- ess |>
    filter(cntry == cc) |>
    select(
      cntry,
      essround,
      name,
      idno,
      clsprty,
      vote,
      matches(paste0("^prtcl[a-z]?", lc, "$")),
      matches(vote_pattern)
    ) |>
    collect() |>
    mutate(
      row = row_number(),
      # Some lists are stored as text
      across(starts_with("prt"), as.integer)
    )

  # Attachment party, long
  att <- raw |>
    select(row, starts_with("prtcl")) |>
    pivot_longer(
      -row,
      names_to = "att_var",
      values_to = "att_value",
      values_drop_na = TRUE
    )

  # Vote party, long
  vot <- raw |>
    select(row, starts_with("prtv")) |>
    pivot_longer(
      -row,
      names_to = "vote_var",
      values_to = "vote_value",
      values_drop_na = TRUE
    )

  raw |>
    select(row, cntry, essround, name, idno, clsprty, vote) |>
    left_join(att, by = "row") |>
    left_join(vot, by = "row")
}

ess_raw <- map(countries_25, read_country) |>
  list_rbind()

# One row per respondent
stopifnot(nrow(ess_raw) == nrow(ess |> filter(cntry %in% countries_25)))
```

# Harmonise to the Tree

**Attachment** is `clsprty`: `1` yes, `2` no, `7`/`8`/`9` DK.
**Vote** is reported if `vote` is `1` and the party answer is a party or an *other* answer; a non-vote label, `66`--`99`, or no voting at all (`vote` `2`, not eligible `3`, or missing) means *no vote reported*.

```{r}
#| label: harmonise

# Label lookups for both items
att_lookup <- party_labels |>
  select(
    att_var = variable,
    att_value = value,
    att_label = label,
    att_norm = norm,
    att_other = is_other
  )
vote_lookup <- party_labels |>
  select(
    vote_var = variable,
    vote_value = value,
    vote_label = label,
    vote_norm = norm,
    vote_other = is_other,
    vote_novote = is_novote
  )

ess_tree <- ess_raw |>
  left_join(att_lookup, by = c("att_var", "att_value")) |>
  left_join(vote_lookup, by = c("vote_var", "vote_value")) |>
  mutate(
    country = cntry,
    # Level 1: attachment
    att = case_when(
      clsprty == 1 ~ "yes",
      clsprty == 2 ~ "no",
      .default = "dk"
    ),
    # Level 2: a party or other vote
    vote = if_else(
      vote == 1 & !is.na(vote_label) & !vote_novote,
      "yes",
      "no",
      missing = "no"
    )
  )
```

# Matching Parties

Attachment and vote are compared by their Party Facts coding (notebook 1.2), not by the codebook names: the two lists of one country and round spell the same party differently ("PP" against "Partido Popular (PP)", "Smer – SD" against "SMER-Social Democracy"), and the coding has already resolved every spelling to one Party Facts party.
The match level comes from `data/02_processed/ess_party_match.parquet` (notebook 1.9), which applies one rule to all four sources:

- **same party**: the same Party Facts party; a party and an electoral alliance or joint list that contains it, in either direction (Fidesz and the Fidesz–KDNP list, PO and Koalicja Obywatelska); or a party and the party it continued in through a rename, refoundation or merger, from the year of the succession on (the UMP and Les Républicains, the CVP and CD&V, the PRL-FDF and the MR);
- **different party**: two different parties, including two members of one alliance, two predecessors of one successor, and a breakaway and its parent;
- **not matchable**: the attachment party or the vote is an *other* answer, or has no party.

```{r}
#| label: match

ess_pf <- read_parquet(here("data", "02_processed", "ess_party_pf.parquet")) |>
  select(cntry, essround, name, idno, att_pf = ext_prtcl_pf_name, vote_pf = ext_prtv_pf_name)
ess_match <- read_parquet(here("data", "02_processed", "ess_party_match.parquet")) |>
  select(cntry, essround, name, idno, pf_match = ext_match_prtv, route = ext_route_prtv)

n_before <- nrow(ess_tree)
ess_tree <- ess_tree |>
  left_join(ess_pf, by = c("cntry", "essround", "name", "idno")) |>
  left_join(ess_match, by = c("cntry", "essround", "name", "idno")) |>
  mutate(
    # Level 3: Party Facts match, with alliances and lineage (notebook 1.9)
    match = case_when(
      att != "yes" | vote != "yes" ~ NA,
      .default = coalesce(pf_match, "unmatchable")
    ),
    route = if_else(match == "same", route, NA)
  )
stopifnot(nrow(ess_tree) == n_before)
```

Pairs that count as the same party only through an alliance or lineage, by route and weight: the audit trail of the tables of notebook 1.9.

```{r}
#| label: route-audit

ess_tree |>
  filter(route %in% c("alias", "alliance", "lineage")) |>
  count(route, cntry, att_pf, vote_pf, sort = TRUE) |>
  print(n = 40, width = Inf)
```

The largest pairs that remain *different party*: most are genuine switches or parties that did not stand.

```{r}
#| label: unmatched-names

ess_tree |>
  filter(match == "different") |>
  count(cntry, att_pf, vote_pf, sort = TRUE) |>
  print(n = 40, width = Inf)
```

```{r}
#| label: crosswalk

# Pair-level crosswalk, stable order
ess_tree |>
  filter(match %in% c("same", "different", "unmatchable")) |>
  count(cntry, att_label, att_pf, vote_label, vote_pf, match, route, name = "n") |>
  arrange(cntry, att_pf, vote_pf, att_label, vote_label) |>
  write_csv(here("data", "02_processed", "ess_party_crosswalk.csv"))
```

The pair-level table is written to `data/02_processed/ess_party_crosswalk.csv`: every attachment/vote label pair among attached voters, with the two Party Facts names, the match, its route (identical, alliance or lineage) and the number of respondents.

```{r}
#| label: round-labels

ess_tree <- ess_tree |>
  mutate(
    # Round with modal fieldwork year
    round = factor(
      essround,
      levels = 1:11,
      labels = c(
        "R1 2002",
        "R2 2004",
        "R3 2006",
        "R4 2008",
        "R5 2011",
        "R6 2012",
        "R7 2014",
        "R8 2016",
        "R9 2018",
        "R10 2021",
        "R11 2023"
      )
    )
  ) |>
  select(
    round,
    essround,
    country,
    name,
    att,
    vote,
    match,
    att_label,
    vote_label
  )
```

# Checks

```{r}
#| label: checks

# Levels by round
ess_tree |>
  count(essround, att, vote, match) |>
  pivot_wider(names_from = essround, values_from = n) |>
  print(n = Inf)

# Same-party share per country-round
ess_tree |>
  filter(match %in% c("same", "different")) |>
  summarise(
    same = round(mean(match == "same"), 2),
    .by = c(country, essround)
  ) |>
  pivot_wider(names_from = essround, values_from = same, names_sort = TRUE) |>
  arrange(country) |>
  print(n = Inf)
```

# Pooled Tree

All 25 countries and all eleven rounds.

```{r}
#| label: tree-pooled
#| output: asis

ess_tree |>
  tree_counts() |>
  mermaid_tree() |>
  emit_mermaid()
```

# Trees by Round

```{r}
#| label: tree-round
#| output: asis

ess_tree |>
  tree_counts(round) |>
  emit_tree_tabset(by = "round", level = 2)
```

# Trees by Country

All rounds pooled within each country.

```{r}
#| label: tree-country
#| output: asis

ess_tree |>
  tree_counts(country) |>
  emit_tree_tabset(by = "country", level = 2)
```

# Country by Round

Each bar is one country-round; blues report an attachment, oranges report none, greys are DK.
Missing bars are rounds in which a country was not surveyed.

```{r}
#| label: fig-leaves
#| fig-width: 11
#| fig-height: 11

# Leaf shares per country-round
ess_tree |>
  mutate(essround = factor(essround, levels = 1:11)) |>
  plot_leaf_shares(x = essround, facet = country) +
  scale_x_discrete(breaks = c("1", "6", "11"), drop = FALSE) +
  labs(
    title = "Partisan composition in the ESS by country and round, 2002-2023",
    subtitle = "Unweighted shares; the vote is the last national election",
    x = "ESS round"
  )
```

```{r}
#| label: tree-country-round
#| output: asis

ess_tree |>
  tree_counts(country, round) |>
  emit_nested_tabset(outer = "country", inner = "round", level = 2)
```

# Leaner Probes

The ESS has no leaner probe.
`clsprty` already asks for a party one feels *closer to than all the other parties*, a comparative wording that other surveys use for the probe itself, not for the base item.
Its only follow-up is the strength of closeness (`prtdgcl`), asked of those who named a party.
So the ESS *Close to a party* box is closer to "attachment or leaning" than the EES or CSES base items are, and no tree here carries a dashed note.

# Caveats

- **The vote is the last national election**, which can lie up to four or five years before the interview.
- **Round 10 pools two modes** (face-to-face and self-completion) on disjoint country sets, see notebook 1.7.
- **Matching is by Party Facts party, with alliances and lineage** (notebook 1.9); the same rule is used in all four composition notebooks, so their *same party* shares are comparable.
  A party and an electoral alliance or joint list containing it count as the same party, and so do a party and the party it continued in through a rename, refoundation or merger (UMP and Les Républicains, CVP and CD&V, PRL-FDF and MR), in surveys from the year of the succession on.
  Because the vote is the last national election, this is where lineage matters most: a respondent interviewed in 2016 who feels close to Les Républicains and voted UMP in 2012 has not switched.
  Two members of one alliance (CDU and CSU), two predecessors of one successor, and a breakaway and its parent stay *different party*.

# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
