---
title: |
  Code Notebook 3.3
subtitle: |
  Party attachment over time in EU member states: Mannheim Eurobarometer Trend File and CSES
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()

# set width of console output
options(width = 80)

# no randomness is used below, but the seed is set so the notebook is
# reproducible if a sampling step is ever added
set.seed(20260804)

# Install and load required packages
p_required <- c(
  "tidyverse",
  "here",
  "ggpubr", # theme_pubr()
  "haven", # read_dta() for the Stata trend file
  "arrow", # parquet cache
  "sessioninfo"
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

This notebook recovers country-level distributions of **felt party attachment** over time
in EU member states from two independent instruments, and keeps them as **side-by-side
series**. Nothing is spliced, bridged or calibrated across sources.

The quantity of primary interest is the share of respondents who **deny any party
attachment yet report a vote** — the historical prevalence of the group this project calls
*vote-anchored*.

## Switches

```{r}
#| label: switches

# "ever": any country that has ever been an EU/EC member enters with all of its
#         available years (the UK is retained for its full run).
# "from_accession": a country enters only from its accession year.
eu_scope <- "ever"

# Eurobarometer `closepty == 3` ("merely a sympathiser") counts as attached.
# Set FALSE to recompute every table with sympathisers as not attached.
sympathizer_as_attached <- TRUE

# Country-year cells below this Kish effective N are flagged unreliable.
min_eff_n <- 100

# Read parsed sources from the parquet cache when it exists.
use_cache <- TRUE
```

## Paths

```{r}
#| label: paths

path_imd <- here(
  "data", "01_raw", "external", "cses", "imd", "cses_imd.rdata"
)
path_mod6 <- here(
  "data", "01_raw", "external", "cses", "mod6", "cses6.rdata"
)
path_eb <- here(
  "data", "01_raw", "external", "eb",
  "mannheim-eurobarometer-trend-file-1970-2002",
  "ZA3521_v2-0-1.dta"
)

# `.gitignore` carries a blanket `*cache*` rule, so this directory is untracked
# by design -- it holds derived artefacts only and is rebuilt from the raw files.
dir_cache <- here("data", "02_processed", "pid_cache")
if (!dir.exists(dir_cache)) {
  dir.create(dir_cache, recursive = TRUE)
}

dir_final <- here("data", "03_final")
```

# Sources and what is in them

Three files were inventoried. **CSES Module 5 is deliberately excluded**: the Integrated
Module Dataset already contains it in full (114,714 rows flagged `IMD1008_MOD_5 == 1`, and
all 56 of its election studies present in IMD), so stacking the two would double-count
every Module 5 election. Module 6 is disjoint from both.

An earlier version of this passage said 55 of 56 IDs matched. That was a naming artefact:
the IMD splits Greece 2015 into `GRC12015` and `GRC22015` while Module 5 carries a single
`GRC_2015`, which is the IMD's `GRC22015` (1,078 rows in both). See
`code/03_explanal/3.5_cses_partisanship.qmd`, which prints the evidence.

| | CSES IMD | CSES Module 6 | Mannheim EB Trend File |
|---|---|---|---|
| file | `cses_imd.rdata` | `cses6.rdata` | `ZA3521_v2-0-1.dta` |
| dim | 395,797 x 406 | 33,871 x 676 | 1,134,384 x 146 |
| name case | UPPER | **lower** | lower |
| unit | respondent x election study | respondent x election study | respondent x EB wave |
| span | elections 1996-2021 | elections 2022-2024 | waves 1970-2002 |

CSES stores its metadata Stata-style at the **data-frame** level — `attr(d, "var.labels")`
runs parallel to `names(d)`, and `attr(d, "label.table")` is a named list of value-label
vectors. The per-column `attr(x, "label")` that `haven` would set is `NULL`, so a naive
label search over the columns finds nothing.

Module 6 **renumbers** the attachment battery to `f3023_*`. In Module 6, `f3005_*` is an
unrelated item ("country better run by"). No variable code is assumed to carry across
releases; each source is read and harmonised on its own terms.

## Item wording and coding

```{r}
#| label: tbl-items
#| tbl-cap: "Source variables carrying the attachment and vote constructs."

tribble(
  ~construct,            ~imd,            ~mod6,          ~eb,
  "close to any party",  "IMD3005_1",     "f3023_1",      "closepty",
  "leaner probe",        "IMD3005_2",     "f3023_2",      NA,
  "which party",         "IMD3005_3",     "f3023_3",      "feelclo",
  "strength",            "IMD3005_4",     "f3023_4",      "(in closepty)",
  "turnout",             "IMD3001_LH",    "f3010",        "(from vote item)",
  "vote, past election", "IMD3002_LH_PL", "f3011_lh_pl",  "lastvote",
  "vote intention",      NA,              NA,             "voteint",
  "ISO3",                "IMD1006_UNALPHA3", "f1006_unalpha3", "(from nation1)",
  "EU member flag",      "IMD1006_EU",    "f1007_eu",     "(accession table)",
  "fieldwork year",      "IMD1013_Y",     "f1019_y",      "year",
  "election year",       "IMD1008_YEAR",  "f1009",        NA,
  "survey id",           "IMD1004",       "f1004",        "eb",
  "respondent id",       "IMD1005",       "f1003_1",      "id",
  "weight used",         "IMD1010_1",     "f1101_1",      "wnation"
) |>
  knitr::kable()
```

The two instruments do **not** ask the same question.

- **CSES** asks a binary — "do you usually think of yourself as close to any particular
  party?" — and then probes those who say no with a leaner follow-up, then asks which
  party and how close. Don't-know and refused are *volunteered and separately coded*
  (`7` refused, `8` don't know), so the three-way harmonisation is directly supported and
  nothing has to be reconstructed.
- **Eurobarometer** asks a single degree-of-closeness item: very close / fairly close /
  merely a sympathiser / not close to any party. Status and strength arrive in one
  variable, and there is no leaner probe.

The Eurobarometer bar is therefore lower and its categories are not nested the way CSES's
are. Consistent with section E of `data/01_raw/external/external-data.md`: **compare
slopes within a source, never levels across sources.**

## Coverage: the binding constraint

Substantive-response counts confirm that the Eurobarometer attachment item runs
**1975-1994 only** — not the full 1970-2002 span of the file. `voteint` runs almost the
whole span; `lastvote` is intermittent; `feelclo` is erratic and carries party codes
rather than a status, so it is used for `pid_party` only and never as the status item.

The consequence for this notebook: the Eurobarometer attachment x vote cross-tab exists
for 1975-1994, and because CSES elections begin in 1996, **the two sources barely
overlap**. The per-country level-offset comparison in the comparability section below is
therefore mostly a documented absence rather than a table of offsets. That is a finding
about the instruments, not a failure of the pipeline, and it is reported as such.

# Scope tables

`nation1` is a bespoke Eurobarometer code that `countrycode` does not know, so the mapping
is written out explicitly. It is also the only variable that keeps **West and East
Germany** and **Great Britain and Northern Ireland** apart — `nation2` pools both pairs,
and is not used.

```{r}
#| label: eb-country-map

eb_country_map <- tribble(
  ~nation1, ~nation1_label,      ~country_unit, ~country_iso3, ~country_name,
  1L,       "france",            "FRA",         "FRA",         "France",
  2L,       "belgium",           "BEL",         "BEL",         "Belgium",
  3L,       "netherlands",       "NLD",         "NLD",         "Netherlands",
  4L,       "GERMANY-WEST",      "DEU-W",       "DEU",         "Germany (West)",
  5L,       "italy",             "ITA",         "ITA",         "Italy",
  6L,       "luxembourg",        "LUX",         "LUX",         "Luxembourg",
  7L,       "denmark",           "DNK",         "DNK",         "Denmark",
  8L,       "ireland",           "IRL",         "IRL",         "Ireland",
  9L,       "GREAT BRITAIN",     "GBR-GB",      "GBR",         "Great Britain",
  10L,      "NORTHERN IRELAND",  "GBR-NI",      "GBR",         "Northern Ireland",
  11L,      "greece",            "GRC",         "GRC",         "Greece",
  12L,      "spain",             "ESP",         "ESP",         "Spain",
  13L,      "portugal",          "PRT",         "PRT",         "Portugal",
  14L,      "GERMANY-EAST",      "DEU-E",       "DEU",         "Germany (East)",
  15L,      "norway",            "NOR",         "NOR",         "Norway",
  16L,      "finland",           "FIN",         "FIN",         "Finland",
  17L,      "sweden",            "SWE",         "SWE",         "Sweden",
  18L,      "austria",           "AUT",         "AUT",         "Austria",
  19L,      "switzerland",       "CHE",         "CHE",         "Switzerland"
)
```

Membership is defined on `country_iso3`, so the German and British units inherit the
membership of their parent state. Norway and Switzerland have never been members and drop
out. `exit_year` is carried because the UK left in 2020 and the `"ever"` scope has to be
an explicit choice rather than an accident of coding.

```{r}
#| label: eu-accession

eu_accession <- tribble(
  ~country_iso3, ~accession_year, ~exit_year,
  "BEL", 1958L, NA_integer_,
  "DEU", 1958L, NA_integer_,
  "FRA", 1958L, NA_integer_,
  "ITA", 1958L, NA_integer_,
  "LUX", 1958L, NA_integer_,
  "NLD", 1958L, NA_integer_,
  "DNK", 1973L, NA_integer_,
  "IRL", 1973L, NA_integer_,
  "GBR", 1973L, 2020L,
  "GRC", 1981L, NA_integer_,
  "ESP", 1986L, NA_integer_,
  "PRT", 1986L, NA_integer_,
  "AUT", 1995L, NA_integer_,
  "FIN", 1995L, NA_integer_,
  "SWE", 1995L, NA_integer_,
  "CYP", 2004L, NA_integer_,
  "CZE", 2004L, NA_integer_,
  "EST", 2004L, NA_integer_,
  "HUN", 2004L, NA_integer_,
  "LVA", 2004L, NA_integer_,
  "LTU", 2004L, NA_integer_,
  "MLT", 2004L, NA_integer_,
  "POL", 2004L, NA_integer_,
  "SVK", 2004L, NA_integer_,
  "SVN", 2004L, NA_integer_,
  "BGR", 2007L, NA_integer_,
  "ROU", 2007L, NA_integer_,
  "HRV", 2013L, NA_integer_
)

eu_ever <- eu_accession$country_iso3
```

Note that East Germany appears in the Eurobarometer from 1990 and inherits Germany's 1958
accession year. That is the correct treatment of the accession switch — the territory
entered the Community through unification, not through an accession of its own — but it
means the `"from_accession"` scope does not blank out the East German series. The two
German units are never pooled by this notebook; the source keeps them apart and so do we.

**CSES has no East/West split.** All German CSES studies are post-unification
(`DEU_1998` onward) and `IMD1007` (sample component) is documented only as "see election
study notes", so no generic respondent-level East/West indicator is available.
Consequently the Germany split exists in the Eurobarometer series only, and CSES German
panels are labelled `DEU`.

# Ingest and cache

Each source is parsed once, stripped of its labelled types (`haven_labelled` and the
Stata-style `label.table` do not round-trip through parquet) and cached. The Eurobarometer
file is 228 MB on disk; reading the twelve columns actually needed via `col_select` takes
about two seconds rather than minutes.

```{r}
#| label: cache-helper

# Read from parquet if cached, otherwise run `parse_fun()`, cache, return.
cached_read <- function(name, parse_fun) {
  f <- file.path(dir_cache, paste0(name, ".parquet"))
  if (use_cache && file.exists(f)) {
    message("cache hit: ", name)
    return(arrow::read_parquet(f))
  }
  message("parsing from source: ", name)
  d <- parse_fun()
  arrow::write_parquet(d, f)
  d
}

# CSES ships value labels in `label.table`; pull one out as a tidy tibble so the
# crosswalk can be emitted from the same object the recoding uses.
cses_value_labels <- function(d, v) {
  lt <- attr(d, "label.table")[[v]]
  if (is.null(lt)) {
    return(tibble(code = integer(), code_label = character()))
  }
  tibble(code = as.integer(unname(lt)), code_label = names(lt))
}
```

```{r}
#| label: read-cses

# One reader for both CSES releases: `vars` is a named vector mapping the shared
# schema name -> the release-specific column name.
read_cses <- function(path, obj_name, vars, expect_rows, expect_cols) {
  env <- new.env()
  load(path, envir = env)
  d <- env[[obj_name]]

  stopifnot(
    is.data.frame(d),
    nrow(d) == expect_rows,
    ncol(d) == expect_cols,
    all(vars %in% names(d))
  )

  # Assert that the codes this notebook recodes are exactly the codes the file
  # declares, so a release that renumbers a category fails loudly here rather
  # than silently falling through to `not_asked`.
  declared <- sort(cses_value_labels(d, vars[["pid_close"]])$code)
  stopifnot(identical(declared, c(0L, 1L, 7L, 8L, 9L)))

  d[, unname(vars)] |>
    as_tibble() |>
    setNames(names(vars)) |>
    mutate(across(everything(), \(x) {
      a <- as.vector(x)
      attributes(a) <- NULL
      a
    }))
}

vars_imd <- c(
  survey_id     = "IMD1004",
  respondent_id = "IMD1005",
  iso3          = "IMD1006_UNALPHA3",
  eu_flag       = "IMD1006_EU",
  election_year = "IMD1008_YEAR",
  field_year    = "IMD1013_Y",
  weight        = "IMD1010_1",
  pid_close     = "IMD3005_1",
  pid_leaner    = "IMD3005_2",
  pid_party     = "IMD3005_3",
  pid_strength  = "IMD3005_4",
  turnout       = "IMD3001_LH",
  vote_party    = "IMD3002_LH_PL",
  mod5_flag     = "IMD1008_MOD_5"
)

vars_mod6 <- c(
  survey_id     = "f1004",
  respondent_id = "f1003_1",
  iso3          = "f1006_unalpha3",
  eu_flag       = "f1007_eu",
  election_year = "f1009",
  field_year    = "f1019_y",
  weight        = "f1101_1",
  pid_close     = "f3023_1",
  pid_leaner    = "f3023_2",
  pid_party     = "f3023_3",
  pid_strength  = "f3023_4",
  turnout       = "f3010",
  vote_party    = "f3011_lh_pl"
)

imd_raw <- cached_read("imd_raw", \() {
  read_cses(path_imd, "cses_imd", vars_imd, 395797L, 406L)
})

mod6_raw <- cached_read("mod6_raw", \() {
  read_cses(path_mod6, "cses6", vars_mod6, 33871L, 676L)
})

dim(imd_raw)
dim(mod6_raw)
```

Module 6 carries three tiers of weight and **the tiers are not interchangeable**.
`f1103_1` ("polity weight: sample") is identically 1 within a polity — it exists to make
polities count equally in pooled analyses, not to correct a sample. Since every estimate
below is a share *within* a country-year, the within-sample weight is the right one:
`f1101_1` for Module 6, `IMD1010_1` for the IMD. Both are asserted to vary.

```{r}
#| label: check-weights

stopifnot(
  n_distinct(imd_raw$weight) > 1,
  n_distinct(mod6_raw$weight) > 1
)

bind_rows(
  tibble(source = "CSES IMD", w = imd_raw$weight),
  tibble(source = "CSES MOD6", w = mod6_raw$weight)
) |>
  group_by(source) |>
  summarise(
    n_distinct = n_distinct(w),
    min = min(w, na.rm = TRUE),
    median = median(w, na.rm = TRUE),
    max = max(w, na.rm = TRUE),
    n_zero = sum(w == 0, na.rm = TRUE),
    .groups = "drop"
  )
```

The IMD is confirmed to contain Module 5 in full, which is why `cses5.rdata` is not read.

```{r}
#| label: mod5-containment

imd_raw |>
  count(in_module_5 = mod5_flag == 1) |>
  mutate(share = n / sum(n))
```

```{r}
#| label: read-eb

read_eb <- function(path, expect_rows) {
  d <- haven::read_dta(
    path,
    col_select = c(
      study_id, id, year, eb, nation1,
      closepty, feelclo, voteint, lastvote, wnation
    )
  )
  stopifnot(nrow(d) == expect_rows)

  d |>
    transmute(
      study_id = as.character(study_id),
      respondent_id = as.numeric(id),
      year = as.integer(year),
      eb = as.numeric(eb),
      nation1 = as.integer(nation1),
      closepty = as.integer(closepty),
      feelclo = as.numeric(feelclo),
      voteint = as.numeric(voteint),
      lastvote = as.numeric(lastvote),
      weight = as.numeric(wnation)
    )
}

eb_raw <- cached_read("eb_raw", \() read_eb(path_eb, 1134384L))

dim(eb_raw)
```

# Harmonisation

## Mapping objects

Every recode below is driven by these tibbles, and `pid_crosswalk.csv` is written **from
the same objects**. The crosswalk therefore cannot drift away from the code that produced
the data.

```{r}
#| label: map-pid

# CSES: the status is built from the pair (close, leaner), so the map is over the
# pair rather than over a single code.
map_cses_pid <- tribble(
  ~pid_close, ~pid_leaner, ~pid_status,    ~pid_strength_lab,
  1L,         NA_integer_, "attached",     "close",
  0L,         1L,          "attached",     "leaner",
  0L,         0L,          "not_attached", NA_character_,
  0L,         7L,          "not_attached", NA_character_,
  0L,         8L,          "not_attached", NA_character_,
  0L,         9L,          "not_attached", NA_character_,
  0L,         NA_integer_, "not_attached", NA_character_,
  7L,         NA_integer_, "dk_refused",   NA_character_,
  8L,         NA_integer_, "dk_refused",   NA_character_,
  9L,         NA_integer_, "not_asked",    NA_character_
)

# the ingest asserted the file declares exactly these codes; assert the map
# covers the same set, so the two checks cannot drift apart
stopifnot(setequal(unique(map_cses_pid$pid_close), c(0L, 1L, 7L, 8L, 9L)))

# Eurobarometer: status and strength arrive in one variable.
map_eb_pid <- tribble(
  ~closepty, ~code_label,    ~pid_status,    ~pid_strength_lab,
  1L,        "VERY CLOSE",   "attached",     "very close",
  2L,        "FAIRLY CLOSE", "attached",     "fairly close",
  3L,        "sympathizer",  "attached",     "sympathizer",
  4L,        "NO PARTY",     "not_attached", NA_character_,
  8L,        "DK,NA",        "dk_refused",   NA_character_,
  9L,        "inap",         "not_asked",    NA_character_,
  NA_integer_, "(absent)",   "not_asked",    NA_character_
)
```

`closepty == 3` ("merely a sympathiser") is coded **attached** and flagged, so the
sensitivity check at the end can move it without touching anything else. It is the closest
Eurobarometer analogue to a CSES leaner, which the CSES branching probe also counts as
attached.

```{r}
#| label: map-vote

# CSES turnout. 9999993 (not registered) is a genuine non-vote; refused/DK/missing
# are unknown; 9999996 means the item does not apply to that study.
map_cses_turnout <- tribble(
  ~turnout,  ~code_label,                        ~voted,
  0L,        "did not cast a ballot",            FALSE,
  1L,        "cast a ballot",                    TRUE,
  9999993L,  "not registered on electoral lists", FALSE,
  9999996L,  "not applicable: no LH election",   NA,
  9999997L,  "volunteered: refused",             NA,
  9999998L,  "volunteered: don't know",          NA,
  9999999L,  "missing",                          NA
)

# Eurobarometer vote items. Anything not listed here is a country-specific party
# code and therefore counts as a reported vote.
map_eb_vote <- tribble(
  ~code,  ~code_label,        ~voted,        ~vote_party_valid,
  0,      "na",               NA,            FALSE,
  990,    "OTHER PARTY",      TRUE,          FALSE,
  995,    "BLANK-SPOIL VOTE", TRUE,          FALSE,
  996,    "WOULD NOT VOTE / NOT VOTED", FALSE, FALSE,
  997,    "refused",          NA,            FALSE,
  998,    "DK",               NA,            FALSE,
  999,    "inap",             NA,            FALSE
)
```

A blank or spoilt ballot (`995`) is counted as *having voted* — the respondent turned out
— but never as a party choice. `990` ("other party") is likewise a reported vote without a
usable party code.

## Harmonisers

```{r}
#| label: harmonise-cses

harmonise_cses <- function(d, source_label) {
  d |>
    mutate(
      # the leaner probe is only asked of those who said "no", so it is NA-coded
      # 9 elsewhere; collapse to NA before the join so the map stays readable
      pid_leaner_j = if_else(pid_close == 0L, pid_leaner, NA_integer_),
      pid_close_j = pid_close
    ) |>
    left_join(
      map_cses_pid |> rename(pid_close_j = pid_close, pid_leaner_j = pid_leaner),
      by = c("pid_close_j", "pid_leaner_j")
    ) |>
    left_join(map_cses_turnout, by = "turnout") |>
    transmute(
      source = source_label,
      country_unit = iso3,
      country_iso3 = iso3,
      # The CSES time index is the ELECTION year, not the fieldwork year.
      # Fieldwork routinely crosses a calendar boundary, so indexing on
      # `IMD1013_Y` / `f1019_y` splits single election studies into two cells,
      # one of them tiny (e.g. Belgium 1999/2000 at n = 4). The election year
      # is what the study is a study of, and keeps one study in one cell.
      year = as.integer(election_year),
      election_year = as.integer(election_year),
      field_year = if_else(field_year > 2100 | field_year < 1900,
                           NA_integer_, as.integer(field_year)),
      survey_id = as.character(survey_id),
      respondent_id = as.character(respondent_id),
      weight = as.numeric(weight),
      eu_at_time = eu_flag == 1,
      pid_status,
      pid_strength = pid_strength_lab,
      is_sympathizer = FALSE,
      pid_party = if_else(pid_party >= 9999900, NA_real_, as.numeric(pid_party)),
      voted,
      vote_item_type = "past_vote",
      vote_party = if_else(vote_party >= 9999900, NA_real_, as.numeric(vote_party))
    )
}

cses_h <- bind_rows(
  harmonise_cses(imd_raw, "CSES IMD"),
  harmonise_cses(mod6_raw, "CSES MOD6")
)

stopifnot(!any(is.na(cses_h$pid_status)))
count(cses_h, source, pid_status)
```

The CSES vote construct is **past vote in the lower-house election the study follows**.
It is not a vote intention and the two are never merged.

Two notes on the CSES time index. First, it is the **election year**: fieldwork routinely
crosses a calendar boundary, so indexing on the fieldwork year splits a single election
study into two cells, one of them near-empty. The fieldwork year is still carried as
`field_year`. Second, a country that held two elections indexed to the same year, or that
ran two studies of one election (`DEU12002` and `DEU22002`), is pooled into a single
country-year — the series is descriptive and the alternative is an arbitrary tie-break.

```{r}
#| label: harmonise-eb

# EB contributes two row-sets, one per vote construct, because the file carries
# both `voteint` (intention) and `lastvote` (past vote). A country-year therefore
# legitimately appears twice, distinguished by `vote_item_type`.
harmonise_eb_one <- function(d, vote_col, vote_item_type_label) {
  d |>
    mutate(vote_code = .data[[vote_col]]) |>
    left_join(map_eb_pid |> select(closepty, pid_status, pid_strength_lab),
              by = "closepty") |>
    left_join(map_eb_vote |> rename(vote_code = code),
              by = "vote_code") |>
    left_join(eb_country_map, by = "nation1") |>
    transmute(
      source = "Mannheim EB",
      country_unit,
      country_iso3,
      year,
      election_year = NA_integer_,
      field_year = year,
      survey_id = paste0("EB", eb),
      respondent_id = as.character(respondent_id),
      weight,
      eu_at_time = NA,
      # rows with a code absent from the map are country-specific party codes:
      # a reported vote, with a usable party code
      pid_status = if_else(is.na(pid_status), "not_asked", pid_status),
      pid_strength = pid_strength_lab,
      is_sympathizer = !is.na(closepty) & closepty == 3L,
      pid_party = if_else(feelclo %in% map_eb_vote$code, NA_real_, feelclo),
      voted = if_else(!is.na(vote_code) & !vote_code %in% map_eb_vote$code,
                      TRUE, voted),
      vote_item_type = vote_item_type_label,
      vote_party = if_else(!is.na(vote_code) & !vote_code %in% map_eb_vote$code,
                           vote_code, NA_real_)
    )
}

eb_h <- bind_rows(
  harmonise_eb_one(eb_raw, "voteint", "vote_intention"),
  harmonise_eb_one(eb_raw, "lastvote", "past_vote")
)

stopifnot(!any(is.na(eb_h$pid_status)))
count(eb_h, vote_item_type, pid_status)
```

## Stack, apply scope, apply the sympathiser switch

```{r}
#| label: stack

pid_long <- bind_rows(cses_h, eb_h) |>
  mutate(
    # the switch is applied here, once, so every downstream table follows it
    pid_status = if_else(
      !sympathizer_as_attached & is_sympathizer, "not_attached", pid_status
    ),
    pid_status = factor(
      pid_status,
      levels = c("attached", "not_attached", "dk_refused", "not_asked")
    )
  ) |>
  left_join(eu_accession, by = "country_iso3")

n_before_scope <- nrow(pid_long)

pid_long <- pid_long |>
  filter(country_iso3 %in% eu_ever) |>
  filter(
    if (eu_scope == "ever") TRUE else year >= accession_year
  )

scope_label <- if (eu_scope == "ever") {
  "EU member states (ever a member), all available years"
} else {
  "EU member states, from accession year onward"
}

tibble(
  rows_before_scope = n_before_scope,
  rows_after_scope = nrow(pid_long),
  countries = n_distinct(pid_long$country_unit),
  scope = scope_label
)
```

An item that a survey never fielded and an item a respondent skipped are different things,
and CSES codes both as `9`. `item_fielded` distinguishes them at the survey level: a
survey counts as having fielded the battery if any respondent in it gave a substantive
answer. Country-years in which the item was never fielded are dropped from the
denominators; item nonresponse inside a fielded survey stays in as `not_asked` and is
reported.

```{r}
#| label: item-fielded

pid_long <- pid_long |>
  group_by(source, survey_id, country_unit) |>
  mutate(
    item_fielded = any(pid_status %in% c("attached", "not_attached")),
    vote_fielded = any(!is.na(voted))
  ) |>
  ungroup()

pid_long |>
  distinct(source, survey_id, country_unit, item_fielded) |>
  count(source, item_fielded)
```

A second, subtler filter is needed on the vote side. In some early Eurobarometer waves the
vote item records **no negative category at all** — the "would not vote" code is absent
from the entire wave, across every country in it. Where that is true the item cannot
distinguish a voter from a non-voter, and the share of non-attached respondents "reporting
a vote" is mechanically 100% rather than estimated.

The test is structural and applied at the **wave** level, not the country level, because
the presence of a response code is a property of the questionnaire rather than of a
country. A country-level test would wrongly flag compulsory-voting countries such as
Belgium, where a near-100% turnout share is genuine.

```{r}
#| label: vote-negative-available

pid_long <- pid_long |>
  group_by(source, survey_id, vote_item_type) |>
  mutate(vote_neg_available = any(voted == FALSE, na.rm = TRUE)) |>
  ungroup()

pid_long |>
  filter(source == "Mannheim EB", vote_fielded) |>
  distinct(year, survey_id, vote_item_type, vote_neg_available) |>
  group_by(year, vote_item_type) |>
  summarise(
    waves = n(),
    waves_without_negative = sum(!vote_neg_available),
    .groups = "drop"
  ) |>
  filter(waves_without_negative > 0)
```

## Emit the crosswalk

```{r}
#| label: write-crosswalk

crosswalk <- bind_rows(
  map_cses_pid |>
    transmute(
      source = "CSES (IMD + MOD6)",
      construct = "pid_status",
      original_variable = "IMD3005_1/f3023_1 x IMD3005_2/f3023_2",
      original_code = paste0("close=", pid_close, ", leaner=",
                             ifelse(is.na(pid_leaner), "-", pid_leaner)),
      harmonised = pid_status,
      note = ifelse(is.na(pid_strength_lab), "", pid_strength_lab)
    ),
  map_eb_pid |>
    transmute(
      source = "Mannheim EB",
      construct = "pid_status",
      original_variable = "closepty",
      original_code = ifelse(is.na(closepty), "NA", as.character(closepty)),
      harmonised = pid_status,
      note = code_label
    ),
  map_cses_turnout |>
    transmute(
      source = "CSES (IMD + MOD6)",
      construct = "voted",
      original_variable = "IMD3001_LH/f3010",
      original_code = as.character(turnout),
      harmonised = ifelse(is.na(voted), "NA", as.character(voted)),
      note = code_label
    ),
  map_eb_vote |>
    transmute(
      source = "Mannheim EB",
      construct = "voted",
      original_variable = "voteint / lastvote",
      original_code = as.character(code),
      harmonised = ifelse(is.na(voted), "NA", as.character(voted)),
      note = code_label
    ),
  tibble(
    source = "Mannheim EB",
    construct = "voted",
    original_variable = "voteint / lastvote",
    original_code = "(any other code)",
    harmonised = "TRUE",
    note = "country-specific party code = a reported vote"
  ),
  eb_country_map |>
    transmute(
      source = "Mannheim EB",
      construct = "country_unit",
      original_variable = "nation1",
      original_code = as.character(nation1),
      harmonised = country_unit,
      note = nation1_label
    )
)

write_csv(crosswalk, file.path(dir_final, "pid_crosswalk.csv"))
nrow(crosswalk)
```

# Series breaks

Breaks are carried in a tibble and drawn onto the plots, rather than described only in
prose.

```{r}
#| label: tbl-breaks
#| tbl-cap: "Documented series breaks and coverage limits."

series_breaks <- tribble(
  ~source,       ~year, ~break_label,
  "Mannheim EB", 1975L, "closepty first fielded",
  "Mannheim EB", 1995L, "closepty dropped from the Standard EB",
  "CSES IMD",    1996L, "CSES series begins",
  "CSES MOD6",   2022L, "Module 6 (advance release)"
)

series_breaks |> knitr::kable()
```

Additional breaks and filters that do not resolve to a single year:

- **Eurobarometer `voteint`** is absent in 1998 and 2001; `lastvote` is fielded only in
  1979 and 1982-1995. Gaps are left blank and never interpolated.
- **No negative vote category in some early waves.** The "would not vote" code is entirely
  absent from `voteint` in 1976 and 1979 (and near-absent in 1975), and from `lastvote` in
  1982. In those waves the item cannot record a non-vote, so the headline share is 100% by
  construction. Those country-years are marked unreliable by `vote_neg_available` and
  dropped from the headline plot; they remain in the export, flagged, so the suppression is
  auditable. This is the single most consequential filter in the notebook — without it the
  headline series opens with a flat, entirely artefactual 100%.
- **Zero recorded non-votes in a cell.** The wave-level test above is a statement about the
  questionnaire, but the code was evidently offered in only some countries of a given wave:
  several country-years still showed a flat 100% because that country recorded no non-vote
  at all. A country-year containing not one non-vote cannot support an estimate of the
  non-vote rate whatever the cause, so `n_no_vote > 0` is required as well. This does not
  penalise genuinely high-turnout countries: Belgium, with compulsory voting, records
  non-votes in every wave and is retained throughout.
- **`feelclo`** (which party) is fielded erratically (1973, 1981, 1985-1994), so
  `pid_party` is sparse for the Eurobarometer. It is not used to derive status.
- **Skip patterns.** The CSES leaner probe (`_2`) is asked only of respondents who said
  "no" to the first item; the "which party" and "how close" items only of those who ended
  up attached. The Eurobarometer vote items are asked of the whole sample, but which of
  `voteint` / `lastvote` was fielded varies by wave.
- **Module 6 is an advance release.** Its country coverage will change on full release, so
  every Module 6 point is provisional.
- **Finland, Sweden and Austria** acceded in 1995, after `closepty` was dropped, so they
  have no Eurobarometer attachment series at all under either scope.

# Aggregation

Shares are weighted; uncertainty uses the **Kish effective N**, $n_{\text{eff}} =
(\sum w)^2 / \sum w^2$, in a Wilson interval. Neither source ships primary sampling units
or strata, so a full `survey::svydesign` could only ever use the weights — which is
exactly what this does. Using the effective rather than the nominal N is what makes the
intervals design-corrected.

```{r}
#| label: helpers-agg

eff_n <- function(w) {
  w <- w[!is.na(w)]
  if (sum(w^2) == 0) return(0)
  sum(w)^2 / sum(w^2)
}

# Wilson score interval evaluated at the effective N
wilson <- function(p, n, z = 1.96) {
  n <- pmax(n, 1e-9)
  denom <- 1 + z^2 / n
  centre <- (p + z^2 / (2 * n)) / denom
  halfw <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / denom
  tibble(lower = pmax(0, centre - halfw), upper = pmin(1, centre + halfw))
}
```

## Attachment shares by country-year

```{r}
#| label: agg-country-year

# the two EB vote row-sets duplicate the attachment item, so the marginals are
# computed on one of them only
pid_marginal_base <- pid_long |>
  filter(item_fielded) |>
  filter(!(source == "Mannheim EB" & vote_item_type == "past_vote"))

cy_n <- pid_marginal_base |>
  group_by(source, country_unit, country_iso3, year) |>
  summarise(
    n_raw = n(),
    n_eff = eff_n(weight),
    w_total = sum(weight, na.rm = TRUE),
    .groups = "drop"
  )

pid_shares <- pid_marginal_base |>
  group_by(source, country_unit, country_iso3, year, pid_status) |>
  summarise(w = sum(weight, na.rm = TRUE), .groups = "drop") |>
  complete(
    nesting(source, country_unit, country_iso3, year),
    pid_status,
    fill = list(w = 0)
  ) |>
  group_by(source, country_unit, country_iso3, year) |>
  mutate(share = w / sum(w)) |>
  ungroup()

pid_country_year <- pid_shares |>
  select(-w) |>
  pivot_wider(
    names_from = pid_status, values_from = share, names_prefix = "share_"
  ) |>
  left_join(cy_n, by = c("source", "country_unit", "country_iso3", "year")) |>
  mutate(reliable = n_eff >= min_eff_n) |>
  arrange(source, country_unit, year)

pid_country_year <- pid_country_year |>
  bind_cols(wilson(pid_country_year$share_attached, pid_country_year$n_eff))

head(pid_country_year)
```

```{r}
#| label: tbl-thin-cells
#| tbl-cap: "Country-years below the minimum effective N."

pid_country_year |>
  filter(!reliable) |>
  select(source, country_unit, year, n_raw, n_eff) |>
  arrange(n_eff) |>
  head(20) |>
  knitr::kable(digits = 1)
```

## Attachment x vote cross-tabulation

This is the key quantity. The cross-tab is computed **separately for each
`vote_item_type`**, never pooled across them.

```{r}
#| label: agg-crosstab

pid_vote_crosstab <- pid_long |>
  filter(item_fielded, vote_fielded, !is.na(voted)) |>
  mutate(
    cell = case_when(
      pid_status == "attached"     &  voted ~ "attached_and_voted",
      pid_status == "attached"     & !voted ~ "attached_and_no_vote",
      pid_status == "not_attached" &  voted ~ "not_attached_but_voted",
      pid_status == "not_attached" & !voted ~ "not_attached_and_no_vote",
      pid_status == "dk_refused"   &  voted ~ "dk_attachment_and_voted",
      pid_status == "dk_refused"   & !voted ~ "dk_attachment_and_no_vote",
      TRUE ~ "not_asked"
    )
  ) |>
  group_by(source, country_unit, country_iso3, year, vote_item_type) |>
  mutate(
    n_eff = eff_n(weight),
    w_total = sum(weight, na.rm = TRUE),
    vote_neg_available = all(vote_neg_available)
  ) |>
  group_by(source, country_unit, country_iso3, year, vote_item_type,
           n_eff, w_total, vote_neg_available, cell) |>
  summarise(w = sum(weight, na.rm = TRUE), n_raw = n(), .groups = "drop") |>
  mutate(share = w / w_total)

# The headline series. Computed directly from the respondent rows rather than
# from the cross-tab, so the effective N is that of the non-attached subgroup
# itself rather than a rescaling of the full-sample effective N.
nonattached_voted <- pid_long |>
  filter(item_fielded, vote_fielded, !is.na(voted), pid_status == "not_attached") |>
  group_by(source, country_unit, country_iso3, year, vote_item_type) |>
  summarise(
    n_denom = n(),
    n_eff = eff_n(weight),
    w_denom = sum(weight, na.rm = TRUE),
    w_voted = sum(weight[voted], na.rm = TRUE),
    # a country-year is usable only if every wave behind it could record a
    # non-vote; otherwise the share is 100% by construction
    vote_neg_available = all(vote_neg_available),
    # and, regardless of the questionnaire, a cell in which not one non-vote was
    # actually recorded cannot support an estimate of the non-vote rate
    n_no_vote = sum(!voted),
    .groups = "drop"
  ) |>
  filter(n_denom > 0, w_denom > 0) |>
  mutate(
    share_voted = w_voted / w_denom,
    reliable = n_eff >= min_eff_n & vote_neg_available & n_no_vote > 0
  )

nonattached_voted <- nonattached_voted |>
  bind_cols(wilson(nonattached_voted$share_voted, nonattached_voted$n_eff))

nonattached_voted |>
  select(source, country_unit, year, vote_item_type, n_denom, n_eff, share_voted) |>
  head(10)
```

## Pooled series

Pooling across countries here is an **unweighted country average**: countries are not
population-weighted, so every country counts once. Labelled as such wherever it appears.

```{r}
#| label: agg-pooled

# The set of countries observed changes from year to year, especially in CSES
# where a country contributes only in its election years. A pooled mean over a
# shifting country set is not comparable across years, so years resting on
# fewer than this many countries are not drawn.
min_countries_pooled <- 3

pooled_attach <- pid_country_year |>
  filter(reliable) |>
  group_by(source, year) |>
  summarise(
    countries = n(),
    share_attached = mean(share_attached),
    share_dk = mean(share_dk_refused),
    .groups = "drop"
  ) |>
  filter(countries >= min_countries_pooled)

pooled_nonattached_voted <- nonattached_voted |>
  filter(reliable) |>
  group_by(source, vote_item_type, year) |>
  summarise(
    countries = n(),
    share_voted = mean(share_voted, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(countries >= min_countries_pooled)

head(pooled_attach)
```

# Comparability across the two instruments

```{r}
#| label: tbl-overlap
#| tbl-cap: "Country-years carrying both an Eurobarometer and a CSES attachment estimate."

overlap <- pid_country_year |>
  mutate(family = if_else(source == "Mannheim EB", "EB", "CSES")) |>
  distinct(country_unit, year, family) |>
  count(country_unit, year) |>
  filter(n > 1)

overlap |> knitr::kable()
```

```{r}
#| label: tbl-spans
#| tbl-cap: "Observed year span per source."

pid_country_year |>
  group_by(source) |>
  summarise(
    first_year = min(year),
    last_year = max(year),
    country_years = n(),
    countries = n_distinct(country_unit),
    .groups = "drop"
  ) |>
  knitr::kable()
```

If the overlap table above is empty, that is the substantive result: the Eurobarometer
attachment item ends in 1994 and the CSES series begins in 1996, so **the per-country
level offset between the two instruments is not estimable from these data**. No bridged or
spliced series is produced, and no constant instrument effect is assumed. Under this
constraint the only defensible cross-source comparison is of *slopes within* each series.

# Plots

```{r}
#| label: plot-setup

pal_source <- c(
  "Mannheim EB" = "#1B4F72",
  "CSES IMD" = "#B7472A",
  "CSES MOD6" = "#E39C34"
)

# Gaps must be visibly blank, but the two instruments have different natural
# cadences: the Eurobarometer fields waves annually, while a CSES country is
# observed only in its election years, typically three to five years apart.
# Filling a common one-year grid would therefore shatter every CSES line into
# unconnected dots, and connecting everything regardless would draw a line
# across waves in which the item was never fielded.
#
# Instead a series is cut into segments wherever the gap between consecutive
# observations exceeds what is normal for that source. Points are always drawn,
# so isolated observations stay visible.
max_gap_years <- c("Mannheim EB" = 2, "CSES IMD" = 6, "CSES MOD6" = 6)

add_segments <- function(d, by = c("source", "country_unit")) {
  d |>
    arrange(across(all_of(c(by, "year")))) |>
    group_by(across(all_of(by))) |>
    mutate(
      gap = year - lag(year),
      seg = cumsum(is.na(gap) | gap > max_gap_years[source]),
      series = paste(!!!syms(by), seg, sep = "|")
    ) |>
    ungroup()
}
```

```{r}
#| label: fig-pid-share-facets
#| fig-cap: "Share reporting party attachment, by country and source. Ribbons are Wilson intervals on the Kish effective N. The dashed line is the don't-know/refused share. Gaps are years in which the item was not fielded and are left blank. Series are never spliced across sources."
#| fig-width: 12
#| fig-height: 14
#| fig-dpi: 500

d_fac <- pid_country_year |>
  filter(reliable) |>
  add_segments()

ggplot(d_fac, aes(x = year, colour = source, fill = source)) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper, group = series),
    alpha = 0.18, colour = NA
  ) +
  geom_line(aes(y = share_attached, group = series), linewidth = 0.6) +
  # several countries contribute a single election study; without points those
  # series would render as nothing at all
  geom_point(aes(y = share_attached), size = 0.7) +
  geom_line(
    aes(y = share_dk_refused, group = series),
    linetype = "dashed", linewidth = 0.4, alpha = 0.8
  ) +
  geom_vline(
    data = series_breaks,
    aes(xintercept = year),
    linetype = "dotted", colour = "grey55", linewidth = 0.3
  ) +
  facet_wrap(~country_unit, ncol = 4) +
  scale_colour_manual(values = pal_source, name = NULL) +
  scale_fill_manual(values = pal_source, name = NULL) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = "share of respondents",
    title = "Reported party attachment over time",
    subtitle = paste0(
      scope_label,
      ". Solid: attached. Dashed: don't know / refused."
    ),
    caption = paste0(
      "Sources kept separate; no splicing. Cells below an effective N of ",
      min_eff_n, " are omitted."
    )
  ) +
  theme_pubr(legend = "top") +
  theme(panel.spacing = unit(0.6, "lines"))
```

```{r}
#| label: fig-pid-composition
#| fig-cap: "Composition of the attachment item over time: attached, not attached, and don't know/refused. Don't know is shown as its own band and is never folded into 'not attached'. Panels are separated by source so that no stack mixes instruments."
#| fig-width: 12
#| fig-height: 14
#| fig-dpi: 500

d_comp <- pid_shares |>
  semi_join(
    pid_country_year |> filter(reliable),
    by = c("source", "country_unit", "year")
  ) |>
  filter(pid_status != "not_asked")

ggplot(d_comp, aes(x = year, y = share, fill = pid_status)) +
  geom_col(width = 1) +
  # sources do not overlap in time, so one panel per country shows all of them
  # without any stack ever mixing two instruments in the same bar
  facet_wrap(~country_unit, ncol = 4) +
  scale_fill_manual(
    values = c(
      attached = "#1B4F72",
      not_attached = "#C7CCD1",
      dk_refused = "#E39C34"
    ),
    name = NULL
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = NULL,
    title = "Attachment / non-attachment / don't know, over time",
    subtitle = scope_label
  ) +
  theme_pubr(legend = "top") +
  theme(panel.spacing = unit(0.3, "lines"))
```

```{r}
#| label: fig-nonattached-voted
#| fig-cap: "The headline quantity: among respondents who denied any party attachment, the share who nonetheless report a vote. Past vote and vote intention are different constructs and are drawn distinctly; they are never merged. Gaps are left blank."
#| fig-width: 12
#| fig-height: 14
#| fig-dpi: 500

d_head <- nonattached_voted |>
  filter(reliable) |>
  add_segments(by = c("source", "country_unit", "vote_item_type"))

ggplot(d_head, aes(x = year, y = share_voted, colour = source)) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper, group = series, fill = source),
    alpha = 0.15, colour = NA
  ) +
  geom_line(aes(linetype = vote_item_type, group = series), linewidth = 0.6) +
  geom_point(aes(shape = vote_item_type), size = 1.1) +
  facet_wrap(~country_unit, ncol = 4) +
  scale_colour_manual(values = pal_source, name = NULL) +
  scale_fill_manual(values = pal_source, name = NULL) +
  scale_linetype_manual(
    values = c(past_vote = "solid", vote_intention = "22"),
    name = "vote item"
  ) +
  scale_shape_manual(
    values = c(past_vote = 16, vote_intention = 1),
    name = "vote item"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = "share reporting a vote",
    title = "Among the non-attached: share who nonetheless report a vote",
    subtitle = paste0(
      scope_label,
      ". Eurobarometer asks vote intention and last vote; CSES asks past vote."
    )
  ) +
  theme_pubr(legend = "top") +
  theme(panel.spacing = unit(0.6, "lines"))
```

```{r}
#| label: fig-pooled
#| fig-cap: "Unweighted country averages. Countries are not population-weighted, so each country counts once and the level is not a population quantity for the EU as a whole. Because a CSES country contributes only in its election years, the set of countries behind each point changes; years resting on fewer than three countries are not drawn, and the remaining year-to-year movement still mixes real change with compositional change."
#| fig-width: 10
#| fig-height: 5
#| fig-dpi: 500

bind_rows(
  pooled_attach |>
    transmute(source, year, value = share_attached,
              panel = "attached (all respondents)"),
  pooled_nonattached_voted |>
    transmute(source, year, value = share_voted,
              panel = paste0("reports a vote (non-attached only), ", vote_item_type))
) |>
  ggplot(aes(x = year, y = value, colour = source, group = interaction(source, panel))) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~panel, nrow = 1) +
  scale_colour_manual(values = pal_source, name = NULL) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = NULL,
    title = "Unweighted country averages",
    subtitle = scope_label,
    caption = "Unweighted across countries -- not an EU population estimate."
  ) +
  theme_pubr(legend = "top")
```

# Coverage grid

```{r}
#| label: tbl-coverage
#| tbl-cap: "Coverage grid: which source carries the attachment item and the vote item, by country and five-year period."

# A (attachment) and V (vote) are marked PER SOURCE. Marking them per
# country-period would let one source's attachment item and another source's
# vote item combine into a spurious "A+V".
coverage <- pid_long |>
  mutate(period = paste0(5 * (year %/% 5), "-", 5 * (year %/% 5) + 4)) |>
  group_by(country_unit, period, source) |>
  summarise(
    attach = any(item_fielded),
    vote = any(vote_fielded),
    .groups = "drop"
  ) |>
  mutate(
    src_short = recode(
      source,
      "Mannheim EB" = "EB", "CSES IMD" = "IMD", "CSES MOD6" = "M6"
    ),
    mark = case_when(
      attach & vote ~ paste0(src_short, "(A+V)"),
      attach ~ paste0(src_short, "(A)"),
      vote ~ paste0(src_short, "(V)"),
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(mark)) |>
  group_by(country_unit, period) |>
  summarise(mark = paste(sort(mark), collapse = " "), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = mark, values_fill = "")

coverage |>
  select(country_unit, sort(setdiff(names(coverage), "country_unit"))) |>
  knitr::kable()
```

# Exports

```{r}
#| label: write-exports

write_csv(pid_country_year, file.path(dir_final, "pid_country_year.csv"))
write_csv(pid_vote_crosstab,
          file.path(dir_final, "pid_vote_crosstab_country_year.csv"))

list.files(dir_final, pattern = "^pid_") |> sort()
```

# Sensitivity: the sympathiser coding

The one genuinely discretionary recode is Eurobarometer `closepty == 3`. Recomputing the
attachment share with sympathisers moved to `not_attached` bounds how much it matters —
and it matters a great deal, so this is reported as a headline result rather than a
footnote.

```{r}
#| label: tbl-sympathizer
#| tbl-cap: "Eurobarometer attachment share with and without sympathisers counted as attached."

eb_sens <- pid_long |>
  filter(source == "Mannheim EB", item_fielded,
         vote_item_type == "vote_intention") |>
  group_by(year) |>
  summarise(
    with_symp = sum(weight[pid_status == "attached"], na.rm = TRUE) /
      sum(weight, na.rm = TRUE),
    without_symp = sum(weight[pid_status == "attached" & !is_sympathizer],
                       na.rm = TRUE) / sum(weight, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(difference = with_symp - without_symp)

eb_sens |> knitr::kable(digits = 3)
```

```{r}
#| label: sympathizer-magnitude

tibble(
  mean_with = mean(eb_sens$with_symp),
  mean_without = mean(eb_sens$without_symp),
  mean_difference = mean(eb_sens$difference),
  min_difference = min(eb_sens$difference),
  max_difference = max(eb_sens$difference)
)
```

The middle category is not a rounding detail: "merely a sympathiser" accounts for roughly
**thirty percentage points** of the Eurobarometer attachment share, and the measured level
of European partisanship over 1975-1994 roughly *halves* depending on which side of the
line it falls. The gap is also stable across the whole span, so it shifts the level of the
series far more than its slope.

This is itself evidence for the paper's argument rather than a nuisance. A single
instrument, fielded unchanged, yields two defensible partisan shares that differ by a
factor of about two, purely from where the analyst draws the line between attachment and
non-attachment. That is precisely step 3 of the argument chain — coding choices shift the
population of inference — visible inside one survey item.

# Findings and caveats

```{r}
#| label: findings-numbers

summary_stats <- list(
  eb_years = range(pid_country_year$year[pid_country_year$source == "Mannheim EB"]),
  cses_years = range(pid_country_year$year[pid_country_year$source != "Mannheim EB"]),
  n_overlap = nrow(overlap),
  n_thin = sum(!pid_country_year$reliable),
  n_country_years = nrow(pid_country_year)
)
summary_stats
```

**Coverage limits.** The Eurobarometer attachment item is fielded 1975-1994 only. The
trend file spans 1970-2002, but `closepty` is absent before 1975 and is dropped from the
Standard Eurobarometer after 1994. The long-run attachment series therefore cannot be
pushed back to the early 1970s, and the claim in
`data/01_raw/external/external-data.md` that the explicit/vote-anchored split is
constructible "back to the early 1970s" holds for the *vote* item alone, not for
attachment. Austria, Finland and Sweden acceded in 1995 and so have no Eurobarometer
attachment series under any scope.

**Series breaks.** Marked on the plots: 1975 (item introduced), 1995 (item dropped), 1996
(CSES begins), 2022 (Module 6). `voteint` is missing in 1998 and 2001; `lastvote` is
fielded only in 1979 and 1982-1995. All gaps are left blank; nothing is interpolated.

**Thin cells.** Country-years below an effective N of `r min_eff_n` are excluded from the
plots and listed above. Effective N is the Kish quantity, so a country-year with a
nominal N of several hundred can still fall below the threshold when the weights are
dispersed.

**Don't know is a category, not noise.** DK/refused is reported in every table and drawn
as its own band. It is never folded into "not attached" — a respondent who declines to
answer the attachment item is not a respondent who denies attachment, and the two move
differently across waves and countries.

**Two vote constructs.** CSES asks about past vote in the lower-house election the study
follows. The Eurobarometer carries both a vote intention ("if there were a general
election tomorrow") and a last-vote recall. These are different constructs, carried in
`vote_item_type` and drawn distinctly throughout. A country-year can appear twice in the
cross-tab exports, once per construct; they are never averaged together.

**What the two instruments do and do not share.** They share a target construct — felt
closeness to a party — and both separate a substantive denial from a don't-know. They do
not share a question form: CSES asks a binary and then probes leaners, while the
Eurobarometer asks a single graded item whose middle category ("merely a sympathiser") has
no CSES equivalent. The Eurobarometer bar is lower. Levels are consequently **not**
comparable across the two; only slopes within a source are. Because the two series do not
overlap in time, this notebook cannot estimate the offset empirically, and does not
pretend to: no bridged line is drawn anywhere.

**Germany.** The Eurobarometer keeps West (`DEU-W`, from 1970) and East (`DEU-E`, from
1990) apart via `nation1`, and this notebook preserves that split rather than pooling to
`nation2`'s single "germany". CSES has no equivalent: all German CSES studies are
post-unification and no generic respondent-level East/West indicator exists, so CSES
German panels are labelled `DEU`. Great Britain and Northern Ireland are kept apart on the
same principle.

**Module 6 is provisional.** It is an advance release; its country coverage will change on
full release and every Module 6 point should be re-run against the final version.

**What this does not do.** Party codes are carried as codes, not resolved to party names —
`IMD3005_3` is a CSES numeric code and Eurobarometer `feelclo` codes are country-specific.
Resolution via Party Facts (`IMD5103_*` / `f5203_*`) is deferred until a section needs
party-level attachment. Nothing here is a causal estimate; this is descriptive input to
step 2 of the argument chain.

# Session Info

```{r}
#| label: session-info

session_info()
```

# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```
