---
title: |
  Code Notebook 2.1
subtitle: |
  Construction of Key Variables
date: last-modified
date-format: MMMM D, YYYY
format:
  html:
    toc: true
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
  eval: true
  message: true
---

# Setup

```{r}
#| label: setup

# To track render duration
start_time <- Sys.time()


# Install and load required packages
p_required <- c(
  "tidyverse", #dplyr, ggplot, tibbles etc
  "here", # relative and sys agnostic file paths
  "sessioninfo" # documentation
)
packages <- rownames(installed.packages())
p_to_install <- p_required[!(p_required %in% packages)]
if (length(p_to_install) > 0) {
  pak::pkg_install(p_to_install)
}
sapply(p_required, require, character.only = TRUE)
rm(p_required, p_to_install, packages)
```

```{r}
#| label: load-data

# Load Data
eu25games2019 <- readRDS(
  here(
    "data",
    "02_processed",
    "eu25games2019_long.rds"
  )
)

```

# Vote and Party ID Variables

The published data harmonises every party-bearing item against Party Facts, so each is
available as an `ext_<item>_pf_name` / `ext_<item>_pf_id` pair. The items we need are:

- Partisan attachment: `ext_q_party_id_pf_*` — *"Do you consider yourself to be close to
  any particular political party? If so, which party do you feel close to?"*
- Vote intention: `ext_q_vote_choice_intended_pf_*` (wave 2, *"Which party will you vote
  for in this election?"*)
- Hypothetical vote: `ext_q_vote_choice_hypo_pf_*` (*"Just assuming you were going
  to vote / had voted, which party would you (have) vote(d) for?"*)
- Reported vote choice: `ext_q_vote_choice_past_pf_*` (*"Which party did you vote for in
  this election?"*)

Given our primary interest in comparing explicit and implicit attachment, we pursue a
recall-maximizing strategy: intention first, then the hypothetical probe, then reported
choice. Note that the published data merges the wave-2 and wave-3 hypothetical items
into a single `q_vote_choice_hypo` column, so this collapses to three sources rather than
the four the previous pipeline coalesced.

## Non-party responses

Party Facts assigns real parties a numeric `pf_id`; non-party answers instead carry a
sentinel string in `pf_name` with `pf_id = NA`. Parties genuinely absent from Party Facts
are marked `"<name> [og]"`, also with `pf_id = NA` — **these are real parties** and must
stay in the partisan groups, so the sentinel set has to be matched literally rather than
by testing `is.na(pf_id)`.

```{r}
#| label: pf-sentinels

# non-party answers in the Party Facts crosswalk
pf_no_party <- c(
  "none",
  "dont-know",
  "refused",
  "did-not-vote",
  "blank-invalid",
  "not-eligible"
)

# answers that name something, but not a codeable party
pf_other <- c("other", "independent")

# which of these actually occur in this sample?
eu25games2019 |>
  distinct(meta_pid, ext_q_party_id_pf_name) |>
  filter(ext_q_party_id_pf_name %in% c(pf_no_party, pf_other)) |>
  count(ext_q_party_id_pf_name, name = "respondents")
```

```{r}
#| label: aggregated-vote-and-pid-variables

eu25games2019 <- eu25games2019 |>
  # Create aggregated/combined variables
  mutate(
    der_vote_combined_name = coalesce(
      ext_q_vote_choice_intended_pf_name,
      ext_q_vote_choice_hypo_pf_name,
      ext_q_vote_choice_past_pf_name
    ),
    der_vote_combined_id = coalesce(
      ext_q_vote_choice_intended_pf_id,
      ext_q_vote_choice_hypo_pf_id,
      ext_q_vote_choice_past_pf_id
    ),
    der_vote_cat = case_when(
      der_vote_combined_name %in% pf_other ~ "Other",
      der_vote_combined_name %in% pf_no_party ~ "Don't know",
      is.na(der_vote_combined_name) ~ "NA",
      .default = "Vote reported"
    ),
    der_pid = case_when(
      ext_q_party_id_pf_name %in% pf_no_party ~ "No Party ID",
      ext_q_party_id_pf_name %in% pf_other ~ "Other",
      is.na(ext_q_party_id_pf_name) ~ "NA",
      .default = "Party ID reported"
    ),
    # Hahm et al.'s `nopid` flag, reconstructed from the harmonized attachment item
    der_nopid = if_else(der_pid == "Party ID reported", 0, 1)
  ) |>
  # Convert to factors and sort by frequency
  mutate(
    across(
      c(der_vote_combined_name, der_vote_combined_id, der_vote_cat, der_pid),
      ~ fct_infreq(as.character(.x))
    )
  )

eu25games2019 |>
  distinct(meta_pid, der_pid, der_vote_cat) |>
  count(der_pid, der_vote_cat)
```

# Variable Combining Partisan Type (Implicit/Explicit) and Group-relationship (Co/Out/None)

The conjoint conditions and the displayed co-player nationality now come straight from
the published data rather than from Hahm et al.'s derived analysis frame:

| previously | now | levels |
|---|---|---|
| `cj_trmnt` | `cj_treatment` | `1_conat_eustance_partisan`, `2_conat_partisan`, `3_eunat_eustance`, `4_somenat` |
| `der_conational` | `cj_nationality_shown` | `own_country`, `eu`, `non_eu` |

`cj_nationality_shown` is the nationality the respondent actually saw. The treatment
branch overrides the background nationality lottery in every condition except
`4_somenat`, so the raw draw (`cj_nationality_drawn`) must not be used here.

```{r}
#| label: implicit-explicit-co-out-interaction-variable

# create implicit/explicit partisan marker variable
eu25games2019 <- eu25games2019 |>
  mutate(
    # combined variable for partisanship
    der_partisanship = case_when(
      der_pid == "Party ID reported" ~ "1_pid_expl",
      der_pid == "No Party ID" & der_vote_cat == "Vote reported" ~ "2_pid_impl",
      der_pid == "No Party ID" & der_vote_cat == "Don't know" ~ "3_nopid",
      der_pid == "No Party ID" & der_vote_cat == "NA" ~ "3_nopid",
      der_pid == "NA" & der_vote_cat == "Vote reported" ~ "2_pid_impl",
      der_pid == "NA" & der_vote_cat == "Don't know" ~ "3_nopid",
      der_pid == "NA" & der_vote_cat == "NA" ~ "3_nopid"
    )
  )

# now let's create a new co/outpartisan variable including implicit partisans too
eu25games2019 <- eu25games2019 |>
  mutate(
    der_outpartisan_comb = case_when(
      # === EXPLICIT PID =======================================================
      # case 1: pid and control group (implies conational)
      der_partisanship == "1_pid_expl" &
        cj_treatment == "4_somenat" &
        cj_nationality_shown == "own_country" ~ "1_control_expl",
      # case 2: pid and copartisan (same)
      der_partisanship == "1_pid_expl" &
        str_detect(cj_treatment, "partisan") &
        (ext_cj_party_pf_name == ext_q_party_id_pf_name) ~ "2_co_expl",
      # case 3: pid and outpartisan
      der_partisanship == "1_pid_expl" &
        str_detect(cj_treatment, "partisan") &
        (ext_cj_party_pf_name != ext_q_party_id_pf_name) ~ "3_out_expl",
      # case 4: pid and eu outnational cj (hence no party shown)
      der_partisanship == "1_pid_expl" &
        !str_detect(cj_treatment, "partisan") &
        str_detect(cj_treatment, "eunat") ~ "97_eunat_expl",
      # case 5: pid and noneu outnational cj (hence no party shown)
      der_partisanship == "1_pid_expl" &
        !str_detect(cj_treatment, "partisan") &
        str_detect(cj_treatment, "somenat") ~ "98_outnat_expl",

      # === IMPLICIT PID =======================================================
      # case 4: pid and control group
      der_partisanship == "2_pid_impl" &
        cj_treatment == "4_somenat" &
        cj_nationality_shown == "own_country" ~ "4_control_impl",
      # case 5: impl pid and copartisan
      der_partisanship == "2_pid_impl" &
        str_detect(cj_treatment, "partisan") &
        (ext_cj_party_pf_name == der_vote_combined_name) ~ "5_co_impl",
      # case 6: pid and outpartisan
      der_partisanship == "2_pid_impl" &
        str_detect(cj_treatment, "partisan") &
        (ext_cj_party_pf_name != der_vote_combined_name) ~ "6_out_impl",
      # case 4: pid and eu outnational cj (hence no party shown)
      der_partisanship == "2_pid_impl" &
        !str_detect(cj_treatment, "partisan") &
        str_detect(cj_treatment, "eunat") ~ "97_eunat_impl",
      # case 5: pid and noneu outnational cj (hence no party shown)
      der_partisanship == "2_pid_impl" &
        !str_detect(cj_treatment, "partisan") &
        str_detect(cj_treatment, "somenat") ~ "98_outnat_impl",

      # === NO PID =============================================================
      der_partisanship == "3_nopid" &
        cj_treatment == "4_somenat" &
        cj_nationality_shown == "own_country" ~ "7_control_nopid",
      der_partisanship == "3_nopid" &
        str_detect(cj_treatment, "partisan") ~ "8_ptycue_nopid",
      der_partisanship == "3_nopid" &
        cj_nationality_shown == "eu" ~ "99_outnatEU_nopid",
      der_partisanship == "3_nopid" &
        cj_nationality_shown == "non_eu" ~ "99_outnat_nopid",
      .default = NA
    )
  )
```


# Generate Key Variables in the Framework: Partisan Type (T), Partisan Anchor (A), Relationship Category (R)

## Partisan Type $T$

Let partisan type $T$ denote the type of subjective attachement reported by respondents, with $T=1$ if they do feel attached to any political party, and $T=0$ if they do *not* feel attached to a party but revealed a partisan preference in voting items.

```{r}
#| label: partisan-type

eu25games2019 <- eu25games2019 |>
  mutate(
    der_partisan_type = case_when(
      der_partisanship == "1_pid_expl" ~ 1,
      der_partisanship == "2_pid_impl" ~ 0,
      .default = NA
    )
  )

```

## Partisan Anchor $A$

The anchor is keyed on `der_partisanship` rather than restated in terms of `der_pid` and
`der_vote_cat`. Restating the conditions is how the two definitions previously drifted
apart: `der_partisanship` counts a respondent whose attachment item is missing but whose
vote is reported as an implicit partisan, whereas the old anchor rule matched only
`der_pid == "No Party ID"` and so left those respondents anchorless (11 respondents in
the current sample). Deriving $A$ from $T$'s own input makes the two consistent by
construction.

```{r}
#| label: partisan-anchor

eu25games2019 <- eu25games2019 |>
  mutate(
    der_partisan_anchor = case_when(
      # explicit partisans are anchored to the party they report feeling close to
      der_partisanship == "1_pid_expl" ~ ext_q_party_id_pf_name,
      # vote-anchored partisans to the party they report voting for
      der_partisanship == "2_pid_impl" ~ der_vote_combined_name
    )
  )

# how many anchors in data?
eu25games2019 |>
  distinct(der_partisan_anchor) |>
  na.omit() |>
  nrow()

# same if grouped by country?
eu25games2019 |>
  group_by(meta_country) |>
  distinct(der_partisan_anchor) |>
  na.omit() |>
  nrow()
```


## Partisan Relationship Category $R$

```{r}
#| label: partisan-relationship-category

eu25games2019 <- eu25games2019 |>
  mutate(
    der_partisan_relationship = case_when(
      # control group (no partisan cue but still conational)
      cj_treatment == "4_somenat" &
        cj_nationality_shown == "own_country" ~ "None",
      # copartisan
      der_partisan_anchor == ext_cj_party_pf_name ~ "Co",
      # outpartisan
      der_partisan_anchor != ext_cj_party_pf_name ~ "Out"
    ),
    der_partisan_relationship = fct_relevel(der_partisan_relationship, "None")
  )
```


# More wrangling worth to be saved

```{r}
#| label: final-wrangling

eu25games2019 <- eu25games2019 |>
  mutate(
    meta_game_lab = case_when(
      meta_game == "dict" ~ "Dictator",
      meta_game == "trust" ~ "Trust",
      .default = NA
    )
  )

```

# Verify the framework variables

The three framework variables are what every later result rests on, so check their joint
distribution rather than trusting the `case_when()` chains silently.

```{r}
#| label: verify-framework-variables

# T x R, restricted to the co-national profiles the design compares
eu25games2019 |>
  filter(cj_nationality_shown == "own_country", !is.na(der_partisan_type)) |>
  count(meta_game, der_partisan_type, der_partisan_relationship)

stopifnot(
  # partisan type is defined exactly for explicit and implicit partisans
  all(
    is.na(eu25games2019$der_partisan_type) ==
      !(eu25games2019$der_partisanship %in% c("1_pid_expl", "2_pid_impl"))
  ),
  # every partisan carries an anchor
  all(!is.na(
    eu25games2019$der_partisan_anchor[!is.na(eu25games2019$der_partisan_type)]
  )),
  # anchors are real parties, never Party Facts sentinels
  !any(
    eu25games2019$der_partisan_anchor %in% c(pf_no_party, pf_other),
    na.rm = TRUE
  ),
  # both games survive, three rounds each
  setequal(eu25games2019$meta_game, c("dict", "trust")),
  setequal(eu25games2019$meta_round, 1:3)
)
```

# Save dataset ready for estimation

```{r}
#| label: save-data-2model

saveRDS(
  eu25games2019,
  file = here(
    "data",
    "03_final",
    "eu25games2019.rds"
  )
)

```


# Session Info

```{r}
#| label: session-info

session_info()
```


# Render Time

```{r}
#| label: render-time

end_time <- Sys.time()

rendering_time <- end_time - start_time

message(paste(
  "Document rendered in:",
  round(as.numeric(rendering_time, units = "secs"), 2),
  "seconds.
"
))
```


  
